### 2016-01-13 00-abstract.tex
Initial commit

% 1. what is the field of your work?
% 2. what is the problem you solve?
% 3. what is your solution/approach?
% 4. what are your results?

### 2016-01-13 01-intro.tex
Initial commit

\section{Introduction}
\cite{DBLP:journals/corr/LiAW15}

### 2016-01-13 04-evaluation.tex
Initial commit

\section{Evaluation}

\subsection{ml-curves}
Show convergence of different data sets
\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/ppx-amazon.eps, width=\columnwidth}
  %\caption{}
  \label{fig-ppx-cpu}
\end{figure}

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/ppx-dblp.eps, width=\columnwidth}
  %\caption{}
  \label{fig-ppx-cpu}
\end{figure}

  \begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/ppx-friendster.eps, width=\columnwidth}
  %\caption{}
  \label{fig-ppx-cpu}
\end{figure}

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/ppx-lj.eps, width=\columnwidth}
  %\caption{}
  \label{fig-ppx-cpu}
\end{figure}

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/ppx-youtube.eps, width=\columnwidth}
  %\caption{}
  \label{fig-ppx-cpu}
\end{figure}

\subsection{sweep-over-K-fixed-np}
\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-fixed-np.eps, width=\columnwidth}
  %\caption{}
  \label{fig-ppx-cpu}
\end{figure}

\subsection{sweep-over-K-proportional-np}
Weak scaling

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\columnwidth}
  %\caption{}
  \label{fig-ppx-cpu}
\end{figure}

\subsection{sweep-over-np-fixed-K}
Strong scaling

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\columnwidth}
  %\caption{}
  \label{fig-ppx-cpu}
\end{figure}

\subsection{hpc-cloud}
Scale up vs scale out
\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/hpc-cloud.eps, width=\columnwidth}
  %\caption{}
  \label{fig-ppx-cpu}
\end{figure}

### 2016-01-13 plots/plot-hpc-cloud.sh

@@ -1,0 +1,29 @@
+#!/bin/bash
+
+grep iteration ../data/hpc-cloud/* \
+  | grep -v minibatch \
+  | sed 's/.*-K\([0-9]\+\).*:/\1/' \
+  | awk '{print "echo " $1 " " $3 " $((16#" $4 "))"}' \
+  | bash \
+  | sort -n >> $$.tmp
+
+cat <<EOF | gnuplot --persist
+
+set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+
+set xlabel "Number of Communities (K)"
+set ylabel "Time (seconds)"
+#set xrange [256:]
+set nokey
+set grid ytics
+
+
+set output 'hpc-cloud.eps'
+set title "Scaling K"
+plot "$$.tmp" u 1:2 w l
+
+EOF
+
+cat $$.tmp
+
+rm $$.tmp

### 2016-01-13 plots/plot-ml-curves.sh

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+cat <<EOF | gnuplot --persist
+
+set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+
+set xlabel "Time (seconds)"
+set ylabel "Perplexity"
+set xrange [0:]
+set nokey
+set grid ytics
+
+
+set output 'ppx-amazon.eps'
+set title "com-Amazon | K75149-m1K-n32-msx1K-np24"
+plot '< grep average_count ../data/ml-curves/log-amazon-K75149-m1K-n32-msx1K-epst1e-04-np24/log.0' u 8:(exp(\$14)) w l
+
+set output 'ppx-lj.eps'
+set title "com-LiveJournal | K192K-flt32-m1K-n32-mxs1K-x1M-np65"
+plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-np65/log.0' u 8:(exp(\$14)) w l
+
+set output 'ppx-dblp.eps'
+set title "com-DBLP | K13477-flt32-m1K-n32-msx4K-x10M-np24"
+plot '< grep average_count ../data/ml-curves/log-dblp-K13477-flt32-m1K-n32-msx1K-x10M-epst1e-05-np24/log.0' u 8:(exp(\$14)) w l
+
+set output 'ppx-friendster.eps'
+set title "com-Friendster | K12K-flt32-m16K-n32-mxs-16K-np65"
+plot '< grep average_count ../data/ml-curves/log-Friendster-K12K-flt32-m16K-n32-mxs-16K-np65/log.0' u 8:(exp(\$14)) w l
+
+set output 'ppx-youtube.eps'
+set title "com-Youtube | K8385-m1K-n32-msx1K-np15"
+plot '< grep average_count ../data/ml-curves/log-youtube-K8385-m1K-n32-msx1K-np15/log.0' u 8:(exp(\$14)) w l
+EOF

### 2016-01-13 plots/plot-sweep-K-fixed-np.sh

@@ -1,0 +1,41 @@
+#!/bin/bash
+
+grep iteration ../data/sweep-over-K-fixed-np-single-buffer/*/* \
+  | grep -v minibatch \
+  | grep -v K10 \
+  | sed 's/.*-K\([0-9K]*\).*log.0:/\1/' \
+  | awk '{print $1 " " $3 " " $4}' \
+  | sed 's/\([0-9]\)K/$[\1*1024]/' \
+  | sed 's/\(.*\)/echo \1/' \
+  | bash \
+  | sort -n > $$.tmp
+
+grep iteration ../data/sweep-over-K-fixed-np/*/* \
+  | grep -v minibatch \
+  | grep -v K10 \
+  | sed 's/.*-K\([0-9K]*\).*log.0:/\1/' \
+  | awk '{print $1 " " $3 " " $4}' \
+  | sed 's/\([0-9]\)K/$[\1*1024]/' \
+  | sed 's/\(.*\)/echo \1/' \
+  | bash \
+  | sort -n > $$.tmp2
+
+cat <<EOF | gnuplot --persist
+
+set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+
+set xlabel "Number of Communities (K)"
+set ylabel "Time (seconds)"
+#set xrange [256:]
+set grid ytics
+
+
+set output 'sweep-over-K-fixed-np.eps'
+set title "Scaling K (fixed np=65)"
+plot "$$.tmp" u 1:2 w l t "Single buffering", "$$.tmp2" u 1:2 w l t "Double buffering"
+
+EOF
+
+cat $$.tmp $$.tmp2
+
+rm $$.tmp $$.tmp2

### 2016-01-13 plots/plot-sweep-K-proportional-np.sh

@@ -1,0 +1,34 @@
+#!/bin/bash
+
+grep iteration ../data/sweep-over-K-proportional-np/*/* \
+  | grep -v minibatch \
+  | grep -v np5 \
+  | sed 's/.*-K\([0-9K]\+\).*-np\([0-9]*\).*log.0:/\1 \2/' \
+  | awk '{print $1 " " $2 " " $4 " " $5}' \
+  | sed 's/\([0-9]\)K/$[\1*1024]/' \
+  | sed 's/\(.*\)/echo \1/' \
+  | bash \
+  | sort -n > $$.tmp
+
+cat $$.tmp
+
+cat <<EOF | gnuplot --persist
+
+set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+
+set ylabel "Time per Iteration (milliseconds)"
+set xlabel "Number of Nodes"
+set xrange [0:]
+set yrange[0:]
+set y2range[0:]
+set nokey
+set grid ytics
+
+set output 'sweep-over-K-proportional-np.eps'
+set title "Weak Scaling (over K)"
+plot "$$.tmp" u (\$2-1):(1000*\$3/\$4) w lp axis x1y1
+#, "" u (\$2-1):((\$1 * 4 * 65e6/1024**3) / ((\$2-1) * 64 * 1024**3)) w lp axis x1y2
+
+EOF
+
+rm $$.tmp

### 2016-01-13 plots/plot-sweep-np-fixed-K.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+grep iteration ../data/sweep-over-np-fixed-K/*/* \
+  | grep -v minibatch \
+  | sed 's/.*-K\([0-9K]\+\).*-np\([0-9]*\).*log.0:/\1 \2/' \
+  | awk '{print $1 " " $2 " " $4 " " $5}' \
+  | sed 's/\([0-9]\)K/$[\1*1024]/' \
+  | sed 's/\(.*\)/echo \1/' \
+  | bash \
+  | sort -n -k 2 > $$.tmp
+
+BASE=`head -n 1 $$.tmp | awk '{print $3}'`
+
+cat $$.tmp
+
+cat <<EOF | gnuplot --persist
+
+set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+
+set ylabel "Time (Seconds)"
+set xlabel "Number of Nodes"
+set xrange [0:]
+set yrange[0:]
+set y2range[0:]
+set y2tics
+set nokey
+set grid ytics
+
+set output 'sweep-over-np-fixed-K.eps'
+set title "Strong Scaling"
+plot "$$.tmp" u (\$2-1):3 w lp axis x1y1
+#  "" u (\$2-1):(${BASE}/\$3) w lp axis x1y2 t "Speedup over 8-nodes"
+
+EOF
+
+rm $$.tmp

### 2016-01-13 04-evaluation.tex
Added evaluation intro, DAS5 description and fixed some captions

@@
 
-\subsection{ml-curves}
-Show convergence of different data sets
-\begin{figure}[t] % [htb]
+This section presents an in-depth empirical analysis of the design discussed in
+Section~\ref{design-section}. To this end, we assess the scalability and
+efficiency of the solution using multiple criteria.
+%
+First, an evaluation of the strong scaling behavior of the distributed
+algorithm is presented.
+%
+Second, the weak scaling of the solution is assessed with respect to increasing
+numbers of targeted latent communities.
+%
+Next, the effects of varying the number of latent communities on the system's
+performance is analyzed.
+%
+Further, we evaluate the efficiency and overhead associated with network
+communication between cluster nodes.
+%
+Finally, we contrast the effectiveness of scaling the computation horizontally
+and vertically.
+
+Empirical results were obtained by performing experiments on the VU and TU/D
+DAS5 clusters which consist of 68 and 48 compute nodes respectively. Each
+compute node is equipped with a dual 8-core Intel Xeon E5-2630v3 CPU clocked
+at 2.40GHz, 64GB of memory and 8TB of storage. Moreover, the compute nodes of
+each site are interconnected by an internal 1~Gbit/s Ethernet and FDR~InfiniBand.
+
+\subsection{Strong scaling}
+
+\begin{figure}[ht] % [htb]
   \centering
-  \epsfig{file=plots/ppx-amazon.eps, width=\columnwidth}
-  %\caption{}
-  \label{fig-ppx-cpu}
+  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\columnwidth}
+  \caption{Strong scaling: varying the number of compute nodes for a fixed
+  problem size.}
+  \label{fig-strong-scaling}
 \end{figure}
 
-\begin{figure}[t] % [htb]
-  \centering
-  \epsfig{file=plots/ppx-dblp.eps, width=\columnwidth}
-  %\caption{}
-  \label{fig-ppx-cpu}
-\end{figure}
-
-  \begin{figure}[t] % [htb]
-  \centering
-  \epsfig{file=plots/ppx-friendster.eps, width=\columnwidth}
-  %\caption{}
-  \label{fig-ppx-cpu}
-\end{figure}
-
-\begin{figure}[t] % [htb]
-  \centering
-  \epsfig{file=plots/ppx-lj.eps, width=\columnwidth}
-  %\caption{}
-  \label{fig-ppx-cpu}
-\end{figure}
-
-\begin{figure}[t] % [htb]
-  \centering
-  \epsfig{file=plots/ppx-youtube.eps, width=\columnwidth}
-  %\caption{}
-  \label{fig-ppx-cpu}
-\end{figure}
-
-\subsection{sweep-over-K-fixed-np}
-\begin{figure}[t] % [htb]
-  \centering
-  \epsfig{file=plots/sweep-over-K-fixed-np.eps, width=\columnwidth}
-  %\caption{}
-  \label{fig-ppx-cpu}
-\end{figure}
-
-\subsection{sweep-over-K-proportional-np}
-Weak scaling
+\subsection{Weak Scaling}
 
@@
   \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\columnwidth}
-  %\caption{}
+  \caption{Weak scaling: varying the number of compute nodes proportionally to
+  the number of latent communities.}
+  \label{fig-weak-scaling}
+\end{figure}
+
+\subsection{Effects of Number of Communities on Performance}
+\begin{figure}[t] % [htb]
+  \centering
+  \epsfig{file=plots/sweep-over-K-fixed-np.eps, width=\columnwidth}
+  \caption{Effects of varying the number of communities on the algorithm's
+  throughput.}
+  \label{fig-throughput}
+\end{figure}
+
+\subsection{Horizontal vs. Vertical Scalability}
+\begin{figure}[t] % [htb]
+  \centering
+  \epsfig{file=plots/hpc-cloud.eps, width=\columnwidth}
+  \caption{FIXME}
   \label{fig-ppx-cpu}
@@
 
-\subsection{sweep-over-np-fixed-K}
-Strong scaling
+\subsection{Cluster Communication Efficiency}
 
-\begin{figure}[t] % [htb]
+\subsection{Convergence of Large Datasets}
+\begin{figure*}[t] % [htb]
   \centering
-  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\columnwidth}
-  %\caption{}
+  \epsfig{file=plots/ppx.eps, width=\textwidth}
+  \caption{FIXME}
   \label{fig-ppx-cpu}
-\end{figure}
-
-\subsection{hpc-cloud}
-Scale up vs scale out
-\begin{figure}[t] % [htb]
-  \centering
-  \epsfig{file=plots/hpc-cloud.eps, width=\columnwidth}
-  %\caption{}
-  \label{fig-ppx-cpu}
-\end{figure}
+\end{figure*}

### 2016-01-13 plots/plot-ml-curves.sh

@@ -2,32 +2,56 @@
 
 cat <<EOF | gnuplot --persist
 
-set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+#set terminal postscript eps enhanced color font ',8' size 3.3,1.6
 
 set xlabel "Time (seconds)"
 set ylabel "Perplexity"
-set xrange [0:]
+#set xrange [0:]
 set nokey
 set grid ytics
 
+w = 0.6
+h = 0.4
+k = 0.0
+nr = 3
+nc = 2
+row(x) = ((nr-x-1)*h)+k
+col(x) = x*w
 
-set output 'ppx-amazon.eps'
-set title "com-Amazon | K75149-m1K-n32-msx1K-np24"
+set terminal postscript eps enhanced color font ',8'
+set size (nc*w),(nr*h+k)
+set output 'ppx.eps'
+set multiplot layout nr,nc
+set size w,h
+
+
+set origin col(0),row(0)
+set size w,h
+#set output 'ppx-amazon.eps'
+set title "(a) com-Amazon | K75149-m1K-n32-msx1K-np24"
 plot '< grep average_count ../data/ml-curves/log-amazon-K75149-m1K-n32-msx1K-epst1e-04-np24/log.0' u 8:(exp(\$14)) w l
 
-set output 'ppx-lj.eps'
-set title "com-LiveJournal | K192K-flt32-m1K-n32-mxs1K-x1M-np65"
+set origin col(1),row(0)
+set size w,h
+#set output 'ppx-lj.eps'
+set title "(b) com-LiveJournal | K192K-flt32-m1K-n32-mxs1K-x1M-np65"
 plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-np65/log.0' u 8:(exp(\$14)) w l
 
-set output 'ppx-dblp.eps'
-set title "com-DBLP | K13477-flt32-m1K-n32-msx4K-x10M-np24"
+set origin col(0),row(1)
+set size w,h
+#set output 'ppx-dblp.eps'
+set title "(c) com-DBLP | K13477-flt32-m1K-n32-msx4K-x10M-np24"
 plot '< grep average_count ../data/ml-curves/log-dblp-K13477-flt32-m1K-n32-msx1K-x10M-epst1e-05-np24/log.0' u 8:(exp(\$14)) w l
 
-set output 'ppx-friendster.eps'
[…]

### 2016-01-13 04-evaluation.tex
Strong scaling section

@@
 
+\begin{table*}
+  \centering
+  \begin{tabular}{l r r l}
+    Name            & \#Vertices &       \#Edges & Description \\
+    \hline
+    com-LiveJournal &  3,997,962 &    34,681,189 & Online blogging social network \\
+    com-Friendster  & 65,608,366 & 1,806,067,135 & Online gaming social network \\
+    com-Orkut       &  3,072,441 &   117,185,083 & Online social network \\
+    com-Youtube     &  1,134,890 &     2,987,624 & Video-sharing social network \\
+    com-DBLP        &    317,080 &     1,049,866 & Computer science bibliography collaboration network \\
+    com-Amazon      &    334,863 &       925,872 & Product co-purchasing network \\
+    \hline
+  \end{tabular}
+  \label{tab-datasets}
+  \caption{Summary of SNAP graph data sets used for evaluation.}
+\end{table*}
+
 Empirical results were obtained by performing experiments on the VU and TU/D
@@
 at 2.40GHz, 64GB of memory and 8TB of storage. Moreover, the compute nodes of
-each site are interconnected by an internal 1~Gbit/s Ethernet and FDR~InfiniBand.
+each site are interconnected by an internal 1~Gbit/s Ethernet and
+FDR~InfiniBand. All of the experiments reported in this section utilized
+publicly available graphs from the Stanford Large Network Dataset
+Collection~(SNAP) \cite{snapnets}. Table~\ref{tab-datasets} lists the
+collection of data sets used.
 
+%################################################
 \subsection{Strong scaling}
 
-\begin{figure}[ht] % [htb]
+\begin{figure*}[t] % [htb]
   \centering
-  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\columnwidth}
-  \caption{Strong scaling: varying the number of compute nodes for a fixed
-  problem size.}
+  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\textwidth}
+  \caption{Strong scaling: Execution time of 2048 algorithm iterations for the
+  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
+  cluster sizes.}
   \label{fig-strong-scaling}
-\end{figure}
+\end{figure*}
 
+In order to evaluate the horizontal scalability of the distributed
+implementation we tested the system's performance across different cluster
+sizes while holding the problem size constant. For this study, we used the
+com-Friendster graph as it contains the largest number of vertices and edges.
+%
+Figure~\ref{fig-strong-scaling}-a presents the execution time of 2048 algorithm
+iterations across multiple cluster sizes. As shown in the figure, the execution
+time steadily decreases by increasing the cluster size.
+A deeper analysis of the individual execution phases of the algorithm provides
+insights into the scalability if its building blocks.
+%
+In addition to the total execution time, Figure~\ref{fig-strong-scaling}-a
+presents the cumulative time spent in individual computational phases across
+iterations. Moreover, Figure~\ref{fig-strong-scaling}-b presents the speedup
+achieved for the same experiments reported in
+Figure~\ref{fig-string-scaling}-a. As is clearly shown, the dominant phase of
+the execution is update\_pi.
+The reported total time for each cluster size is significantly less than the
+sum of the execution times of the individual phases. This is due to the
+overlapping execution of the two most expensive phases, namely, update\_pi and
+deploy\_minibatch. Both of these phases initially gain significant speedup with the
+addition of compute nodes. However, the speedup curve gradually slows down for
+larger cluster sizes as the work granularity of each worker node decreases
+limiting their resource utilization.
+%
+The time spent in update\_beta remains relatively constant across cluster
+sizes as it is an insignificant amount of work compared to the synchronization
+overhead required to perform a collective MPI reduction operation to complete
+it.
+
+%################################################
 \subsection{Weak Scaling}

### 2016-01-13 plots/plot-sweep-np-fixed-K.sh

@@ -9,28 +9,79 @@ grep iteration ../data/sweep-over-np-fixed-K/*/* \
   | bash \
   | sort -n -k 2 > $$.tmp
 
+grep "deploy minibatch" ../data/sweep-over-np-fixed-K/*/* \
+  | sed 's/.*-K\([0-9K]\+\).*-np\([0-9]*\).*log.0:/\1 \2/' \
+  | awk '{print $1 " " $2 " " $5}' \
+  | sed 's/\([0-9]\)K/$[\1*1024]/' \
+  | sed 's/\(.*\)/echo \1/' \
+  | bash \
+  | sort -n -k 2 > $$.tmp.2
+
+grep "update_phi_pi" ../data/sweep-over-np-fixed-K/*/* \
+  | sed 's/.*-K\([0-9K]\+\).*-np\([0-9]*\).*log.0:/\1 \2/' \
+  | awk '{print $1 " " $2 " " $4}' \
+  | sed 's/\([0-9]\)K/$[\1*1024]/' \
+  | sed 's/\(.*\)/echo \1/' \
+  | bash \
+  | sort -n -k 2 > $$.tmp.3
+
+grep "update_beta_theta" ../data/sweep-over-np-fixed-K/*/* \
+  | sed 's/.*-K\([0-9K]\+\).*-np\([0-9]*\).*log.0:/\1 \2/' \
+  | awk '{print $1 " " $2 " " $4}' \
+  | sed 's/\([0-9]\)K/$[\1*1024]/' \
+  | sed 's/\(.*\)/echo \1/' \
+  | bash \
+  | sort -n -k 2 > $$.tmp.4
+
 BASE=`head -n 1 $$.tmp | awk '{print $3}'`
+BASE2=`head -n 1 $$.tmp.2 | awk '{print $3}'`
+BASE3=`head -n 1 $$.tmp.3 | awk '{print $3}'`
+BASE4=`head -n 1 $$.tmp.4 | awk '{print $3}'`
 
 cat $$.tmp
 
 cat <<EOF | gnuplot --persist
 
-set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+w = 0.6
+h = 0.4
+k = 0.0
+nr = 1
+nc = 2
+row(x) = ((nr-x-1)*h)+k
+col(x) = x*w
 
-set ylabel "Time (Seconds)"
+set terminal postscript eps enhanced color font ',8'
+set size (nc*w),(nr*h+k)
+set output 'sweep-over-np-fixed-K.eps'
[…]

### 2016-01-14 03-design.tex
Unripe design text

\section{System Design}
\label{design-section}

\begin{comment}
- distributed parallelisation:
  cut up into stages like the CC-Grid paper:
  pipeline:
    1) sample minibatch
    2) sample neighbors
    3) update_phi
  4) update_pi
  5) update_theta
  5a) calc_grads
  5b) sum_grads
  5c) update_theta
  6) update_beta
  7) perplexity

- DKV RDMA store

- data dependencies:
  mb = nodes in the minibatch
  nb = nodes in the neighbor sets
  H  = held-out set

  	pi	phi	beta	theta	grad.th	G	prev.minibatch
  1) 	-	-	-	-		R	-
  2) 	-	-	-	-		R	R
  3) 	R(mb)	R	R	-		R	-
	R(nb)
  4)	R/W(mb)	R	-	-		-	-
  5a)	R(mb)	-	R	R	W	-	-
  5b)	-	-	-	-	R/W	-	-
  5c)	-	-	-	R/W	R	-	-
  6)	-	-	W	R	-	-	-
  7)	R(H)	-	R(H)	-	-	H	-

- parallelization:
  1) sequential at master's, multithreaded
  2) parallel over mb; at workers', multithreaded
  3) parallel over mb; at workers', multithreaded
  4) parallel over mb; at workers', multithreaded
  5a) parallel at workers', multithreaded
  5b) omp_reduce(+) at workers', MPI reduce global
  5c) at master's, multithreaded
  6) broadcast theta by master, parallel update at workers'
  7) parallel over H; at workers', multithreaded
     reduce(+) at workers, MPI reduce global

- distributed parallelization:
  trade memory size for communication costs
  pity: the algorithm is already memory-dominant

- synchronization:
  in principle, each step synchronized
  but: pipelining in sample minibatch / load pi/phi / update_phi

- data sizes:
  G has N vertices and |E| undirected edges
  |H| is 1..10% of |E|
  pi, phi, beta, theta are float32
  pi[N][K + 1] if phi folded into pi
  beta[K]
  theta[K][2]
  so, pi is the memory eater
   . beta/theta are replicated
   . phi is folded into pi
   . pi is partitioned across the workers
  G is stored at the master only: save the memory for pi so we can handle
  larger K. The relevant slices of G (m's adjacency lists) are scattered
  together with the minibatch.
  G is stored as a Google SparseHash because c++ std::(unordered_)set has a
  huge overhead per item. Workers unpack their slice of G into a
  std::unordered_set because that is somewhat faster.

- pi storage: DKV
   . pi/phi is read in 3), 4), 5a), 7)
   . pi is written in 4); read[i], then write[i], so no other data dependencies
   . so access is very synchronous: either read-only, or write without
     concurrent readers
   . DKV store properties:
      - single-sized keys
      - contiguous key space (integers 0..N-1)
      - update-only writes: no new KVs, no deletes
     so: no load imbalance, no hashing that is worth its name
      - no read/write or write/write concurrency
     so:
      - build RDMA store with remote reads and remote writes only, with exactly
        one RDMA transaction per read or write
\end{comment}

% rationale -- here or in the intro?

% http://stanford.edu/~rezab/nips2014workshop/slides/jure.pdf

The SC-MCMC algorithm described in the previous section has an abundance of
opportunities for parallelism, for the multithreaded, shared-memory type as
well as for distributed-memory parallelism. The benefit of multithreaded
parallelism is speed-up of the computation. A distributed implementation
additionaly allows to store data in the collective memory of the cluster
machine and increases memory bandwidth; these both scale with the number of
machines. The downside of a distributed implementation is that it requires
considerably more programming effort.

% intro parallelisation, distribution

In this section, we describe how we parallelize each of the stages of the
algorithm's main loop. In most places, the usage of multithreaded parallelism
is straightforward; we will discuss details only where necessary. The
distributed design follows a master-worker paradigm, where in essence
the master controls the parallel operations and the workers perform the
calculations. For thread parallelism, we annotate the program with
OpenMP~\cite{OpenMP}. For distributed communication, we use MPI~\cite{MPI}.

% largest data structures

\subsection{Data distribution}

The largest data structures of the algorithm are $G$, $\pi$ and~$\phi$.
For the largest dataset in this paper, com-Friendster, $G$ has 1.8~billion
undirected edges. In our representation with directed edges, this takes
up 13.5GB. Our design lets $G$ reside only at the master. We observe that
the calculations in the update stages only require the subset of $G$ that
is accessed by the minibatch nodes, so the master scatters that subset to
the workers together with the scattering of the minibatch nodes. $\pi$~and
$\phi$ are float arrays of size $K \times N$. For our largest distributed
experiment, com-Friendster with $N$=64M and $K$=12K, each takes up~3TB. We
decided to store only $\pi$ and $\pi^{sum}$, and recalculate $\phi$ from these
whenever appropriate. $\pi$ is partitioned across the workers, and $\pi$
values are accessed via a DKV (distributed key-value) store. In both our
decisions to transport a subset of $G$ in each iteration and to recalculate
$\phi$, we gain a larger available memory at the cost of communication and
computation.

% lack of locality

{\Large\bf Discuss the lack of locality}

% each of the stages

\subsection{Implementation of distributed parallelism}

Now, we describe the parallelization of each of the stages of the algorithm.

Exceptionally, the first stage, minibatch selection, is not itself
parallelized. However, in the distributed version its execution at the
master's is fully overlapped, using pipeline parallelism, while the workers are
performing \textit{update\_phi}. The minibatch is partitioned equally over the
workers; the relevant section of $G$ is distributed together with the
minibatch subsets.

After a worker has received its subset of the minibatch, it samples a
neighbor set for each of its minibatch nodes, using thread parallelism.

The next stage, \textit{update\_phi}, is the algorithm's dominant stage, not
only in calculation but also in memory accesses. It can be fully parallelized
because it is a data-parallel operation over each of the minibatch nodes.
\textit{update\_phi} first loads the $\pi$ values for its minibatch nodes and
their neighbors using the DKV store. The updates to $\phi$ for the minibatch
nodes are calculated independently using thread parallelism.

\textit{update\_pi} uses the value of $\pi/\phi$ for the minibatch nodes. The
update to $\pi$ can be done fully in parallel. Because of memory consistency,
this stage awaits completion of \textit{update\_phi} with an MPI barrier.
After the calculation, the updated values of $\pi/\pi^{sum}$ are written
through the DKV store.

\subsection{Pipelining of $\pi$ staging and computation}

\subsection{RDMA Distributed Key-Value Store}

### 2016-01-14 04-evaluation.tex
Weak scaling + some strong scaling edits

@@
   \end{tabular}
+  \caption{Summary of SNAP graph data sets used for evaluation.}
   \label{tab-datasets}
-  \caption{Summary of SNAP graph data sets used for evaluation.}
 \end{table*}
@@
   \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\textwidth}
-  \caption{Strong scaling: Execution time of 2048 algorithm iterations for the
+  \caption{(a) Execution time of 2048 algorithm iterations for the
   same problem size (com-Friendster, K=1024, M=16384 n=32) across different
-  cluster sizes.}
+  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
+  8-nodes}
   \label{fig-strong-scaling}
@@
 Figure~\ref{fig-strong-scaling}-a presents the execution time of 2048 algorithm
-iterations across multiple cluster sizes. As shown in the figure, the execution
+iterations across multiple cluster sizes. The x-axis starts from 8 worker nodes
+as the data set is too large to fit into the collective memory of a smaller
+cluster.
+%
+As shown in the figure, the execution
 time steadily decreases by increasing the cluster size.
@@
 achieved for the same experiments reported in
-Figure~\ref{fig-string-scaling}-a. As is clearly shown, the dominant phase of
+Figure~\ref{fig-strong-scaling}-a. As is clearly shown, the dominant phase of
 the execution is update\_pi.
@@
 %
-The time spent in update\_beta remains relatively constant across cluster
-sizes as it is an insignificant amount of work compared to the synchronization
-overhead required to perform a collective MPI reduction operation to complete
-it.
+The time spent in update\_beta remains relatively constant across cluster sizes
+as it performs an insignificant amount of work compared to the synchronization
+overhead of a collective MPI reduction operation contained within it.
 
@@
 
-\begin{figure}[t] % [htb]
+\begin{figure*}[t] % [htb]
   \centering
-  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\columnwidth}
-  \caption{Weak scaling: varying the number of compute nodes proportionally to
-  the number of latent communities.}
+  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\textwidth}
+  \caption{(a) Average execution time per algorithm iteration varying
+  the number of latent communities proportionally to the number of compute
+  nodes. (b) The exact number of communities used for each data point in (a).}
   \label{fig-weak-scaling}
-\end{figure}
+\end{figure*}
 
+A study of a system's weak scaling aids in the assessment of the communication
+and synchronization overheads associated with the management of large clusters.
+Similarly, it can expose complex issues that hinder performance at scale such
+as load imbalance. To fairly evaluate the algorithm's weak scaling behavior we
+conducted several executions varying the cluster size and the number of latent
+communities proportionally. This methodology ensures that each compute node
+performs a relatively constant amount of computational work across the all
+configurations. However, the number and size of messages exchanged between the
+nodes would vary significantly. Figure~\ref{fig-weak-scaling}-a presents the
+average execution
+time per algorithm iteration across different cluster sizes. On the other hand,
+Figure~\ref{fig-weak-scaling}-b reports the set number of communities for each
+cluster configuration. The relative change in the average execution time per
+iteration is insignificant even though the communication intensity increases
+for larger cluster sizes. This observation verifies that the system's overall
+overhead is minimal and scales well.
+
+%################################################
 \subsection{Effects of Number of Communities on Performance}
+
 \begin{figure}[t] % [htb]
@@
   \epsfig{file=plots/sweep-over-K-fixed-np.eps, width=\columnwidth}
-  \caption{Effects of varying the number of communities on the algorithm's
-  throughput.}
-  \label{fig-throughput}
+  \caption{Performance effect of varying the number of communities on the
+    algorithm's execution time on 64-nodes.}
+  \label{fig-k}
 \end{figure}
@@
   \caption{FIXME}
-  \label{fig-ppx-cpu}
+  \label{fig-scale-up}
 \end{figure}
@@
   \caption{FIXME}
-  \label{fig-ppx-cpu}
+  \label{fig-ppx}
 \end{figure*}

### 2016-01-14 plots/plot-sweep-K-proportional-np.sh

@@ -10,24 +10,60 @@ grep iteration ../data/sweep-over-K-proportional-np/*/* \
   | bash \
   | sort -n > $$.tmp
 
+ls ../data/sweep-over-K-proportional-np/*/* \
+  | grep -wv np5 \
+  | sed 's/.*-K\([0-9K]\+\).*-np\([0-9]\+\).*/\2 \1/' \
+  | sed 's/\([0-9]\+\)K/$[\1*1024]/' \
+  | sed 's/\(.*\)/echo \1/' \
+  | bash \
+  | sort -n \
+  | awk '{print $1-1 " " $2}' > $$.tmp.2
+
 cat $$.tmp
 
 cat <<EOF | gnuplot --persist
 
-set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+w = 0.6
+h = 0.4
+k = 0.0
+nr = 1
+nc = 2
+row(x) = ((nr-x-1)*h)+k
+col(x) = x*w
 
-set ylabel "Time per Iteration (milliseconds)"
+set terminal postscript eps enhanced color font ',8'
+set size (nc*w),(nr*h+k)
+set output 'sweep-over-K-proportional-np.eps'
+set multiplot layout nr,nc
+set size w,h
+
 set xlabel "Number of Nodes"
 set xrange [0:]
-set yrange[0:]
-set y2range[0:]
+set yrange[0:500]
 set nokey
 set grid ytics
 
-set output 'sweep-over-K-proportional-np.eps'
-set title "Weak Scaling (over K)"
+set origin col(0),row(0)
+set size w,h
+set title "(a) Weak Scaling (over K)"
+set ylabel "Time per Iteration (milliseconds)"
 plot "$$.tmp" u (\$2-1):(1000*\$3/\$4) w lp axis x1y1
-#, "" u (\$2-1):((\$1 * 4 * 65e6/1024**3) / ((\$2-1) * 64 * 1024**3)) w lp axis x1y2
+
+set origin col(1),row(0)
+set size w,h
+set style data histogram 
+set style histogram cluster gap 1
+
+set style fill solid noborder
+set auto x
+set ytics
+set mytics
+set grid ytics mytics
+
+set yrange[0:*]
+set title "(b) Number of communities used per cluster size"
+set ylabel "#K"
+plot "$$.tmp.2" u 2:xtic(1)
 
 EOF
 

### 2016-01-14 04-evaluation.tex
Pipeline efficiency + scale-up vs scale-out figure edit

@@
 %################################################
-\subsection{Effects of Number of Communities on Performance}
+\subsection{Pipeline efficiency}
 
@@
   \caption{Performance effect of varying the number of communities on the
-    algorithm's execution time on 64-nodes.}
-  \label{fig-k}
+    algorithm's execution time on 64-nodes when using single- or
+    double-buffering.}
+  \label{fig-pipeline}
 \end{figure}
+
+As discussed in Section~\ref{design-section}, the collective memory of all
+worker nodes serves as the storage for the state of the computation. As such,
+the rows of $\pi$ are equally and statically partitioned across all workers.
+Given that most $\pi$ accesses are random, a node in a cluster of size $C$ must
+fetch $(C-1)/C$ of all read requests over the network. Therefore, large cluster
+configurations exhibit higher bandwidth demands and are more sensitive
+to network latency. To reduce the negative effects of network latency on the
+computation, a pipelining scheme was devised to prefetch data dependencies.
+Figure~\ref{fig-pipeline} presents the execution time of 1024 algorithm
+iterations on a 64-node cluster with double-bufferring enabled and disabled.
+Naturally, increasing the number of communities causes a proportional increase
+in execution time. However, when double-buffering is enabled, some of the
+incurred network latency is hidden by overlapping it with computation.
+Moreover, since both computation time and network latency increase with larger
+$K$, the benefit of pipelining increases. This can be observed through the
+widening gap between both lines depicted in Figure~\ref{fig-pipeline}.
 

### 2016-01-14 plots/plot-hpc-cloud.sh

@@ -3,27 +3,39 @@
 grep iteration ../data/hpc-cloud/* \
   | grep -v minibatch \
   | sed 's/.*-K\([0-9]\+\).*:/\1/' \
-  | awk '{print "echo " $1 " " $3 " $((16#" $4 "))"}' \
+  | awk '{print "echo " $1 " " $3 " $((16#" $4 "))" }' \
   | bash \
-  | sort -n >> $$.tmp
+  | awk '{print $1 " " $2/$3}' \
+  | sort -n > $$.tmp
+
+
+grep iteration ../data/sweep-over-K-fixed-np/*/* \
+  | grep -v minibatch \
+  | sed 's/.*-K\([0-9K]\+\).*:/\1/' \
+  | sed 's/\([0-9]\+\)K/$[\1*1024]/' \
+  | sed 's/\(.*\)/echo \1/' \
+  | bash \
+  | awk '{print $1 " " $3/$4}' \
+  | sort -n \
+  | head -n 5 > $$.tmp.2
 
 cat <<EOF | gnuplot --persist
 
 set terminal postscript eps enhanced color font ',8' size 3.3,1.6
 
 set xlabel "Number of Communities (K)"
-set ylabel "Time (seconds)"
+set ylabel "Time per iteration (seconds)"
 #set xrange [256:]
-set nokey
 set grid ytics
+set key below
 
 
 set output 'hpc-cloud.eps'
-set title "Scaling K"
-plot "$$.tmp" u 1:2 w l
+set title "Scale-up vs Scale-out"
+plot "$$.tmp" u 1:(1000*\$2) w lp t 'HPC Cloud', "$$.tmp.2" u 1:(1000*\$2) w lp t '64-nodes on DAS5'
 
 EOF
 
-cat $$.tmp
+cat $$.tmp*
 
-rm $$.tmp
+rm $$.tmp*

### 2016-01-14 plots/plot-ml-curves.sh

@@ -13,8 +13,8 @@ set grid ytics
 w = 0.6
 h = 0.4
 k = 0.0
-nr = 3
-nc = 2
+nr = 2
+nc = 3
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
@@ -37,19 +37,19 @@ set size w,h
 set title "(b) com-LiveJournal | K192K-flt32-m1K-n32-mxs1K-x1M-np65"
 plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-np65/log.0' u 8:(exp(\$14)) w l
 
-set origin col(0),row(1)
+set origin col(2),row(0)
 set size w,h
 #set output 'ppx-dblp.eps'
 set title "(c) com-DBLP | K13477-flt32-m1K-n32-msx4K-x10M-np24"
 plot '< grep average_count ../data/ml-curves/log-dblp-K13477-flt32-m1K-n32-msx1K-x10M-epst1e-05-np24/log.0' u 8:(exp(\$14)) w l
 
-set origin col(1),row(1)
+set origin col(0),row(1)
 set size w,h
 #set output 'ppx-friendster.eps'
 set title "(d) com-Friendster | K12K-flt32-m16K-n32-mxs-16K-np65"
 plot '< grep average_count ../data/ml-curves/log-Friendster-K12K-flt32-m16K-n32-mxs-16K-np65/log.0' u 8:(exp(\$14)) w l
 
-set origin col(0),row(2)
+set origin col(1),row(1)
 set size w,h
 #set output 'ppx-youtube.eps'
 set title "(e) com-Youtube | K8385-m1K-n32-msx1K-np15"

### 2016-01-14 plots/plot-sweep-K-fixed-np.sh

@@ -32,7 +32,7 @@ set grid ytics
 
 set output 'sweep-over-K-fixed-np.eps'
 set title "Scaling K (fixed np=65)"
-plot "$$.tmp" u 1:2 w l t "Single buffering", "$$.tmp2" u 1:2 w l t "Double buffering"
+plot "$$.tmp" u 1:2 w lp t "Single buffering", "$$.tmp2" u 1:2 w lp t "Double buffering"
 
 EOF
 

### 2016-01-14 plots/plot-sweep-K-proportional-np.sh

@@ -67,4 +67,4 @@ plot "$$.tmp.2" u 2:xtic(1)
 
 EOF
 
-rm $$.tmp
+rm $$.tmp*

### 2016-01-14 plots/plot-ml-curves.sh
Import Orkut + graph generator

@@ -54,4 +54,10 @@ set size w,h
 #set output 'ppx-youtube.eps'
 set title "(e) com-Youtube | K8385-m1K-n32-msx1K-np15"
 plot '< grep average_count ../data/ml-curves/log-youtube-K8385-m1K-n32-msx1K-np15/log.0' u 8:(exp(\$14)) w l
+
+set origin col(2),row(1)
+set size w,h
+#set output 'ppx-orkut.eps'
+set title "(f) com-Orkut | K256K-m1K-n32-msx1K-np65"
+plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst1e-05-np65/log.0' u 8:(exp(\$14)) w l
 EOF

### 2016-01-14 03-design.tex
Work on the design section

@@
 
-In this section, we describe how we parallelize each of the stages of the
-algorithm's main loop. In most places, the usage of multithreaded parallelism
-is straightforward; we will discuss details only where necessary. The
+This section describes how we parallelize the algorithm.
+In most places, the usage of multithreaded parallelism
+is straightforward; we will discuss details only where appropriate. The
 distributed design follows a master-worker paradigm, where in essence
@@
 
-The largest data structures of the algorithm are $G$, $\pi$ and~$\phi$.
+The largest data structures of the algorithm are $G$, $H$, $\pi$ and~$\phi$.
 For the largest dataset in this paper, com-Friendster, $G$ has 1.8~billion
@@
 the calculations in the update stages only require the subset of $G$ that
-is accessed by the minibatch nodes, so the master scatters that subset to
-the workers together with the scattering of the minibatch nodes. $\pi$~and
-$\phi$ are float arrays of size $K \times N$. For our largest distributed
-experiment, com-Friendster with $N$=64M and $K$=12K, each takes up~3TB. We
-decided to store only $\pi$ and $\pi^{sum}$, and recalculate $\phi$ from these
-whenever appropriate. $\pi$ is partitioned across the workers, and $\pi$
-values are accessed via a DKV (distributed key-value) store. In both our
-decisions to transport a subset of $G$ in each iteration and to recalculate
-$\phi$, we gain a larger available memory at the cost of communication and
-computation.
+is touched by the minibatch nodes, so the master scatters that subset to
+the workers together with the scattering of the minibatch nodes. This way,
+we trade a reduction in memory at the workers against the cost of limited
+communication. In contrast, $H$ can be partitioned statically over all
+machines for the parallel perplexity calculation.
+
+$\pi$~and $\phi$ are 32-bit float arrays of size $K \times N$. For our
+largest distributed experiment, com-Friendster with $N$=64M and $K$=12K, each
+requires~3TB. We decided to store only $\pi$ and $\pi^{sum}$, and recalculate
+$\phi$ from these whenever appropriate. Here, we trade a substantial reduction
+in memory against limited computation. $\pi$ is partitioned across the workers,
+and $\pi$ values are accessed via a DKV (distributed key-value) store.
 
@@
 
-{\Large\bf Discuss the lack of locality}
+{\Large\bf NOTE: Discuss the lack of locality? Here?}
 
@@
 
-Now, we describe the parallelization of each of the stages of the algorithm.
+This section describes the parallelization of each of the stages of the
+algorithm's main loop. All but the first stage justify parallelisation for
+high values of combinations of the  parameters $M$, $K$, $n$, and $|H|$.
 
-Exceptionally, the first stage, minibatch selection, is not itself
-parallelized. However, in the distributed version its execution at the
-master's is fully overlapped, using pipeline parallelism, while the workers are
-performing \textit{update\_phi}. The minibatch is partitioned equally over the
-workers; the relevant section of $G$ is distributed together with the
+The first stage, minibatch selection, is done at the master's, and it is
+not itself parallelized. However, in the distributed implementation its
+execution is overlapped using pipeline parallelism, while the workers are
+performing \textit{update\_phi}. The minibatch is partitioned equally over
+the workers; the relevant sections of $G$ are distributed together with the
 minibatch subsets.
 
-After a worker has received its subset of the minibatch, it samples a
-neighbor set for each of its minibatch nodes, using thread parallelism.
+After a worker has received its subset of the minibatch, it samples a neighbor
+set of size $n$ for each of its minibatch nodes, using thread parallelism.
 
 The next stage, \textit{update\_phi}, is the algorithm's dominant stage, not
-only in calculation but also in memory accesses. It can be fully parallelized
-because it is a data-parallel operation over each of the minibatch nodes.
-\textit{update\_phi} first loads the $\pi$ values for its minibatch nodes and
-their neighbors using the DKV store. The updates to $\phi$ for the minibatch
-nodes are calculated independently using thread parallelism.
+only in calculation but also in memory accesses. It performs $M\times n\times
+K$ operations. It can be fully parallelized because it is a data-parallel
+operation over each of the minibatch nodes. \textit{update\_phi} loads the
+$\pi$ values for the local minibatch nodes and their neighbors from the
+DKV store. The updates to $\phi$ for the minibatch nodes are calculated
+independently using thread parallelism.
 
-\textit{update\_pi} uses the value of $\pi/\phi$ for the minibatch nodes. The
-update to $\pi$ can be done fully in parallel. Because of memory consistency,
-this stage awaits completion of \textit{update\_phi} with an MPI barrier.
-After the calculation, the updated values of $\pi/\pi^{sum}$ are written
-through the DKV store.
+These updates are used by \textit{update\_pi} locally, besides the value of
+$\pi/\phi$ for the minibatch nodes. The update to $\pi$ requires $M\times
+K$ operations, and it is done in parallel over minibatch nodes. Because of
+memory consistency, this stage awaits completion of \textit{update\_phi} with
+an MPI barrier. After the calculation, the updated values of $\pi+\pi^{sum}$
+are written through the DKV store.
 
-\subsection{Pipelining of $\pi$ staging and computation}
+\textit{update\_beta} requires $M\times K$ operations. It is preceded by an MPI
+barrier to ensure that up-to-date $\pi$ values are read. \textit{update\_beta}
+is split into four steps. The first stage partitions the minibatch across
+machines to calculate contributions to $\nabla\theta$. The values of $\pi$ for
+the local minibatch nodes are loaded from the DKV store. These calculations
+are done with thread parallelism per minibatch node. In the second step,
+a multithreaded summation is performed to calculate the contributions per
+machine, followed by a distributed summation using an MPI reduce operation. The
+third step calculates $\beta$ from $\theta$, sequentially at the master's; this
+operation takes only $K$ steps. The resulting $\beta$ is broadcast with MPI
+to the workers.
+
+The calculation of perplexity requires $|H|\times K$ steps. Each of the
+machines owns a subset of $H$. It loads up-to-date values of $\pi$ for each
+node in its part of $H$, then for each of its edges in $H$ calculates the
+contribution to the perplexity using thread parallelism. The resulting
+contributions are summed to the global perplexity value in two stages,
+with a local OpenMP reduction followed by a global MPI reduction.
+
+Note that the calculation requirements per iteration do not depend
+on $N$ or $|E|$, the size of the network graph. However, larger graphs are
+expected to require more iterations to achieve convergence.
+
+\subsection{Pipelining of computation and loading $\pi$}
+
+As will be shown in Section~\ref{eval-pipeline}, loading $\pi$ is the
+dominant contribution within the dominant stage \textit{update\_phi}. This
+is the performance bottleneck of the distributed algorithm. We introduce
+pipeline parallelism in order to shorten the critical path. Loading $\pi$
+and calculating \textit{update\_pi} are partitioned into chunks, so the
+calculation of a chunk is overlapped with loading the next chunk of $\pi$
+values, and also with sampling the minibatch at the master's and deploying
+it to the workers.
 
 \subsection{RDMA Distributed Key-Value Store}
+
+Modern network technology allows access to the memory of remote machines
+through the network cards (NICs) without involvement from the remote host
+altogether~\cite{Infiniband,RDMA,Roce}. This has opened up the development of
+extremely efficient remote memory services, often in the form of DKV stores.
+This project stores $\pi$ in a DKV store in the collective memory of the
+cluster, where $\pi[i]+\pi^{sum}[i]$ is the value for key $i$. We decided not
+to use any of the existing implementations of RDMA DKV stores, and in stead
+build our own DKV store on top of the ib-verbs API~\cite{ib-verbs}. The
+main consideration is that our use case is unusually simple in a number
+of important aspects. The KV layout is static in the sense that there are
+no inserts or deletes after the initial population, which allows a static
+partitioning of KV pairs over the machines. The values are all of the same size,
+a vector of $K+1$ floats. The access pattern is very regular, because the
+computation is partitioned into stages that are separated by barriers; a
+stage either reads values or it updates values that it 'owns', so there
+is no read/write or write/write contention. The existing DKV stores incur
+overhead for insert/delete flexibility, concurrency control, or variable-sized
+values. Our use case allows us to do any operation in exactly one RDMA read
+or RDMA write. The Herd paper~\cite{Herd} argues that replacing RDMA reads by
+an RPC to the server, followed by an unreliable write, is faster. However,
+they show this only holds for small payload packets, up to 256B. A $\pi$
+packet in our application is typically thousands to hundreds of thousands
+of 4-byte floats.

### 2016-01-14 04-evaluation.tex
Scale-up vs scale-out: needs heavy polishing still

@@
   \epsfig{file=plots/hpc-cloud.eps, width=\columnwidth}
-  \caption{FIXME}
+  \caption{Performance comparison between the distributed implementation
+  running on DAS5 and the multi-threaded solution on machine with 40-cores and
+  1TB of RAM.}
   \label{fig-scale-up}
 \end{figure}
+One of the main drawbacks of designing a distributed solution for a given
+algorithm is the inherent complexity of communication and synchronization. A
+slightly simpler approach would be developing a multi-threaded version and
+running it on a machine with abundant memory and CPU cores. In such a context,
+access to all of the algorithm's state would be an order of magnitude faster
+than RDMA. Additionally, the overhead associated with synchronizing threads is
+negligible compared to using MPI primitives. To evaluate the efficacy of both
+approaches we utilized SurfSara's HPC Cloud system to instantiate a virtual
+machine with 40~CPU cores and 1TB of memory. The physical machine underlying
+the HPC Cloud system contains 40-CPU cores and does not oversubscribe
+resources. Therefore, by provisioning all 40~cores we ensured that there is no
+resource contention from other users of the system. Figure~\ref{fig-scale-up}
+reports the execution time per algorithm iteration for two experimental setups.
+Namely, using the HPC Cloud system compared to using 64-nodes of the DAS5
+cluster. Clearly, the parallel and distributed implementation vastly
+outperforms the single-node multi-threaded solution. Moreover, the trajectory
+of both curves shows a widening gap between them suggesting that the relative
+performance difference will increase for larger $K$.
 

### 2016-01-14 plots/plot-hpc-cloud.sh

@@ -24,7 +24,7 @@ cat <<EOF | gnuplot --persist
 set terminal postscript eps enhanced color font ',8' size 3.3,1.6
 
 set xlabel "Number of Communities (K)"
-set ylabel "Time per iteration (seconds)"
+set ylabel "Time per iteration (milliseconds)"
 #set xrange [256:]
 set grid ytics
 set key below

### 2016-01-14 04-evaluation.tex
Henri's comments on draft 00

@@
 %
-First, an evaluation of the strong scaling behavior of the distributed
+First, an evaluation of the \textit{strong scaling} behavior of the distributed
 algorithm is presented.
 %
-Second, the weak scaling of the solution is assessed with respect to increasing
+Second, the \textit{weak scaling} of the solution is assessed with respect to increasing
 numbers of targeted latent communities.
 %
-Next, the effects of varying the number of latent communities on the system's
+Next, the effects of \textit{pipelining the computation} and varying the number of latent communities on the system's
 performance is analyzed.
 %
-Further, we evaluate the efficiency and overhead associated with network
-communication between cluster nodes.
+Further, we evaluate the efficiency and overhead associated with \textit{network
+communication} between cluster nodes.
 %
-Finally, we contrast the effectiveness of scaling the computation horizontally
+Finally, we \textit{contrast} the effectiveness of scaling the computation horizontally
 and vertically.
@@
 at 2.40GHz, 64GB of memory and 8TB of storage. Moreover, the compute nodes of
-each site are interconnected by an internal 1~Gbit/s Ethernet and
+each site are interconnected by
 FDR~InfiniBand. All of the experiments reported in this section utilized
@@
 addition of compute nodes. However, the speedup curve gradually slows down for
-larger cluster sizes as the work granularity of each worker node decreases
+larger cluster sizes as the work granularity of each worker node decreases,
 limiting their resource utilization.
@@
 communities proportionally. This methodology ensures that each compute node
-performs a relatively constant amount of computational work across the all
+performs a relatively constant amount of computational work across all
 configurations. However, the number and size of messages exchanged between the
@@
 for larger cluster sizes. This observation verifies that the system's overall
-overhead is minimal and scales well.
+overhead is minimal. Additionally, the experimental results shows that the
+implementation is capable of achieving good speedups provided the input problem
+is large enough for the given cluster size.
 
@@
 the rows of $\pi$ are equally and statically partitioned across all workers.
-Given that most $\pi$ accesses are random, a node in a cluster of size $C$ must
+Given that $\pi$ accesses are random, a node in a cluster of size $C$ must
 fetch $(C-1)/C$ of all read requests over the network. Therefore, large cluster
@@
 
+%################################################
 \subsection{Horizontal vs. Vertical Scalability}
@@
 negligible compared to using MPI primitives. To evaluate the efficacy of both
-approaches we utilized SurfSara's HPC Cloud system to instantiate a virtual
+approaches we utilized SURFsara's HPC Cloud system to instantiate a virtual
 machine with 40~CPU cores and 1TB of memory. The physical machine underlying
@@
 performance difference will increase for larger $K$.
+%
+In conclusion, the overhead of network communication in the distributed version
+is more than compensated by the increasing compute power compared to a
+a single-node implementation.
 
+%################################################
 \subsection{Cluster Communication Efficiency}
 
+%################################################
 \subsection{Convergence of Large Datasets}

### 2016-01-15 04-evaluation.tex
Adding DBLP-np5 to plot + intro to convergence subsection

@@
   \epsfig{file=plots/ppx.eps, width=\textwidth}
-  \caption{FIXME}
+  \caption{Convergence time of 6 different data sets.}
   \label{fig-ppx}
 \end{figure*}
+
+The previous sections focused on the computational performance of the
+distributed implementation. However, the algorithm's throughput does not
+necessarily indicate how fast it can converge to a solution. For instance, it
+remains unclear how many iterations are needed for the algorithm to reach a
+stable state and terminate. Therefore, we now shift our focus to the
+convergence time to assess the system's utility. Figure~\ref{fig-ppx} presents
+the convergence time of 6 different data sets with a diverse set of properties.

### 2016-01-15 04-evaluation.tex
A table w/ breakdown for the pipelining

@@
 overlapping execution of the two most expensive phases, namely, update\_pi and
-deploy\_minibatch. Both of these phases initially gain significant speedup with the
+deploy\_mini-batch. Both of these phases initially gain significant speedup with the
 addition of compute nodes. However, the speedup curve gradually slows down for
@@
 
+\begin{table}[htbp]
+  \centering
+  \begin{tabular}{l d{5.1} d{5.1}}
+    \textbf{Iteration stages}  & \multicolumn{1}{c}{non-pipelined}
+                                          & \multicolumn{1}{c}{pipelined} \\
+    \hline
+    total                      & 450    & 365 \\
+    draw/deploy mini-batch      &  45.6  &      \\
+    update\_phi                & 285    & 241 \\
+    update\_pi                 &   3.8  &   4.6 \\
+    update $\beta/\theta$      &  25.9  &   33.6 \\
+    \hline
+    \textbf{Substages within update\_phi} \\
+    draw/deploy mini-batch      &        &  26.2 \\
+    load $\pi$                 & 205    & 209 \\
+    update $\phi$              &  74    &  74 \\
+    \hline
+  \end{tabular}
+  \caption{The most important stages in the execution; com-Friendster on 65
+  compute nodes, with 12K communities. Times in ms per iteration. The lower
+  half of the table shows substages within \textit{update\_phi}. In the
+  pipelining version, these components are done in parallel.}
+  \label{table-pipeline}
+\end{table}
+
+Table~\ref{table-pipeline} shows a breakdown of the most important time
+consumers; the left-hand column is without pipelining, the right-hand
+column has pipelining. \textit{update\_phi} is the dominant contribution,
+and within that, loading $\pi$ from the DKV store takes by far the most time.
+The pipelining optimization overlaps loading $\pi$ with generation and
+deployment of the mini-batch, and with the calculation of \textit{update\_phi}.
+
 %################################################

### 2016-01-15 04-evaluation.tex
More on convergence + plot fixes

--- 2016-01-15 04-evaluation.tex
@@
 In conclusion, the overhead of network communication in the distributed version
-is more than compensated by the increasing compute power compared to a
+is more than compensated by the increasing compute power compared to
 a single-node implementation.
@@
 distributed implementation. However, the algorithm's throughput does not
-necessarily indicate how fast it can converge to a solution. For instance, it
+necessarily indicate how fast it can converge to a solution. More specifically, it
 remains unclear how many iterations are needed for the algorithm to reach a
@@
 convergence time to assess the system's utility. Figure~\ref{fig-ppx} presents
-the convergence time of 6 different data sets with a diverse set of properties.
+the convergence time of the data sets listed in Table~\ref{tab-datasets}.
+%
+The results in sub-figures a, b and c where obtained by using 64 compute nodes
+of DAS5. In Figure~\ref{fig-ppx}-a, the number of communities was set to 12K
+which fully occupied the aggregate memory resources of all 64 machines since
+com-Friendster has roughly 65 million vertices. In this case, the algorithm
+reached a stable state after 3-4 hours. Next, Figures~\ref{fig-ppx}-b and
+\ref{fig-ppx}-c present the convergence of com-LiveJournal and com-Orkut
+respectively. As the number of vertices in these data sets is an order of
+magnitude smaller than com-Friendster, we used a larger number of
+communities. Naturally, the convergence time was extended as the complexity of
+the algorithm increases dramatically with larger K. However, the system was
+still capable of reporting results after \textbf{XXX time}.
+%
+Figures~\ref{fig-ppx}-d, \ref{fig-ppx}-e and \ref{fig-ppx}-f were all
+configured to use the total number of ground-truth communities associated with
+their respective data sets, com-Youtube, com-DBLP and com-Amazon. However,
+these experiments were conducted on 14, 24 and 24 cluster nodes respectively.
+
+The results reported in Figure~\ref{fig-ppx} verify that the distributed
+implementation of the algorithm is capable of detecting overlapping communities
+within real graphs in reasonable time given the available compute resources.
+The time required to reach convergence may vary depending on a graph's properties and
+the targeted number of communities. To the best of our knowledge, the data sets
+used in this study are the largest publicly available organic graphs.

### 2016-01-15 03-design.tex

@@
   pipeline:
-    1) sample minibatch
+    1) sample mini-batch
     2) sample neighbors
@@
 - data dependencies:
-  mb = nodes in the minibatch
+  mb = nodes in the mini-batch
   nb = nodes in the neighbor sets
@@
 
-  	pi	phi	beta	theta	grad.th	G	prev.minibatch
+  	pi	phi	beta	theta	grad.th	G	prev.mini-batch
   1) 	-	-	-	-		R	-
@@
   in principle, each step synchronized
-  but: pipelining in sample minibatch / load pi/phi / update_phi
+  but: pipelining in sample mini-batch / load pi/phi / update_phi
 
@@
   larger K. The relevant slices of G (m's adjacency lists) are scattered
-  together with the minibatch.
+  together with the mini-batch.
   G is stored as a Google SparseHash because c++ std::(unordered_)set has a
@@
 the calculations in the update stages only require the subset of $G$ that
-is touched by the minibatch nodes, so the master scatters that subset to
-the workers together with the scattering of the minibatch nodes. This way,
+is touched by the mini-batch nodes, so the master scatters that subset to
+the workers together with the scattering of the mini-batch nodes. This way,
 we trade a reduction in memory at the workers against the cost of limited
@@
 
-The first stage, minibatch selection, is done at the master's, and it is
+The first stage, mini-batch selection, is done at the master's, and it is
 not itself parallelized. However, in the distributed implementation its
 execution is overlapped using pipeline parallelism, while the workers are
-performing \textit{update\_phi}. The minibatch is partitioned equally over
+performing \textit{update\_phi}. The mini-batch is partitioned equally over
 the workers; the relevant sections of $G$ are distributed together with the
-minibatch subsets.
+mini-batch subsets.
 
-After a worker has received its subset of the minibatch, it samples a neighbor
-set of size $n$ for each of its minibatch nodes, using thread parallelism.
+After a worker has received its subset of the mini-batch, it samples a neighbor
+set of size $n$ for each of its mini-batch nodes, using thread parallelism.
 
@@
 K$ operations. It can be fully parallelized because it is a data-parallel
-operation over each of the minibatch nodes. \textit{update\_phi} loads the
-$\pi$ values for the local minibatch nodes and their neighbors from the
-DKV store. The updates to $\phi$ for the minibatch nodes are calculated
+operation over each of the mini-batch nodes. \textit{update\_phi} loads the
+$\pi$ values for the local mini-batch nodes and their neighbors from the
+DKV store. The updates to $\phi$ for the mini-batch nodes are calculated
 independently using thread parallelism.
@@
 These updates are used by \textit{update\_pi} locally, besides the value of
-$\pi/\phi$ for the minibatch nodes. The update to $\pi$ requires $M\times
-K$ operations, and it is done in parallel over minibatch nodes. Because of
+$\pi/\phi$ for the mini-batch nodes. The update to $\pi$ requires $M\times
+K$ operations, and it is done in parallel over mini-batch nodes. Because of
 memory consistency, this stage awaits completion of \textit{update\_phi} with
@@
 barrier to ensure that up-to-date $\pi$ values are read. \textit{update\_beta}
-is split into four steps. The first stage partitions the minibatch across
+is split into four steps. The first stage partitions the mini-batch across
 machines to calculate contributions to $\nabla\theta$. The values of $\pi$ for
-the local minibatch nodes are loaded from the DKV store. These calculations
-are done with thread parallelism per minibatch node. In the second step,
+the local mini-batch nodes are loaded from the DKV store. These calculations
+are done with thread parallelism over the mini-batch nodes. In the second step,
 a multithreaded summation is performed to calculate the contributions per
@@
 machines owns a subset of $H$. It loads up-to-date values of $\pi$ for each
-node in its part of $H$, then for each of its edges in $H$ calculates the
+node in its part of $H$, then for each of its edges in $H$ it calculates the
 contribution to the perplexity using thread parallelism. The resulting
@@
 calculation of a chunk is overlapped with loading the next chunk of $\pi$
-values, and also with sampling the minibatch at the master's and deploying
+values, and also with sampling the mini-batch at the master's and deploying
 it to the workers.

### 2016-01-15 04-evaluation.tex

@@
 widening gap between both lines depicted in Figure~\ref{fig-pipeline}.
+
+\begin{table}[htbp]
+  \centering
+  \begin{tabular}{l d{5.1} d{5.1}}
+    \textbf{Iteration stages}  & \multicolumn{1}{c}{non-pipelined}
+                                          & \multicolumn{1}{c}{pipelined} \\
+    \hline
+    total                      & 450    & 365 \\
+    draw/deploy mini-batch      &  45.6  &      \\
+    update\_phi                & 285    & 241 \\
+    update\_pi                 &   3.8  &   4.6 \\
+    update $\beta/\theta$      &  25.9  &   33.6 \\
+    \hline
+    \textbf{Substages within update\_phi} \\
+    draw/deploy mini-batch      &        &  26.2 \\
+    load $\pi$                 & 205    & 209 \\
+    update $\phi$              &  74    &  74 \\
+    \hline
+  \end{tabular}
+  \caption{The most important stages in the execution; com-Friendster on 65
+  compute nodes, with 12K communities. Times in ms per iteration. The lower
+  half of the table shows substages within \textit{update\_phi}. In the
+  pipelining version, these components are done in parallel.}
+  \label{table-pipeline}
+\end{table}
+
+Table~\ref{table-pipeline} shows a breakdown of the most important time
+consumers; the left-hand column is without pipelining, the right-hand
+column has pipelining. \textit{update\_phi} is the dominant contribution,
+and within that, loading $\pi$ from the DKV store takes by far the most time.
+The pipelining optimization overlaps loading $\pi$ with generation and
+deployment of the mini-batch, and with the calculation of \textit{update\_phi}.
 

### 2016-01-15 03-design.tex
Add some citations already

@@
 calculations. For thread parallelism, we annotate the program with
-OpenMP~\cite{OpenMP}. For distributed communication, we use MPI~\cite{MPI}.
+OpenMP~\cite{OpenMPSpec}. For distributed communication, we use
+MPI~\cite{Forum:1994:MMI:898758}.
 
@@
 through the network cards (NICs) without involvement from the remote host
-altogether~\cite{Infiniband,RDMA,Roce}. This has opened up the development of
+altogether~\cite{Hamada_infinibandtrade,RDMA,Beck:2011:PER:2043535.2043537}. This has opened up the development of
 extremely efficient remote memory services, often in the form of DKV stores.

### 2016-01-15 02-related.tex
Adding intro and background from overleaf

\section{Background}
% In this section, we briefly review the Bayesian model we are considering in this paper.

% In this section, we briefly review the Bayesian model we are considering in this paper.
\subsection{Assortative mixed-membership stochastic blockmodel (a-MMSB)}
Assortative mixed-membership stochastic blockmodel (a-MMSB) ~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group-structure in a network of $N$ nodes. We assume $\cV*$ is the set containing all the nodes in the graph, and $\cE*$ is the set containing all the linked edges between pairs of nodes.  For each node $a$ in the node set $\cV^*$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.  For every possible peer $b$ in the network, each node $a$ randomly draws a community $z_{ab}$. If a pair of nodes ($a,b$) in the edge set $\cE^*$ are in the same community: $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$. Otherwise this probability is small. Each community has its connection strength $\beta_{k} \in (0,1)$ which explains how likely its members are linked to each other. \\

The whole generative process of a-MMSB is given by,

\benum
\item For each community $k$, draw community strength $\bt_k \sim \mbox{Beta}(\eta)$
\item For each node $a$, draw community memberships $\pi_a \sim \mbox{Dirichlet}(\al)$
\item For each pair of nodes $a$ and $b$,
\benum
	\item Draw interaction indicator $z_{ab} \sim \pi_a$
    \item Draw interaction indicator $z_{ba} \sim \pi_b$
    \item Draw link $y_{ab} \sim \mbox{Bernoulli}(r)$, where $r = \bt_k$ if $z_{ab}=z_{ba}=k$, and $r=\dt$ otherwise.
\eenum
\eenum

\subsection{Stochastic Gradient MCMC Algorithms}

The algorithm we are working on in this paper is based on the stochastic gradient Langevin dynamics (SGLD) ~\cite{welling2011bayesian}. SLGD applies the following update rule to sample from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ given $N$ i.i.d. data points $\cX=\{x_i\}_{i=1}^{N}$

\bea
\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\cD_n)\right) + \xi, \label{eqn:sgld_update}
\eea

where $\xi \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\cD_n$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\cD_n) = \frac{1}{|\cD_n|}\sum_{x\in \cD_n}^{}\grad_{\ta} \log p(x|\ta)$. As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution. One benefit of using SGLD is that we don't need Metropolis-Hastings (MH) accept-reject tests since the rejection probability goes to zero as the step size collapses to zero.  Although the finite step size still results in some bias in practice, we can reduce the overall error by drawing many more samples per unit time.

SGLD originated from the Langevin Monte Carlo (LMC)~\cite{girolami2011riemann} , where the gradient is computed exactly using all data points and then a Metropolis-Hastings accept-reject test is applied. Comparing to LMC,  SGLD only requires to process a mini-batch $\cD_n$ at each iteration and ignores the MH test, which substantially reduces the computational complexity from  $\cO(N)$ to $\cO(n)$ .

The stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex. By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch estimator in Eqn. \ref{eqn:sgld_update}, it achieved the state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}. In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$.
Then, the update rule becomes

\bea
\ta_{k}^* \law \left| \ta_{k} + \f{\ep}{2} \left( \al - \ta_{k} + \f{N}{|\cD_n|} \sum_{d \in \cD_n} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{} \right|. \label{eqn:sgrld_update}
\eea
here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \cD_n$.

\subsection{Scalable MCMC for a-MMSB}

The algorithm iterates updating local parameter  $\pi$ and a global parameter $\beta$.  Since both parameters lie on the probability simplex, SGRLD is applied to make the sampling process more efficient.  Here, the parameters $\phi$ and $\ta$ are used to re-parameterize $\pi$ and $\beta$ respectively.  After updating  $\phi$ and $\ta$,  then we can obtain $\pi$ and $\beta$ by normalizing them.  In the following, we briefly sketch the iterative updating steps for a-MMSB.  For more details about the algorithm, please refer to~\cite{LiAW15}

\paragraph{Sampling global parameters} The update rule for global parameter $\theta$ is
\bea
\ta_{ki}^* \law \left| \ta_{ki} + \f{\ep}{2} \left\{ \eta - \ta_{ki} + h(\cE_{n_t})\sum_{(a,b) \in \cE_{n_t}}g_{ab}(\ta_{ki})\right\} \right.\nn \\ \left.+ (\ta_{ki})^{\ha}\xi_{ki} \right|, \label{eqn:global_update}
\eea

where 
\bea
g_{ab}(\ta_{ki})
&=& \f{f_{ab}^{(y)}(k,k)}{Z_{ab}^{(y)}} \left(\f{|1-i-y|}{\ta_{ki}} - \f{1}{\ta_k} \right).
\eea

Here $\cE_{n_t}$ is a mini-batch of $n_t$ node pairs sampled from $\cE^*$, and $h(\cE_{n_t})$ is the weighting terms. \\

\paragraph{Sampling local parameters}
The update rule for the local parameters $\phi$ is
\bea
\phi_{ak}^* \law \left| \phi_{ak} + \f{\ep}{2} \left( \al - \phi_{ak} + \f{N}{|\cV_n|} \sum_{b \in \cV_n} g_{ab}(\phi_{ak})\right) \right. \nn \\ \left. + (\phi_{ak})^{\ha} \xi_{ak}\right|.\label{eqn:local_update}
\eea

where \bea
g_{ab}(\phi_{ak}) = \f{f_{ab}^{(y)}(k)}{Z_{ab}^{(y)}\phi_{ak}} - \f{1}{\phi_a}.
\eea

Here, the $\cV_n$ is a random mini-batch of $n$ nodes sampled from $\cV^*$. Note that $|\cV_n| \ll |\cV^*|=N$.

The overall pseudo code is shown in Algorithm 1.

\begin{algorithm}[t]
\caption{Pseudo Code}\label{alg}
\begin{algorithmic}[1]
\STATE Initialize $\pi , \bt, \phi, \ta$
\WHILE {sampling}
\STATE Sample a mini-batch of node pairs, $\cE_t$, from $E$
\FOR {each node in $\cE_t$}
	\STATE Sample a mini-batch of nodes, $\cV_t$, from $V$
    \STATE Update $\phi_{a}$ using Eqn~\ref{eqn:local_update}
  %  $\phi_{a}^* \law \left| \phi_{a} + \f{\ep_{\phi}}{2} \left( \al - \phi_{a} + \f{N}{|\cV_t|} \sum_{b \in \cV_t} G_{\phi_{a}}\right) + (\phi_{a})^{\ha} \xi_{a}\right|$
    \STATE Obtain $\pi_a$ from $\phi_a^*$ \ENDFOR
\FOR {$k = 1,\dots,K$}
    \STATE Update $\ta_{k}$ using Eqn~\ref{eqn:global_update}
   % $\ta_{k}^* \law \left| \ta_{k} + \f{\ep_\ta}{2} \left( \eta - \ta_{k} + h(\cE_t) \sum_{(a,b) \in \cE_t} G_{\ta_{k}}(a,b) \right) + (\ta_{k})^{\ha}\xi_{k} \right|$
    \STATE Obtain $\bt_k$ from $\ta_k^*$
\ENDFOR
\ENDWHILE
\end{algorithmic}
\end{algorithm}

\begin{comment}
Assortative mixed-membership stochastic blockmodel (a-MMSB) [1] is a special case of MMSB [8] that models the group-structure in a network of $N$ nodes. In particular, each node $a$ in the node set $\cV^*$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$. For every possible peer $b$ in the network, each node $a$ randomly draws a community $z_{ab}$. If a pair of nodes ($a,b$) in the edge set $\cE^*$ are in the same community: $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$. Otherwise this probability is small. Each community has its connection strength $\beta_{k} \in (0,1)$ which explains how likely its members are linked to each other.

The generative process of a-MMSB is then given by,

\benum
\item For each community $k$, draw community strength $\bt_k \sim \mbox{Beta}(\eta)$
\item For each node $a$, draw community memberships $\pi_a \sim \mbox{Dirichlet}(\al)$
\item For each pair of nodes $a$ and $b$,
\benum
	\item Draw interaction indicator $z_{ab} \sim \pi_a$
    \item Draw interaction indicator $z_{ba} \sim \pi_b$
    \item Draw link $y_{ab} \sim \mbox{Bernoulli}(r)$, where $r = \bt_k$ if $z_{ab}=z_{ba}=k$, and $r=\dt$ otherwise.
\eenum
\eenum

Unlike the a-MMSB, the original MMSB maintains pair-wise community strength $\bt_{k,k'}$ for all pairs of the communities. Note that it is trivial to extend the results that we obtain in this paper to the general MMSB model. The joint probability of the above process can be written as:
\begin{align}
p(y,z,\pi ,\bt | \al, \eta) &= \prod_{a=1}^{N} \prod_{b > a}^{N} p(y_{ab} | z_{ab}, z_{ba}, \bt) p(z_{ab}|\pi_a)\nonumber \\ &p(z_{ba}|\pi_b) \pd{a}{N}p(\pi_a|\al) \pd{k}{K} p(\bt_k|\eta).\label{eqn:joint}
\end{align}

Both variational inference [1,4,16] and collapsed Gibbs sampling algorithms [11] have been used successfully for small to medium scale problems. However, the $\cO(N^2)$ computational complexity per update prevents it from being applied to large scale networks. A stochastic variational algorithm was developed in [1] to address this issue, where each update only depends on a small mini-batch of the nodes in the network.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Stochastic Gradient MCMC Algorithms}
Our algorithm will be based on the stochastic gradient Langevin dynamics (SGLD) [3]. To sample from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ given $N$ i.i.d. data points $\cX=\{x_i\}_{i=1}^{N}$, SGLD applies the following update rule:
\bea
\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\cD_n)\right) + \xi, \label{eqn:sgld_update}
\eea
where $\xi \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\cD_n$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\cD_n) = \frac{1}{|\cD_n|}\sum_{x\in \cD_n}^{}\grad_{\ta} \log p(x|\ta)$. As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution. In SGLD, the Metropolis-Hastings (MH) accept-reject tests are ignored since the rejection probability goes to zero as the step size collapses to zero. While for a finite step size this results in some bias, the overall error is reduced by the reduction of variance due to the ability to draw many more samples per unit time.

SGLD originated from the Langevin Monte Carlo (LMC) [15] where, unlike SGLD, the gradient is computed exactly using all data points and then a Metropolis-Hastings accept-reject test is applied. Because at each iteration SGLD requires to process only a mini-batch $\cD_n$ and ignores the MH test, the computation complexity per iteration is only $\cO(n)$ as opposed to $\cO(N)$ of LMC. Any mini-batch sampling algorithm in the form of Eqn. \eqref{eqn:sgld_update} is called valid SGLD as long as it guarantees the gradient estimator to be unbiased, i.e.,
$\eE_{\cD_n} \left[N\bar{g}(\ta;\cD_n)\right] = \grad_{\ta} \log p(\cX|\ta)\label{eqn:sgld_grad_estm}$ and the variance to be finite [5].

%%%%%%%%%%%%%%
%\subsection{Stochastic Gradient Riemannian Langevin Dynamics (SGRLD)}

The stochastic gradient Riemannian Langevin dynamics (SGRLD) [2] is a subclass of SGLD which is developed to sample from the probability simplex. By applying Riemannian geometry [15] and using the mini-batch estimator in Eqn. \ref{eqn:sgld_update}, it achieved the state-of-the-art performance for latent Dirichlet allocation (LDA). In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$.
Then, the update rule becomes

%==============================================================

\bea
\ta_{k}^* \law \left| \ta_{k} + \f{\ep}{2} \left( \al - \ta_{k} + \f{N}{|\cD_n|} \sum_{d \in \cD_n} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{} \right|. \label{eqn:sgrld_update}
\eea
here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \cD_n$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scalable MCMC for a-MMSB}
Our algorithm iterates updating local parameters $\pi$ and a global parameter $\beta$. Because both parameters lie on the probability simplex, we start from the SGRLD and modify it to be more efficient. Also, we introduce parameters $\phi$ and $\ta$ to re-parameterize $\pi$ and $\beta$ respectively. Thus, we alternatingly sample in the $\phi$ and $\ta$ spaces, and then obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$. From Eqn. \ref{eqn:joint}, summing over the latent variable $z$, we obtain the following joint probability,
\bea
&p(y, \pi , \bt | \al , \eta) = \prod_{a} p(\pi_a | \al ) + \prod_{a} p(\bt_k | \eta)\nn \\
&+ \prod_{a} \prod_{b > a}\sum_{z_{ab}, z_{ba}} p(y_{ab}, z_{ab}, z_{ba} | \bt , \pi_a, \pi_b)
\label{eqn:log_joint}.
\eea

%==============================================================
\subsection{Sampling the global parameter}

By the re-parameterization, we have $\beta_k=\ta_{k1}/(\ta_{k0}+\ta_{k1})$, where $\ta_{ki}\sim$ Gamma$(\eta)\propto \ta_{ki}^{\eta-1}e^{-\ta_{ki}}$. Because $p(y, \pi , \bt | \al , \eta)$ decomposes into $p(y,\bt|\pi,\eta) p(\pi|\al)$, replacing $\bt$ by $\ta$, we compute the derivative of log of Eqn. \ref{eqn:log_joint} w.r.t. $\ta_{ki}$ for $i=\{0,1\}$ as follows:
\bea
\fp{\ln p(y ,\ta | \pi, \eta )}{\ta_{ki}} =
\fp{}{\ta_{ki}} \ln p(\ta_{ki} | \eta) + \sum_{a}\sum_{b>a} g_{ab}(\ta_{ki}),
\label{eqn:grad_bt}
\eea
where $g_{ab}(\ta_{ki}) = \fp{}{\ta_{ki}} \ln \sum_{z_{ab},z_{ba}} p(y_{ab},z_{ab},z_{ba} | \ta , \pi_a, \pi_b)$ which, similar to SGRLD for LDA [2], we can rewrite as
\bea
g_{ab}(\ta_{ki}) = \eE \left[ \eI[z_{ab}=z_{ba}=k] \left( \f{|1-i-y_{ab}|}{\ta_{ki}} - \f{1}{\ta_{k}} \right)\right]. \label{eqn:G}
\eea
where $\ta_{k}=\sum_i \ta_{ki}$ and $\eI[S]$ is equal to $1$ if a condition $S$ is TRUE and $0$ otherwise. The expectation is w.r.t. the posterior distribution of latent variables $z_{ab}$ and $z_{ba}$,
\bea
&&p(z_{ab}=k,z_{ba}=l | y_{ab}, \pi_a, \pi_b, \bt )\\ &&\propto f_{ab}^{(y)}(k,l)=
\begin{cases}
\bt_k^y(1-\bt_k)^{(1-y)}\pi_{ak}\pi_{bk}, & \mbox{if } k=l\\
\dt^{y}(1-\dt)^{(1-y)} \pi_{ak} \pi_{bl}, & \mbox{if } k \neq l \nn
\end{cases}
\label{eqn:case}
\eea
here we used simple notation $y$ instead of $y_{ab}$. Unlike the SGRLD for LDA [2], we compute the expectation in Eqn. \eqref{eqn:G} analytically by computing the normalization constant $Z_{ab}^{(y)} = \sm{k}{K}\sm{l}{K} f_{ab}^{(y)}(k,l)$ which can be reduced to $\cO(K)$ computation as follows
\begin{align}
&Z_{ab}^{(y)} = \dt^{y}(1-\dt)^{(1-y)} \nonumber\\&+\sm{k}{K} \left( \bt_k^{y}(1-\bt_k)^{(1-y)} - \dt^{y}(1-\dt)^{(1-y)}\right)\pi_{ak}\pi_{bk}
\end{align}

Then Eqn. \eqref{eqn:G} becomes
\bea
g_{ab}(\ta_{ki}) 
&=& \f{f_{ab}^{(y)}(k,k)}{Z_{ab}^{(y)}} \left(\f{|1-i-y|}{\ta_{ki}} - \f{1}{\ta_k} \right).
\eea
Plugging this into Eqn. \ref{eqn:sgrld_update}, we obtain the update rule for the global parameter,

\bea
\ta_{ki}^* \law \left| \ta_{ki} + \f{\ep}{2} \left\{ \eta - \ta_{ki} + h(\cE_{n_t})\sum_{(a,b) \in \cE_{n_t}}g_{ab}(\ta_{ki})\right\} \right.\nn \\ \left.+ (\ta_{ki})^{\ha}\xi_{ki} \right|, \label{eqn:global_update}
\eea
here $\cE_{n_t}$ is a mini-batch of $n_t$ node pairs sampled from $\cE^*$ for which we use the following strategy.

\textbf{Stratified sampling:} considering that the number of links is much smaller than that of non-links, we can reduce the variance of the gradient using stratified sampling, similar to the method used in [1]. For this, at every iteration we first randomly select a node $a$ and then toss a coin with probability 0.5 to decide whether to sample link edges or non-link edges for node $a$. If it is a link, we assign all of the link edges of node $a$ to $\cE_{n_t}$. Otherwise, i.e. if it is non-link, we uniformly sample a mini-batch of $N/m$ non-link edges from the entire set of non-link edges and assign it to $\cE_{n_t}$. Here, the $m$ is a hyper-parameter. Note that the size of $|\cE_{n_t}|$ will thus be much smaller than the total number of $N(N-1)/2$ edges when $m$ is reasonably large. Then, to ensure that the gradient is unbiased, a \textit{scaling parameter} $h(\cE_{n_t})$ is multiplied. Specifically, $h(\cE_{n_t})$ is set to $N$ when $\cE_{n_t}$ is a set of link edges and to $mN$ otherwise.

Because the global parameters $\{\beta_k\}$ does not change very fast compared to the local parameters $\{\pi_{ak}\}$, in practice we update only a random subset of the $\{\beta_k\}$ at each iteration.

\subsection{Sampling the local parameters}

Similar to the global parameter, we re-parameterize the local parameter $\pi_a$ such that $\pi_{ak} = \phi_{ak} / \sum_{j=1}^{K}\phi_{aj}$, with $\phi_{ak}\sim$ Gamma$(\alpha)\propto \phi_{ak}^{\alpha-1}e^{-\phi_{ak}}$. Then, taking the derivative of the log of Eqn. \ref{eqn:log_joint} w.r.t. $\phi_{ak}$, we obtain
\bea
\fp{\ln p(y , \phi | \bt, \al)}{\phi_{ak}} = \fp{}{\phi_{ak}} \ln p(\phi_{ak} | \al) +
\sum_{b} g_{ab}(\phi_{ak})
\eea
where $g_{ab}(\phi_{ak}) = \fp{}{\phi_{ak}} \ln \sum_{z_{ab}, z_{ba}} p(y_{ab}, z_{ab}, z_{ba} | \bt , \phi_a, \phi_b)$ which can be written as 
\bea 
g_{ab}(\phi_{ak}) = \eE\left[ \f{\eI [z_{ab} = k]}{ \phi_{ak}} - \f{1}{\phi_{a\cdot}} \right].
\label{eqn:grad_local}
\eea
Here the expectation is w.r.t. the distribution in Eqn. \eqref{eqn:case}. To compute the expectation analytically, we first integrate out $z_{ba}$ from Eqn. \eqref{eqn:case} because the expectation depends only on $z_{ab}$, and obtain the following probability up to a normalization constant

\begin{align}
&f_{ab}^{(y)}(k) = \sum_{l=1}^K f_{ab}^{(y)}(k,l)\nonumber \\ =& \pi_{ak}
\left\{ \bt_k^{y}(1-\bt_k)^{(1-y)} \pi_{bk} + \dt^{y}(1-\dt)^{(1-y)} (1- \pi_{bk}) \right\}.\label{eqn:local_f}
\end{align}

Then we obtain the normalization term by $Z_{ab}^{(y)} = \sm{k}{K}f_{ab}^{(y)}(k)$. 
Integrating out the expectation in Eqn. \eqref{eqn:grad_local}, we obtain 
\bea
g_{ab}(\phi_{ak}) = \f{f_{ab}^{(y)}(k)}{Z_{ab}^{(y)}\phi_{ak}} - \f{1}{\phi_a}.
\eea 
Plugging this to Eqn. \ref{eqn:sgrld_update}, we obtain the SGRLD update rule for the local parameter $\phi_{ak}$
\bea
\phi_{ak}^* \law \left| \phi_{ak} + \f{\ep}{2} \left( \al - \phi_{ak} + \f{N}{|\cV_n|} \sum_{b \in \cV_n} g_{ab}(\phi_{ak})\right) \right. \nn \\ \left. + (\phi_{ak})^{\ha} \xi_{ak}\right|.\label{eqn:local_update}
\eea
Here, the $\cV_n$ is a random mini-batch of $n$ nodes sampled from $\cV^*$. Note that $|\cV_n| \ll |\cV^*|=N$.

\end{comment}

### 2016-01-15 plots/plot-ml-curves.sh
Increase font in convergence plot

@@ -2,8 +2,6 @@
 
 cat <<EOF | gnuplot --persist
 
-#set terminal postscript eps enhanced color font ',8' size 3.3,1.6
-
 set xlabel "Time (hours)"
 set ylabel "Perplexity"
 #set xrange [0:]
@@ -18,7 +16,7 @@ nc = 3
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
-set terminal postscript eps enhanced color font ',8'
+set terminal postscript eps enhanced color font ',10'
 set size (nc*w),(nr*h+k)
 set output 'ppx.eps'
 set multiplot layout nr,nc

### 2016-01-15 04-evaluation.tex
Add some citations already

@@
 \subsection{Pipeline efficiency}
+\label{eval-pipeline}
 
@@
 \subsection{Horizontal vs. Vertical Scalability}
-\begin{figure}[t] % [htb]
+\begin{figure*}[t] % [htb]
   \centering
-  \epsfig{file=plots/hpc-cloud.eps, width=\columnwidth}
+  \epsfig{file=plots/hpc-cloud.eps, width=\textwidth}
   \caption{Performance comparison between the distributed implementation
@@
   \label{fig-scale-up}
-\end{figure}
+\end{figure*}
 One of the main drawbacks of designing a distributed solution for a given
@@
 approaches we utilized SURFsara's HPC Cloud system to instantiate a virtual
-machine with 40~CPU cores and 1TB of memory. The physical machine underlying
+machine with 40 Intel Xeon E7-4850 cores clocked at 2.00GHz cores and 1TB of
+memory. The physical machine underlying
 the HPC Cloud system contains 40-CPU cores and does not oversubscribe
@@
 reports the execution time per algorithm iteration for two experimental setups.
-Namely, using the HPC Cloud system compared to using 64-nodes of the DAS5
-cluster. Clearly, the parallel and distributed implementation vastly
+%
+First, Figure~\ref{fig-scale-up}-a shows the performance of executing the
+algorithm on the HPC Cloud
+system with 40 and 16 cores compared to a single DAS5 node with 16-cores. This
+test uses the com-DBLP data set to enable the use of a large number of
+communities without running out of memory. It is clear from the results that
+the performance can benefit from the additional cores provided by HPC Cloud system.
+%
+Further, Figure~\ref{fig-scale-up} tests the performance of the HPC Cloud
+system compared to 64-nodes of the DAS5
+cluster using the com-Friendster data set. Clearly, the parallel and
+distributed implementation vastly
 outperforms the single-node multi-threaded solution. Moreover, the trajectory

### 2016-01-15 01-intro.tex

@@
 \section{Introduction}
-\cite{DBLP:journals/corr/LiAW15}
+The tremendous amount of data that we generate through our daily applications such as social networking services, online shopping, and news recommendations, provides us a big opportunity to take advantages of the big raw data, when we can extract hidden but invaluable and useful information.
+Realizing this opportunity, however, also requires significant amount of efforts because the efficiency of traditional machine learning algorithms often become extremely inefficient by the large amount of data.
+
+There have been two main approaches to this end; machine learning researchers have developed new scalable algorithms [] while systems and networking researchers have worked on developing new general infra-systems to operate many machine learning algorithms more efficiently [].
+However, it is clear that, in most cases we see the best performance by carefully integrating both of these approaches in one system.
+
+One such big data problem is analyzing graph data such as social networks where it is not unusual to see a network consisting of billions of edges and tens of million of nodes~\cite{}.
+In particular, we are interested in the overlapping community detection problem~\cite{}, where the goal is to learn the probability distribution of each node to participate in each community, provided a set of nodes, the link information between the nodes (which is usually very sparse), and the number of latent communities.
+A community can be seen as a densely connected group of nodes that are only sparsely connect to the rest of the network.
+
+This problem is modeled by the mixed membership stochastic blockmodels (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of the MMSB, called assortative-MMSB~\cite{gopalan2012scalable} (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
+
+The MMSB model is a probabilistic graphical model [] that represents a convenient paradigm for modeling complex relationships between a potentially very large number of random variables.
+Bayesian graphical models [], where we define priors and infer posteriors over parameters also allow us to quantify model uncertainty and facilitate model selection and averaging.
+But an increasingly urgent question is whether these models and their inference procedures will be up to the challenge of handling very large ``big data'' problems.
+
+There has been two main recent advances in this direction of scalable Bayesian inference methods based on stochastic variational Bayes (SVB)~\cite{hoffman2013stochastic,gopalan2013efficient,gopalan2012scalable} and stochastic gradient Markov chain Monte Carlo (SG-MCMC)~\cite{welling2011bayesian,patterson2013stochastic,ahn2014distributed}.
+Both methods have the important property that they only require a small subset of the data-items for every iteration.
+In other words, they can be applied to (infinite) streaming data.
+
+In this paper, we are particularly interested in the SG-MCMC method applied to the a-MMSB model introduced in \cite{LiAW15}, which turned out to be faster and more accurate than the SVB method.
+Although \cite{LiAW15} proposed a scalable algorithm for the problem, there still exist some area to improve the performance further by considering an efficient implementation at system level.
+...
+
+[@Ismail Contribution of this paper]
+
+% -------\\
+% Community detection is the central problem in network analysis, with the goal of identifying the groups of related nodes that are densely connected within this group but sparsely connected to the rest of the network.
+Different from classical community detection problem where we assume each node belongs to one single community, our paper considers overlapping communities where each of the nodes might belong to multiple communities.
+In particular, we consider the model called a-MMSB(Assortive Mixed Membership Stochastic Blockmodel) in this paper, which was first introduced in~\cite{gopalan2012scalable}.\\
+
+% a-MMSB, as a probabilistic graphical model, represents a convenient paradigm for modeling complex relationships between a potentially large number of random variables.
+It also uses priors and posteriors to quantify model uncertainty and facilitate model selection and averaging.
+While a-MMSB provides rich representation power, like other Bayesian models, the inference procedures of handling big data which is common in real world is still a very challenging problem.
+\\
+
+% Consider the large networks such as a social network, it easily runs into billions of edges and tens of million of nodes. In addition to that, the number of communities might exceed few millions. 
+% There were two types of scalable algorithm for a-MMSB have been introduced recently \cite{gopalan2012scalable}, stochastic variational Bayesian inference (SVB) and stochastic gradient Markov chain Monte Carlo(SG-MCMC), respectively.
+Both methods have the important property that each iteration only relies on a small subset of the data.
+Although both methods can work for some large networks with many hundreds of nodes, the performance is far less satisfactory when it applies to the so-called "big data" such as facebook network with millions of nodes and billions of edges.
+Given the facts that \cite{LiAW15}, SG-MCMC runs faster and converges to the better local minima, in this paper, we mainly consider the problem of scaling up SG-MCMC to the big data set.
+\textit{As far as we know, this is the first paper that studies community detection on the frencter dataset with few billion's of edges.}

### 2016-01-15 04-evaluation.tex
Add plots + text on DKV vs qperf

@@
 
+\begin{figure}[b] % [htbp]
+  \centering
+  \epsfig{file=plots/qperf.eps, width=0.5\textwidth}
+  \caption{Performance of DKV store compared to qperf.}
+  \label{fig-qperf}
+\end{figure}
+
+A thorough performance evaluation of our RDMA DKV store is outside the scope of
+this paper. In figure~\ref{fig-qperf}, we provide maximum bandwidth numbers for
+read operations between one server and one client for a range of payload sizes,
+and compare these to the bandwidth achieved by \textit{qperf}, which shows
+the best achievable preformance for Infiniband. We present \textit{qperf}
+bandwidth for both RDMA read and RDMA write operations; these are nearly
+identical, which corroborates the results from the Herd project for payloads
+upwards from 256B. The bandwidth achieved by our DKV store mostly falls short
+of the \textit{qperf} performance. For smallish packets, this is attributed
+to additional per-request overhead for the DKV store. For the largest packet
+size, the DKV store performance is hampered by the fact that its values
+are spread over a larger memory area, whereas \textit{qperf} always reads
+from the same memory locations. For packets between 8KB and 512KB, our DKV
+achieves performance very close to \textit{qperf}.
+
 %################################################

### 2016-01-15 plots/plot-qperf.sh

@@ -1,0 +1,16 @@
+cat <<EOI | gnuplot --persist
+
+set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+set output 'qperf.eps'
+
+set logscale xy
+set xrange [:2000000]
+set yrange [0.25:10]
+set key at 1000000,1
+set xlabel "payload size (bytes)"
+set ylabel "throughput (GB/s)"
+plot \
+	"../data/qperf/qperf/qperf-write-bw" with lp title "qperf RDMA write", \
+	"../data/qperf/qperf/qperf-read-bw" with lp title "qperf RDMA read", \
+	"../data/qperf/dkv/thrp" with lp title "qperf DKV store read"
+EOI

### 2016-01-15 plots/plot-sweep-K-fixed-np.sh

@@ -32,6 +32,7 @@ set grid ytics
 
 set output 'sweep-over-K-fixed-np.eps'
 set title "Scaling K (fixed np=65)"
+set key left
 plot "$$.tmp" u 1:2 w lp t "Single buffering", "$$.tmp2" u 1:2 w lp t "Double buffering"
 
 EOF

### 2016-01-15 01-intro.tex
Edits to sections 1 and 2

--- 2016-01-13 01-intro.tex
@@
 \section{Introduction}
-\cite{DBLP:journals/corr/LiAW15}
+The tremendous amount of data that we generate through our daily applications such as social networking services, online shopping, and news recommendations, provides us with an opportunity to extract hidden but invaluable and useful information.
+Realizing this opportunity, however, requires a significant amount of effort because traditional machine learning algorithms often become extremely inefficient with large amounts of data.
+
+There have been two main approaches to this end; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed} while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning algorithms more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
+However, it is clear that, in most cases we see the best performance by carefully integrating both of these approaches in one system.
+
+One such big data problem is analyzing large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of million of vertices~\cite{yang2015defining}.
+In particular, we are interested in the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, provided a set of vertices, the link information between them (which is usually very sparse), and the number of latent communities.
+A community can be seen as a densely connected group of vertices that are only sparsely connect to the rest of the network.
+
+This problem is modeled by the mixed membership stochastic blockmodels (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of the MMSB, called assortative-MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
+
+The MMSB model is a probabilistic graphical model~\cite{koller2009probabilistic} that represents a convenient paradigm for modeling complex relationships between a potentially large number of random variables.
+Bayesian graphical models, where we define priors and infer posteriors over parameters also allow us to quantify model uncertainty and facilitate model selection and averaging.
+However, an increasingly urgent question is whether these models and their inference procedures will be up to the challenge of handling very large graphs.
+
+There have been two main recent advances in this direction of scalable Bayesian inference methods based on stochastic variational Bayes (SVB)~\cite{hoffman2013stochastic,gopalan2013efficient,gopalan2012scalable} and stochastic gradient Markov chain Monte Carlo (SG-MCMC)~\cite{welling2011bayesian,patterson2013stochastic,ahn2014distributed,ahn2012bayesian}.
+Both methods have the important property that they only require a small subset of the data for every iteration.
+In other words, they can be applied to (infinite) streaming data.
+
+In this paper, we are particularly interested in the SG-MCMC method applied to the a-MMSB model introduced in \cite{LiAW15}, which turned out to be faster and more accurate than the SVB method.
+Although \cite{LiAW15} proposed a scalable algorithm for the problem, there is still room to improve the performance further by considering a custom high-performance implementation.
+To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem.
+In particular, we use a mixture of OpenMP, MPI and RDMA in order to efficiently scale and accelerate the algorithm's computation.
+
+Achieving this goal necessitated overcoming several challenges.
+First, the algorithm's state grows rapidly with larger graphs and number of latent communities.
+Since the full state information is too large to fit in a single machine's memory, it is partitioned and distributed across a cluster of machines.
+Second, to access the full state, each cluster node must read remote memory hosted by its peers.
+We leveraged RDMA to limit the high latency of such operations and increase the communication bandwidth.
+To reduce the latency further, we pipelined the algorithms computations such that data can be fetched in advance over the network.
+Finally, the algorithm's computation had to be effectively distributed across the cluster nodes and parallelized further within each node to exploit their multi-core CPUs.
+
+The remainder of this paper is organized as follows.
+Section~2 provides and overview of the algorithm and its theoretical foundation.
+The design and implementation of the parallel and distributed solution is presented in Section~3.
+Section~4 evaluates the efficacy of the system and analyzes its performance.
+Finally, Section~5 provides concluding remarks.
+
+% -------\\
+% Community detection is the central problem in network analysis, with the goal of identifying the groups of related nodes that are densely connected within this group but sparsely connected to the rest of the network.
+Different from classical community detection problem where we assume each node belongs to one single community, our paper considers overlapping communities where each of the nodes might belong to multiple communities.
+In particular, we consider the model called a-MMSB(Assortive Mixed Membership Stochastic Blockmodel) in this paper, which was first introduced in~\cite{gopalan2012scalable}.\\
+
+% a-MMSB, as a probabilistic graphical model, represents a convenient paradigm for modeling complex relationships between a potentially large number of random variables.
+It also uses priors and posteriors to quantify model uncertainty and facilitate model selection and averaging.
+While a-MMSB provides rich representation power, like other Bayesian models, the inference procedures of handling big data which is common in real world is still a very challenging problem.
+\\
+
+% Consider the large networks such as a social network, it easily runs into billions of edges and tens of million of nodes. In addition to that, the number of communities might exceed few millions. 
+% There were two types of scalable algorithm for a-MMSB have been introduced recently \cite{gopalan2012scalable}, stochastic variational Bayesian inference (SVB) and stochastic gradient Markov chain Monte Carlo(SG-MCMC), respectively.
+Both methods have the important property that each iteration only relies on a small subset of the data.
+Although both methods can work for some large networks with many hundreds of nodes, the performance is far less satisfactory when it applies to the so-called "big data" such as facebook network with millions of nodes and billions of edges.
+Given the facts that \cite{LiAW15}, SG-MCMC runs faster and converges to the better local minima, in this paper, we mainly consider the problem of scaling up SG-MCMC to the big data set.
+\textit{As far as we know, this is the first paper that studies community detection on the frencter dataset with few billion's of edges.}

### 2016-01-15 02-related.tex

@@
 % In this section, we briefly review the Bayesian model we are considering in this paper.
-\subsection{Assortative mixed-membership stochastic blockmodel (a-MMSB)}
-Assortative mixed-membership stochastic blockmodel (a-MMSB) ~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group-structure in a network of $N$ nodes.
-We assume $\cV*$ is the set containing all the nodes in the graph, and $\cE*$ is the set containing all the linked edges between pairs of nodes.
-For each node $a$ in the node set $\cV^*$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
-For every possible peer $b$ in the network, each node $a$ randomly draws a community $z_{ab}$.
-If a pair of nodes ($a,b$) in the edge set $\cE^*$ are in the same community: $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
+\subsection{Assortative Mixed-membership Stochastic Blockmodels (a-MMSB)}
+The assortative mixed membership stochastic blockmodel (a-MMSB) ~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group-structure in a network of $N$ vertices.
+Consider a set  $\cV^*$ containing all the vertices in the graph, and a set $\cE^*$ containing all the linked edges between pairs of vertices.
+Each vertex $a$ in the vertex set $\cV^*$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
+For every possible peer $b$ in the network, each vertex $a$ randomly draws a community $z_{ab}$.
+If a pair of vertices ($a,b$) in the edge set $\cE^*$ are in the same community, i.e., $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
 Otherwise this probability is small.
 Each community has its connection strength $\beta_{k} \in (0,1)$ which explains how likely its members are linked to each other.
-\\
-
-The whole generative process of a-MMSB is given by,
-
+The whole generative process of a-MMSB is then described by,
 \benum
 \item For each community $k$, draw community strength $\bt_k \sim \mbox{Beta}(\eta)$
-\item For each node $a$, draw community memberships $\pi_a \sim \mbox{Dirichlet}(\al)$
-\item For each pair of nodes $a$ and $b$,
+\item For each vertex $a$, draw community memberships $\pi_a \sim \mbox{Dirichlet}(\al)$
+\item For each pair of vertices $a$ and $b$,
 \benum
@@
 
-\subsection{Stochastic Gradient MCMC Algorithms}
+\subsection{Stochastic Gradient Markov Chain Monte Carlo}
+The algorithm we are working on in this paper is based on the stochastic gradient Langevin dynamics (SGLD) ~\cite{welling2011bayesian}.
+SLGD applies the following update rule to obtain samples from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ of $N$ i.i.d.
+data points $\cX=\{x_i\}_{i=1}^{N}$:
+\bea
+\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\cD_n)\right) + \xi_t, \label{eqn:sgld_update}
+\eea
+where $\xi_t \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\cD_n$ a minibatch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\cD_n)$ the mean stochastic gradient, i.e., $\frac{1}{|\cD_n|}\sum_{x\in \cD_n}^{}\grad_{\ta} \log p(x|\ta)$.
+As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution.
+One benefit of using SGLD is that we do not need the Metropolis-Hastings (MH) accept-reject tests since the rejection probability goes to zero as the step size collapses to zero.
+Although the finite step size, which we use in practice rather than converging to zero, still results in some bias, we can reduce the overall error by drawing many more samples per unit time.
 
-The algorithm we are working on in this paper is based on the stochastic gradient Langevin dynamics (SGLD) ~\cite{welling2011bayesian}.
-SLGD applies the following update rule to sample from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ given $N$ i.i.d.
-data points $\cX=\{x_i\}_{i=1}^{N}$
-
-\bea
-\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\cD_n)\right) + \xi, \label{eqn:sgld_update}
-\eea
-
-where $\xi \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\cD_n$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\cD_n) = \frac{1}{|\cD_n|}\sum_{x\in \cD_n}^{}\grad_{\ta} \log p(x|\ta)$.
-As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution.
-One benefit of using SGLD is that we don't need Metropolis-Hastings (MH) accept-reject tests since the rejection probability goes to zero as the step size collapses to zero.
-Although the finite step size still results in some bias in practice, we can reduce the overall error by drawing many more samples per unit time.
-
-SGLD originated from the Langevin Monte Carlo (LMC)~\cite{girolami2011riemann} , where the gradient is computed exactly using all data points and then a Metropolis-Hastings accept-reject test is applied.
-Comparing to LMC,  SGLD only requires to process a mini-batch $\cD_n$ at each iteration and ignores the MH test, which substantially reduces the computational complexity from  $\cO(N)$ to $\cO(n)$ .
+SGLD originated from the Langevin Monte Carlo (LMC)~\cite{girolami2011riemann}, where unlike SGLD the gradient is computed exactly by using all data points.
+Then, Metropolis-Hastings accept-reject tests are applied.
+Comparing to LMC, SGLD only requires to process a minibatch $\cD_n$ at each iteration and ignores the MH test, and thus the computational complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
 
 The stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex.
-By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch estimator in Eqn.
+By applying Riemannian geometry~\cite{girolami2011riemann} and using the minibatch-based estimator in Eqn.
 \ref{eqn:sgld_update}, it achieved the state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}.
 In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$.
-Then, the update rule becomes
-
+Then, the update rule of SGRLD is:
 \bea
-\ta_{k}^* \law \left| \ta_{k} + \f{\ep}{2} \left( \al - \ta_{k} + \f{N}{|\cD_n|} \sum_{d \in \cD_n} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{} \right|. \label{eqn:sgrld_update}
+\ta_{k}^* \law \left| \ta_{k} + \f{\ep_t}{2} \left( \al - \ta_{k} + \f{N}{|\cD_n|} \sum_{d \in \cD_n} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{t} \right|. \label{eqn:sgrld_update}
 \eea
@@
 \subsection{Scalable MCMC for a-MMSB}
-
+This section describes the SG-MCMC algorithm for a-MMSB.
+Please refer \cite{LiAW15} for more details.
 The algorithm iterates updating local parameter $\pi$ and a global parameter $\beta$.
@@
 Here, the parameters $\phi$ and $\ta$ are used to re-parameterize $\pi$ and $\beta$ respectively.
-After updating  $\phi$ and $\ta$,  then we can obtain $\pi$ and $\beta$ by normalizing them.
-In the following, we briefly sketch the iterative updating steps for a-MMSB.
-For more details about the algorithm, please refer to~\cite{LiAW15}
+After updating $\phi$ and $\ta$, we can obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$, respectively.
+In the following, we briefly sketch the iterative update steps.
 
-\paragraph{Sampling global parameters} The update rule for global parameter $\theta$ is
+\paragraph{\textbf{Sampling global parameters}} The update rule for global parameter $\theta$ is
 \bea
@@
 \eea
-
 where 
@@
 \eea
+here $\cE_{n_t}$ is a mini-batch of $n_t$ vertex pairs sampled from $\cE^*$, and $h(\cE_{n_t})$ is the weighting terms and 
+\bea
+f_{ab}^{(y)}(k,l)=
+\begin{cases}
+\bt_k^y(1-\bt_k)^{(1-y)}\pi_{ak}\pi_{bk}, & \mbox{if } k=l\\
+\dt^{y}(1-\dt)^{(1-y)} \pi_{ak} \pi_{bl}, & \mbox{if } k \neq l \nn
+\end{cases}
+\label{eqn:case}
+\eea
+$Z_{ab}^{(y)}$ is the normalization constant which we can compute in $\cO(K)$ time \cite{LiAW15}.
 
-Here $\cE_{n_t}$ is a mini-batch of $n_t$ node pairs sampled from $\cE^*$, and $h(\cE_{n_t})$ is the weighting terms. \\
-
-\paragraph{Sampling local parameters}
+\paragraph{\textbf{Sampling local parameters}}
 The update rule for the local parameters $\phi$ is
@@
 \eea
-
 where \bea
@@
 \eea
+Here, the $\cV_n$ is a random mini-batch of $n$ vertices sampled from $\cV^*$. Note that $|\cV_n| \ll |\cV^*|=N$.
 
-Here, the $\cV_n$ is a random mini-batch of $n$ nodes sampled from $\cV^*$. Note that $|\cV_n| \ll |\cV^*|=N$.
-
-The overall pseudo code is shown in Algorithm 1.
+The overall pseudo-code of the sequential algorithm is described in Algorithm 1.
 
 \begin{algorithm}[t]
-\caption{Pseudo Code}\label{alg}
+\caption{Sequential version of SG-MCMC for a-MMSB}\label{alg}
 \begin{algorithmic}[1]
@@
 \WHILE {sampling}
-\STATE Sample a mini-batch of node pairs, $\cE_t$, from $E$
-\FOR {each node in $\cE_t$}
-	\STATE Sample a mini-batch of nodes, $\cV_t$, from $V$
+\STATE Sample a mini-batch of vertex pairs, $\cE_t$, from $E$
+\FOR {each vertex in $\cE_t$}
+	\STATE Sample a mini-batch of vertices, $\cV_t$, from $V$
     \STATE Update $\phi_{a}$ using Eqn~\ref{eqn:local_update}

### 2016-01-15 03-design.tex

@@
 through the network cards (NICs) without involvement from the remote host
-altogether~\cite{Hamada_infinibandtrade,RDMA,Beck:2011:PER:2043535.2043537}. This has opened up the development of
+altogether~\cite{Hamada_infinibandtrade,Beck:2011:PER:2043535.2043537}.
+This has opened up the development of
 extremely efficient remote memory services, often in the form of DKV stores.
@@
 values. Our use case allows us to do any operation in exactly one RDMA read
-or RDMA write. The Herd paper~\cite{Herd} argues that replacing RDMA reads by
+or RDMA write. The Herd paper~\cite{Kalia:2014:URE:2740070.2626299,} argues
+that replacing RDMA reads by
 an RPC to the server, followed by an unreliable write, is faster. However,

### 2016-01-15 04-evaluation.tex

@@
 identical, which corroborates the results from the Herd project for payloads
-upwards from 256B. The bandwidth achieved by our DKV store mostly falls short
-of the \textit{qperf} performance. For smallish packets, this is attributed
+upwards from 256B. The bandwidth achieved by our DKV store falls short
+of the \textit{qperf} performance for packets less than 4KB; this is attributed
 to additional per-request overhead for the DKV store. For the largest packet

### 2016-01-15 03-design.tex
Some typos, one or another citation

@@
 values. Our use case allows us to do any operation in exactly one RDMA read
-or RDMA write. The Herd paper~\cite{Kalia:2014:URE:2740070.2626299,} argues
+or RDMA write. The Herd paper~\cite{Kalia:2014:URE:2740070.2626299} argues
 that replacing RDMA reads by

### 2016-01-15 04-evaluation.tex
Some typos, one or another citation

@@
 
-Empirical results were obtained by performing experiments on the VU and TU/D
-DAS5 clusters which consist of 68 and 48 compute nodes respectively. Each
+Empirical results were obtained by performing experiments on the VU and
+Leiden University
+DAS5 clusters which consist of 68 and 24 compute nodes respectively. Each
 compute node is equipped with a dual 8-core Intel Xeon E5-2630v3 CPU clocked
@@
 each site are interconnected by
-FDR~InfiniBand. All of the experiments reported in this section utilized
+FDR~InfiniBand. The MPI implementation used is MVAPICH2, which is optimally
+tuned for Infiniband~\cite{mvapich}. All of the experiments reported in this section utilized
 publicly available graphs from the Stanford Large Network Dataset
@@
 A deeper analysis of the individual execution phases of the algorithm provides
-insights into the scalability if its building blocks.
+insights into the scalability of its building blocks.
 %
@@
 Figure~\ref{fig-strong-scaling}-a. As is clearly shown, the dominant phase of
-the execution is update\_pi.
+the execution is \textit{update\_pi}.
 The reported total time for each cluster size is significantly less than the
 sum of the execution times of the individual phases. This is due to the
-overlapping execution of the two most expensive phases, namely, update\_pi and
-deploy\_mini-batch. Both of these phases initially gain significant speedup with the
+overlapping execution of the two most expensive phases, namely,
+\textit{update\_pi} and
+\textit{deploy\_minibatch}. Both of these phases initially gain significant speedup with the
 addition of compute nodes. However, the speedup curve gradually slows down for
@@
 %
-The time spent in update\_beta remains relatively constant across cluster sizes
+The time spent in \textit{update\_beta} remains relatively constant across cluster sizes
 as it performs an insignificant amount of work compared to the synchronization
@@
 time per algorithm iteration across different cluster sizes. On the other hand,
-Figure~\ref{fig-weak-scaling}-b reports the set number of communities for each
+Figure~\ref{fig-weak-scaling}-b reports the number of communities for each
 cluster configuration. The relative change in the average execution time per
@@
 Figure~\ref{fig-pipeline} presents the execution time of 1024 algorithm
-iterations on a 64-node cluster with double-bufferring enabled and disabled.
+iterations on a 64-node cluster with double-buffering enabled and disabled.
 Naturally, increasing the number of communities causes a proportional increase
@@
 algorithm is the inherent complexity of communication and synchronization. A
-slightly simpler approach would be developing a multi-threaded version and
+considerably simpler approach would be developing a multi-threaded version and
 running it on a machine with abundant memory and CPU cores. In such a context,
@@
 memory. The physical machine underlying
-the HPC Cloud system contains 40-CPU cores and does not oversubscribe
+the HPC Cloud system contains 40~CPU cores and does not oversubscribe
 resources. Therefore, by provisioning all 40~cores we ensured that there is no
@@
 algorithm on the HPC Cloud
-system with 40 and 16 cores compared to a single DAS5 node with 16-cores. This
+system with 40 and 16 cores compared to a single DAS5 node with 16~cores. This
 test uses the com-DBLP data set to enable the use of a large number of
@@
 Further, Figure~\ref{fig-scale-up} tests the performance of the HPC Cloud
-system compared to 64-nodes of the DAS5
+system compared to 64~nodes of the DAS5
 cluster using the com-Friendster data set. Clearly, the parallel and
@@
 %
-The results in sub-figures a, b and c where obtained by using 64 compute nodes
+The results in sub-figures a, b and c were obtained by using 64 compute nodes
 of DAS5. In Figure~\ref{fig-ppx}-a, the number of communities was set to 12K
@@
 the algorithm increases dramatically with larger K. However, the system was
-still capable of reporting results after \textbf{XXX time}.
+capable of reporting results for both after 40 hours.
 %

### 2016-01-15 03-design.tex

@@
 
-The SC-MCMC algorithm described in the previous section has an abundance of
+The SG-MCMC algorithm described in the previous section has an abundance of
 opportunities for parallelism, for the multithreaded, shared-memory type as
@@
 parallelism is speed-up of the computation. A distributed implementation
-additionaly allows to store data in the collective memory of the cluster
-machine and increases memory bandwidth; these both scale with the number of
+additionally allows to store data in the collective memory of the cluster
+machine and increases memory bandwidth which scales with the number of
 machines. The downside of a distributed implementation is that it requires
@@
 In most places, the usage of multithreaded parallelism
-is straightforward; we will discuss details only where appropriate. The
+is straightforward, therefore, we will discuss details only where appropriate. The
 distributed design follows a master-worker paradigm, where in essence
@@
 
+{\Large\bf NOTE: introduce perplexity}
+
+As is common in machine learning problems, the edges of the graph are divided
+into two sets a training set $G$ and a held-out set $H$.
 The largest data structures of the algorithm are $G$, $H$, $\pi$ and~$\phi$.
@@
 the calculations in the update stages only require the subset of $G$ that
-is touched by the mini-batch nodes, so the master scatters that subset to
-the workers together with the scattering of the mini-batch nodes. This way,
+is touched by the mini-batch vertices, so the master scatters that subset to
+the workers together with the scattering of the mini-batch vertices. This way,
 we trade a reduction in memory at the workers against the cost of limited
@@
 This section describes the parallelization of each of the stages of the
-algorithm's main loop. All but the first stage justify parallelisation for
+algorithm's main loop. All but the first stage justify parallelization for
 high values of combinations of the  parameters $M$, $K$, $n$, and $|H|$.
 
-The first stage, mini-batch selection, is done at the master's, and it is
-not itself parallelized. However, in the distributed implementation its
-execution is overlapped using pipeline parallelism, while the workers are
+The first stage, mini-batch selection, is done at the master and it is
+not itself parallelized. However, the distributed implementation overlaps
+its execution using pipeline parallelism while the workers are
 performing \textit{update\_phi}. The mini-batch is partitioned equally over
-the workers; the relevant sections of $G$ are distributed together with the
+the workers and the relevant sections of $G$ are distributed together with the
 mini-batch subsets.
@@
 After a worker has received its subset of the mini-batch, it samples a neighbor
-set of size $n$ for each of its mini-batch nodes, using thread parallelism.
+set of size $n$ for each of its mini-batch vertices, using thread parallelism.
 
@@
 K$ operations. It can be fully parallelized because it is a data-parallel
-operation over each of the mini-batch nodes. \textit{update\_phi} loads the
-$\pi$ values for the local mini-batch nodes and their neighbors from the
-DKV store. The updates to $\phi$ for the mini-batch nodes are calculated
+operation over each of the mini-batch vertices. \textit{update\_phi} loads the
+$\pi$ values for the local mini-batch vertices and their neighbors from the
+DKV store. The updates to $\phi$ for the mini-batch vertices are calculated
 independently using thread parallelism.
@@
 These updates are used by \textit{update\_pi} locally, besides the value of
-$\pi/\phi$ for the mini-batch nodes. The update to $\pi$ requires $M\times
-K$ operations, and it is done in parallel over mini-batch nodes. Because of
+$\pi/\phi$ for the mini-batch vertices. The update to $\pi$ requires $M\times
+K$ operations, and it is done in parallel over mini-batch vertices. Because of
 memory consistency, this stage awaits completion of \textit{update\_phi} with
@@
 machines to calculate contributions to $\nabla\theta$. The values of $\pi$ for
-the local mini-batch nodes are loaded from the DKV store. These calculations
-are done with thread parallelism over the mini-batch nodes. In the second step,
-a multithreaded summation is performed to calculate the contributions per
+the local mini-batch vertices are loaded from the DKV store. These calculations
+are done with thread parallelism over the mini-batch vertices. In the second step,
+a multi-threaded summation is performed to calculate the contributions per
 machine, followed by a distributed summation using an MPI reduce operation. The
-third step calculates $\beta$ from $\theta$, sequentially at the master's; this
+third step calculates $\beta$ from $\theta$ sequentially at the master. This
 operation takes only $K$ steps. The resulting $\beta$ is broadcast with MPI
@@
 machines owns a subset of $H$. It loads up-to-date values of $\pi$ for each
-node in its part of $H$, then for each of its edges in $H$ it calculates the
+vertex in its part of $H$, then for each of its edges in $H$ it calculates the
 contribution to the perplexity using thread parallelism. The resulting
@@
 calculation of a chunk is overlapped with loading the next chunk of $\pi$
-values, and also with sampling the mini-batch at the master's and deploying
+values, and also with sampling the mini-batch at the master and deploying
 it to the workers.
@@
 cluster, where $\pi[i]+\pi^{sum}[i]$ is the value for key $i$. We decided not
-to use any of the existing implementations of RDMA DKV stores, and in stead
+to use any of the existing implementations of RDMA DKV stores, and instead
 build our own DKV store on top of the ib-verbs API~\cite{ib-verbs}. The
@@
 partitioning of KV pairs over the machines. The values are all of the same size,
-a vector of $K+1$ floats. The access pattern is very regular, because the
-computation is partitioned into stages that are separated by barriers; a
-stage either reads values or it updates values that it 'owns', so there
-is no read/write or write/write contention. The existing DKV stores incur
+a vector of $K+1$ floats. The access pattern is well controlled because the
+computation is partitioned into stages that are separated by barriers. A
+stage either reads values or updates values. The algorithm ensures that updates
+are always targeting unique elements,
+therefore, there are no read/write or write/write hazards. The existing DKV stores incur
 overhead for insert/delete flexibility, concurrency control, or variable-sized
 values. Our use case allows us to do any operation in exactly one RDMA read
-or RDMA write. The Herd paper~\cite{Kalia:2014:URE:2740070.2626299} argues
+or RDMA write. Herd~\cite{Kalia:2014:URE:2740070.2626299} argues
 that replacing RDMA reads by

### 2016-01-15 04-evaluation.tex
More evaluation edits

--- 2016-01-15 04-evaluation.tex
@@
 %
-Further, Figure~\ref{fig-scale-up} tests the performance of the HPC Cloud
+Further, Figure~\ref{fig-scale-up}-b tests the performance of the HPC Cloud
 system compared to 64-nodes of the DAS5
@@
 
-A thorough performance evaluation of our RDMA DKV store is outside the scope of
-this paper. In figure~\ref{fig-qperf}, we provide maximum bandwidth numbers for
+Since the distributed algorithm is data-intensive, a key aspect in improving
+its performance is to maximize the utility of the network resources.
+In figure~\ref{fig-qperf}, we provide maximum bandwidth numbers for
 read operations between one server and one client for a range of payload sizes,
 and compare these to the bandwidth achieved by \textit{qperf}, which shows
-the best achievable preformance for Infiniband. We present \textit{qperf}
+the best achievable performance for Infiniband. We present \textit{qperf}
 bandwidth for both RDMA read and RDMA write operations; these are nearly
@@
 the algorithm increases dramatically with larger K. However, the system was
-still capable of reporting results after \textbf{XXX time}.
+still capable of reporting results after 25-30 hours.
 %

### 2016-01-15 04-evaluation.tex

--- 2016-01-15 04-evaluation.tex
@@
 %
-Further, Figure~\ref{fig-scale-up} tests the performance of the HPC Cloud
-system compared to 64~nodes of the DAS5
+Further, Figure~\ref{fig-scale-up}-b tests the performance of the HPC Cloud
+system compared to 64-nodes of the DAS5
 cluster using the com-Friendster data set. Clearly, the parallel and

### 2016-01-15 02-related.tex
Steal the paragraph on perplexity

@@
 \subsection{Assortative Mixed-membership Stochastic Blockmodels (a-MMSB)}
-The assortative mixed membership stochastic blockmodel (a-MMSB) ~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group-structure in a network of $N$ vertices.
-Consider a set  $\cV^*$ containing all the vertices in the graph, and a set $\cE^*$ containing all the linked edges between pairs of vertices.
+The assortative mixed membership stochastic blockmodel (a-MMSB) ~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group structure in a network of $N$ vertices.
+Consider a set  $\cV^*$ containing all the vertices in the graph $G$, and a set $\cE^*$ containing all the linked edges between pairs of vertices.
 Each vertex $a$ in the vertex set $\cV^*$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
@@
 Each community has its connection strength $\beta_{k} \in (0,1)$ which explains how likely its members are linked to each other.
-The whole generative process of a-MMSB is then described by,
+The whole generative process of a-MMSB is then described by:
 \benum
@@
 \subsection{Stochastic Gradient Markov Chain Monte Carlo}
-The algorithm we are working on in this paper is based on the stochastic gradient Langevin dynamics (SGLD) ~\cite{welling2011bayesian}.
+The algorithm we are working on in this paper is based on stochastic gradient Langevin dynamics (SGLD) ~\cite{welling2011bayesian}.
 SLGD applies the following update rule to obtain samples from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ of $N$ i.i.d.
@@
 \eea
-where $\xi_t \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\cD_n$ a minibatch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\cD_n)$ the mean stochastic gradient, i.e., $\frac{1}{|\cD_n|}\sum_{x\in \cD_n}^{}\grad_{\ta} \log p(x|\ta)$.
+where $\xi_t \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\cD_n$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\cD_n)$ the mean stochastic gradient, i.e., $\frac{1}{|\cD_n|}\sum_{x\in \cD_n}^{}\grad_{\ta} \log p(x|\ta)$.
 As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution.
@@
 Then, Metropolis-Hastings accept-reject tests are applied.
-Comparing to LMC, SGLD only requires to process a minibatch $\cD_n$ at each iteration and ignores the MH test, and thus the computational complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
+Comparing to LMC, SGLD only requires to process a mini-batch $\cD_n$ at each iteration and ignores the MH test, and thus the computational complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
 
 The stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex.
-By applying Riemannian geometry~\cite{girolami2011riemann} and using the minibatch-based estimator in Eqn.
+By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch-based estimator in Eqn.
 \ref{eqn:sgld_update}, it achieved the state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}.
@@
 Here, the $\cV_n$ is a random mini-batch of $n$ vertices sampled from $\cV^*$. Note that $|\cV_n| \ll |\cV^*|=N$.
+
+As the performance metric we use perplexity, which is defined as the exponential
+of the negative average log-likelihood of the held-out set $\cE_h$. Given a collection of $T$
+samples of the model parameters $\{\bt_t\}$ and $\{\pi_t\}$, the averaged
+perplexity on the held-out test set $\cE_h$ is 
+\begin{align}
+&\text{perp}_{\text{avg}}(\cE_h|\{\beta_t\},\{\pi_t\})\nonumber \\=&
+\exp\left(-\frac{\sum_{(a,b)\in \cE_h}^{}\log
+\{(1/T)\sm{t}{T}p(y_{ab}|\beta_t,\pi_t)\}}{|\cE_{h}|}\right)
+\end{align}
+The perplexity is not evaluated at every iteration, but at regular intervals.
 
@@
 By applying Riemannian geometry [15] and using the mini-batch estimator in Eqn.
-\ref{eqn:sgld_update}, it achieved the state-of-the-art performance for latent Dirichlet allocation (LDA).
+\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA).
 In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$.

### 2016-01-15 03-design.tex
Start to switch using their symbols. Add ground truth values

@@
 As is common in machine learning problems, the edges of the graph are divided
-into two sets a training set $G$ and a held-out set $H$.
-The largest data structures of the algorithm are $G$, $H$, $\pi$ and~$\phi$.
+into two sets, a training set $G$ and a held-out set $\cE_h$.
+The largest data structures of the algorithm are $G$, $\cE_h$, $\pi$ and~$\phi$.
 For the largest dataset in this paper, com-Friendster, $G$ has 1.8~billion
@@
 we trade a reduction in memory at the workers against the cost of limited
-communication. In contrast, $H$ can be partitioned statically over all
+communication. In contrast, $\cE_h$ can be partitioned statically over all
 machines for the parallel perplexity calculation.
@@
 
-{\Large\bf NOTE: Discuss the lack of locality? Here?}
+The distribution within the graph of the nodes of the mini-batch as well as
+the neighbor sets is completely random. That means that there is no locality
+in $\pi$ access patterns, especially in the accesses to the neighbor sets.
+Hence, there is no opportunity to exploit caching of $\pi$ through data
+locality.
 
@@
 algorithm's main loop. All but the first stage justify parallelization for
-high values of combinations of the  parameters $M$, $K$, $n$, and $|H|$.
+high values of combinations of the  parameters $|\cV_n|$, $K$, $n$, and $|\cE_h|$.
 
-The first stage, mini-batch selection, is done at the master and it is
+The first stage, mini-batch selection (line 3 in Algorithm~1), is done at the master and it is
 not itself parallelized. However, the distributed implementation overlaps
@@
 After a worker has received its subset of the mini-batch, it samples a neighbor
-set of size $n$ for each of its mini-batch vertices, using thread parallelism.
+set of size $n$ for each of its mini-batch vertices, using thread parallelism
+(line 5 in Algorithm~1).
 
-The next stage, \textit{update\_phi}, is the algorithm's dominant stage, not
+The next stage, \textit{update\_phi} (line 6 in Algorithm~1), is the algorithm's dominant stage, not
 only in calculation but also in memory accesses. It performs $M\times n\times
@@
 
-These updates are used by \textit{update\_pi} locally, besides the value of
+These updates are used by \textit{update\_pi} (line 7 in Algorithm~1) locally, besides the value of
 $\pi/\phi$ for the mini-batch vertices. The update to $\pi$ requires $M\times
@@
 
-\textit{update\_beta} requires $M\times K$ operations. It is preceded by an MPI
+\textit{update\_beta} (lines 10 and 11 in Algorithm~1) requires $M\times K$ operations. It is preceded by an MPI
 barrier to ensure that up-to-date $\pi$ values are read. \textit{update\_beta}
@@
 
-The calculation of perplexity requires $|H|\times K$ steps. Each of the
-machines owns a subset of $H$. It loads up-to-date values of $\pi$ for each
-vertex in its part of $H$, then for each of its edges in $H$ it calculates the
+The calculation of perplexity requires $|\cE_h|\times K$ steps. Each of the
+machines owns a subset of $\cE_h$. It loads up-to-date values of $\pi$ for each
+vertex in its part of $\cE_h$, then for each of its edges in $\cE_h$ it calculates the
 contribution to the perplexity using thread parallelism. The resulting

### 2016-01-15 04-evaluation.tex
Start to switch using their symbols. Add ground truth values

--- 2016-01-15 04-evaluation.tex
@@
   \centering
-  \begin{tabular}{l r r l}
-    Name            & \#Vertices &       \#Edges & Description \\
+  \begin{tabular}{l r r r l}
+    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
+                    &            &               & communities    &             \\
     \hline
-    com-LiveJournal &  3,997,962 &    34,681,189 & Online blogging social network \\
-    com-Friendster  & 65,608,366 & 1,806,067,135 & Online gaming social network \\
-    com-Orkut       &  3,072,441 &   117,185,083 & Online social network \\
-    com-Youtube     &  1,134,890 &     2,987,624 & Video-sharing social network \\
-    com-DBLP        &    317,080 &     1,049,866 & Computer science bibliography collaboration network \\
-    com-Amazon      &    334,863 &       925,872 & Product co-purchasing network \\
+    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
+    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
+    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
+    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
+    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
+    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
     \hline
@@
 \textit{update\_pi} and
-\textit{deploy\_minibatch}. Both of these phases initially gain significant speedup with the
+minibatch deployment. Both of these phases initially gain significant speedup with the
 addition of compute nodes. However, the speedup curve gradually slows down for

### 2016-01-15 00-abstract.tex
Adding abstract

@@
 % 4. what are your results?
+
+Recent advancements in machine learning algorithms have transformed the data
+analytics domain and provided innovative solutions to inherently difficult
+problems. However, training models at scale over large data sets remains a
+daunting challenge. One such problem is the detection of overlapping
+communities within graphs. For example, a social network can be modeled as a
+graph where the vertices and edges represent individuals and their
+relationships. In this paper, we present and evaluate an efficient parallel and
+distributed implementation of a Stochastic Gradient Markov Chain Monte Carlo
+algorithm that solves the overlapping community detection problem. We show that
+the algorithm can scale and process graphs consisting of billions of edges and
+tens of millions of vertices on a compute cluster of 65 nodes.

### 2016-01-16 05-conclusion.tex
Adding conclusion

\section{Conclusion}

The recent advancements of machine learning algorithms make them ideal
candidates to solve complex problems. However, using these solutions in order
to process problems at scale is still a daunting task. In particular, knowledge
of parallel and distributed computing techniques is necessary to facilitate the
development of such systems and streamline their execution time.

In this paper, we shared our experience in developing a highly scalable and
efficient solution for a stochastic gradient Markov chain Monte Carlo algorithm
that detects overlapping communities in graphs. The system design had to
overcome several obstacles in order to achieve high performance. Specifically,
we discussed how the algorithm was structured to facilitate its
parallelization. Moreover, we evaluated the efficacy of overlapping computation
with communication to hide latency.  Further, we demonstrated the use of a
mixture of MPI and RDMA primitives to speedup the communication between cluster
nodes.

We conducted a thorough empirical evaluation of the system to study its strong
and weak scalability on 64 cluster nodes using large data sets.  Additionally,
we assessed the efficiency of the algorithm's resource utilization. Finally, a
demonstration of the implementation's utility was provided by processing 6
different organic data sets.

### 2016-01-16 04-evaluation.tex
Refactor strong scaling plot and text: update_pi -> update_phi_pi, update_beta -> update_beta_theta

@@
 Figure~\ref{fig-strong-scaling}-a. As is clearly shown, the dominant phase of
-the execution is \textit{update\_pi}.
+the execution is \textit{update\_phi\_pi}.
 The reported total time for each cluster size is significantly less than the
@@
 overlapping execution of the two most expensive phases, namely,
-\textit{update\_pi} and
+\textit{update\_phi\_pi} and
 minibatch deployment. Both of these phases initially gain significant speedup with the
@@
 %
-The time spent in \textit{update\_beta} remains relatively constant across cluster sizes
+The time spent in \textit{update\_beta\_theta} remains relatively constant across cluster sizes
 as it performs an insignificant amount of work compared to the synchronization

### 2016-01-16 01-intro.tex
Some other parallel/distributed community detection projects; some commas

@@
 \section{Introduction}
-The tremendous amount of data that we generate through our daily applications such as social networking services, online shopping, and news recommendations, provides us with an opportunity to extract hidden but invaluable and useful information.
+The tremendous amount of data that we generate through our daily applications such as social networking services, online shopping, and news recommendations, provides us with an opportunity to extract hidden but useful, even invaluable information.
 Realizing this opportunity, however, requires a significant amount of effort because traditional machine learning algorithms often become extremely inefficient with large amounts of data.
 
-There have been two main approaches to this end; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed} while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning algorithms more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
-However, it is clear that, in most cases we see the best performance by carefully integrating both of these approaches in one system.
+There have been two main approaches to this end; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning algorithms more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
+However, it is clear that in most cases we see the best performance by carefully integrating both of these approaches in one system.
 
@@
 
-This problem is modeled by the mixed membership stochastic blockmodels (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of the MMSB, called assortative-MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
+This problem is modeled by the mixed membership stochastic blockmodels (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of MMSB, called assortative MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
 
 The MMSB model is a probabilistic graphical model~\cite{koller2009probabilistic} that represents a convenient paradigm for modeling complex relationships between a potentially large number of random variables.
-Bayesian graphical models, where we define priors and infer posteriors over parameters also allow us to quantify model uncertainty and facilitate model selection and averaging.
+Bayesian graphical models, where we define priors and infer posteriors over parameters, also allow us to quantify model uncertainty and facilitate model selection and averaging.
 However, an increasingly urgent question is whether these models and their inference procedures will be up to the challenge of handling very large graphs.
@@
 We leveraged RDMA to limit the high latency of such operations and increase the communication bandwidth.
-To reduce the latency further, we pipelined the algorithms computations such that data can be fetched in advance over the network.
-Finally, the algorithm's computation had to be effectively distributed across the cluster nodes and parallelized further within each node to exploit their multi-core CPUs.
+To reduce the latency further, we pipelined the algorithm's computations such that data can be fetched in advance over the network.
+Finally, the algorithm's computation is effectively distributed across the cluster nodes and parallelized further within each node to exploit their multi-core CPUs.
 
 The remainder of this paper is organized as follows.
-Section~2 provides and overview of the algorithm and its theoretical foundation.
+Section~2 provides an overview of the algorithm and its theoretical foundation.
 The design and implementation of the parallel and distributed solution is presented in Section~3.

### 2016-01-16 02-related.tex
Some other parallel/distributed community detection projects; some commas

@@
 The assortative mixed membership stochastic blockmodel (a-MMSB) ~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group structure in a network of $N$ vertices.
-Consider a set  $\cV^*$ containing all the vertices in the graph $G$, and a set $\cE^*$ containing all the linked edges between pairs of vertices.
+Consider a set  $\cV^*$ containing all the vertices in the graph, and a set $\cE^*$ containing all the linked edges between pairs of vertices.
 Each vertex $a$ in the vertex set $\cV^*$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
@@
 \end{comment}
+
+\subsection{Related work on distributed MCMC}
+
+Among previous work on parallelising MMSB algorithms, we mention the Online
+Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13} on GPUs, and our previous
+parallel implementation of SG-MCMC on GPUs and multicore CPUs (submitted for
+publication). There are a few projects that did a distributed implementation
+for community detection on very large network graphs. In contrast to our work,
+they all detect non-overlapping communities: \cite{Bu2013246} investigate a
+large number of networks; \cite{2015arXiv150302115L} investigate hierarchical
+stochastic blockmodels; \cite{Prat-Perez:2014:HQS:2566486.2568010} use
+multi-core machines.
+
+\begin{comment}
+In this paper, we describe our custom RDMA D-KV (Distributed Key-Value)
+store. Current RMDA D-KV store implementations are RamCloud~\cite{RamCloud},
+Pilaf~\cite{Pilaf}, Herd~\cite{Herd} and FaRM~\cite{FaRM}. All these systems
+use RDMA to implement a D-KV store. However, all of them are far more powerful
+than our custom implementation -- and this power comes at a cost that we
+can avoid. They implement a generic D-KV store that controls concurrency,
+supports dynamic inserts and deletes, supports variable-sized values
+(whose size may change at an update), and keys of arbitrary type. Because
+of the nature of our distributed algorithm, we have to deal with none of
+these issues. For us, values are fixed-size, allocated only at the initial
+population, and remain alive forever. We have no concurrency between writes
+and reads or other writes. Our keys are a contiguous range of integers. All
+these properties together allow an extremely low-overhead implementation
+that does not involve the remote host in any transaction.
+\end{comment}

### 2016-01-16 03-design.tex
Some other parallel/distributed community detection projects; some commas

@@
 
-{\Large\bf NOTE: introduce perplexity}
-
 As is common in machine learning problems, the edges of the graph are divided
@@
 
-The distribution within the graph of the nodes of the mini-batch as well as
+The distribution within the graph of the vertices of the mini-batch as well as
 the neighbor sets is completely random. That means that there is no locality
@@
 to use any of the existing implementations of RDMA DKV stores, and instead
-build our own DKV store on top of the ib-verbs API~\cite{ib-verbs}. The
+build our own DKV store on top of the Infiniband ib-verbs~API. The
 main consideration is that our use case is unusually simple in a number
@@
 are always targeting unique elements,
-therefore, there are no read/write or write/write hazards. The existing DKV stores incur
+therefore, there are no read/write or write/write hazards. The existing DKV
+stores (RamCloud~\cite{Ousterhout:2015:RSS:2818727.2806887}, FaRM~\cite{179767},
+and reputedly the fastest, Herd~\cite{Kalia:2014:URE:2740070.2626299}) incur
 overhead for insert/delete flexibility, concurrency control, or variable-sized
 values. Our use case allows us to do any operation in exactly one RDMA read
-or RDMA write. Herd~\cite{Kalia:2014:URE:2740070.2626299} argues
+or RDMA write. Herd argues
 that replacing RDMA reads by

### 2016-01-16 04-evaluation.tex
Some other parallel/distributed community detection projects; some commas

--- 2016-01-15 04-evaluation.tex
@@
   cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
-  8-nodes}
+  8 nodes}
   \label{fig-strong-scaling}
@@
   \caption{Performance effect of varying the number of communities on the
-    algorithm's execution time on 64-nodes when using single- or
+    algorithm's execution time on 64 nodes when using single- or
     double-buffering.}
@@
 Further, Figure~\ref{fig-scale-up}-b tests the performance of the HPC Cloud
-system compared to 64-nodes of the DAS5
+system compared to 64 nodes of the DAS5
 cluster using the com-Friendster data set. Clearly, the parallel and

### 2016-01-16 04-evaluation.tex
Run time for LJ. Some hyphens.

@@
 for larger cluster sizes. This observation verifies that the system's overall
-overhead is minimal. Additionally, the experimental results shows that the
+overhead is minimal. Additionally, the experimental results show that the
 implementation is capable of achieving good speedups provided the input problem
@@
   \caption{Performance comparison between the distributed implementation
-  running on DAS5 and the multi-threaded solution on machine with 40-cores and
+  running on DAS5 and the multi-threaded solution on machine with 40 cores and
   1TB of RAM.}
@@
 the algorithm increases dramatically with larger K. However, the system was
-capable of reporting results for both after 40 hours.
+capable of reporting results for both in around 40 hours.
 %

### 2016-01-16 02-related.tex
Typo fixes

@@
 The algorithm we are working on in this paper is based on stochastic gradient Langevin dynamics (SGLD) ~\cite{welling2011bayesian}.
-SLGD applies the following update rule to obtain samples from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ of $N$ i.i.d.
+SGLD applies the following update rule to obtain samples from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ of $N$ i.i.d.
 data points $\cX=\{x_i\}_{i=1}^{N}$:
@@
 
-Among previous work on parallelising MMSB algorithms, we mention the Online
+Among previous work on parallelizing MMSB algorithms, we mention the Online
 Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13} on GPUs, and our previous
-parallel implementation of SG-MCMC on GPUs and multicore CPUs (submitted for
+parallel implementation of SG-MCMC on GPUs and multi-core CPUs (submitted for
 publication). There are a few projects that did a distributed implementation

### 2016-01-16 03-design.tex
Typo fixes

@@
 The SG-MCMC algorithm described in the previous section has an abundance of
-opportunities for parallelism, for the multithreaded, shared-memory type as
-well as for distributed-memory parallelism. The benefit of multithreaded
+opportunities for parallelism, for the multi-threaded, shared-memory type as
+well as for distributed-memory parallelism. The benefit of multi-threaded
 parallelism is speed-up of the computation. A distributed implementation
@@
 This section describes how we parallelize the algorithm.
-In most places, the usage of multithreaded parallelism
+In most places, the usage of multi-threaded parallelism
 is straightforward, therefore, we will discuss details only where appropriate. The

### 2016-01-16 01-intro.tex
Typos, phrasings

@@
 
-There have been two main approaches to this end; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning algorithms more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
+There have been two main approaches to this issue; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning algorithms more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
 However, it is clear that in most cases we see the best performance by carefully integrating both of these approaches in one system.
@@
 One such big data problem is analyzing large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of million of vertices~\cite{yang2015defining}.
-In particular, we are interested in the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, provided a set of vertices, the link information between them (which is usually very sparse), and the number of latent communities.
+In particular, we are interested in the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, provided a set of vertices, the links between them (which are usually very sparse), and the number of latent communities.
 A community can be seen as a densely connected group of vertices that are only sparsely connect to the rest of the network.
@@
 
-There have been two main recent advances in this direction of scalable Bayesian inference methods based on stochastic variational Bayes (SVB)~\cite{hoffman2013stochastic,gopalan2013efficient,gopalan2012scalable} and stochastic gradient Markov chain Monte Carlo (SG-MCMC)~\cite{welling2011bayesian,patterson2013stochastic,ahn2014distributed,ahn2012bayesian}.
+There have been two main recent advances in this direction of scalable Bayesian inference: methods based on stochastic variational Bayes (SVB)~\cite{hoffman2013stochastic,gopalan2013efficient,gopalan2012scalable} and stochastic gradient Markov chain Monte Carlo (SG-MCMC)~\cite{welling2011bayesian,patterson2013stochastic,ahn2014distributed,ahn2012bayesian}.
 Both methods have the important property that they only require a small subset of the data for every iteration.
@@
 To reduce the latency further, we pipelined the algorithm's computations such that data can be fetched in advance over the network.
-Finally, the algorithm's computation is effectively distributed across the cluster nodes and parallelized further within each node to exploit their multi-core CPUs.
+Finally, the algorithm's computation is effectively distributed across the cluster nodes and parallelized further within each node by exploiting their multi-core CPUs.
 

### 2016-01-16 02-related.tex
Typos, phrasings

@@
 Consider a set  $\cV^*$ containing all the vertices in the graph, and a set $\cE^*$ containing all the linked edges between pairs of vertices.
-Each vertex $a$ in the vertex set $\cV^*$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
+Each vertex $a$ in the vertex set $\cV^*$ has a $K$-dimensional probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
 For every possible peer $b$ in the network, each vertex $a$ randomly draws a community $z_{ab}$.
@@
 Otherwise this probability is small.
-Each community has its connection strength $\beta_{k} \in (0,1)$ which explains how likely its members are linked to each other.
+Each community has its connection strength $\beta_{k} \in (0,1)$ which reflects how likely its members are linked to each other.
 The whole generative process of a-MMSB is then described by:
@@
 
-\subsection{Related work on distributed MCMC}
+\subsection{Related work on parallel community detection}
 

### 2016-01-16 03-design.tex
Typos, phrasings

@@
 the neighbor sets is completely random. That means that there is no locality
-in $\pi$ access patterns, especially in the accesses to the neighbor sets.
+in $\pi$ access patterns, especially in the accesses to the neighbor sets
+$\cV_t$.
 Hence, there is no opportunity to exploit caching of $\pi$ through data
@@
 algorithm's main loop. All but the first stage justify parallelization for
-high values of combinations of the  parameters $|\cV_n|$, $K$, $n$, and $|\cE_h|$.
+high values of combinations of the  parameters $|\cV_n|$, $K$, $\cV_t$, and $|\cE_h|$.
 
@@
 After a worker has received its subset of the mini-batch, it samples a neighbor
-set of size $n$ for each of its mini-batch vertices, using thread parallelism
+set $\cV_t$ for each of its mini-batch vertices, using thread parallelism
 (line 5 in Algorithm~1).

### 2016-01-16 03-design.tex
Make figs 1x2 i.s.o. 2x1. Move them around

@@
 
+% horrible hack to move the table up
+\begin{table*}
+  \centering
+  \begin{tabular}{l r r r l}
+    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
+                    &            &               & communities    &             \\
+    \hline
+    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
+    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
+    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
+    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
+    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
+    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
+    \hline
+  \end{tabular}
+  \caption{Summary of SNAP graph data sets used for evaluation.}
+  \label{tab-datasets}
+\end{table*}
+
 \subsection{RDMA Distributed Key-Value Store}

### 2016-01-16 04-evaluation.tex
Typo fixes
Make figs 1x2 i.s.o. 2x1. Move them around

@@
 
-\begin{table*}
-  \centering
-  \begin{tabular}{l r r r l}
-    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
-                    &            &               & communities    &             \\
-    \hline
-    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
-    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
-    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
-    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
-    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
-    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
-    \hline
-  \end{tabular}
-  \caption{Summary of SNAP graph data sets used for evaluation.}
-  \label{tab-datasets}
-\end{table*}
-
 Empirical results were obtained by performing experiments on the VU and
@@
 
-\begin{figure*}[t] % [htb]
+\begin{figure}[htb] % [htb]
   \centering
-  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\textwidth}
+  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=0.48\textwidth}
   \caption{(a) Execution time of 2048 algorithm iterations for the
@@
   \label{fig-strong-scaling}
-\end{figure*}
+\end{figure}
 
@@
 \textit{update\_phi\_pi} and
-minibatch deployment. Both of these phases initially gain significant speedup with the
+mini-batch deployment. Both of these phases initially gain significant speedup with the
 addition of compute nodes. However, the speedup curve gradually slows down for
@@
 
-\begin{figure*}[t] % [htb]
+\begin{figure}[htb] % [htb]
   \centering
-  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\textwidth}
+  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=0.48\textwidth}
   \caption{(a) Average execution time per algorithm iteration varying
@@
   \label{fig-weak-scaling}
-\end{figure*}
+\end{figure}
 
@@
   compute nodes, with 12K communities. Times in ms per iteration. The lower
-  half of the table shows substages within \textit{update\_phi}. In the
+  half of the table shows sub-stages within \textit{update\_phi}. In the
   pipelining version, these components are done in parallel.}
@@
 \subsection{Horizontal vs. Vertical Scalability}
-\begin{figure*}[t] % [htb]
+\begin{figure}[htbp] % [htb]
   \centering
-  \epsfig{file=plots/hpc-cloud.eps, width=\textwidth}
+  \epsfig{file=plots/hpc-cloud.eps, width=0.48\textwidth}
   \caption{Performance comparison between the distributed implementation
@@
   \label{fig-scale-up}
-\end{figure*}
+\end{figure}
 One of the main drawbacks of designing a distributed solution for a given
@@
 
-\begin{figure}[b] % [htbp]
+\begin{figure}[t] % [htbp]
   \centering
@@
 \end{figure}
+
+\begin{figure*}[htbp] % [htb]
+  \centering
+  \epsfig{file=plots/ppx.eps, width=\textwidth}
+  \caption{Convergence time of 6 different data sets.}
+  \label{fig-ppx}
+\end{figure*}
 
@@
 \subsection{Convergence of Large Datasets}
-\begin{figure*}[t] % [htb]
-  \centering
-  \epsfig{file=plots/ppx.eps, width=\textwidth}
-  \caption{Convergence time of 6 different data sets.}
-  \label{fig-ppx}
-\end{figure*}
 

### 2016-01-16 plots/plot-hpc-cloud.sh

@@ -63,8 +63,8 @@ set grid ytics
 w = 0.6
 h = 0.4
 k = 0.0
-nr = 1
-nc = 2
+nr = 2
+nc = 1
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
@@ -89,7 +89,7 @@ plot \
      "$$.tmp.hpc16" u 1:(1000*\$2) w lp t 'HPC Cloud 16-cores', \
      "$$.tmp.4" u 1:(1000*\$2) w lp t 'Single DAS5 node 16-cores'
 
-set origin col(1),row(0)
+set origin col(0),row(1)
 set size w,h
 set title '(b) Performance of HPC Cloud (40 cores) vs. 64 DAS5 nodes (64*16 cores) using com-Friendster'
 plot \

### 2016-01-16 plots/plot-sweep-K-proportional-np.sh

@@ -26,8 +26,8 @@ cat <<EOF | gnuplot --persist
 w = 0.6
 h = 0.4
 k = 0.0
-nr = 1
-nc = 2
+nr = 2
+nc = 1
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
@@ -49,7 +49,7 @@ set title "(a) Weak Scaling (over K)"
 set ylabel "Time per Iteration (milliseconds)"
 plot "$$.tmp" u (\$2-1):(1000*\$3/\$4) w lp axis x1y1
 
-set origin col(1),row(0)
+set origin col(0),row(1)
 set size w,h
 set style data histogram 
 set style histogram cluster gap 1

### 2016-01-16 plots/plot-sweep-np-fixed-K.sh

@@ -45,8 +45,8 @@ cat <<EOF | gnuplot --persist
 w = 0.6
 h = 0.4
 k = 0.0
-nr = 1
-nc = 2
+nr = 2
+nc = 1
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
@@ -72,7 +72,7 @@ plot \
      "$$.tmp.3" u (\$2-1):3 w lp axis x1y1 t 'update\_phi\_pi', \
      "$$.tmp.4" u (\$2-1):3 w lp axis x1y1 t 'update\_beta\_theta'
 
-set origin col(1),row(0)
+set origin col(0),row(1)
 set size w,h
 set ylabel "Speedup"
 set title "(b) Speedup over 8-nodes"

### 2016-01-16 04-evaluation.tex
More little textual thingies

@@
 %
-The results in sub-figures a, b and c were obtained by using 64 compute nodes
+The results in sub-figures a, b and c were obtained by using 65 compute nodes
 of DAS5. In Figure~\ref{fig-ppx}-a, the number of communities was set to 12K
-which fully occupied the aggregate memory resources of all 64 machines since
+which fully occupied the aggregate memory resources of all 64 worker nodes since
 com-Friendster has roughly 65 million vertices. In this case, the algorithm
@@
 respectively. As the number of vertices in these data sets is an order of
-magnitude smaller than com-Friendster, we used a larger number of
-communities. Naturally, the convergence time was extended as the complexity of
+magnitude smaller than com-Friendster, we could use a larger number of
+communities to fill up the collective memory. Naturally, the convergence time was extended as the complexity of
 the algorithm increases dramatically with larger K. However, the system was
@@
 Figures~\ref{fig-ppx}-d, \ref{fig-ppx}-e and \ref{fig-ppx}-f were all
-configured to use the total number of ground-truth communities associated with
-their respective data sets, com-Youtube, com-DBLP and com-Amazon. However,
+configured to use the number of ground-truth communities associated with
+their respective data sets, com-Youtube, com-DBLP and com-Amazon. Since this
+yields a much smaller storage requirement,
 these experiments were conducted on 14, 24 and 24 cluster nodes respectively.

### 2016-01-17 01-intro.tex
Some of Henri's comments got lost

@@
 One such big data problem is analyzing large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of million of vertices~\cite{yang2015defining}.
-In particular, we are interested in the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, provided a set of vertices, the links between them (which are usually very sparse), and the number of latent communities.
-A community can be seen as a densely connected group of vertices that are only sparsely connect to the rest of the network.
+In particular, we are interested in the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, given a set of vertices, the links between them (which are usually very sparse), and the number of latent communities.
+A community can be seen as a densely connected group of vertices that are only sparsely connected to the rest of the network.
 
@@
 
-There have been two main recent advances in this direction of scalable Bayesian inference: methods based on stochastic variational Bayes (SVB)~\cite{hoffman2013stochastic,gopalan2013efficient,gopalan2012scalable} and stochastic gradient Markov chain Monte Carlo (SG-MCMC)~\cite{welling2011bayesian,patterson2013stochastic,ahn2014distributed,ahn2012bayesian}.
+There have been two main recent advances in this direction of scalable Bayesian inference: methods based on stochastic variational Bayes (SVB)~\cite{gopalan2012scalable,hoffman2013stochastic,gopalan2013efficient} and stochastic gradient Markov chain Monte Carlo (SG-MCMC)~\cite{welling2011bayesian,patterson2013stochastic,ahn2014distributed,ahn2012bayesian}.
 Both methods have the important property that they only require a small subset of the data for every iteration.

### 2016-01-17 03-design.tex
Use \sum\phi iso \phi^{sum}

@@
 
+%\looseness=-1
 $\pi$~and $\phi$ are 32-bit float arrays of size $K \times N$. For our
 largest distributed experiment, com-Friendster with $N$=64M and $K$=12K, each
-requires~3TB. We decided to store only $\pi$ and $\pi^{sum}$, and recalculate
+requires~3TB. We decided to store only $\pi$ and $\sum\phi$, and recalculate
 $\phi$ from these whenever appropriate. Here, we trade a substantial reduction
@@
 
+%\looseness=-1
 These updates are used by \textit{update\_pi} (line 7 in Algorithm~1) locally, besides the value of
@@
 memory consistency, this stage awaits completion of \textit{update\_phi} with
-an MPI barrier. After the calculation, the updated values of $\pi+\pi^{sum}$
+an MPI barrier. After the calculation, the updated values of $\pi+\sum\phi$
 are written through the DKV store.
@@
 
+%\looseness=-1
 Modern network technology allows access to the memory of remote machines
@@
 This project stores $\pi$ in a DKV store in the collective memory of the
-cluster, where $\pi[i]+\pi^{sum}[i]$ is the value for key $i$. We decided not
+cluster, where $\pi[i]+\sum\phi[i]$ is the value for key $i$. We decided not
 to use any of the existing implementations of RDMA DKV stores, and instead

### 2016-01-17 02-related.tex
One more place to fix 02 and 03 notation (grad theta)

@@
 \subsection{Assortative Mixed-membership Stochastic Blockmodels (a-MMSB)}
-The assortative mixed membership stochastic blockmodel (a-MMSB) ~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group structure in a network of $N$ vertices.
+The assortative mixed membership stochastic blockmodel (a-MMSB)~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group structure in a network of $N$ vertices.
 Consider a set  $\cV^*$ containing all the vertices in the graph, and a set $\cE^*$ containing all the linked edges between pairs of vertices.
@@
 \subsection{Stochastic Gradient Markov Chain Monte Carlo}
-The algorithm we are working on in this paper is based on stochastic gradient Langevin dynamics (SGLD) ~\cite{welling2011bayesian}.
+The algorithm we are working on in this paper is based on stochastic gradient Langevin dynamics (SGLD)~\cite{welling2011bayesian}.
 SGLD applies the following update rule to obtain samples from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ of $N$ i.i.d.
@@
 By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch-based estimator in Eqn.
-\ref{eqn:sgld_update}, it achieved the state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}.
+\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}.
 In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$.

### 2016-01-18 03-design.tex
One more place to fix 02 and 03 notation (grad theta)
Line of text. Float placement

@@
 algorithm's main loop. All but the first stage justify parallelization for
-high values of combinations of the  parameters $|\cV_n|$, $K$, $\cV_t$, and $|\cE_h|$.
+high values of combinations of the parameters mini-batch size $|\cV_n|$,
+number of communities $K$, neighbor sample size $|\cV_t|$, and held-out set
+size $|\cE_h|$.
 
@@
 is split into four steps. The first stage partitions the mini-batch across
-machines to calculate contributions to $\nabla\theta$. The values of $\pi$ for
+machines to calculate contributions to $g_{ab}(\theta)$. The values of $\pi$ for
 the local mini-batch vertices are loaded from the DKV store. These calculations

### 2016-01-19 03-design.tex
Forgot to iron out some instances of M and n

@@
 The next stage, \textit{update\_phi} (line 6 in Algorithm~1), is the algorithm's dominant stage, not
-only in calculation but also in memory accesses. It performs $M\times n\times
+only in calculation but also in memory accesses. It performs $|\cV_n| \times |\cV_t| \times
 K$ operations. It can be fully parallelized because it is a data-parallel
@@
 These updates are used by \textit{update\_pi} (line 7 in Algorithm~1) locally, besides the value of
-$\pi/\phi$ for the mini-batch vertices. The update to $\pi$ requires $M\times
+$\pi/\phi$ for the mini-batch vertices. The update to $\pi$ requires $|\cV_n| \times
 K$ operations, and it is done in parallel over mini-batch vertices. Because of
@@
 
-\textit{update\_beta} (lines 10 and 11 in Algorithm~1) requires $M\times K$ operations. It is preceded by an MPI
+\textit{update\_beta} (lines 10 and 11 in Algorithm~1) requires $|\cV_n| \times K$ operations. It is preceded by an MPI
 barrier to ensure that up-to-date $\pi$ values are read. \textit{update\_beta}

### 2016-01-20 03-design.tex
More RDMA DKV upward. Rephrase pipeline discussion.

--- 2016-01-18 03-design.tex
@@
 
-This section describes how we parallelize the algorithm.
+This section describes how we parallelize the algorithm, which had been
+implemented sequentially in~C++.
 In most places, the usage of multi-threaded parallelism
@@
 
-As is common in machine learning problems, the edges of the graph are divided
-into two sets, a training set $G$ and a held-out set $\cE_h$.
-The largest data structures of the algorithm are $G$, $\cE_h$, $\pi$ and~$\phi$.
-For the largest dataset in this paper, com-Friendster, $G$ has 1.8~billion
+The largest data structures of the algorithm are \Edges, $\cE_h$, $\pi$
+and~$\phi$, see Table~\ref{tbl-symbols}.
+For the largest dataset in this paper, com-Friendster, \Edges has 1.8~billion
 undirected edges. In our representation with directed edges, this takes
-up 13.5GB. Our design lets $G$ reside only at the master. We observe that
-the calculations in the update stages only require the subset of $G$ that
+up 13.5GB. Our design lets \Edges reside only at the master's. We observe that
+the calculations in the update stages only require the subset of \Edges that
 is touched by the mini-batch vertices, so the master scatters that subset to
 the workers together with the scattering of the mini-batch vertices. This way,
-we trade a reduction in memory at the workers against the cost of limited
-communication. In contrast, $\cE_h$ can be partitioned statically over all
-machines for the parallel perplexity calculation.
+we trade a reduction in memory usage at the workers against limited
+communication costs. In contrast, $\cE_h$ is be partitioned statically over
+all machines for the parallel perplexity calculation.
 
@@
 requires~3TB. We decided to store only $\pi$ and $\sum\phi$, and recalculate
-$\phi$ from these whenever appropriate. Here, we trade a substantial reduction
-in memory against limited computation. $\pi$ is partitioned across the workers,
+$\phi$ from these whenever appropriate. Here, we trade a substantial gain
+in memory usage against some computation. $\pi$ is partitioned across the workers,
 and $\pi$ values are accessed via a DKV (distributed key-value) store.
@@
 in $\pi$ access patterns, especially in the accesses to the neighbor sets
-$\cV_t$.
-Hence, there is no opportunity to exploit caching of $\pi$ through data
-locality.
-
-% each of the stages
-
-\subsection{Implementation of distributed parallelism}
-
-This section describes the parallelization of each of the stages of the
-algorithm's main loop. All but the first stage justify parallelization for
-high values of combinations of the parameters mini-batch size $|\cV_n|$,
-number of communities $K$, neighbor sample size $|\cV_t|$, and held-out set
-size $|\cE_h|$.
-
-The first stage, mini-batch selection (line 3 in Algorithm~1), is done at the master and it is
-not itself parallelized. However, the distributed implementation overlaps
-its execution using pipeline parallelism while the workers are
-performing \textit{update\_phi}. The mini-batch is partitioned equally over
-the workers and the relevant sections of $G$ are distributed together with the
-mini-batch subsets.
-
-After a worker has received its subset of the mini-batch, it samples a neighbor
-set $\cV_t$ for each of its mini-batch vertices, using thread parallelism
-(line 5 in Algorithm~1).
-
-The next stage, \textit{update\_phi} (line 6 in Algorithm~1), is the algorithm's dominant stage, not
-only in calculation but also in memory accesses. It performs $M\times n\times
-K$ operations. It can be fully parallelized because it is a data-parallel
-operation over each of the mini-batch vertices. \textit{update\_phi} loads the
-$\pi$ values for the local mini-batch vertices and their neighbors from the
-DKV store. The updates to $\phi$ for the mini-batch vertices are calculated
-independently using thread parallelism.
-
-%\looseness=-1
-These updates are used by \textit{update\_pi} (line 7 in Algorithm~1) locally, besides the value of
-$\pi/\phi$ for the mini-batch vertices. The update to $\pi$ requires $M\times
-K$ operations, and it is done in parallel over mini-batch vertices. Because of
-memory consistency, this stage awaits completion of \textit{update\_phi} with
-an MPI barrier. After the calculation, the updated values of $\pi+\sum\phi$
-are written through the DKV store.
-
-\textit{update\_beta} (lines 10 and 11 in Algorithm~1) requires $M\times K$ operations. It is preceded by an MPI
-barrier to ensure that up-to-date $\pi$ values are read. \textit{update\_beta}
-is split into four steps. The first stage partitions the mini-batch across
-machines to calculate contributions to $g_{ab}(\theta)$. The values of $\pi$ for
-the local mini-batch vertices are loaded from the DKV store. These calculations
-are done with thread parallelism over the mini-batch vertices. In the second step,
-a multi-threaded summation is performed to calculate the contributions per
-machine, followed by a distributed summation using an MPI reduce operation. The
-third step calculates $\beta$ from $\theta$ sequentially at the master. This
-operation takes only $K$ steps. The resulting $\beta$ is broadcast with MPI
-to the workers.
-
-The calculation of perplexity requires $|\cE_h|\times K$ steps. Each of the
-machines owns a subset of $\cE_h$. It loads up-to-date values of $\pi$ for each
-vertex in its part of $\cE_h$, then for each of its edges in $\cE_h$ it calculates the
-contribution to the perplexity using thread parallelism. The resulting
-contributions are summed to the global perplexity value in two stages,
-with a local OpenMP reduction followed by a global MPI reduction.
-
-Note that the calculation requirements per iteration do not depend
-on $N$ or $|E|$, the size of the network graph. However, larger graphs are
-expected to require more iterations to achieve convergence.
-
-\subsection{Pipelining of computation and loading $\pi$}
-
-As will be shown in Section~\ref{eval-pipeline}, loading $\pi$ is the
-dominant contribution within the dominant stage \textit{update\_phi}. This
-is the performance bottleneck of the distributed algorithm. We introduce
-pipeline parallelism in order to shorten the critical path. Loading $\pi$
-and calculating \textit{update\_pi} are partitioned into chunks, so the
-calculation of a chunk is overlapped with loading the next chunk of $\pi$
-values, and also with sampling the mini-batch at the master and deploying
-it to the workers.
-
-% horrible hack to move the table up
-\begin{table*}
-  \centering
-  \begin{tabular}{l r r r l}
-    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
-                    &            &               & communities    &             \\
-    \hline
-    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
-    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
-    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
-    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
-    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
-    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
-    \hline
-  \end{tabular}
-  \caption{Summary of SNAP graph data sets used for evaluation.}
-  \label{tab-datasets}
-\end{table*}
+$\Neighbors$.
+Hence, there is no opportunity to exploit data locality through caching of
+$\pi$.
 
@@
 of 4-byte floats.
+
+% each of the stages
+
+\subsection{Implementation of distributed parallelism}
+
+This section describes the parallelization of each of the stages of the
+algorithm's main loop. All but the first stage justify parallelization for
+high values of combinations of the parameters mini-batch size $|\Minibatch|$,
+number of vertices in the mini-batch $M$,
+number of communities $K$, neighbor sample size $|\Neighbors|$, and held-out set
+size $|\cE_h|$.
+
+The first stage, mini-batch selection (line 3 in Algorithm~\ref{algorithm}),
+is done by the master and it is
+not itself parallelized. However, the distributed implementation overlaps
+its execution using pipeline parallelism while the workers are
+performing \textit{update\_phi}. The mini-batch is partitioned equally over
+the workers and the relevant sections of \Edges are distributed together with the
+mini-batch subsets.
+
+After a worker has received its subset of the mini-batch, it samples a neighbor
+set $\Neighbors$ for each of its mini-batch vertices, using thread parallelism
+(line 5 in Algorithm~\ref{algorithm}).
+
+The next stage, \textit{update\_phi} (line 6 in Algorithm~\ref{algorithm}), is the algorithm's dominant stage, not
+only in calculation but also in memory accesses. It performs $M \times
+|\Neighbors| \times
+K$ operations. It can be fully parallelized because it is a data-parallel
+operation over each of the mini-batch vertices. \textit{update\_phi} loads the
+$\pi$ values for the local mini-batch vertices and their neighbors from the
+DKV store. The updates to $\phi$ for the mini-batch vertices are calculated
+independently using thread parallelism.
+
+%\looseness=-1
+These updates are used by \textit{update\_pi} (line 7 in
+Algorithm~\ref{algorithm}) locally, besides the value of
+$\pi/\phi$ for the mini-batch vertices. The update to $\pi$ requires
+$M \times K$ operations, and it is done in parallel over mini-batch vertices. Because of
+memory consistency, this stage awaits completion of \textit{update\_phi} with
+an MPI barrier. After the calculation, the updated values of $\pi+\sum\phi$
+are written through the DKV store.
+
+\textit{update\_beta} (lines 10 and 11 in Algorithm~\ref{algorithm}) requires $|\Minibatch| \times K$ operations. It is preceded by an MPI
+barrier to ensure that up-to-date $\pi$ values are read. \textit{update\_beta}
+is split into four steps. The first stage partitions the mini-batch across
+machines to calculate contributions to $g_{ab}(\theta)$. The values of $\pi$ for
+the local mini-batch vertices are loaded from the DKV store. These calculations
+are done with thread parallelism over the mini-batch vertices. In the second step,
+a multi-threaded summation is performed to calculate the contributions per
+machine, followed by a distributed summation using an MPI reduce operation. The
+third step calculates $\beta$ from $\theta$ sequentially at the master. This
+operation takes only $K$ steps. The resulting $\beta$ is broadcast with MPI
+to the workers.
+
+The calculation of perplexity requires $|\cE_h|\times K$ steps. Each of the
+machines owns a subset of $\cE_h$. It loads up-to-date values of $\pi$ for each
+vertex in its part of $\cE_h$, then for each of its edges in $\cE_h$ it calculates the
+contribution to the perplexity using thread parallelism. The resulting
+contributions are summed to the global perplexity value in two stages,
+with a local OpenMP reduction followed by a global MPI reduction.
+
+Note that the calculation requirements per iteration do not depend
+on $N$ or $|\cE^*|$, the size of the network graph. However, larger graphs are
+expected to require more iterations to achieve convergence.
+
+\subsection{Pipelining of computation and loading $\pi$}
+
+The distributed design features two instances of pipeline parallelism. First,
+at the process level, sampling of the minibatch by the master is decoupled
+from \textit{update\_phi} by the workers, so the master computes the minibatch
+for the next iteration while the workers compute \textit{update\_phi} for the
+current iteration. Second, at the threads level within the workers, loading of
+$\pi$ and \textit{update\_phi} are split up into chunks, and the calculation
+in \textit{update\_pi} for the current chunk is done simultaneously with
+loading $\pi$ for the next chunk. The rationale for implementing pipeline
+parallelism in these places of the program, is that loading $\pi$ is the
+dominant contribution within the dominant stage \textit{update\_phi}, as
+will be shown in Section~\ref{eval-pipeline}.  Since loading $\pi$ is the
+performance bottleneck of the distributed algorithm, performing other work
+in parallel shortens the critical path.
+
+% horrible hack to move the table up
+\begin{table*}
+  \centering
+  \begin{tabular}{l r r r l}
+    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
+                    &            &               & communities    &             \\
+    \hline
+    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
+    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
+    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
+    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
+    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
+    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
+    \hline
+  \end{tabular}
+  \caption{Summary of SNAP graph data sets used for evaluation.}
+  \label{tab-datasets}
+\end{table*}

### 2016-01-20 02-related.tex
Add table for symbol definitions. Nitpicks in the texts. Start orthogonising some symbols.

@@
 \eenum
+$\eta$ and $\alpha$ are parameters to the Beta and Dirichlet distribution functions. $\dt$ (usually small) is a parameter for the algorithm.
+
+\newcommand{\Vertices}{\ensuremath{\cV^*{}}\xspace}
+\newcommand{\Edges}{\ensuremath{\cE^*{}}\xspace}
+\newcommand{\Heldout}{\ensuremath{\cE_h{}}\xspace}
+\newcommand{\SGLDMinibatch}{\ensuremath{\cD_n{}}\xspace}
+\newcommand{\Minibatch}{\ensuremath{\cE_n{}}\xspace}
+\newcommand{\Minibatcht}{\ensuremath{\cE_{n_t}{}}\xspace}
+\newcommand{\Neighbors}{\ensuremath{\cV_n{}}\xspace}
 
@@
 \bea
-\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\cD_n)\right) + \xi_t, \label{eqn:sgld_update}
+\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\SGLDMinibatch)\right) + \xi_t, \label{eqn:sgld_update}
 \eea
-where $\xi_t \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\cD_n$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\cD_n)$ the mean stochastic gradient, i.e., $\frac{1}{|\cD_n|}\sum_{x\in \cD_n}^{}\grad_{\ta} \log p(x|\ta)$.
+where $\xi_t \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\SGLDMinibatch$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\SGLDMinibatch)$ the mean stochastic gradient, i.e., $\frac{1}{|\SGLDMinibatch|}\sum_{x\in \SGLDMinibatch}^{}\grad_{\ta} \log p(x|\ta)$.
 As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution.
@@
 Then, Metropolis-Hastings accept-reject tests are applied.
-Comparing to LMC, SGLD only requires to process a mini-batch $\cD_n$ at each iteration and ignores the MH test, and thus the computational complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
+Comparing to LMC, SGLD only requires to process a mini-batch $\SGLDMinibatch$ at each iteration and ignores the MH test, and thus the computational complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
 
-The stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex.
+Stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex.
 By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch-based estimator in Eqn.
@@
 \bea
-\ta_{k}^* \law \left| \ta_{k} + \f{\ep_t}{2} \left( \al - \ta_{k} + \f{N}{|\cD_n|} \sum_{d \in \cD_n} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{t} \right|. \label{eqn:sgrld_update}
+\ta_{k}^* \law \left| \ta_{k} + \f{\ep_t}{2} \left( \al - \ta_{k} + \f{N}{|\SGLDMinibatch|} \sum_{d \in \SGLDMinibatch} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{t} \right|, \label{eqn:sgrld_update}
 \eea
-here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \cD_n$.
+here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \SGLDMinibatch$.
 
@@
 Please refer \cite{LiAW15} for more details.
-The algorithm iterates updating local parameter $\pi$ and a global parameter $\beta$.
+The algorithm iterates updating local parameter $\pi$ and global parameter $\beta$.
 Since both parameters lie on the probability simplex, SGRLD is applied to make the sampling process more efficient.
@@
 \bea
-\ta_{ki}^* \law \left| \ta_{ki} + \f{\ep}{2} \left\{ \eta - \ta_{ki} + h(\cE_{n_t})\sum_{(a,b) \in \cE_{n_t}}g_{ab}(\ta_{ki})\right\} \right.\nn \\ \left.+ (\ta_{ki})^{\ha}\xi_{ki} \right|, \label{eqn:global_update}
+\ta_{ki}^* \law \left| \ta_{ki} + \f{\ep}{2} \left( \eta - \ta_{ki} + h(\Minibatcht)\sum_{(a,b) \in \Minibatcht}g_{ab}(\ta_{ki})\right) \right.\nn \\ \left.+ (\ta_{ki})^{\ha}\xi_{ki} \right|, \label{eqn:global_update}
 \eea
-where 
+where the gradient in $\theta$
 \bea
 g_{ab}(\ta_{ki})
-&=& \f{f_{ab}^{(y)}(k,k)}{Z_{ab}^{(y)}} \left(\f{|1-i-y|}{\ta_{ki}} - \f{1}{\ta_k} \right).
+&=& \f{f_{ab}^{(y)}(k,k)}{Z_{ab}^{(y)}} \left(\f{|1-i-y|}{\ta_{ki}} - \f{1}{\sum_j\ta_{kj}} \right),
 \eea
-here $\cE_{n_t}$ is a mini-batch of $n_t$ vertex pairs sampled from $\cE^*$, and $h(\cE_{n_t})$ is the weighting terms and 
+here $\Minibatcht$ is a mini-batch of $n_t$ vertex pairs sampled from $\Edges$ or $E$; $h(\Minibatcht)$ is a weight factor to scale the effect of a mini-batch towards the full network; and 
 \bea
@@
 \bt_k^y(1-\bt_k)^{(1-y)}\pi_{ak}\pi_{bk}, & \mbox{if } k=l\\
-\dt^{y}(1-\dt)^{(1-y)} \pi_{ak} \pi_{bl}, & \mbox{if } k \neq l \nn
+\dt^{y}(1-\dt)^{(1-y)} \pi_{ak} \pi_{bl}, & \mbox{if } k \neq l \nn.
 \end{cases}
@@
 \bea
-\phi_{ak}^* \law \left| \phi_{ak} + \f{\ep}{2} \left( \al - \phi_{ak} + \f{N}{|\cV_n|} \sum_{b \in \cV_n} g_{ab}(\phi_{ak})\right) \right.
-\nn \\ \left.
-+ (\phi_{ak})^{\ha} \xi_{ak}\right|.\label{eqn:local_update}
+\phi_{ak}^* \law \left| \phi_{ak} + \f{\ep}{2} \left( \al - \phi_{ak} + \f{N}{|\Neighbors|} \sum_{b \in \Neighbors} g_{ab}(\phi_{ak})\right) \right. \nn \\ \left. + (\phi_{ak})^{\ha} \xi_{ak}\right|,
+\label{eqn:local_update}
 \eea
-where \bea
-g_{ab}(\phi_{ak}) = \f{f_{ab}^{(y)}(k)}{Z_{ab}^{(y)}\phi_{ak}} - \f{1}{\phi_a}.
+where the gradient in $\phi$
+\bea
+g_{ab}(\phi_{ak}) = \f{f_{ab}^{(y)}(k)}{Z_{ab}^{(y)}\phi_{ak}} - \f{1}{\sum_j\phi_{aj}}.
 \eea
-Here, the $\cV_n$ is a random mini-batch of $n$ vertices sampled from $\cV^*$. Note that $|\cV_n| \ll |\cV^*|=N$.
+Here, $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV^*$. Note that $|\Neighbors| \ll |\cV^*|=N$.
 
+As is common in machine learning problems, the edges of the graph are divided
+into two sets, the training set and a held-out set \Heldout.
 As the performance metric we use perplexity, which is defined as the exponential
-of the negative average log-likelihood of the held-out set $\cE_h$. Given a collection of $T$
+of the negative average log-likelihood of the held-out set \Heldout. Given a collection of $T$
 samples of the model parameters $\{\bt_t\}$ and $\{\pi_t\}$, the averaged
-perplexity on the held-out test set $\cE_h$ is 
+perplexity on the held-out test set $\Heldout$ is 
 \begin{align}
-&\text{perp}_{\text{avg}}(\cE_h|\{\beta_t\},\{\pi_t\})\nonumber \\=&
-\exp\left(-\frac{\sum_{(a,b)\in \cE_h}^{}\log
+&\text{perp}_{\text{avg}}(\Heldout|\{\beta_t\},\{\pi_t\})\nonumber \\=&
+\exp\left(-\frac{\sum_{(a,b)\in \Heldout}^{}\log
 \{(1/T)\sm{t}{T}p(y_{ab}|\beta_t,\pi_t)\}}{|\cE_{h}|}\right)
@@
 
-The overall pseudo-code of the sequential algorithm is described in Algorithm 1.
+The pseudo-code of the sequential algorithm is presented in Algorithm~\ref{algorithm}.
 
@@
 \WHILE {sampling}
-\STATE Sample a mini-batch of vertex pairs, $\cE_t$, from $E$
-\FOR {each vertex in $\cE_t$}
-	\STATE Sample a mini-batch of vertices, $\cV_t$, from $V$
+\STATE Sample a mini-batch of vertex pairs, \Minibatch, from $E$
+\FOR {each vertex in \Minibatch}
+	\STATE Sample a mini-batch of vertices, $\Neighbors$, from $\Vertices$
     \STATE Update $\phi_{a}$ using Eqn~\ref{eqn:local_update}
-  %  $\phi_{a}^* \law \left| \phi_{a} + \f{\ep_{\phi}}{2} \left( \al - \phi_{a} + \f{N}{|\cV_t|} \sum_{b \in \cV_t} G_{\phi_{a}}\right) + (\phi_{a})^{\ha} \xi_{a}\right|$
+  %  $\phi_{a}^* \law \left| \phi_{a} + \f{\ep_{\phi}}{2} \left( \al - \phi_{a} + \f{N}{|\Neighbors|} \sum_{b \in \Neighbors} G_{\phi_{a}}\right) + (\phi_{a})^{\ha} \xi_{a}\right|$
     \STATE Obtain $\pi_a$ from $\phi_a^*$ \ENDFOR
@@
 \end{algorithmic}
+\label{algorithm}
 \end{algorithm}
+
+\addtolength{\tabcolsep}{-3pt}
+\renewcommand{\arraystretch}{1.5}
+\begin{table}[b] % [htbp]
+\center
+\begin{tabular}{c C{0.075\textwidth} c L{0.26\textwidth}}
+\textbf{symbol}
+	 & \textbf{type} & \textbf{size} & \textbf{description} \\
+         % & \multicolumn{1}{c}{\textbf{type}}
+                           % & \multicolumn{1}{c}{\textbf{description}} \\
+\hline
+$\cK$      &              & $K$        & set of communities \\
+\Vertices  & \{vertex\}   & $N$          & vertices in the graph \\
+\Edges     & \{edge\}     &              & linked edges in the graph \\
+$E$        & \{edge\}     &              & $\Vertices\times\Vertices$: linked and nonlinked edges \\
+\Heldout   & \{edge\}     &              & held-out subset of the graph \\
+\Minibatch & \{edge\}     &              & sampled mini-batch of edges in~$E$ \\
+$M$        &              &              & number of vertices in \Minibatch \\
+\Neighbors & \{vertex\}   &              & sampled neighbor set for a vertex in~\Minibatch \\
+$\theta$   & float~vector 2-D & $K\times{}2$ & global latent variables for community strength \\
+$\beta$    & float vector & $K$          & $\beta[k] = \theta[k][0] / \sum_j\theta[k][j]$ is the community strength\\
+$\phi$     & float~vector 2-D & $N\times{}K$ & local latent variables; $\phi[i][k]$ reflects
+                            probability that vertex $i$ is in community~$k$ \\
+$\pi$      & float~vector 2-D & $N\times{}K$ &
+					$\pi[i][k]$~is~probability that $i$ is in $k$
+					$\pi[i][k] = \phi[i][k] / \sum_j\phi[i][j]$ \\
+\hline
+\end{tabular}
+\caption{Definition of most important symbols}
+\label{tbl-symbols}
+\end{table}
+\addtolength{\tabcolsep}{3pt}
+\renewcommand{\arraystretch}{1.0}
 
@@
 \bea
-\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\cD_n)\right) + \xi, \label{eqn:sgld_update}
+\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\SGLDMinibatch)\right) + \xi, \label{eqn:sgld_update}
 \eea
-where $\xi \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\cD_n$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\cD_n) = \frac{1}{|\cD_n|}\sum_{x\in \cD_n}^{}\grad_{\ta} \log p(x|\ta)$.
+where $\xi \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\SGLDMinibatch$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\SGLDMinibatch) = \frac{1}{|\SGLDMinibatch|}\sum_{x\in \SGLDMinibatch}^{}\grad_{\ta} \log p(x|\ta)$.
 As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution.
@@
 SGLD originated from the Langevin Monte Carlo (LMC) [15] where, unlike SGLD, the gradient is computed exactly using all data points and then a Metropolis-Hastings accept-reject test is applied.
-Because at each iteration SGLD requires to process only a mini-batch $\cD_n$ and ignores the MH test, the computation complexity per iteration is only $\cO(n)$ as opposed to $\cO(N)$ of LMC.
+Because at each iteration SGLD requires to process only a mini-batch $\SGLDMinibatch$ and ignores the MH test, the computation complexity per iteration is only $\cO(n)$ as opposed to $\cO(N)$ of LMC.
 Any mini-batch sampling algorithm in the form of Eqn.
 \eqref{eqn:sgld_update} is called valid SGLD as long as it guarantees the gradient estimator to be unbiased, i.e.,
-$\eE_{\cD_n} \left[N\bar{g}(\ta;\cD_n)\right] = \grad_{\ta} \log p(\cX|\ta)\label{eqn:sgld_grad_estm}$ and the variance to be finite [5].
+$\eE_{\SGLDMinibatch} \left[N\bar{g}(\ta;\SGLDMinibatch)\right] = \grad_{\ta} \log p(\cX|\ta)\label{eqn:sgld_grad_estm}$ and the variance to be finite [5].
 
@@
 \bea
-\ta_{k}^* \law \left| \ta_{k} + \f{\ep}{2} \left( \al - \ta_{k} + \f{N}{|\cD_n|} \sum_{d \in \cD_n} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{} \right|. \label{eqn:sgrld_update}
+\ta_{k}^* \law \left| \ta_{k} + \f{\ep}{2} \left( \al - \ta_{k} + \f{N}{|\SGLDMinibatch|} \sum_{d \in \SGLDMinibatch} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{} \right|. \label{eqn:sgrld_update}
 \eea
-here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \cD_n$.
+here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \SGLDMinibatch$.
 
@@
 \bea
-\ta_{ki}^* \law \left| \ta_{ki} + \f{\ep}{2} \left\{ \eta - \ta_{ki} + h(\cE_{n_t})\sum_{(a,b) \in \cE_{n_t}}g_{ab}(\ta_{ki})\right\} \right.\nn \\ \left.+ (\ta_{ki})^{\ha}\xi_{ki} \right|, \label{eqn:global_update}
+\ta_{ki}^* \law \left| \ta_{ki} + \f{\ep}{2} \left\{ \eta - \ta_{ki} + h(\Minibatcht)\sum_{(a,b) \in \Minibatcht}g_{ab}(\ta_{ki})\right\} \right.\nn \\ \left.+ (\ta_{ki})^{\ha}\xi_{ki} \right|, \label{eqn:global_update}
 \eea
-here $\cE_{n_t}$ is a mini-batch of $n_t$ node pairs sampled from $\cE^*$ for which we use the following strategy.
+here $\Minibatcht$ is a mini-batch of $n_t$ node pairs sampled from $\cE^*$ for which we use the following strategy.
 
@@
 \bea
-\phi_{ak}^* \law \left| \phi_{ak} + \f{\ep}{2} \left( \al - \phi_{ak} + \f{N}{|\cV_n|} \sum_{b \in \cV_n} g_{ab}(\phi_{ak})\right) \right.
+\phi_{ak}^* \law \left| \phi_{ak} + \f{\ep}{2} \left( \al - \phi_{ak} + \f{N}{|\Neighbors|} \sum_{b \in \Neighbors} g_{ab}(\phi_{ak})\right) \right.
 \nn \\ \left.
@@
 \eea
-Here, the $\cV_n$ is a random mini-batch of $n$ nodes sampled from $\cV^*$. Note that $|\cV_n| \ll |\cV^*|=N$.
+Here, the $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV^*$. Note that $|\Neighbors| \ll |\cV^*|=N$.
 

### 2016-01-20 02-related.tex
updated notations in section 2

@@
 The assortative mixed membership stochastic blockmodel (a-MMSB)~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group structure in a network of $N$ vertices.
-Consider a set  $\cV^*$ containing all the vertices in the graph, and a set $\cE^*$ containing all the linked edges between pairs of vertices.
-Each vertex $a$ in the vertex set $\cV^*$ has a $K$-dimensional probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
+Consider a set  $\cV$ containing all the vertices in the graph, and a set $\cE$ containing all the linked edges between pairs of vertices.
+Each vertex $a$ in the vertex set $\cV$ has a $K$-dimensional probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
 For every possible peer $b$ in the network, each vertex $a$ randomly draws a community $z_{ab}$.
-If a pair of vertices ($a,b$) in the edge set $\cE^*$ are in the same community, i.e., $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
+If a pair of vertices ($a,b$) in the edge set $\cE$ are in the same community, i.e., $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
 Otherwise this probability is small.
@@
 
-\newcommand{\Vertices}{\ensuremath{\cV^*{}}\xspace}
-\newcommand{\Edges}{\ensuremath{\cE^*{}}\xspace}
+\newcommand{\Vertices}{\ensuremath{\cV{}}\xspace}
+\newcommand{\Edges}{\ensuremath{\cE{}}\xspace}
 \newcommand{\Heldout}{\ensuremath{\cE_h{}}\xspace}
@@
 \newcommand{\Minibatch}{\ensuremath{\cE_n{}}\xspace}
-\newcommand{\Minibatcht}{\ensuremath{\cE_{n_t}{}}\xspace}
+\newcommand{\Minibatcht}{\ensuremath{\cE_n{}}\xspace}
 \newcommand{\Neighbors}{\ensuremath{\cV_n{}}\xspace}
@@
 As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution.
-One benefit of using SGLD is that we do not need the Metropolis-Hastings (MH) accept-reject tests since the rejection probability goes to zero as the step size collapses to zero.
-Although the finite step size, which we use in practice rather than converging to zero, still results in some bias, we can reduce the overall error by drawing many more samples per unit time.
+One benefit of using SGLD is that we do not need the Metropolis-Hastings (MH) accept-reject tests since the rejection probability goes to zero as the step size tends to zero.
+Although in practice we often do not reduce the stepsize to zero to obtain better mixing (thus resulting in some bias), we obtain good performance by drawing many more samples per unit time.
 
@@
 Then, Metropolis-Hastings accept-reject tests are applied.
-Comparing to LMC, SGLD only requires to process a mini-batch $\SGLDMinibatch$ at each iteration and ignores the MH test, and thus the computational complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
+Comparing to LMC, SGLD only requires to process a mini-batch $\SGLDMinibatch$ at each iteration and ignores the MH test, and thus the computation complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
 
@@
 \eea
-here $\Minibatcht$ is a mini-batch of $n_t$ vertex pairs sampled from $\Edges$ or $E$; $h(\Minibatcht)$ is a weight factor to scale the effect of a mini-batch towards the full network; and 
+here $\Minibatcht$ is a mini-batch of $n_t$ vertex pairs sampled from $\Edges$; $h(\Minibatcht)$ is a weight factor to scale the effect of a mini-batch towards the full network; and 
 \bea
@@
 \eea
-Here, $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV^*$. Note that $|\Neighbors| \ll |\cV^*|=N$.
+Here, $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV$. Note that $|\Neighbors| \ll |\cV|=N$.
 
@@
 \WHILE {sampling}
-\STATE Sample a mini-batch of vertex pairs, \Minibatch, from $E$
+\STATE Sample a mini-batch of vertex pairs, \Minibatch, from $\cE$
 \FOR {each vertex in \Minibatch}
@@
 Assortative mixed-membership stochastic blockmodel (a-MMSB) [1] is a special case of MMSB [8] that models the group-structure in a network of $N$ nodes.
-In particular, each node $a$ in the node set $\cV^*$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
+In particular, each node $a$ in the node set $\cV$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
 For every possible peer $b$ in the network, each node $a$ randomly draws a community $z_{ab}$.
-If a pair of nodes ($a,b$) in the edge set $\cE^*$ are in the same community: $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
+If a pair of nodes ($a,b$) in the edge set $\cE$ are in the same community: $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
 Otherwise this probability is small.
@@
 \eea
-here $\Minibatcht$ is a mini-batch of $n_t$ node pairs sampled from $\cE^*$ for which we use the following strategy.
+here $\Minibatcht$ is a mini-batch of $n_t$ node pairs sampled from $\cE$ for which we use the following strategy.
 
@@
 For this, at every iteration we first randomly select a node $a$ and then toss a coin with probability 0.5 to decide whether to sample link edges or non-link edges for node $a$.
-If it is a link, we assign all of the link edges of node $a$ to $\cE_{n_t}$.
+If it is a link, we assign all of the link edges of node $a$ to $\cE_n$.
 Otherwise, i.e.
-if it is non-link, we uniformly sample a mini-batch of $N/m$ non-link edges from the entire set of non-link edges and assign it to $\cE_{n_t}$.
+if it is non-link, we uniformly sample a mini-batch of $N/m$ non-link edges from the entire set of non-link edges and assign it to $\cE_n$.
 Here, the $m$ is a hyper-parameter.
-Note that the size of $|\cE_{n_t}|$ will thus be much smaller than the total number of $N(N-1)/2$ edges when $m$ is reasonably large.
-Then, to ensure that the gradient is unbiased, a \textit{scaling parameter} $h(\cE_{n_t})$ is multiplied.
-Specifically, $h(\cE_{n_t})$ is set to $N$ when $\cE_{n_t}$ is a set of link edges and to $mN$ otherwise.
+Note that the size of $|\cE_n|$ will thus be much smaller than the total number of $N(N-1)/2$ edges when $m$ is reasonably large.
+Then, to ensure that the gradient is unbiased, a \textit{scaling parameter} $h(\cE_n)$ is multiplied.
+Specifically, $h(\cE_n)$ is set to $N$ when $\cE_n$ is a set of link edges and to $mN$ otherwise.
 
@@
 \eea
-Here, the $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV^*$. Note that $|\Neighbors| \ll |\cV^*|=N$.
+Here, the $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV$. Note that $|\Neighbors| \ll |\cV|=N$.
 

### 2016-01-21 02-related.tex
more update on notation
Straighten out symbols a bit (again)

@@
+\newcommand{\Vertices}{\ensuremath{\cV{}}\xspace}
+\newcommand{\Edges}{\ensuremath{\cE{}}\xspace}
+\newcommand{\Heldout}{\ensuremath{\cE_h{}}\xspace}
+\newcommand{\SGLDMinibatch}{\ensuremath{\cD_n{}}\xspace}
+\newcommand{\Minibatch}{\ensuremath{\cE_n{}}\xspace}
+\newcommand{\Minibatcht}{\ensuremath{\cE_n{}}\xspace}
+\newcommand{\Neighbors}{\ensuremath{\cV_n{}}\xspace}
+
 \section{Background}
@@
 The assortative mixed membership stochastic blockmodel (a-MMSB)~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group structure in a network of $N$ vertices.
-Consider a set  $\cV$ containing all the vertices in the graph, and a set $\cE$ containing all the linked edges between pairs of vertices.
-Each vertex $a$ in the vertex set $\cV$ has a $K$-dimensional probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
+Consider a set  \Vertices containing all the vertices in the graph, and a set \Edges containing all the linked edges between pairs of vertices.
+Each vertex $a$ in the vertex set \Vertices has a $K$-dimensional probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
 For every possible peer $b$ in the network, each vertex $a$ randomly draws a community $z_{ab}$.
-If a pair of vertices ($a,b$) in the edge set $\cE$ are in the same community, i.e., $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
+If a pair of vertices ($a,b$) in the edge set \Edges are in the same community, i.e., $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
 Otherwise this probability is small.
@@
 $\eta$ and $\alpha$ are parameters to the Beta and Dirichlet distribution functions. $\dt$ (usually small) is a parameter for the algorithm.
-
-\newcommand{\Vertices}{\ensuremath{\cV{}}\xspace}
-\newcommand{\Edges}{\ensuremath{\cE{}}\xspace}
-\newcommand{\Heldout}{\ensuremath{\cE_h{}}\xspace}
-\newcommand{\SGLDMinibatch}{\ensuremath{\cD_n{}}\xspace}
-\newcommand{\Minibatch}{\ensuremath{\cE_n{}}\xspace}
-\newcommand{\Minibatcht}{\ensuremath{\cE_n{}}\xspace}
-\newcommand{\Neighbors}{\ensuremath{\cV_n{}}\xspace}
 
@@
 Stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex.
-By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch-based estimator in Eqn.
-\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}.
+By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch-based estimator in Eq.~\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}.
 In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$.
@@
 This section describes the SG-MCMC algorithm for a-MMSB.
-Please refer \cite{LiAW15} for more details.
+Please refer to~\cite{LiAW15} for more details.
 The algorithm iterates updating local parameter $\pi$ and global parameter $\beta$.
@@
 After updating $\phi$ and $\ta$, we can obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$, respectively.
+That is, $\bt_k = \ta_{k1}/(\ta_{k0} + \ta_{k1})$ and $\pi_{ak} = \phi_{ak}/\sm{j}{K}\phi_{aj}$.
 In the following, we briefly sketch the iterative update steps.
@@
 \eea
-where the gradient in $\theta$
+where the gradient in $\theta$ is
 \bea
@@
 \eea
-here $\Minibatcht$ is a mini-batch of $n_t$ vertex pairs sampled from $\Edges$; $h(\Minibatcht)$ is a weight factor to scale the effect of a mini-batch towards the full network; and 
+here $\Minibatcht$ is a mini-batch of $n_t$ vertex pairs sampled from $\Edges$ or $E$; $h(\Minibatcht)$ is a weight factor to scale the effect of a mini-batch towards the full network; and 
 \bea
@@
 \eea
-where the gradient in $\phi$
+where the gradient in $\phi$ is
 \bea
@@
 \eea
-Here, $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV$. Note that $|\Neighbors| \ll |\cV|=N$.
+Here, $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from \Vertices. Note that $|\Neighbors| \ll |\Vertices|=N$.
 
@@
 \WHILE {sampling}
-\STATE Sample a mini-batch of vertex pairs, \Minibatch, from $\cE$
+\STATE Sample mini-batch of vertex pairs \Minibatch from \Edges or $E$
 \FOR {each vertex in \Minibatch}
-	\STATE Sample a mini-batch of vertices, $\Neighbors$, from $\Vertices$
-    \STATE Update $\phi_{a}$ using Eqn~\ref{eqn:local_update}
+    \STATE Sample mini-batch of vertices $\Neighbors$ from $\Vertices$
+    \STATE Update $\phi_{a}$ using Eq.~\ref{eqn:local_update}
   %  $\phi_{a}^* \law \left| \phi_{a} + \f{\ep_{\phi}}{2} \left( \al - \phi_{a} + \f{N}{|\Neighbors|} \sum_{b \in \Neighbors} G_{\phi_{a}}\right) + (\phi_{a})^{\ha} \xi_{a}\right|$
@@
 \FOR {$k = 1,\dots,K$}
-    \STATE Update $\ta_{k}$ using Eqn~\ref{eqn:global_update}
+    \STATE Update $\ta_{k}$ using Eq.~\ref{eqn:global_update}
    % $\ta_{k}^* \law \left| \ta_{k} + \f{\ep_\ta}{2} \left( \eta - \ta_{k} + h(\cE_t) \sum_{(a,b) \in \cE_t} G_{\ta_{k}}(a,b) \right) + (\ta_{k})^{\ha}\xi_{k} \right|$
@@
 \Neighbors & \{vertex\}   &              & sampled neighbor set for a vertex in~\Minibatch \\
-$\theta$   & float~vector 2-D & $K\times{}2$ & global latent variables for community strength \\
-$\beta$    & float vector & $K$          & $\beta[k] = \theta[k][0] / \sum_j\theta[k][j]$ is the community strength\\
-$\phi$     & float~vector 2-D & $N\times{}K$ & local latent variables; $\phi[i][k]$ reflects
-                            probability that vertex $i$ is in community~$k$ \\
+$\theta$   & float~vector 2-D & $K\times{}2$ & reparameterization of $\beta$.
+					$\beta[k] = \theta[k][1] / \sum_j\theta[k][j]$ \\
+$\beta$    & float vector & $K$          & community strength \\
+$\phi$     & float~vector 2-D & $N\times{}K$ & reparameterization of $\pi$.
+					$\pi[i][k] = \phi[i][k] / \sum_j\phi[i][j]$ \\
 $\pi$      & float~vector 2-D & $N\times{}K$ &
-					$\pi[i][k]$~is~probability that $i$ is in $k$
-					$\pi[i][k] = \phi[i][k] / \sum_j\phi[i][j]$ \\
+					$\pi[i][k]$~is~probability that vertex $i$ is in community $k$ \\
 \hline
@@
 Assortative mixed-membership stochastic blockmodel (a-MMSB) [1] is a special case of MMSB [8] that models the group-structure in a network of $N$ nodes.
-In particular, each node $a$ in the node set $\cV$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
+In particular, each node $a$ in the node set \Vertices has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
 For every possible peer $b$ in the network, each node $a$ randomly draws a community $z_{ab}$.
-If a pair of nodes ($a,b$) in the edge set $\cE$ are in the same community: $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
+If a pair of nodes ($a,b$) in the edge set \Edges are in the same community: $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
 Otherwise this probability is small.
@@
 Because at each iteration SGLD requires to process only a mini-batch $\SGLDMinibatch$ and ignores the MH test, the computation complexity per iteration is only $\cO(n)$ as opposed to $\cO(N)$ of LMC.
-Any mini-batch sampling algorithm in the form of Eqn.
-\eqref{eqn:sgld_update} is called valid SGLD as long as it guarantees the gradient estimator to be unbiased, i.e.,
+Any mini-batch sampling algorithm in the form of Eq.~\eqref{eqn:sgld_update} is called valid SGLD as long as it guarantees the gradient estimator to be unbiased, i.e.,
 $\eE_{\SGLDMinibatch} \left[N\bar{g}(\ta;\SGLDMinibatch)\right] = \grad_{\ta} \log p(\cX|\ta)\label{eqn:sgld_grad_estm}$ and the variance to be finite [5].
@@
 The stochastic gradient Riemannian Langevin dynamics (SGRLD) [2] is a subclass of SGLD which is developed to sample from the probability simplex.
-By applying Riemannian geometry [15] and using the mini-batch estimator in Eqn.
-\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA).
+By applying Riemannian geometry [15] and using the mini-batch estimator in Eq.~\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA).
 In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$.
@@
 Thus, we alternatingly sample in the $\phi$ and $\ta$ spaces, and then obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$.
-From Eqn.
-\ref{eqn:joint}, summing over the latent variable $z$, we obtain the following joint probability,
+From Eq.~\ref{eqn:joint}, summing over the latent variable $z$, we obtain the following joint probability,
 \bea
@@
 By the re-parameterization, we have $\beta_k=\ta_{k1}/(\ta_{k0}+\ta_{k1})$, where $\ta_{ki}\sim$ Gamma$(\eta)\propto \ta_{ki}^{\eta-1}e^{-\ta_{ki}}$.
-Because $p(y, \pi , \bt | \al , \eta)$ decomposes into $p(y,\bt|\pi,\eta) p(\pi|\al)$, replacing $\bt$ by $\ta$, we compute the derivative of log of Eqn.
-\ref{eqn:log_joint} w.r.t.
+Because $p(y, \pi , \bt | \al , \eta)$ decomposes into $p(y,\bt|\pi,\eta) p(\pi|\al)$, replacing $\bt$ by $\ta$, we compute the derivative of log of Eq.~\ref{eqn:log_joint} w.r.t.
 $\ta_{ki}$ for $i=\{0,1\}$ as follows:
@@
 here we used simple notation $y$ instead of $y_{ab}$.
-Unlike the SGRLD for LDA [2], we compute the expectation in Eqn.
-\eqref{eqn:G} analytically by computing the normalization constant $Z_{ab}^{(y)} = \sm{k}{K}\sm{l}{K} f_{ab}^{(y)}(k,l)$ which can be reduced to $\cO(K)$ computation as follows
+Unlike the SGRLD for LDA [2], we compute the expectation in Eq.~\eqref{eqn:G} analytically by computing the normalization constant $Z_{ab}^{(y)} = \sm{k}{K}\sm{l}{K} f_{ab}^{(y)}(k,l)$ which can be reduced to $\cO(K)$ computation as follows
 \begin{align}
@@
 
-Then Eqn. \eqref{eqn:G} becomes
+Then Eq.~\eqref{eqn:G} becomes
 \bea
@@
 \eea
-Plugging this into Eqn. \ref{eqn:sgrld_update}, we obtain the update rule for the global parameter,
+Plugging this into Eq.~\ref{eqn:sgrld_update}, we obtain the update rule for the global parameter,
 
@@
 \eea
-here $\Minibatcht$ is a mini-batch of $n_t$ node pairs sampled from $\cE$ for which we use the following strategy.
+here $\Minibatcht$ is a mini-batch of $n_t$ node pairs sampled from \Edges for which we use the following strategy.
 
@@
 Similar to the global parameter, we re-parameterize the local parameter $\pi_a$ such that $\pi_{ak} = \phi_{ak} / \sum_{j=1}^{K}\phi_{aj}$, with $\phi_{ak}\sim$ Gamma$(\alpha)\propto \phi_{ak}^{\alpha-1}e^{-\phi_{ak}}$.
-Then, taking the derivative of the log of Eqn.
-\ref{eqn:log_joint} w.r.t.
+Then, taking the derivative of the log of Eq.~\ref{eqn:log_joint} w.r.t.
 $\phi_{ak}$, we obtain
@@
 Here the expectation is w.r.t.
-the distribution in Eqn.
-\eqref{eqn:case}.
-To compute the expectation analytically, we first integrate out $z_{ba}$ from Eqn.
-\eqref{eqn:case} because the expectation depends only on $z_{ab}$, and obtain the following probability up to a normalization constant
+the distribution in Eq.~\eqref{eqn:case}.
+To compute the expectation analytically, we first integrate out $z_{ba}$ from Eq.~\eqref{eqn:case} because the expectation depends only on $z_{ab}$, and obtain the following probability up to a normalization constant
 
@@
 Then we obtain the normalization term by $Z_{ab}^{(y)} = \sm{k}{K}f_{ab}^{(y)}(k)$. 
-Integrating out the expectation in Eqn. \eqref{eqn:grad_local}, we obtain 
+Integrating out the expectation in Eq.~\eqref{eqn:grad_local}, we obtain 
 \bea
@@
 \eea 
-Plugging this to Eqn. \ref{eqn:sgrld_update}, we obtain the SGRLD update rule for the local parameter $\phi_{ak}$
+Plugging this to Eq.~\ref{eqn:sgrld_update}, we obtain the SGRLD update rule for the local parameter $\phi_{ak}$
 \bea
@@
 \eea
-Here, the $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV$. Note that $|\Neighbors| \ll |\cV|=N$.
+Here, the $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from \Vertices or $E$. Note that $|\Neighbors| \ll |\Vertices|=N$.
 

### 2016-01-21 04-evaluation.tex
Allow floats at the bottom by setting bottomfraction
Line of text. Float placement
Work on 04 pipeline. Remove 04 weak scaling Fig 2(b)

@@
 %################################################
-\subsection{Strong scaling}
+\subsection{Strong Scaling}
 
-\begin{figure}[htb] % [htb]
+\begin{figure}[b] % [htb]
   \centering
@@
 
-\begin{figure}[htb] % [htb]
+\begin{figure}[htbp] % [htb]
   \centering
   \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=0.48\textwidth}
-  \caption{(a) Average execution time per algorithm iteration varying
+  \caption{Average execution time per algorithm iteration varying
   the number of latent communities proportionally to the number of compute
-  nodes. (b) The exact number of communities used for each data point in (a).}
+  nodes.}
   \label{fig-weak-scaling}
@@
 conducted several executions varying the cluster size and the number of latent
-communities proportionally. This methodology ensures that each compute node
+communities proportionally. The number of communities at 64 workers was 12K.
+This methodology ensures that each compute node
 performs a relatively constant amount of computational work across all
 configurations. However, the number and size of messages exchanged between the
-nodes would vary significantly. Figure~\ref{fig-weak-scaling}-a presents the
+nodes would vary significantly. Figure~\ref{fig-weak-scaling} presents the
 average execution
-time per algorithm iteration across different cluster sizes. On the other hand,
-Figure~\ref{fig-weak-scaling}-b reports the number of communities for each
-cluster configuration. The relative change in the average execution time per
-iteration is insignificant even though the communication intensity increases
-for larger cluster sizes. This observation verifies that the system's overall
+time per algorithm iteration across different cluster sizes.
+This figure shows that the system's overall
 overhead is minimal. Additionally, the experimental results show that the
@@
 
-\begin{figure}[t] % [htb]
+As discussed in Section~\ref{design-section}, we implemented pipeline
+parallelism at
+two places in the algorithm. The first is an inter-process pipeline that
+overlaps drawing of the mini-batch at the master's with \textit{update\_phi}
+at the workers. The second pipeline overlaps loading of $\pi$ with the
+computation within \textit{update\_pi} within each worker.
+
+\begin{table}[htbp]
+  \centering
+  \begin{tabular}{l d{5.1} d{5.1}}
+    \textbf{Iteration (sub)stages}  & \multicolumn{1}{c}{non-pipelined}
+                                          & \multicolumn{1}{c}{pipelined} \\
+    \hline
+    draw/deploy mini-batch     &  45.6  &  26.2 \\
+    load $\pi$                 & 205    & 209 \\
+    update $\phi$              &  74    &  74 \\
+    \hline
+    total                      & 331    & 241 \\
+    \hline
+  \end{tabular}
+  \caption{Stages that profit from the two pipelining optimizations.
+  com-Friendster on 65
+  compute nodes, with 12K communities. Times in ms per iteration.}
+  \label{table-pipeline}
+\end{table}
+
+Table~\ref{table-pipeline} shows a breakdown of the (sub)stages that are
+overlapped by both pipelining optimizations. The left-hand column has
+all pipelining disabled, the right-hand
+column has pipelining enabled. Within \textit{update\_phi},
+loading $\pi$ from the DKV store is domininant.
+The pipelining optimization significantly overlaps loading $\pi$ both with generation and
+deployment of the mini-batch, and with the calculation of \textit{update\_phi}.
+
+\begin{figure}[htb]
   \centering
@@
   \caption{Performance effect of varying the number of communities on the
-    algorithm's execution time on 64 nodes when using single- or
-    double-buffering.}
+    algorithm's execution time on 64 nodes when disabling or enabling both
+    pipelining optimizations.}
   \label{fig-pipeline}
@@
 
-As discussed in Section~\ref{design-section}, the collective memory of all
-worker nodes serves as the storage for the state of the computation. As such,
-the rows of $\pi$ are equally and statically partitioned across all workers.
-Given that $\pi$ accesses are random, a node in a cluster of size $C$ must
-fetch $(C-1)/C$ of all read requests over the network. Therefore, large cluster
-configurations exhibit higher bandwidth demands and are more sensitive
-to network latency. To reduce the negative effects of network latency on the
-computation, a pipelining scheme was devised to prefetch data dependencies.
 Figure~\ref{fig-pipeline} presents the execution time of 1024 algorithm
-iterations on a 64-node cluster with double-buffering enabled and disabled.
+iterations on a 64-node cluster with the pipelines enabled and disabled.
 Naturally, increasing the number of communities causes a proportional increase
-in execution time. However, when double-buffering is enabled, some of the
+in execution time. However, when pipelining is enabled, some of the
 incurred network latency is hidden by overlapping it with computation.
@@
 widening gap between both lines depicted in Figure~\ref{fig-pipeline}.
-
-\begin{table}[htbp]
-  \centering
-  \begin{tabular}{l d{5.1} d{5.1}}
-    \textbf{Iteration stages}  & \multicolumn{1}{c}{non-pipelined}
-                                          & \multicolumn{1}{c}{pipelined} \\
-    \hline
-    total                      & 450    & 365 \\
-    draw/deploy mini-batch      &  45.6  &      \\
-    update\_phi                & 285    & 241 \\
-    update\_pi                 &   3.8  &   4.6 \\
-    update $\beta/\theta$      &  25.9  &   33.6 \\
-    \hline
-    \textbf{Substages within update\_phi} \\
-    draw/deploy mini-batch      &        &  26.2 \\
-    load $\pi$                 & 205    & 209 \\
-    update $\phi$              &  74    &  74 \\
-    \hline
-  \end{tabular}
-  \caption{The most important stages in the execution; com-Friendster on 65
-  compute nodes, with 12K communities. Times in ms per iteration. The lower
-  half of the table shows sub-stages within \textit{update\_phi}. In the
-  pipelining version, these components are done in parallel.}
-  \label{table-pipeline}
-\end{table}
-
-Table~\ref{table-pipeline} shows a breakdown of the most important time
-consumers; the left-hand column is without pipelining, the right-hand
-column has pipelining. \textit{update\_phi} is the dominant contribution,
-and within that, loading $\pi$ from the DKV store takes by far the most time.
-The pipelining optimization overlaps loading $\pi$ with generation and
-deployment of the mini-batch, and with the calculation of \textit{update\_phi}.
+Large cluster
+configurations exhibit higher bandwidth demands and are more sensitive
+to network latency:
+given that $\pi$ accesses are random, a node in a cluster of size $C$ must
+fetch $(C-1)/C$ of all read requests over the network.
 
@@
 \subsection{Horizontal vs. Vertical Scalability}
-\begin{figure}[htbp] % [htb]
+\begin{figure}[b] % [htb]
   \centering
@@
 
-\begin{figure}[t] % [htbp]
+\begin{figure}[b] % [htbp]
   \centering

### 2016-01-21 plots/plot-sweep-K-fixed-np.sh

@@ -33,7 +33,7 @@ set grid ytics
 set output 'sweep-over-K-fixed-np.eps'
 set title "Scaling K (fixed np=65)"
 set key left
-plot "$$.tmp" u 1:2 w lp t "Single buffering", "$$.tmp2" u 1:2 w lp t "Double buffering"
+plot "$$.tmp" u 1:2 w lp t "Pipelining disabled", "$$.tmp2" u 1:2 w lp t "Pipelining enabled"
 
 EOF
 

### 2016-01-21 plots/plot-sweep-K-proportional-np.sh

@@ -10,31 +10,15 @@ grep iteration ../data/sweep-over-K-proportional-np/*/* \
   | bash \
   | sort -n > $$.tmp
 
-ls ../data/sweep-over-K-proportional-np/*/* \
-  | grep -wv np5 \
-  | sed 's/.*-K\([0-9K]\+\).*-np\([0-9]\+\).*/\2 \1/' \
-  | sed 's/\([0-9]\+\)K/$[\1*1024]/' \
-  | sed 's/\(.*\)/echo \1/' \
-  | bash \
-  | sort -n \
-  | awk '{print $1-1 " " $2}' > $$.tmp.2
-
 cat $$.tmp
 
 cat <<EOF | gnuplot --persist
 
 w = 0.6
 h = 0.4
-k = 0.0
-nr = 2
-nc = 1
-row(x) = ((nr-x-1)*h)+k
-col(x) = x*w
 
 set terminal postscript eps enhanced color font ',8'
-set size (nc*w),(nr*h+k)
 set output 'sweep-over-K-proportional-np.eps'
-set multiplot layout nr,nc
 set size w,h
 
 set xlabel "Number of Nodes"
@@ -43,27 +27,9 @@ set yrange[0:500]
 set nokey
 set grid ytics
 
-set origin col(0),row(0)
-set size w,h
-set title "(a) Weak Scaling (over K)"
+set title "Weak Scaling (over K)"
 set ylabel "Time per Iteration (milliseconds)"
 plot "$$.tmp" u (\$2-1):(1000*\$3/\$4) w lp axis x1y1
-
-set origin col(0),row(1)
-set size w,h
-set style data histogram 
-set style histogram cluster gap 1
-
-set style fill solid noborder
-set auto x
-set ytics
-set mytics
-set grid ytics mytics
-
-set yrange[0:*]
-set title "(b) Number of communities used per cluster size"
-set ylabel "#K"
-plot "$$.tmp.2" u 2:xtic(1)
 
 EOF
 

### 2016-01-21 01-intro.tex
Polish on 01. Add 06.

@@
 A community can be seen as a densely connected group of vertices that are only sparsely connected to the rest of the network.
+This problem is much harder than the related domain of detecting disjoint communities.
 
@@
 In this paper, we are particularly interested in the SG-MCMC method applied to the a-MMSB model introduced in \cite{LiAW15}, which turned out to be faster and more accurate than the SVB method.
-Although \cite{LiAW15} proposed a scalable algorithm for the problem, there is still room to improve the performance further by considering a custom high-performance implementation.
+\cite{LiAW15} further proposes a heuristic to scale up the number of communities at the cost of less precision; the work in this paper scales the algorithm without that heuristic with a custom high-performance implementation.
 To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem.

### 2016-01-21 04-evaluation.tex
Polish a bit on 02, 04, 04, 05

@@
 %################################################
+\subsection{Weak Scaling}
+
+\begin{figure}[t] % [htbp]
+  \centering
+  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=0.48\textwidth}
+  \caption{Average execution time per algorithm iteration when varying
+  the number of latent communities proportionally to the number of compute
+  nodes.}
+  \label{fig-weak-scaling}
+\end{figure}
+
+A study of a system's weak scaling aids in the assessment of the communication
+and synchronization overheads associated with the management of large clusters.
+Similarly, it can expose complex issues that hinder performance at scale such
+as load imbalance. To fairly evaluate the algorithm's weak scaling behavior we
+conducted several executions varying the cluster size and the number of latent
+communities proportionally. The number of communities is 192 times the number
+of worker nodes in all runs.
+This methodology ensures that each compute node
+performs an approximately constant amount of computational work across all
+configurations. However, the number and size of messages exchanged between the
+nodes would vary significantly. Figure~\ref{fig-weak-scaling} presents the
+average execution
+time per algorithm iteration across different cluster sizes.
+This figure shows that the system's overall
+overhead is minimal. Additionally, the experimental results show that the
+implementation is capable of achieving good speedups provided the input problem
+is large enough for the given cluster size.
+
+%################################################
 \subsection{Strong Scaling}
 
-\begin{figure}[b] % [htb]
+\begin{figure}[t] % [htb]
   \centering
@@
 %################################################
-\subsection{Weak Scaling}
-
-\begin{figure}[htbp] % [htb]
-  \centering
-  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=0.48\textwidth}
-  \caption{Average execution time per algorithm iteration varying
-  the number of latent communities proportionally to the number of compute
-  nodes.}
-  \label{fig-weak-scaling}
-\end{figure}
-
-A study of a system's weak scaling aids in the assessment of the communication
-and synchronization overheads associated with the management of large clusters.
-Similarly, it can expose complex issues that hinder performance at scale such
-as load imbalance. To fairly evaluate the algorithm's weak scaling behavior we
-conducted several executions varying the cluster size and the number of latent
-communities proportionally. The number of communities at 64 workers was 12K.
-This methodology ensures that each compute node
-performs a relatively constant amount of computational work across all
-configurations. However, the number and size of messages exchanged between the
-nodes would vary significantly. Figure~\ref{fig-weak-scaling} presents the
-average execution
-time per algorithm iteration across different cluster sizes.
-This figure shows that the system's overall
-overhead is minimal. Additionally, the experimental results show that the
-implementation is capable of achieving good speedups provided the input problem
-is large enough for the given cluster size.
-
-%################################################
 \subsection{Pipeline efficiency}
@@
 
-\begin{table}[htbp]
+\begin{table}[t] % [htbp]
   \centering
@@
 column has pipelining enabled. Within \textit{update\_phi},
-loading $\pi$ from the DKV store is domininant.
+loading $\pi$ from the DKV store is dominant.
 The pipelining optimization significantly overlaps loading $\pi$ both with generation and
@@
 
-\begin{figure}[htb]
+\begin{figure}[t] % [htb]
   \centering
@@
 \subsection{Horizontal vs. Vertical Scalability}
-\begin{figure}[b] % [htb]
+\begin{figure}[t] % [htb]
   \centering

### 2016-01-21 05-conclusion.tex
More little textual thingies
Polish a bit on 02, 04, 04, 05

@@
 We conducted a thorough empirical evaluation of the system to study its strong
-and weak scalability on 64 cluster nodes using large data sets.  Additionally,
+and weak scalability on 65 cluster nodes using large data sets.  Additionally,
 we assessed the efficiency of the algorithm's resource utilization. Finally, a
@@
 different organic data sets.
+To the best of
+our knowledge, this is the first time that the problem of deducing overlapping
+communities has been learned for problems of such a large scale.

### 2016-01-21 plots/plot-sweep-np-fixed-K.sh

@@ -74,8 +74,8 @@ plot \
 
 set origin col(0),row(1)
 set size w,h
-set ylabel "Speedup"
-set title "(b) Speedup over 8-nodes"
+set ylabel "Speedup w.r.t. 8-node time"
+set title "(b) Speedup w.r.t. 8-node time"
 plot \
      "$$.tmp" u (\$2-1):(${BASE}/\$3) w lp axis x1y1 t "total", \
      "$$.tmp.2" u (\$2-1):(${BASE2}/\$3) w lp axis x1y1 t 'deploy minibatch', \

### 2016-01-22 00-abstract.tex
Polish on 01. Add 06.
Abstract edit: point out overlapping community detection != clustering

@@
 graph where the vertices and edges represent individuals and their
-relationships. In this paper, we present and evaluate an efficient parallel and
+relationships. As opposed to the problem graph partitioning or clustering, an
+individual can be part of multiple communities which significantly increases
+the problem complexity.
+In this paper, we present and evaluate an efficient parallel and
 distributed implementation of a Stochastic Gradient Markov Chain Monte Carlo
@@
 the algorithm can scale and process graphs consisting of billions of edges and
-tens of millions of vertices on a compute cluster of 65 nodes.
+tens of millions of vertices on a compute cluster of 65 nodes. To the best of
+our knowledge, this is the first time that the problem of deducing overlapping
+communities has been learned for problems of such a large scale.

### 2016-01-22 01-intro.tex
intro edits

@@
 
-There have been two main approaches to this issue; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning algorithms more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
-However, it is clear that in most cases we see the best performance by carefully integrating both of these approaches in one system.
+There have been two main approaches to resolving this issue; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning solutions more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
+However, the best possible performance is often achieved by carefully integrating both of these approaches in a single solution.
 
@@
 A community can be seen as a densely connected group of vertices that are only sparsely connected to the rest of the network.
-This problem is much harder than the related domain of detecting disjoint communities.
+This problem is significantly more complex than the related domain of detecting disjoint communities.
 
-This problem is modeled by the mixed membership stochastic blockmodels (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of MMSB, called assortative MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
-
+The problem of detecting overlapping communities is modeled by the mixed membership stochastic blockmodels (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of MMSB, called assortative MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
 The MMSB model is a probabilistic graphical model~\cite{koller2009probabilistic} that represents a convenient paradigm for modeling complex relationships between a potentially large number of random variables.
@@
 
-In this paper, we are particularly interested in the SG-MCMC method applied to the a-MMSB model introduced in \cite{LiAW15}, which turned out to be faster and more accurate than the SVB method.
+In this paper, we focus on the SG-MCMC method applied to the a-MMSB model.
+Recent work\cite{LiAW15} showed that this methodology is faster and more accurate than the SVB method.
 \cite{LiAW15} further proposes a heuristic to scale up the number of communities at the cost of less precision; the work in this paper scales the algorithm without that heuristic with a custom high-performance implementation.

### 2016-01-22 04-evaluation.tex
Add MPI roundtrip times

@@
   \epsfig{file=plots/qperf.eps, width=0.5\textwidth}
-  \caption{Performance of DKV store compared to qperf.}
+  \caption{Performance of DKV store compared to qperf and roundtrip MPI.}
   \label{fig-qperf}
@@
 read operations between one server and one client for a range of payload sizes,
-and compare these to the bandwidth achieved by \textit{qperf}, which shows
+and compare these to the bandwidth achieved by an MPI roundtrip test,
+and by \textit{qperf} which reports
 the best achievable performance for Infiniband. We present \textit{qperf}
@@
 identical, which corroborates the results from the Herd project for payloads
-upwards from 256B. The bandwidth achieved by our DKV store falls short
+upwards from 256B.
+The MVAPICH2 implementation
+is highly optimized for Infiniband. It requires an interaction between
+the server network card and the server host, which increases the latency and
+thus lowers the bandwidth
+significantly. This is ample justification for using a low-level RDMA
+implementation for the DKV store.
+The bandwidth achieved by our DKV store falls short
 of the \textit{qperf} performance for packets less than 4KB; this is attributed
@@
 size, the DKV store performance is hampered by the fact that its values
-are spread over a larger memory area, whereas \textit{qperf} always reads
+are spread over a larger memory area, whereas \textit{qperf} and MPI always reads
 from the same memory locations. For packets between 8KB and 512KB, our DKV

### 2016-01-22 plots/plot-qperf.sh

@@ -3,14 +3,15 @@ cat <<EOI | gnuplot --persist
 set terminal postscript eps enhanced color font ',8' size 3.3,1.6
 set output 'qperf.eps'
 
-set logscale xy
-set xrange [:2000000]
-set yrange [0.25:10]
-set key at 1000000,1
+set logscale x
+# set xrange [:2000000]
+# set yrange [0.25:10]
+set key at 1000000,3
 set xlabel "payload size (bytes)"
 set ylabel "throughput (GB/s)"
 plot \
 	"../data/qperf/qperf/qperf-write-bw" with lp title "qperf RDMA write", \
 	"../data/qperf/qperf/qperf-read-bw" with lp title "qperf RDMA read", \
-	"../data/qperf/dkv/thrp" with lp title "qperf DKV store read"
+	"../data/qperf/dkv/thrp" with lp title "DKV store read", \
+	"../data/qperf/mpi/thrp" with lp title "MPI read"
 EOI

### 2016-01-22 03-design.tex
Polish a bit on 02, 04, 04, 05
First half of design edits

@@
 The SG-MCMC algorithm described in the previous section has an abundance of
-opportunities for parallelism, for the multi-threaded, shared-memory type as
+opportunities for parallelism, for the multi-threaded shared-memory type as
 well as for distributed-memory parallelism. The benefit of multi-threaded
-parallelism is speed-up of the computation. A distributed implementation
-additionally allows to store data in the collective memory of the cluster
-machine and increases memory bandwidth which scales with the number of
+parallelism is the speedup of the computation. Additionally, a distributed implementation
+allows the storage of data in the collective memory of a cluster
+and increases aggregate memory bandwidth. Both of these parameters scale with the number of
 machines. The downside of a distributed implementation is that it requires
@@
 
-This section describes how we parallelize the algorithm, which had been
+This section describes how we parallelized the algorithm, which had been
 implemented sequentially in~C++.
-In most places, the usage of multi-threaded parallelism
-is straightforward, therefore, we will discuss details only where appropriate. The
-distributed design follows a master-worker paradigm, where in essence
-the master controls the parallel operations and the workers perform the
+The usage of multi-threaded parallelism
+is straightforward, therefore, we will discuss its details only where appropriate. The
+distributed design follows a master-worker paradigm, where
+the master controls the sequence of parallel operations and the workers perform the
 calculations. For thread parallelism, we annotate the program with
@@
 The distribution within the graph of the vertices of the mini-batch as well as
-the neighbor sets is completely random. That means that there is no locality
+the neighbor sets is completely random. That means that there is \textit{no
+locality}
 in $\pi$ access patterns, especially in the accesses to the neighbor sets
@@
 %\looseness=-1
-Modern network technology allows access to the memory of remote machines
-through the network cards (NICs) without involvement from the remote host
-altogether~\cite{Hamada_infinibandtrade,Beck:2011:PER:2043535.2043537}.
+Modern network technology allows for Remote Direct Memory Access (RDMA)
+through the network cards (NICs) with negligible
+overhead~\cite{Hamada_infinibandtrade,Beck:2011:PER:2043535.2043537}. The use
+of RDMA enables communicating peers to avoid the high latency associated with
+interrupting a processor and waiting for it to act upon a request. Instead, a
+process would register a region of memory that can be directly accessed by the
+NIC. However, RDMA only provides primitive operations and limited abstractions.
 This has opened up the development of
-extremely efficient remote memory services, often in the form of DKV stores.
-This project stores $\pi$ in a DKV store in the collective memory of the
-cluster, where $\pi[i]+\sum\phi[i]$ is the value for key $i$. We decided not
+efficient remote memory services, often in the form of distributed key-value
+(DKV) stores.
+
+The distributed algorithm implementation stores $\pi$ in a DKV store in the
+collective memory of the
+cluster, where $\pi[i]+\sum\phi[i]$ is the value for key $i$.
+%
+We decided not
 to use any of the existing implementations of RDMA DKV stores, and instead
-build our own DKV store on top of the Infiniband ib-verbs~API. The
-main consideration is that our use case is unusually simple in a number
-of important aspects. The KV layout is static in the sense that there are
-no inserts or deletes after the initial population, which allows a static
-partitioning of KV pairs over the machines. The values are all of the same size,
-a vector of $K+1$ floats. The access pattern is well controlled because the
+build our own DKV store on top of the Infiniband ib-verbs~API.
+%
+This stems from the fact that our use case is unique in a number
+of important aspects. First, the number of keys and values to be stored are
+constant throughout the execution of the algorithm. For instance, no records
+are dynamically inserted or deleted from the DKV after their initial insertion.
+However, the value associated with each key can be updated. Additionally, all
+keys are of the same type and represent a unique vertex in the graph.
+Therefore, keys can be represented as integer identifiers. Moreover, each key
+points to a vector of $K+1$ floats as its value.
+%
+This data representation can be exploited to simplify the management of
+read and write requests. Similarly, the keys can be deterministically
+partitioned across the cluster using a fast hash function.
+
+The algorithm's access pattern is well controlled because the
 computation is partitioned into stages that are separated by barriers. A
-stage either reads values or updates values. The algorithm ensures that updates
+single stage either reads values or updates values. The algorithm ensures that updates
 are always targeting unique elements,
@@
 stores (RamCloud~\cite{Ousterhout:2015:RSS:2818727.2806887}, FaRM~\cite{179767},
-and reputedly the fastest, Herd~\cite{Kalia:2014:URE:2740070.2626299}) incur
+and Herd~\cite{Kalia:2014:URE:2740070.2626299}) incur
 overhead for insert/delete flexibility, concurrency control, or variable-sized
 values. Our use case allows us to do any operation in exactly one RDMA read
-or RDMA write. Herd argues
-that replacing RDMA reads by
-an RPC to the server, followed by an unreliable write, is faster. However,
-they show this only holds for small payload packets, up to 256B. A $\pi$
+or RDMA write.
+%
+Herd argues that replacing RDMA reads by
+an RPC to the server, followed by an unreliable write, is faster.
+%
+However,
+the paper shows this only holds for small payload packets, up to 256B. A $\pi$
 packet in our application is typically thousands to hundreds of thousands
 of 4-byte floats.
+
+% horrible hack to move the table up
+\begin{table*}
+  \centering
+  \begin{tabular}{l r r r l}
+    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
+                    &            &               & communities    &             \\
+    \hline
+    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
+    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
+    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
+    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
+    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
+    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
+    \hline
+  \end{tabular}
+  \caption{Summary of SNAP graph data sets used for evaluation.}
+  \label{tab-datasets}
+\end{table*}
 
@@
 in parallel shortens the critical path.
-
-% horrible hack to move the table up
-\begin{table*}
-  \centering
-  \begin{tabular}{l r r r l}
-    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
-                    &            &               & communities    &             \\
-    \hline
-    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
-    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
-    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
-    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
-    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
-    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
-    \hline
-  \end{tabular}
-  \caption{Summary of SNAP graph data sets used for evaluation.}
-  \label{tab-datasets}
-\end{table*}

### 2016-01-22 plots/plot-ml-curves.sh
Fix font and line width in all plots

@@ -11,8 +11,8 @@ set grid ytics
 w = 0.6
 h = 0.4
 k = 0.0
-nr = 2
-nc = 3
+nr = 3
+nc = 2
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
@@ -22,40 +22,39 @@ set output 'ppx.eps'
 set multiplot layout nr,nc
 set size w,h
 
-
 set origin col(0),row(0)
 set size w,h
 
 set title "(a) com-Friendster | 12K communities on 64-nodes"
-plot '< grep average_count ../data/ml-curves/log-Friendster-K12K-flt32-m16K-n32-mxs-16K-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l
+plot '< grep average_count ../data/ml-curves/log-Friendster-K12K-flt32-m16K-n32-mxs-16K-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
 set origin col(1),row(0)
 set size w,h
 
 set title "(b) com-LiveJournal | 192K communities on 64-nodes"
-plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l
-
-set origin col(2),row(0)
-set size w,h
-
-set title "(c) com-Orkut | 256K communities on 64-nodes"
-plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst1e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l
+plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
 set origin col(0),row(1)
 set size w,h
 
-set title "(d) com-Youtube | 8,385 communities on 14-nodes"
-plot '< grep average_count ../data/ml-curves/log-youtube-K8385-m1K-n32-msx1K-np15/log.0' u (\$8 / 3600):(exp(\$14)) w l
+set title "(c) com-Orkut | 256K communities on 64-nodes"
+plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst1e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
[…]

### 2016-01-22 plots/plot-qperf.sh

@@ -1,7 +1,12 @@
 cat <<EOI | gnuplot --persist
 
-set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+set terminal postscript eps enhanced color font ',10' size 3.3,1.6
 set output 'qperf.eps'
+
+set style line 1 lt 1 lw 2 linecolor rgb "red"
+set style line 2 lt 2 lw 2 linecolor rgb "blue"
+set style line 3 lt 3 lw 2 linecolor rgb "#006400"
+set style line 4 lt 4 lw 2 linecolor rgb "magenta"
 
 set logscale x
 # set xrange [:2000000]
@@ -10,8 +15,8 @@ set key at 1000000,3
 set xlabel "payload size (bytes)"
 set ylabel "throughput (GB/s)"
 plot \
-	"../data/qperf/qperf/qperf-write-bw" with lp title "qperf RDMA write", \
-	"../data/qperf/qperf/qperf-read-bw" with lp title "qperf RDMA read", \
-	"../data/qperf/dkv/thrp" with lp title "DKV store read", \
-	"../data/qperf/mpi/thrp" with lp title "MPI read"
+	"../data/qperf/qperf/qperf-write-bw" with lp ls 1 title "qperf RDMA write", \
+	"../data/qperf/qperf/qperf-read-bw"  with lp ls 2 title "qperf RDMA read", \
+	"../data/qperf/dkv/thrp"             with lp ls 3 title "DKV store read", \
+	"../data/qperf/mpi/thrp"             with lp ls 4 title "MPI read"
 EOI

### 2016-01-22 plots/plot-sweep-np-fixed-K.sh

@@ -50,7 +50,7 @@ nc = 1
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
-set terminal postscript eps enhanced color font ',8'
+set terminal postscript eps enhanced color font ',10'
 set size (nc*w),(nr*h+k)
 set output 'sweep-over-np-fixed-K.eps'
 set multiplot layout nr,nc
@@ -62,25 +62,30 @@ set yrange[0:]
 set key below
 set grid ytics
 
+set style line 1 lt 1 lw 2 linecolor rgb "red"
+set style line 2 lt 2 lw 2 linecolor rgb "blue"
+set style line 3 lt 3 lw 2 linecolor rgb "#006400"
+set style line 4 lt 4 lw 2 linecolor rgb "magenta"
+
 set origin col(0),row(0)
 set size w,h
 set ylabel "Time (Seconds)"
 set title "(a) Execution time"
 plot \
-     "$$.tmp"   u (\$2-1):3 w lp axis x1y1 t 'total', \
-     "$$.tmp.2" u (\$2-1):3 w lp axis x1y1 t 'deploy minibatch', \
-     "$$.tmp.3" u (\$2-1):3 w lp axis x1y1 t 'update\_phi\_pi', \
-     "$$.tmp.4" u (\$2-1):3 w lp axis x1y1 t 'update\_beta\_theta'
+     "$$.tmp"   u (\$2-1):3 w lp axis x1y1 ls 1 t 'total', \
+     "$$.tmp.2" u (\$2-1):3 w lp axis x1y1 ls 2 t 'deploy minibatch', \
+     "$$.tmp.3" u (\$2-1):3 w lp axis x1y1 ls 3 t 'update\_phi\_pi', \
+     "$$.tmp.4" u (\$2-1):3 w lp axis x1y1 ls 4 t 'update\_beta\_theta'
 
 set origin col(0),row(1)
 set size w,h
 set ylabel "Speedup w.r.t. 8-node time"
 set title "(b) Speedup w.r.t. 8-node time"
 plot \
-     "$$.tmp" u (\$2-1):(${BASE}/\$3) w lp axis x1y1 t "total", \
-     "$$.tmp.2" u (\$2-1):(${BASE2}/\$3) w lp axis x1y1 t 'deploy minibatch', \
[…]

### 2016-01-22 04-evaluation.tex
Make plots fit column width + fix evaluation intro to mention subsections in correct order

@@
 %
-First, an evaluation of the \textit{strong scaling} behavior of the distributed
+First, the \textit{weak scaling} of the solution is assessed with respect to increasing
+numbers of targeted latent communities.
+%
+Second, an evaluation of the \textit{strong scaling} behavior of the distributed
 algorithm is presented.
 %
-Second, the \textit{weak scaling} of the solution is assessed with respect to increasing
-numbers of targeted latent communities.
+Third, the effects of \textit{pipelining the computation} and varying the number of latent communities on the system's
+performance are analyzed.
 %
-Next, the effects of \textit{pipelining the computation} and varying the number of latent communities on the system's
-performance is analyzed.
+Next, we \textit{contrast} the effectiveness of scaling the computation horizontally
+and vertically.
 %
@@
 %
-Finally, we \textit{contrast} the effectiveness of scaling the computation horizontally
-and vertically.
+Finally, we apply the algorithm to large data sets and comment on its execution time.
 
@@
   \centering
-  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=0.48\textwidth}
+  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\columnwidth}
   \caption{Average execution time per algorithm iteration when varying
@@
   \centering
-  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=0.48\textwidth}
+  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\columnwidth}
   \caption{(a) Execution time of 2048 algorithm iterations for the
@@
   \centering
-  \epsfig{file=plots/hpc-cloud.eps, width=0.48\textwidth}
+  \epsfig{file=plots/hpc-cloud.eps, width=\columnwidth}
   \caption{Performance comparison between the distributed implementation
@@
   \centering
-  \epsfig{file=plots/qperf.eps, width=0.5\textwidth}
+  \epsfig{file=plots/qperf.eps, width=\columnwidth}
   \caption{Performance of DKV store compared to qperf and roundtrip MPI.}

### 2016-01-22 04-evaluation.tex
Tiny rephrasing in qperf

--- 2016-01-22 04-evaluation.tex
@@
 
-\begin{figure}[t] % [htbp]
+\begin{figure}[b] % [htbp]
   \centering
   \epsfig{file=plots/qperf.eps, width=0.5\textwidth}
-  \caption{Performance of DKV store compared to qperf and roundtrip MPI.}
+  \caption{Performance of DKV store compared to qperf and MPI roundtrip.}
   \label{fig-qperf}
@@
 and compare these to the bandwidth achieved by an MPI roundtrip test,
-and by \textit{qperf} which reports
+and by \textit{qperf}~\cite{qperf-mellanox} which reports
 the best achievable performance for Infiniband. We present \textit{qperf}

### 2016-01-22 plots/plot-qperf.sh

@@ -13,5 +13,5 @@ plot \
 	"../data/qperf/qperf/qperf-write-bw" with lp title "qperf RDMA write", \
 	"../data/qperf/qperf/qperf-read-bw" with lp title "qperf RDMA read", \
 	"../data/qperf/dkv/thrp" with lp title "DKV store read", \
-	"../data/qperf/mpi/thrp" with lp title "MPI read"
+	"../data/qperf/mpi/thrp" with lp title "MPI roundtrip read"
 EOI

### 2016-01-22 04-evaluation.tex

--- 2016-01-22 04-evaluation.tex
@@
   \epsfig{file=plots/qperf.eps, width=\columnwidth}
-  \caption{Performance of DKV store compared to qperf and roundtrip MPI.}
+  \caption{Performance of DKV store vs.\ \textit{qperf} and MPI roundtrip.}
   \label{fig-qperf}
@@
 and compare these to the bandwidth achieved by an MPI roundtrip test,
-and by \textit{qperf} which reports
+and by \textit{qperf}~\cite{qperf-mellanox} which reports
 the best achievable performance for Infiniband. We present \textit{qperf}
-bandwidth for both RDMA read and RDMA write operations; these are nearly
-identical, which corroborates the results from the Herd project for payloads
+bandwidth for both RDMA read and RDMA write operations; these are quite
+close, which corroborates the results from the Herd project for payloads
 upwards from 256B.
-The MVAPICH2 implementation
+The MVAPICH2 MPI implementation
 is highly optimized for Infiniband. It requires an interaction between
@@
 size, the DKV store performance is hampered by the fact that its values
-are spread over a larger memory area, whereas \textit{qperf} and MPI always reads
+are spread over a larger memory area, whereas \textit{qperf} and MPI always read
 from the same memory locations. For packets between 8KB and 512KB, our DKV

### 2016-01-22 plots/plot-sweep-K-fixed-np.sh
remove unnecessary plot title

@@ -35,7 +35,6 @@ set style line 3 lt 3 lw 2 linecolor rgb "#006400"
 set style line 4 lt 4 lw 2 linecolor rgb "magenta"
 
 set output 'sweep-over-K-fixed-np.eps'
-set title "Scaling K (fixed np=65)"
 set key left
 plot "$$.tmp" u 1:2 w lp ls 1 t "Pipelining disabled", "$$.tmp2" u 1:2 w lp ls 2 t "Pipelining enabled"
 

### 2016-01-22 plots/plot-hpc-cloud.sh
HPC cloud plot subtitles were too long.. extended beyond column width

@@ -88,7 +88,7 @@ set key below
 
 set origin col(0),row(0)
 set size w,h
-set title '(a) Performance of HPC Cloud (16 and 40 cores) vs. 1 DAS5 node (16 cores) using com-DBLP'
+set title '(a) HPC Cloud (16 and 40 cores) vs. 1 DAS5 node (16 cores) using com-DBLP'
 plot \
      "$$.tmp.hpc40" u 1:(1000*\$2) w lp ls 1 t 'HPC Cloud 40-cores', \
      "$$.tmp.hpc16" u 1:(1000*\$2) w lp ls 2 t 'HPC Cloud 16-cores', \
@@ -96,7 +96,7 @@ plot \
 
 set origin col(0),row(1)
 set size w,h
-set title '(b) Performance of HPC Cloud (40 cores) vs. 64 DAS5 nodes (64*16 cores) using com-Friendster'
+set title '(b) HPC Cloud (40 cores) vs. 64 DAS5 nodes (64*16 cores) using com-Friendster'
 plot \
      "$$.tmp" u 1:(1000*\$2)   w lp ls 1 t 'HPC Cloud', \
      "$$.tmp.2" u 1:(1000*\$2) w lp ls 4 t '64 DAS5 nodes'

### 2016-01-24 03-design.tex
Ai, typo update_pi <-> update_phi in 03

@@
 $\pi$ and \textit{update\_phi} are split up into chunks, and the calculation
-in \textit{update\_pi} for the current chunk is done simultaneously with
+in \textit{update\_phi} for the current chunk is done simultaneously with
 loading $\pi$ for the next chunk. The rationale for implementing pipeline

### 2016-02-17 03-design.tex
Add Henri's COMMIT ack

@@
 we trade a reduction in memory usage at the workers against limited
-communication costs. In contrast, $\cE_h$ is be partitioned statically over
-all machines for the parallel perplexity calculation.
+communication costs. In contrast, $\cE_h$ is replicated at
+all machines for the neighbor sampling and the parallel perplexity calculation.
 
@@
 memory consistency, this stage awaits completion of \textit{update\_phi} with
-an MPI barrier. After the calculation, the updated values of $\pi+\sum\phi$
-are written through the DKV store.
+an MPI barrier. After the calculation, the updated values of $\pi;\sum\phi$
+are written to the DKV store.
 

### 2016-02-17 04-evaluation.tex
Add Henri's COMMIT ack

@@
 its performance is to maximize the utility of the network resources.
-In figure~\ref{fig-qperf}, we provide maximum bandwidth numbers for
+In Figure~\ref{fig-qperf}, we provide maximum bandwidth numbers for
 read operations between one server and one client for a range of payload sizes,

### 2016-03-03 00-abstract.tex
Add Henri's COMMIT ack
fix abstract typo

--- 2016-01-15 00-abstract.tex
@@
 % 4. what are your results?
-
+%
 Recent advancements in machine learning algorithms have transformed the data
@@
 graph where the vertices and edges represent individuals and their
-relationships. In this paper, we present and evaluate an efficient parallel and
+relationships. As opposed to the problem of graph partitioning or clustering, an
+individual can be part of multiple communities which significantly increases
+the problem complexity.
+In this paper, we present and evaluate an efficient parallel and
 distributed implementation of a Stochastic Gradient Markov Chain Monte Carlo

### 2016-03-07 plots/plot-ml-curves.sh
Import new ML curves

@@ -25,7 +25,7 @@ set size w,h
 set origin col(0),row(0)
 set size w,h
 
-set title "(a) com-Friendster | 12K communities on 64 nodes"
+set title "(a) com-Friendster | 12K communities on 64+1 nodes"
 set yrange [105:]
 plot '< grep average_count ../data/ml-curves/log-Friendster-K12K-flt32-m16K-n32-mxs-16K-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
@@ -33,33 +33,33 @@ set origin col(1),row(0)
 set size w,h
 
 set yrange [420:]
-set title "(b) com-LiveJournal | 192K communities on 64 nodes"
-plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
+set title "(b) com-LiveJournal | 192K communities on 64+1 nodes"
+plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-epst5e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
 set origin col(0),row(1)
 set size w,h
 
 set yrange [*:]
-set title "(c) com-Orkut | 256K communities on 64 nodes"
-plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst1e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
+set title "(c) com-Orkut | 256K communities on 64+1 nodes"
+plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst3e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
 set origin col(1),row(1)
 set size w,h
 
-set title "(d) com-Youtube | 8,385 communities on 14 nodes"
-plot '< grep average_count ../data/ml-curves/log-youtube-K8385-m1K-n32-msx1K-np15/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
+set title "(d) com-Youtube | 8,385 communities on 12+1 nodes"
[…]

### 2016-03-07 03-design.tex
Move HPC Cloud descr to beginning of 04

@@
 overhead~\cite{Hamada_infinibandtrade,Beck:2011:PER:2043535.2043537}. The use
-of RDMA enables communicating peers to avoid the high latency associated with
+of RDMA enables communicating peers to avoid the high latency for
 interrupting a processor and waiting for it to act upon a request. Instead, a
-process would register a region of memory that can be directly accessed by the
-NIC. However, RDMA only provides primitive operations and limited abstractions.
+process registers a region of memory that can be directly accessed by the
+NIC. % RDMA only provides primitive operations and limited abstractions.
 This has opened up the development of
@@
 %
-This stems from the fact that our use case is unique in a number
+This is motivated by the fact that our use case is unique in a number
 of important aspects. First, the number of keys and values to be stored are
 constant throughout the execution of the algorithm. For instance, no records
-are dynamically inserted or deleted from the DKV after their initial insertion.
-However, the value associated with each key can be updated. Additionally, all
+are dynamically inserted or deleted from the DKV after their initial insertion,
+although the value associated with each key can be updated. Additionally, all
 keys are of the same type and represent a unique vertex in the graph.
@@
 memory consistency, this stage awaits completion of \textit{update\_phi} with
-an MPI barrier. After the calculation, the updated values of $\pi;\sum\phi$
+an MPI barrier. After the calculation, the updated values of $\pi+\sum\phi$
 are written to the DKV store.

### 2016-03-07 04-evaluation.tex
Move HPC Cloud descr to beginning of 04

@@
 
-Empirical results were obtained by performing experiments on the VU and
+Empirical results for the distributed implementation were obtained from experiments on the VU and
 Leiden University
@@
 each site are interconnected by
-FDR~InfiniBand. The MPI implementation used is MVAPICH2, which is optimally
-tuned for Infiniband~\cite{mvapich}. All of the experiments reported in this section utilized
+FDR~InfiniBand. The MPI implementation used is OpenMPI, which supports
+user-space Infiniband~\cite{gabriel04:_open_mpi}.
+
+We compared the distributed performance against a shared-memory virtual machine
+at SURFsara's HPC Cloud system with 40~cores and 1TB of memory; its underlying
+physical system contains 40 Intel Xeon E7-4850 cores clocked at 2.00GHz, and
+it does not oversubscribe resources.
+
+All of the experiments reported in this section utilized
 publicly available graphs from the Stanford Large Network Dataset
@@
 configurations. However, the number and size of messages exchanged between the
-nodes would vary significantly. Figure~\ref{fig-weak-scaling} presents the
+nodes varies significantly. Figure~\ref{fig-weak-scaling} presents the
 average execution
-time per algorithm iteration across different cluster sizes.
+time per algorithm iteration for different cluster sizes.
 This figure shows that the system's overall
@@
 algorithm is the inherent complexity of communication and synchronization. A
-considerably simpler approach would be developing a multi-threaded version and
+considerably simpler approach is developing a multi-threaded version and
 running it on a machine with abundant memory and CPU cores. In such a context,
@@
 negligible compared to using MPI primitives. To evaluate the efficacy of both
-approaches we utilized SURFsara's HPC Cloud system to instantiate a virtual
-machine with 40 Intel Xeon E7-4850 cores clocked at 2.00GHz cores and 1TB of
-memory. The physical machine underlying
-the HPC Cloud system contains 40~CPU cores and does not oversubscribe
-resources. Therefore, by provisioning all 40~cores we ensured that there is no
-resource contention from other users of the system. Figure~\ref{fig-scale-up}
+approaches we used the virtual
+machine with 40~cores and 1TB of
+memory at SURFSara's HPC Cloud system. By provisioning all 40~cores
+we ensured that there is no
+resource contention from other users. Figure~\ref{fig-scale-up}
 reports the execution time per algorithm iteration for two experimental setups.
@@
 com-Friendster has roughly 65 million vertices. In this case, the algorithm
-reached a stable state after 3-4 hours. Next, Figures~\ref{fig-ppx}-b and
+reached a stable state after 8-10 hours. Next, Figures~\ref{fig-ppx}-b and
 \ref{fig-ppx}-c present the convergence of com-LiveJournal and com-Orkut
@@
 the algorithm increases dramatically with larger K. However, the system was
-capable of reporting results for both in around 40 hours.
+capable of reporting results for both in around 60 hours.
 %
@@
 yields a much smaller storage requirement,
-these experiments were conducted on 14, 24 and 24 cluster nodes respectively.
+these experiments were conducted on 13, 21 and 21 cluster nodes respectively.
 

### 2016-03-07 04-evaluation.tex
Use OpenMPI also for the qperf-style benchmark

@@
 upwards from 256B.
-The MVAPICH2 MPI implementation
-is highly optimized for Infiniband. It requires an interaction between
+The OpenMPI implementation is highly optimized for Infiband because it has an
+ib-verbs implmentation. It requires an interaction between
 the server network card and the server host, which increases the latency and
 thus lowers the bandwidth
-significantly. This is ample justification for using a low-level RDMA
+significantly.
+Note that the comparison is
+lopsided: \textit{qperf} and MPI are low-level point-to-point benchmarks without
+any DKV implementation on top.
+This is ample justification for using a low-level RDMA
 implementation for the DKV store.

### 2016-03-07 plots/plot-qperf.sh

@@ -18,5 +18,5 @@ plot \
 	"../data/qperf/qperf/qperf-write-bw" with lp ls 1 title "qperf RDMA write", \
 	"../data/qperf/qperf/qperf-read-bw"  with lp ls 2 title "qperf RDMA read", \
 	"../data/qperf/dkv/thrp"             with lp ls 3 title "DKV store read", \
-	"../data/qperf/mpi/thrp"             with lp ls 4 title "MPI roundtrip read"
+	"../data/qperf/mpi/read-bw-openmpi"  with lp ls 4 title "MPI roundtrip read"
 EOI

### 2016-03-07 04-evaluation.tex
Typos, rephrase a bit qperf section

--- 2016-03-07 04-evaluation.tex
@@
   \caption{Performance comparison between the distributed implementation
-  running on DAS5 and the multi-threaded solution on machine with 40 cores and
+  running on DAS5 and the multi-threaded solution on one machine with 40~cores and
   1TB of RAM.}
@@
 upwards from 256B.
-The MVAPICH2 MPI implementation
-is highly optimized for Infiniband. It requires an interaction between
+The OpenMPI implementation has an ib-verbs implementation for Infiniband,
+hence it has very good performance.
+The MPI benchmark requires an interaction between
 the server network card and the server host, which increases the latency and
 thus lowers the bandwidth
-significantly. This is ample justification for using a low-level RDMA
+significantly.
+Note that the comparison is
+lopsided: \textit{qperf} and MPI are low-level point-to-point benchmarks without
+any DKV store implementation on top.
+The performance presented in this Figure
+is ample justification for using a low-level RDMA
 implementation for the DKV store.
@@
 from the same memory locations. For packets between 8KB and 512KB, our DKV
-achieves performance very close to \textit{qperf}.
+store achieves performance very close to \textit{qperf}.
 

### 2016-03-07 plots/plot-ml-curves.sh

@@ -27,7 +27,9 @@ set size w,h
 
 set title "(a) com-Friendster | 12K communities on 64+1 nodes"
 set yrange [105:]
+set xrange [:20]
 plot '< grep average_count ../data/ml-curves/log-Friendster-K12K-flt32-m16K-n32-mxs-16K-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
+set xrange [:*]
 
 set origin col(1),row(0)
 set size w,h
@@ -35,13 +37,15 @@ set size w,h
 set yrange [420:]
 set title "(b) com-LiveJournal | 192K communities on 64+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-epst5e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
+set yrange [*:]
 
 set origin col(0),row(1)
 set size w,h
 
-set yrange [*:]
+set xrange [:64]
 set title "(c) com-Orkut | 256K communities on 64+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst3e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
+set xrange [:*]
 
 set origin col(1),row(1)
 set size w,h

### 2017-05-03 distr-02-related.tex
Import all texts from IPDPS2016 and ParLearning2016

--- 2016-01-21 02-related.tex
@@
 If a pair of vertices ($a,b$) in the edge set \Edges are in the same community, i.e., $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
-Otherwise this probability is small.
+Otherwise this probability is a small value $\dt$.
 Each community has its connection strength $\beta_{k} \in (0,1)$ which reflects how likely its members are linked to each other.
@@
 \eenum
-$\eta$ and $\alpha$ are parameters to the Beta and Dirichlet distribution functions. $\dt$ (usually small) is a parameter for the algorithm.
+$\eta$ and $\alpha$ are parameters to the Beta and Dirichlet distribution functions.
 
@@
 Here, the parameters $\phi$ and $\ta$ are used to re-parameterize $\pi$ and $\beta$ respectively.
-After updating $\phi$ and $\ta$, we can obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$, respectively.
-That is, $\bt_k = \ta_{k1}/(\ta_{k0} + \ta_{k1})$ and $\pi_{ak} = \phi_{ak}/\sm{j}{K}\phi_{aj}$.
+After updating $\phi$ and $\ta$, we can obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$ respectively, see Table~\ref{tbl-symbols}.
 In the following, we briefly sketch the iterative update steps.
@@
 samples of the model parameters $\{\bt_t\}$ and $\{\pi_t\}$, the averaged
-perplexity on the held-out test set $\Heldout$ is 
+perplexity on the held-out set $\Heldout$ is 
 \begin{align}
@@
 \exp\left(-\frac{\sum_{(a,b)\in \Heldout}^{}\log
-\{(1/T)\sm{t}{T}p(y_{ab}|\beta_t,\pi_t)\}}{|\cE_{h}|}\right)
+\{(1/T)\sm{t}{T}p(y_{ab}|\beta_t,\pi_t)\}}{|\cE_{h}|}\right).
 \end{align}
 The perplexity is not evaluated at every iteration, but at regular intervals.
+Convergence is reached when the perplexity changes by less than a small
+threshold.
 

### 2017-05-03 gpu-01-intro.tex
Import all texts from IPDPS2016 and ParLearning2016

\section{Introduction}

The past decade has witnessed a tremendous increase in the applicability and
usefulness of artificial intelligence in our daily lives. Much of this activity
was fueled by the mainstream adoption of machine learning approaches as they
provide tools to solve inherently difficult problems.
%
However, many of these techniques require a huge amount of computation
which severely limits the
scale of problems that can be tackled.

In this paper, we present the methodology followed in designing and
implementing a high-performance version of an existing Stochastic Gradient Markov Chain
Monte Carlo (SG-MCMC) machine learning algorithm that detects overlapping
communities in graphs. The algorithm analyzes pair-wise interactions between
entities in order to discover hidden attributes.
%
For instance, consider a social network represented as an undirected graph
where the vertices represent individuals and edges represent relations between
them. Given the relation information, the algorithm can identify latent groups
of individuals that represent shared interests.
%
This problem structure differs from graph partitioning or clustering as there
is a many-to-many relationship between individuals and interests. For example,
each individual can have multiple interests. Simultaneously, each interest group can
span multiple individuals.
%
Formally, this problem domain is known as Mixed-Membership Stochastic
Blockmodels (MMSB). The theory behind the algorithm is discussed in more detail
in~\cite{DBLP:journals/corr/LiAW15}.

The focus of this work is on the computational efficiency and parallel
performance of the SG-MCMC algorithm. More specifically, we discuss the process
of accelerating the algorithm by developing aggressive optimizations targeting
multi-core CPUs and GPUs. The parallel algorithm achieves speedup factors up to 86, compared to a
well-tuned sequential C++ program which itself is a factor~1000-1500 faster than
the original Python/Numpy program developed by the algorithm's authors. From a
computational point of view, this algorithm differs from widespread machine
learning algorithms in several ways. First, it is highly data-intensive which
makes GPU acceleration particularly challenging. Second, owing
to the algorithm's stochastic nature, the majority of its memory access
patterns and data dependencies are non-deterministic. As a result,
straightforward
optimization attempts of the memory access patterns either fail or lead
to non-intuitive results.

Through careful analysis of the computation and data structures we show that
the algorithm's full state can be reduced by roughly 75\%. Compressing the
state significantly reduces the data intensity and allows for tackling larger
problems while maintaining all state in memory.
%
Further, by cataloguing and accounting for the various load and store
operations, we identified the highest priority locations of data reuse. In
order to circumvent the unclear optimization landscape, we developed an
effective kernel code generation mechanism that explores the benefits of
exploiting all permutations of the available optimization opportunities. These
optimizations include caching in shared memory, caching in the register file,
loop unrolling and explicit vectorization.

In summary, the contributions of this
work are:
\begin{itemize}
  \item Decreasing the algorithm's data intensity by eliminating 75\% of its
    memory footprint
  \item Tuning the algorithm's performance by maximizing data reuse and
    identifying the highest performing combination of optimizations through
    dynamic kernel code generation
  \item Achieving speedup factors of 21 and 86 over an optimized sequential
    program using a multi-core CPU and a GPU respectively
  \item A comparative performance analysis of the accelerated algorithm
    versions on a multi-core CPU and four GPUs, highlighting the particular
    optimization combinations that were successful per device.
\end{itemize}

The remainder of this paper is organized as follows. A description of the
sequential version of the algorithm and its data structures is provided in
Section~2. Section~3 delves into the design of the parallel algorithm and
discusses its contributions.  Section~4 provides an empirical evaluation of the
contributions of this work.  Section~5 presents an overview of related works.
Finally, Section~6 concludes.

### 2017-05-03 gpu-02-related.tex
Import all texts from IPDPS2016 and ParLearning2016

\section{Related Work}

In this section, we discuss related machine learning projects that use
accelerators.
Immediately when GPUs evolved into generally programmable accelerators, machine
learners recognized the potential of GPUs for their science.
Much research
has been devoted to deep-learning implementations on GPUs. However, Mixed
Membership deduction (MMSB) is usually considered untractable for deep-learning
algorithms. In contrast, our algorithm belongs to the family of Bayesian
(approximative, stochastic) methods, where efficient algorithms for large-scale
data have recently been developed. A broad survey of Bayesian methods is given
in~\cite{DBLP:journals/corr/Zhu0H14}. Broadly speaking, Bayesian approximative
algorithms are subdivided into variational and Monte Carlo methods; Markov
Chain Monte Carlo (MCMC) is a subgroup of the latter. GPUs have been put to
use in MCMC algorithms, e.g.\ \cite{journals/bioinformatics/MedlarGSBK13}
uses GPUs to analyse parental linkage patterns in a biology context and
\cite{DBLP:journals/csda/WhiteP14} does the same to model terrorist activity.
Variational inference on MMSB is described
in~\cite{DBLP:conf/nips/GopalanMGFB12}, but they don't use GPUs. Latent
Dirichlet Allocation, another variety of Bayesian Approximation, is used on
GPUs by~\cite{DBLP:conf/nips/YanXQ09}.

Still, for high-dimension problems, the standard MCMC algorithms fall short in
performance. Recently, varieties of MCMC algorithms have been developed where
gradient information is used to speed up convergence. Langevin and Hamiltonian
dynamics are representatives of these
varieties~\cite{Girolami_riemannmanifold}.  Our algorithm uses Riemann Manifold
Langevin dynamics. In \cite{beam2014fast}, GPUs are used to perform Hamiltonian
descent, using Python interfaces to access the standard cuBLAS
library~\cite{cuBLAS}. They limit their optimizations for the GPU to
saving on data transfers between host and device memory. They do not apply
their framework to MMSB.

A different MMSB algorithm with stochastic gradient descent on the GPU is
the Online Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13}. They deduce
overlapping communities for many of the same datasets as this paper, and they
report fast convergence. Their implementation uses the cuBLAS
library, and, unlike our work, there is no attempt to hand-optimize the GPU
kernels. Additionally, the algorithmic approach to solve the problem is
fundamentally different.
Since they target GPUs only, the datasets they
can handle are limited by the device memory of the GPU. Our implementation converges
quickly on the GPU but can also function with reduced speedup on a multicore
CPU. In the latter case, the dataset
size is only limited by the memory of the host machine.

\begin{comment}
Since our SG-MCMC algorithm is completely different from deep-learning
algorithms, it makes no real sense to compare to the learning frameworks
Theano, Caffe, cuDNN, or NVidia Digits, even though they target GPUs.

SG-MCMC work NOT on GPUs:

GPU work:
MCMC Hamiltonian:
	1. Andrew L. Beam, Sujit K. Ghosh, Jon Doyle
		Fast Hamiltonian Monte Carlo Using GPU Computing
		http://arxiv.org/pdf/1402.4089.pdf
MMSB tensor:

MCMC (not SG):
    1. Alan Medlar, Dorota Głowacka, Horia Stanescu, Kevin Bryson, Robert Kleta
       SwiftLink: Parallel MCMC linkage analysis utilising multicore CPU and GPU
       http://bioinformatics.oxfordjournals.org/content/early/2012/12/13/bioinformatics.bts704.full.pdf
    2. Marc Suchard, Chris Holmes, Mike West
       Some of the What?, Why?, How?, Who?  and Where?  of Graphics Processing Unit Computing for Bayesian Analysis
       https://stat.duke.edu/gpustatsci/GPU-ISBABull2010.pdf
\end{comment}

\begin{comment}
In this paper, we describe our custom RDMA D-KV (Distributed Key-Value)
store. Current RMDA D-KV store implementations are RamCloud~\cite{RamCloud},
Pilaf~\cite{Pilaf}, Herd~\cite{Herd} and FaRM~\cite{FaRM}. All these systems
use RDMA to implement a D-KV store. However, all of them are far more powerful
than our custom implementation -- and this power comes at a cost that we
can avoid. They implement a generic D-KV store that controls concurrency,
supports dynamic inserts and deletes, supports variable-sized values
(whose size may change at an update), and keys of arbitrary type. Because
of the nature of our distributed algorithm, we have to deal with none of
these issues. For us, values are fixed-size, allocated only at the initial
population, and remain alive forever. We have no concurrency between writes
and reads or other writes. Our keys are a contiguous range of integers. All
these properties together allow an extremely low-overhead implementation
that does not involve the remote host in any transaction.
\end{comment}

### 2017-05-03 gpu-03-02-algorithm.tex
Import all texts from IPDPS2016 and ParLearning2016

\begin{comment}
Stochastic: iteratively:
 - sample a mini-batch $M$ (from the graph, or not from G)
 - for each node in the mini-batch:
   sample $n$ random 'neighbors' (not from G)
   for each edge (node, neighbor), calculate its contribution to the gradient
   add noise to the gradient contributions to be able to jump out of local minima
 - sum the gradients for each node in the mini-batch
 - using the gradients, infer new value of $\phi$ for each node in the mini-batch
 - update $\pi$ accordingly
 - calculate the gradients for $\beta$ for the edges in the mini-batch, and
   update $\beta$ and $\theta$
 - every so many iterations:
    + verify the quality of $\pi$ and $\beta$ against the validation set; metric is
      perplexity
    + terminate when perplexity changes less than $\epsilon$
\end{comment}

% \begin{figure}[h]
\begin{lstlisting}[mathescape,caption=SG-MCMC MMSB algorithm,label=algorithm-pseudo-code]
iterate
   sample a mini-batch $M$
   // $M$ may be a subset of $G$, or disjoint with $G$
   for each vertex $i$ in $M$:
      sample $n$ random 'neighbors' (not from $G$)
      for each edge ($i$, neighbor)
         calculate its contribution to the gradient in $\phi$
      // to be able to jump out of local minima:
      add noise to the gradient contributions
      update $\phi[i]$ using stepsize $s$
   for each vertex $i$ in $M$:
      update $\pi[i]$ according to changed $\phi[i]$
   for the edges in $M$:
      calculate gradients in $\theta$ and update $\theta$
   for the edges in $M$:
      update $\beta$ accordingly
   every so many iterations:
      // metric is perplexity:
      verify the quality of $\pi$ and $\beta$ against the validation set
until perplexity change is less than $\epsilon$
\end{lstlisting}
% \end{figure}

### 2017-05-03 gpu-03-problem-statement.tex
Import all texts from IPDPS2016 and ParLearning2016

\section{SG-MCMC Algorithm Overview} 
\label{sec-algorithm}

In this section we describe the computational aspects of the SG-MCMC MMSB
algorithm. Moreover, we will introduce the data structures and notation that
will be used throughout this paper.
%
Kindly refer to~\cite{DBLP:journals/corr/LiAW15} for a detailed explanation of
the statistical concepts of the algorithm.

The network graph $G$ consists of the undirected edges~$E$ and has $N$ vertices.
The algorithm starts by partitioning $G$ into the \textit{training
set}, the \textit{validation set} $H$ and the test set (the latter is not used
in our implementations). $H$ and the test set
are much smaller than $G$, typically between 1\% and 10\% of the edges
in~$G$.
The number of communities $K$
is specified as a model parameter to the algorithm.

The algorithm progresses by iteratively improving the global state of the learning
problem, using the training set. There are two pairs of data structures that
hold the global state. $\theta$, a $K{\times}2$ matrix, is used to
calculate
the community strength. The community strenghth represents the probability that two members in a community share
an edge. The actual community strength
is the vector $\beta$ of length $K$. As is shown in Equation~\ref{eqn-beta}, it is the
normalized version of~$\theta_{k,2}$.
The $\pi$ matrix of dimensions $N{\times}K$ represents the probability for
each vertex in $G$ to be a member of each community. It is the normalized
equivalent of the $\phi$ matrix of dimensions $N{\times}K$, on which the
calculations occur. The
definitions of $\beta$ and $\pi$ are:
%
\begin{eqnarray}
    \label{eqn-beta}
    \beta_{k} = \frac {\theta_{k,2}} {\theta^{sum}_k} &
    	\textrm{where} & \theta^{sum}_k = \sum_{j=1}^{2} \theta_{kj} \\
    \label{eqn-pi}
    \pi_{ik} = \frac {\phi_{ik}} {\phi^{sum}_i} &
    	\textrm{where} & \phi^{sum}_i = \sum_{k=1}^{K} \phi_{ik}
\end{eqnarray}

Pseudo-code for the algorithm is presented in
Listing~\ref{algorithm-pseudo-code}.

An iteration in the algorithm consists of 6 compute stages.
We will highlight the data accessed in the stages as this
determines the opportunities for parallelism.

\begin{figure}[tb]
\input{03-02-algorithm}
\end{figure}

The first stage (line~2 in the listing) randomly draws a mini-batch $M$,
using the "stratified random node" strategy~\cite{DBLP:journals/corr/LiAW15}.
In this strategy, a coin toss is used to decide between two sample types.
%
The first sample type
chooses one random vertex $i$ and selects all of its edges to constitute the
mini-batch $M$. This sample type is referred to as \textit{link edges}. On the other
hand, the second sample type randomly draws a vertex $i$ and generates
random edges of the form ($i$, $j$) such that the edges are not in $G$. This
sample type is referred to as \textit{nonlink edges}. The set of vertices that
constitute the edges of the mini-batch $M$ are represented by $m$.

In the second stage (line~4), for each vertex $i$ in~$m$, a neighbor set of
size $n$ is randomly sampled with edges of the form $(i,j)$.

Stage~3, \textit{update\_phi} (line~6-10), calculates a gradient vector
$\nabla\phi_i$ for each vertex $i$ in the mini-batch, by iterating over
the edges~${(i,j)}$ in $i$'s neighbor set; the data that is used is
$\pi_i$, $\pi_j$,
and~$\beta$. The gradient $\nabla\phi_i$ is used to update~$\phi_i$. Random
noise is added to the update to prevent the algorithm from getting confined
to a local minimum.
Stage 4 (line~11-13), \textit{update\_pi}, updates $\pi_i$ so it remains
the normalized version of~$\phi_i$.

Stage~5, \textit{update\_theta} (line~14),
uses $\beta$ and $\pi_a$, $\pi_b$ for the edges $(a,b)$ in the mini-batch~$M$ to
calculate a
gradient vector~$\nabla\theta$. $\theta$ is updated using $\nabla\theta$, again
with the addition of noise. Stage~6, \textit{update\_beta} (line~16),
recalculates $\beta$ as the normalized
version of~$\theta_{k,2}$.

At regular intervals, the algorithm's global state is assessed by
evaluating the perplexity over the edges in the validation set $H$. The
\textit{perplexity} is a metric that represents the quality of the algorithm solution at
a given point in time. It is used to detect the algorithm's
convergence.
The
perplexity, as defined in~\cite{DBLP:journals/corr/LiAW15}, is the exponential
of the average over time of the negative
log-likelihood of meeting a link edge. In this 7th stage, $\beta$ is used,
as are $\pi_a$ and $\pi_b$ for each edge $(a,b)$ in~$H$.

The graph $G$ is queried for membership in the stages \textit{update\_phi},
\textit{update\_beta}, and \textit{perplexity}. The validation set~$H$ is
traversed in \textit{perplexity}.

### 2017-05-03 gpu-04-02-sequential.tex
Import all texts from IPDPS2016 and ParLearning2016

\subsection{An Efficient Sequential Baseline Version}
\label{sec-design-sequential}

The initial implementation of the algorithm was developed by its authors in
Python as is common with machine learning researchers and practitioners. It
relied on numpy~\cite{numpy} to perform the numerical computations without the
overhead associated with Python's interpreter. However, the algorithm also
depended on native Python data structures such as sets, dictionaries and
lists, which are far less efficient.
This section presents how the Python implementation was
converted into an efficient sequential C++ application which serves as a
baseline for the parallel implementations.

Initially, the Python code base was ported into equivalent C++ code. Namely,
the program structure remained unchanged and the Python data structures were
simply substituted with their corresponding alternatives in C++.

The next step was to remove a number of inefficiencies in the C++
implementation. For example, one recurring idiom in the Python implementation
was an expression of the form $a^y b^{1-y}$, implemented with floating-point
exponentiation. A study of the algorithm showed that $y$ is a variable that can
only hold values of 0 or~1. Therefore, the expression is simply a choice
between $a$ for $y=1$ and $b$ for $y=0$. However, as the compiler cannot infer
the value of $y$, both expressions $a$ and $b$ were computed and their
resulting values were raised to the power $y$ and $1-y$ respectively.
We transformed such expressions into conditional expressions which
compute either the expression $a$ or $b$ and avoid exponentiation.
%
% Similarly, we transformed some expressions to a more efficient form.
% For example, we replaced expressions of the form $a^{0.5}\times{b}^{0.5}$
% with $sqrt(ab)$.
Other optimizations were loop strength
reduction and common subexpression lifting.

In summary, a multitude of localized optimizations were applied to achieve
an efficient sequential baseline version, which is 1000..1500 times faster
than the original Python code, see Section~\ref{sec-evaluation}.
All transformations were rigorously tested to ensure they do
not modify the behavior of the algorithm.

\begin{comment}
Further, we replaced calls to the system's random
functions with a custom implementation of the random generator
\textit{xorshift\_128}~\cite{Marsaglia:2003:XR}. This way, random calls no longer
involve system calls, so we can support easy and fast multi-threaded random
calls by providing each thread with its private, differently seeded, random
generator.
\end{comment}

### 2017-05-03 gpu-04-03-accelerator.tex
Import all texts from IPDPS2016 and ParLearning2016

\subsection{Restructuring for Parallelism}

% Accelerator friendliness
The design of an accelerated version of the MCMC algorithm necessitated several
crucial modifications to allow for efficient parallelization. This section
discusses the key design contributions to attain two optimized
parallel versions, catering for multi-core CPUs and GPUs. First, we present
common structural changes that provide a foundation for compute device specific
optimizations. Next, we provide an in-depth description of the optimized CPU
and GPU versions respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Using OpenCL and CUDA}
%OpenCL/CUDA
Initially, the system was developed using OpenCL~\cite{opencl} in order to
accelerate the computations of the algorithm. OpenCL was chosen as it provides
a common abstraction for a variety of compute devices which fulfilled our
requirements of employing CPUs or GPUs. However, Nvidia's OpenCL SDK limits the
total memory allocations within a context to 4GB which severely limits the
problem sizes that can be tackled on GPUs. Therefore, we migrated the system to
use the abstraction layer CLCudaAPI~\cite{claduc}
to support both OpenCL and CUDA~\cite{Nickolls:2008:SPP:1365490.1365500}
as back-ends.
%
CLCudaAPI
hides the difference between OpenCL and the CUDA driver API.
For example, the library provides abstractions for Device, Context, Queue and
Buffer. Additionally, the implementations use a combination of inline functions
and preprocessor macros to abstract away the differences between OpenCL's and
CUDA's primitive types, address spaces and functions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Fast Lookup of Graph Edges}
% Set/Cuckoo Hash

The algorithm relies on a set data structure to store the edges of the graph.
This set is queried frequently with randomly generated edges to check for their
membership. In order to improve the performance of such lookups, a custom set
implementation was developed that restricts the set's features based on an
observation of its usage patterns. For example, the set is used as a container
for the graph edges which are known in advance. Therefore, the set's maximum
size can be specified at creation time and need not be adjustable or provide
thread-safety guarantees as the graph is immutable. Further, the set data
structure is built as a variant of a cuckoo
hash~\cite{Pagh:2004:CH:1006424.1006426} where each element is a 64-bit
field containing a tuple of the two 32-bit vertices. The cuckoo hash
uses two hash functions to index their corresponding spaces. Additionally, the
cuckoo hash allows for 4 keys per bucket. This hash design allows for a loading
factor upwards of 90\% which reduces the hash space overhead
significantly.
%% FIXME: Talk lookup hit/miss with 4 keys per bucket?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Parallelization \& Data Dependencies}
% Division into kernels
The original algorithm was structurally reorganized into 4 main sections with
one or more kernels in each depending on data dependencies and the applicable
synchronization requirements.
%
First, the sampling of a mini-batch of edges is done on the host as it is a
cheap operation. The mini-batch sampling is followed by the neighbor sampling
kernel which generates uniformly random neighbors for each vertex in the
mini-batch. 
%
Second, the predominant kernel, \textit{update\_phi} is invoked to calculate the
gradients for each vertex $i$ in the mini-batch $m$ and updates the value of $ \phi_i
\mid i \in m$. Next, the \textit{update\_pi} kernel is invoked to normalize the
individual $\phi_i$ and store the result in the corresponding~$\pi_i$.
%
Third, the kernels \textit{update\_theta} and \textit{update\_beta} are invoked to modify the
global parameters.
%
Finally, the perplexity calculation is performed in a dedicated kernel.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Memory Footprint Reduction}
% Memory footprint halved

A key structural change applied to the algorithm is the lossless compression of
its state. This modification enabled the algorithm to process larger data sets
while maintaining all program state in memory. Moreover, as the algorithm is
data-intensive, a reduction of the state is accompanied by a decrease in its
data-intensity.
%
The data structures that occupy most memory are $\pi$ and~$\phi$, both of
which are matrices of dimensions ${N}\times{K}$. However, the storage of both
matrices is redundant as $\pi$ is a row-normalized copy of~$\phi$, see
Equation~\ref{eqn-pi} in Section~\ref{sec-algorithm}. The full storage for
$\phi$ is discarded; $\phi_{ik}$ values are calculated as
$\pi_{ik}/\phi^{sum}_i$, which requires maintenance of a vector
$\phi^{sum}$ of size $N$.
Moreover, the calculation for $\phi_{ik}$ can be cached usefully.
%
The $\phi$ matrix is required in two kernels only, namely, \textit{update\_phi} and
\textit{update\_pi}. Both kernels access $\phi_i$ only for vertices~$i$ in the
mini-batch~$m$, so for each iteration, the calculated values for $\phi$ are cached
in a smaller temporary matrix of size
${|m|}\times{K}$.
%
% In this context, the initialization of the algorithm's state is performed by
% normalizing the full $\pi$ matrix in-place without the need for an additional
% matrix. Moreover, the normalization denominator of each row is cached in
% $\phi^{sum}_{i}$.
%
% Thus each read-only access to $\phi_{i,j}$ can be substituted with
% $\pi_{i,j}\phi_{i}^{sum}$. Given that the value of $\phi_{i,j}$ is consistently
% used in conjunction with $\pi_{i,j}$ and that $\phi_{i}^{sum}$ is read once and
% reused multiple times, this
This transformation trades memory storage and
bandwidth for a minimal computation overhead.
%
% On the other hand, the newly computed $\phi$ values can be stored in the
% temporary matrix and later be normalized into their corresponding rows of
% $\pi$.

Thus, the original memory requirement for $\phi$, ${N}\times{K}$, is
reduced to a vector $\phi_{sum}$ of length~$N$ and a comparatively small
${|m|}\times{K}$ matrix. For sufficiently large~$K$, this transformation roughly
halves the memory footprint of the algorithm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Single floating point vs double
Another modification to the original algorithm is migrating all of its
computation from double to single precision floating-point.
%
This change affects the algorithm in multiple ways. For instance, it reduces
both the computation and data intensity. It again reduces the memory
footprint of the algorithm and frees registers, enabling more
effective data reuse.
%
It has been previously shown that stochastic learning algorithms do not require
high precision in the presence of statistical approximations and the addition
of random noise~\cite{DBLP:journals/corr/GuptaAGN15}.

%Finally, the algorithm requires randomly generated values from 3 different
%distributions.

\subsection*{CPU-Specific Optimization}
The multi-core
CPU version of the algorithm uses OpenCL to perform its computations. The
work decomposition of the CPU kernels ensures that each thread performs
independent computation in order to avoid expensive synchronization operations.
Edge-centric kernels, which operate over mini-batch edges perform computations
over every edge in parallel.  Similarly, vertex-centric kernels that operate
over vertices in the mini-batch exploit parallelism across the selected
vertices.
% vectorization
Additionally, the kernels were vectorized to decrease instruction overhead and
utilize the SSE capabilities of the CPU cores.

\subsection*{GPU-Specific Optimization}
\label{gpu-design}
The GPU implementation builds on top of the CPU work decomposition scheme.
However, instead of having every thread perform independent computations, each
block of threads shares the work associated with the single edge or vertex for
edge-centric and vertex-centric kernels.

Since all of the kernels are data-bound, several memory organization strategies
of varying complexities were investigated to speedup the algorithm by
exploiting the GPU's memory hierarchy.
% Update_phi
As a case study, a discussion of the \textit{update\_phi} kernel is provided as it is
the predominant component of the algorithm.

The \textit{update\_phi} kernel operates over every vertex~$i$ in the mini-batch and
requires two temporary vectors of length~$K$ to perform its computation. For
each vertex, it iterates over the randomly generated neighbors and computes a
vector of probabilities $Probs_i$ of length~$K$. The individual $Probs_i$ vectors
of each ($i$, neighbor) tuple are then used to update the gradients vector
$Grads_i$ for each vertex~$i$. Finally, the $\phi_i$ row is updated to
reflect the changes that were accumulated in $Grads_i$ for each vertex $i$ in the
mini-batch~$m$.

A deeper analysis of the memory access patterns of the \textit{update\_phi} kernel
revealed the frequency and modality of access to the data structures.
%
The read-only access of $\pi_j$ of the randomly generated neighbors are unique
with a high probability. More precisely, each $vertex_i$ in the mini-batch
randomly samples a set of neighbor identifiers from the uniform distribution.
Given that the total number of sampled neighbors is much smaller than $N$,
there is a low likelihood of having duplicate samples which eliminates the
potential of data reuse. Therefore, these accesses provide limited or no
opportunities for optimization without interfering with the algorithm's
entropy.
%
The data structure usage patterns that are deterministic and most frequently
accessed in read/write mode are $Probs$ and $Grads$ respectively. Similarly,
$\pi_i$ for each vertex in the mini-batch is read repeatedly for the
calculation of $Probs$ per neighbor and again for the updates of $Grads$.

The following strategies present alternative methods of handling the
deterministic memory usage patterns of $Probs_i$, $Grads_i$ and $\pi_i$ for each
vertex in the mini-batch.

% Naive strategy
\subsubsection*{The naive strategy} simply allocates temporary vectors in thread
local memory to store $Probs$ and $Grads$ which physically resides in
device memory. All memory accesses are coalesced to achieve the highest
possible bandwidth to and from device memory.

% Shared memory caching strategy
\subsubsection*{The shared memory strategy} allocates the temporary vectors
$Probs_i$ and $Grads_i$
in shared memory. Furthermore, this strategy copies the $\pi_i$ of the selected
mini-batch vertex to shared memory in order to avoid repeated reads of the same
memory locations in device memory

% Register and shared memory using code generation and loop unrolling 
\subsubsection*{The code generation strategy} dynamically generates the code of the kernel to
custom tailor its properties. For example, this strategy can control whether a
vector should be placed or cached in shared memory. Additionally, it can
control which vectors explicitly reside in registers by allocating space on the
stack frame, unrolling all inner loops of the kernel and substituting all
vector addressing with static values. The code generation strategy allows this
flexibility for the three vectors of concern, namely, $Probs_i$, $Grads_i$
and~$\pi_i$.
Hence, this strategy allows for 8 possible configurations denoted by
three letters each of which is a choice between $R$ or $S$ to represent
$Register$ or $Shared$ respectively. For example, $SSR$ denotes that $Probs_i$,
$Grads_i$ and $\pi_i$ were placed in $Shared$, $Shared$ and $Register$
respectively.

% vectorization
Additionally, we investigated the use of vector data types to decrease the
instruction overhead and increase memory bandwidth for all strategies.

### 2017-05-03 gpu-04-04-distributed.tex
Import all texts from IPDPS2016 and ParLearning2016

\begin{comment}
3 subsections:
   1. intro and motivation
   2. design
   3. evaluation

1. intro and motivation
     if K * N exceeds single-host memory need solution: use distributed memory, by
     partitioning pi/phi
     expect things to be memory-access dominated, so:
      - because of speed, cannot use disk or even SSD; fully random accesses, no data
        reuse or locality
      - need fastest remote memory available; introduce RDMA
     trade speed for memory in many places...

2. design
   - use master/worker paradigm

   - memory allocations:
     + distributed key-value store for pi/phi
       choose RDMA read/write; no synch, no concurrency, fixed size for reads
       as well as writes
     + distribute pi/phi over the workers
     + keep the graph at the master
     + distribute the held-out set over the master + workers

   - implementation:
     iterate:
     + master draws mini-batches
     + master scatters mini-batch and graph slice over the W workers
     + workers load pi
     + workers draw neighbors (OpenMP)
     + workers calculate updated phi (m/W * n) (OpenMP)
     + barrier
     + workers calculate and store updated pi (OpenMP)
     + barrier
     + master calculates beta (m) (OpenMP)
     + if perplexity iteration:
       barrier
       master and workers calculate perplexity, summed w/ reduce (OpenMP/MPI)

   - details
     + use google sparseset because std::unordered_set uses 32 bytes of overhead
       per item
     + have a look at RDMA bandwidth as a function of N and W

\end{comment}

\subsection{Distributed-memory solution}

The previous subsection described how an optimized MCMC a-MMSB solver
is implemented, but that implementation is limited to the memory of one
GPU or CPU. This means that the largest problems in the SNAP network
overview cannot be tackled in this way. The network graph requires to
store at least $N$ nodes and $E$ edges, and this requires at least $N +
2E$ 4-byte integers. The storage required for the~$\pi$ field can easily
exceed this: $N \cdot K$ 8-byte doubles. For example, to process the
Friendster graph for $K=4096$ communities, the memory for~$\pi$ alone
requires $N \cdot K$ doubles, i.e.\ 2TB of storage, which far exceeds
the memory of most current servers. Note that computing time can be
traded for memory because only one of the~$\pi$ and~$\phi$ fields needs
to be stored in full; the stored one can be used to recompute the other,
which is a fast, trivially parallel operation.

There are two important observations for a solution to this problem. First,
this type of solver performs a limited amount of computation per data item
that is loaded; therefore it is imperative to provide the fastest possible
access to~$\phi/\pi$ rows that are stored outside main memory. Second,
there is little or no locality in the rows of~$\phi/\pi$ that are accessed,
so opportunities for caching data accesses are very small. Consider the
dominant \texttt{update\_phi} calculation, which requires a row of~$\phi$
for each of the mini-batch members and $n$ rows of neighbor~$\phi$ per
mini-batch node. Only the mini-batch node~$\phi$ accesses can be cached, $1/n$
of all accesses.

In any case, fulfilling the memory requirements will be time-costly. Out
of hand, we discarded the option to use disks; disks, even SSDs, are
orders of magnitude slower than local memory access, and because of the
nature of the algorithm, execution speed will suffer proportionally.

The solution we investigate in this paper is to use the aggregate
memory of a cluster of machines by partitioning~$\pi$ over the memory of
the cluster machines. Our implementation uses Infiniband RDMA (Remote
Direct Memory Access)~\cite{RDMA}, a technology that allows a machine to
read or write the memory of another machine through their NICs (network
interface cards). The fastest RDMA primitives do not even involve the
remote machine host {\em at all}; all memory transfers are performed by
the remote NIC through DMA operations.

\subsection{The RDMA distributed key-value store}

The distributed storage of~$\pi$ is implemented as a D-KV (distributed
key-value) store: the key is the index, the value is a row of~$\pi$. The
key-value store is narrowly tailored to the distributed solver, because
that allows a very fast implementation:
\begin{itemize}
\item the values are fixed-size: an array of $K$ doubles
\item the layout of the data is static: at startup, the key-value store is
filled with~$\pi[node]$ values; during the algorithm's lifetime, there are
only reads and overwrites, no deletes or new inserts
\item the keys are contiguous in the range $0..N-1$, so the value address
can be calculated trivially from the key
\item the computation is separated into phases of reading~$\pi$ and phases
of writing~$\pi$: there are no concurrent write or read/write accesses, so
there is no need for concurrency control at all
\end{itemize}
These properties allow our D-KV store to use only RDMA reads and RDMA writes,
which means that there is never any involvement from the remote host.

\subsection{Design of the distributed algorithm}

The distributed algorithm uses MVAPICH2\cite{mvapich2} for all communication
except the RDMA D-KV store, and OpenMP\cite{openmp} for thread parallelism. The
implementation follows a master/slave paradigm. The master stores the full
network graph~$G$, the workers each store a partition of~$\pi$. \textit{ASIDE:
We decided not to scatter the graph over the workers, so all their memory can
be used to store~$\pi$.} In each iteration, the master draws the mini-batch
and partitions it evenly over the workers. The master also sends the slices
of $G$ accessed by each part of the mini-batch to its owner worker. The
workers independently draw a neighbor set for each mini-batch node that they
own. Then each worker loads~$\pi$ from the RDMA key-value store for its
mini-batch and neighbors nodes. Each worker then independently calculates,
with threads in parallel, an updated value for~$\phi$ for its mini-batch
nodes.  A barrier before the update to~$\pi$ through the D-KV store ensures
consistency. The master then calculates the update to $\beta/\theta$,
using thread parallelism; the amount of work in \texttt{update\_beta} is
too little to warrant a distributed implementation. Then, if applicable,
a perplexity calculation is performed in parallel by all machines in the
cluster. Each machine has its partition of the held-out set. It calculates
its contribution, with thread parallelism, to the perplexity value, and those
contributions are summed through an MPI reduce operation.

As we pointed out in Section~\ref{sec-algorithm}, the computational complexity for
\textbf{update\_phi} is $m \cdot n \cdot K$, with $m$ the number of nodes in the
mini-batch and $n$ the size of the neighbor sampler, which is independent of the size
of $G$. The perplexity calculation has complexity $h \cdot K$, with $h$ the size of
the held-out set. This is a fraction of the edges of $G$: 1\%~for huge network graphs,
otherwise~10\%.
The computation in our algorithm lends itself well for a distributed
implementation, but the remote $\pi$ accesses will shift the balance between
memory accesses and computation significantly. We will report quantitatively
in our evaluation Section~\ref{section-evaluation-distributed}.

### 2017-05-03 gpu-04-design.tex
Import all texts from IPDPS2016 and ParLearning2016

\section{System Design and Implementation}

The process of designing an accelerated version of the SG-MCMC MMSB algorithm
involved multiple transformations. First, an efficient C++ baseline version
was created. Further, changes to the algorithm and the data structures
were carried out, for both efficient resource utilization and parallelization.
This section provides an overview of the system's evolution in incremental
phases, identifying the key contributions and differences between consecutive
states.

\input{04-02-sequential}

\input{04-03-accelerator}

### 2017-05-03 gpu-05-02-sequential.tex
Import all texts from IPDPS2016 and ParLearning2016

\subsection{Achieving an Efficient Baseline Sequential Version}

This section presents the results of transforming the algorithm implementation
from Python to C++, and then removing obvious inefficiencies, to achieve a
good baseline implementation for comparison with the parallel implementation.
%
We also investigate the performance impact of using
32-bit floating point numbers in stead of 64-bit doubles.
This hardly reduces the computation intensity for the Xeon, but it very much
affects data intensity because the data sizes are halved.
%
% It has been previously shown that stochastic learning
% algorithms do not require high precision in the presence
% of statistical approximations and the addition of random
% noise [?].

Due to the performance limitations of the Python implementation, the
experimental configuration of this section is minimal.
%
Namely, we use the CA-HepPh dataset; the number of
iterations is 1000.
%Note that these trial runs stop very far before convergence; using a small
%number of
%iterations is valid because the time spent in an iteration does not vary
%during the lifetime of the algorithm.
The selected number of communities is~1024, the mini-batch size~32, the
neighbor sample size~32.

The mini-batch sampling as described in Section~\ref{sec-algorithm} randomly
chooses either a mini-batch of link edges, whose size is the degree of one
randomly selected vertex, or a mini-batch of nonlink edges, whose size is
specified as a model parameter. In this evaluation, we only select batches
of nonlink edges because that makes the mini-batch size, and hence the
execution times, deterministic. We separately validated that the time spent
\textit{per mini-batch vertex} for samples of link edges and nonlink edges
is fully consistent.

\begin{table}[b]
\center\begin{tabular}{l d{5.1} d{2.1} d{2.1} d{2.1}}
	&	& \multicolumn{3}{c}{C++} \\
\cline{3-5}
	& & 
            \multicolumn{1}{c}{Python} &
                \multicolumn{1}{c}{float64} &
		    \multicolumn{1}{c}{float32} \\
\multicolumn{1}{l}{Algorithm stage} &\multicolumn{1}{c}{Python} &
            \multicolumn{1}{c}{idiom} &
                \multicolumn{1}{c}{(baseline)} &
		    \multicolumn{1}{c}{baseline} \\
\hline
sample mini-batch       &      0.03 &  0.03 &  0.03 & 0.02 \\
sample neighbor sets    &      8.5  &  0.5  &  0.4  & 0.35 \\
update\_phi             &  9,014    & 52.7  &  8.9  & 4.9  \\
update\_pi              &     93.1  & 40.3  &  0.11 & 0.05 \\
update\_theta\_beta     &    339    &  2.1  &  0.76 & 0.55 \\
perplexity*             &    178    &  0.18 &  0.18 & 0.26 \\
\hline
\end{tabular}
\caption{Performance comparison between Python, C++ that follows the Python
idioms, and our baseline C++ implementations. 1000 iterations, times in
\textrm{ms} per iteration; *perplexity time divided by 100 iterations.}
\label{table-python-comparison}
\end{table}

Table~\ref{table-python-comparison} compares performance between Python,
the C++ version labeled 'Python Idiom' with the inefficiencies inherited from
Python, then our baseline C++ version which has the inefficiencies removed,
and finally the same with 32-bit floats.
%
We present timings for the compute stages of the algorithm described in
Section~\ref{sec-algorithm}, with the exception of
\textit{update\_theta\_beta} which combines \textit{update\_theta}
and \textit{update\_beta}.
%
The numbers are times in ms. For all stages except perplexity, the times
are per iteration.
%
Once every so many iterations (hundreds or thousands in production runs),
the perplexity calculation is invoked. The perplexity time in the table
has been divided by 100 iterations; per perplexity invocation, it is large in
comparison with the iteration stage times, but it is amortized over those
iterations.

In all implementations, \textit{update\_phi} dominates the computation.
The translation from Python into 'Python Idiom' C++ speeds up this stage
by a factor~171, even though the Python implementation uses numpy for its data
structures. Comparison of the C++ table entries for 'Python idiom' and
64-bit float baseline shows that our C++ optimizations speed up this stage
by another factor of~6. Reducing the floating-point precision to 32~bits
doubles this factor, which shows that the algorithm is very much data-dominant.
%
\textit{update\_pi} has a disproportionately large speed
gain after removing inefficiencies. This was attained by limiting the
update to $\pi$ to only those values of $\pi$ that have changed in this
iteration, in stead of the full $\pi$ array (which, incidentally, is handled
rather efficiently by numpy).
\textit{update\_theta\_beta} gave fewer opportunities for C++ optimization,
in the sampling stages there was none. In \textit{mini-batch-sampling} there
is no speedup compared to Python because it consists of a call to
a fast Python random primitive that cannot be bettered from C++.
The total performance gain between Python and the 64-bit float baseline C++
implementation exceeds a factor of~1000, and this grows to a factor of over~1500
if the floating-point precision is reduced to 32~bits. The 32-bit precision
implementation will be the baseline for the parallel performance comparison.

\begin{comment}
The introduction of a custom user-space random generator brings at most
a very small
benefit. We show it, because it is necessary for the multi-threaded
implementations described in the next section, and this measurement serves to
prove that it does not harm execution speed.
\end{comment}

### 2017-05-03 gpu-05-03-accelerator.tex
Import all texts from IPDPS2016 and ParLearning2016

\subsection{Analysis of CPU Parallelism}

\begin{comment}
  \begin{table}[t]
    \centering
    \begin{tabular}{l l c c c c}
      Comparison & Data Set        & K    & M    & n  & Iterations \\
      \hline
      Python/Baseline    & CA-HepPh        & 1024 & 32   & 32 & 1000 \\
      Parallel CPU/GPU   & com-DBLP        & 1024 & 4096 & 32 & 1000 \\
      Parallel CPU       & com-LiveJournal & 1024 & 4096 & 32 & 1000 \\
    \end{tabular}
    \caption{Experiment parameter reference.}
    \label{exp-params}
  \end{table}
\end{comment}

\begin{table}[b]	% [htb]
  \centering
  \begin{tabular}{l d{2.8} d{2.8}}
    Kernel & Time (seconds) \\
    \hline
    PPX CALC    &  0.0364737 \\
    PPX ACCUM   &  0.083 \\
    SAMPLING    &  0.535599 \\
    UPDATE\_PHI & 25.6598 \\
    UPDATE\_PI  &  0.645875 \\
    THETA SUM   &  0.0483902 \\
    GRADS PAR   &  1.92919 \\
    GRADS SUM   &  9.31122 \\
    UPDATE THETA&  0.0548013 \\
    NORM THETA  &  0.001 \\
    \hline
    TOTAL    & 38.5858 \\
  \end{tabular}
  \caption{Performance break-down of multi-core CPU version without
  vectorization. Model parameters: dataset com-DBLP; K=1024, m=4096, n=32.}
  \label{TABLE-CPU}
\end{table}

This section discusses the use of the multi-core CPU available on the DAS5
cluster. The parallel OpenCL version divides the work across the CPU cores and
performs independent calculations concurrently. As shown in
Table~\ref{TABLE-CPU}, the dominant kernel in the computation is
\textit{update\_phi},
which accounts for 66.5\% of the computation time. Without exploiting
the dual 8-core processor's vectorization capabilities, the speedup relative to
the baseline sequential C++ version is~9.8.

% CPU
\begin{figure}[htb]
\centering
\epsfig{file=plots/cpu-vector.eps, width=\columnwidth}
\caption{Performance of CPU for varying vector width.}
\label{FIG-CPU-VECTOR}
\end{figure}

In addition to applying computations in parallel, we investigated the use of
the CPU's SIMD instructions to maximize its resource utilization.
Figure~\ref{FIG-CPU-VECTOR} presents the performance obtained by vectorizing
the kernels with varying vector widths. A key aspect in this figure is the
diminishing performance benefit for higher vector widths. As the computational
performance increases, the memory throughput becomes the leading performance
bottleneck. Moreover, using 16-wide SIMD instructions gave a slight performance
penalty compared to 8-wide SIMD. The 8-wide vector version improves the speedup
relative to the baseline version from 9.8 to~20.9.

\begin{table}[b]	% [htb]
% \center\begin{tabular}{l c c c c c c c c d{3.1}}
% Device & Architecture & cores & Clock & \multicolumn{2}{c}{GFlops} & L2 Cache & Memory & Bandwidth \\ 
%        &              &       & MHz   & single & double            & KB       & GB     & (GB/s) \\ 
% \hline
% K20m        & Tesla   &  2496 &  706        & 3520 & 1170 &      &  5 & 208 \\
% K40c        & Tesla   &  2880 &  745        & 4290 & 1430 &      & 12 & 208 \\
% GTX 980     & Maxwell &  2816 & 1126 (1216) & 4612 &  144 &      &  4 & 336.5 \\
% GTX Titan-X & Maxwell &  3072 & 1000 (1075) & 6144 &  192 & 2048 & 12 & 336.5 \\
%
%
%###########################################################################################
% SPECS FROM:
% http://www.geforce.com/hardware/desktop-gpus/geforce-gtx-titan-x/specifications
% http://www.geforce.com/hardware/desktop-gpus/geforce-gtx-980/specifications
% http://www.nvidia.com/content/PDF/kepler/Tesla-K40-Active-Board-Spec-BD-06949-001_v03.pdf
% http://www.nvidia.com/content/PDF/kepler/Tesla-K20-Active-BD-06499-001-v04.pdf
%
% FLOPS DATA:
% K40: http://international.download.nvidia.com/pdf/kepler/TeslaK80-datasheet.pdf
% K20 & K40: http://www.nvidia.com/content/tesla/pdf/nvidia-tesla-kepler-family-datasheet.pdf
% GTX980 & TITANX: https://en.wikipedia.org/wiki/List_of_Nvidia_graphics_processing_units
\center\begin{tabular}{lc c c c c}
                         && \multicolumn{4} {c}{Device} \\
\cline{3-6}
Specification            && K20m  & K40c  & GTX 980 & GTX Titan-X \\
\cline{1-1}\cline{3-6}
Number of Cores          && 2496  & 2880  & 2048    & 3072  \\
Clock (MHz)              && 706   & 745   & 1126    & 1000  \\
GFlops (single)          && 3520  & 4290  & 4612    & 6144  \\
GFlops (double)          && 1170  & 1430  & 144     & 192   \\
Memory (GB)              && 5     & 12    & 4       & 12    \\
Bandwidth (GB/s)         && 208   & 288   & 224     & 336.5 \\
\cline{1-1}\cline{3-6}
\end{tabular}
\caption{Properties of the GPUs used in the evaluation}
\label{table-gpus}
\end{table}

\subsection{Analysis of GPU Parallelism}
As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the
GPU version of the algorithm. Specifically, there is the naive strategy, the
shared strategy and 8 variations of the code generation strategy. This
section investigates the effectiveness of each flavor on Nvidia's GTX TitanX,
GTX980, K40c and K20m.

% TITANX
\begin{figure*}[t]	% [htb]
  \centering
  \epsfig{file=plots/titanx-1024-strategies-w1.eps, width=\textwidth}
  \caption{Execution time of 1000 \textit{update\_phi} invocations using the TitanX GPU,
  without explicit kernel vectorization, across a sweep of
  \textit{update\_phi} thread block
  sizes. Relevant model parameters: K=1024, M=4096, n=32.}
  \label{titanx-w1-sweep}
\end{figure*}

% TITANX K=2K
\begin{figure*}[t]	% [htb]
  \centering
  \epsfig{file=plots/titanx-2048-strategies-w1.eps, width=\textwidth}
  \caption{Execution time of 1000 \textit{update\_phi} invocations using the TitanX GPU,
  without explicit kernel vectorization, across a sweep of
  \textit{update\_phi} thread block
  sizes. Relevant model parameters: K=2048, M=4096, n=32.}
  \label{titanx-w1-sweep-2k}
\end{figure*}

Figure \ref{titanx-w1-sweep}(a) presents the performance of the GTX TitanX GPU
for all 10 strategy flavors, without explicitly vectorizing the kernels. The
x-axis represents different \textit{update\_phi} thread block sizes while the y-axis presents
the total execution time of 1000 invocations of the \textit{update\_phi} kernel. The
naive and shared strategies are labeled NAIVE and SHARED respectively. Further,
each flavor of the code generation strategy is labeled by GEN followed by the 3
choices that identify it.
%
The wide performance range inhibits the readability of the figure,
therefore, a more focused Figure~\ref{titanx-w1-sweep}(b) is provided. As would
be expected, the naive strategy exhibits the worst performance over all
thread block
sizes, as it does not explicitly cache repeated device memory read operations.
The SHARED and \textit{GEN-SSS} strategies come next in terms of performance.
Both strategies cache $Grads_i$, $Probs_i$ and $\pi_i$ in shared memory but
differ in one aspect. Namely, \textit{GEN-SSS} explicitly unrolls the internal
loops of the kernel. However, there is no significant performance difference
between them. The other flavors of the code generation strategy attain higher
performance as they unroll internal loops as well as cache data in
registers. The TitanX obtains the best performance with the \textit{GEN-RSS}
strategy and a thread block size of 64. The results of explicitly vectorized kernels
are omitted as they obtain worse performance on the TitanX.

A key model parameter that affects the behavior of the optimization strategies
is the number of communities $K$. Figure~\ref{titanx-w1-sweep-2k} presents the
same model configuration as in Figure~\ref{titanx-w1-sweep} but $K=2048$
instead of $K=1024$. One key difference between the two figures is the optimal
thread block size. An increase in $K$ comes with a proportional increase in
the size of shared memory space that is used by each thread block for the
strategies that employ shared memory. Similarly, GEN strategies that use the
register file will require additional space. Therefore, the number of
concurrent thread blocks that can execute on a single streaming multiprocessor
will decrease, minimizing the GPU's occupancy and utilization. This effect is
most clear when comparing the NAIVE with the SHARED and GEN-SSS strategies. In
this case, the NAIVE strategy outperforms both SHARED and GEN-SSS for a
thread block
size of 32 due to their low occupancy. This limitation can be counteracted by
selecting a larger thread block size which in turn increases the computation
concurrency and occupancy.

% K40
\begin{figure*}[t]	% [hbt]
  \centering
  \epsfig{file=plots/k40-1024-strategies-w2.eps, width=\textwidth}
  \caption{Execution time of 1000 \textit{update\_phi} invocations using the K40c GPU, with
  explicit kernel vectorization of width 2, across a sweep of
  \textit{update\_phi} thread block
  sizes. Relevant model parameters: K=1024, M=4096, n=32.}
  \label{k40-w2-sweep}
\end{figure*}

% \looseness=-1
Figures~\ref{k40-w2-sweep} (a) and (b) present the performance of the K40c GPU
for the same experimental configuration as shown for the TitanX in
Figure~\ref{titanx-w1-sweep} with one exception. Namely, the code version used
for this plot is explicitly vectorized with a vector width of~2. The results
for the other code versions with vector width 4 and no vectorization are
omitted as they exhibit lower performance.
%
Surprisingly, Figure~\ref{k40-w2-sweep} shows that the NAIVE strategy
outperforms SHARED and some of the GEN strategies. This can be explained by the
unique properties of the Tesla Super Computing line of products to which the
K40c belongs. These GPUs include enhanced L2 caching mechanisms that
accelerate repeated and sparse memory accesses. This is especially advantageous
as it caches repeated reads across streaming multiprocessors. However, the
highest performance is attained by \textit{GEN-RRS} which explicitly employs
registers for both $Probs_i$ and $Grads_i$.

The performance results obtained from the GTX TitanX and Tesla K40c GPUs
reinforce the importance of customizing compute kernels to each GPU's specific
architecture and capabilities. For instance, each GPU achieved its highest
performance by employing a different strategy. Moreover, each GPU displayed
different strategy-performance orderings.

\begin{figure}[b]	% [htb]
  \centering
  \epsfig{file=plots/gpus.eps, width=0.9\columnwidth}
  \caption{Speedup comparison of best performing parallel configurations of
  each compute device normalized over baseline C++ version. Relevant model
  parameters: K=1024, M=4096, n=32.}
  \label{gpus-fig}
\end{figure}

\begin{figure}[b]	% [htb]
  \centering
  \epsfig{file=plots/gpu-vector-sweep.eps, width=0.9\columnwidth}
  \caption{Performance of the best performing strategy per GPU, varying the
  vector width, normalized over the non-vectorized kernel. Relevant model
  parameters: K=1024, M=4096, n=32.}
  \label{gpus-v-sweep}
\end{figure}

\subsection{Comparison of Compute Devices}
Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
GTX TitanX, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
sequential C++ version. These results are consistent with the relative
capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
the TitanX achieves the highest speedup of 86 relative to the baseline. At the
other end, the parallel CPU version attains a speedup factor of~20.9.

As vectorization is a popular GPU optimization, it is important to note that
both GTX GPUs achieved their highest performance with non-vectorized kernels.
On the other hand, the Tesla GPUs obtained the best performance with a vector
width of~2. Figure~\ref{gpus-v-sweep} presents the execution time of the best
performing strategy for each of the 4 GPUs. In this figure, the performance is
normalized over the non-vectorized kernel version for each GPU. For instance,
the TitanX exhibits an overhead factor of approximately 1.8 when using a vector
width of~4. On the other hand, the K40c improves its performance by roughly
20\% when it uses a vector width of~2 compared to the non-vectorized kernel.
Therefore, explicit vectorization of the kernels can be either useful or
harmful depending on the GPU architecture and the specific problem it is
applied to.

### 2017-05-03 gpu-05-04-distributed.tex
Import all texts from IPDPS2016 and ParLearning2016

\begin{comment}
3. evaluation
   - do we speak again about sampler strategies? no need, I think, or a
     one-liner that we use the best strategy from above, and did very short
     evaluation with large data set
   - measurements:
     + Friendster:
       - sweep over K and discuss
     + dblp???? soc-lj????
       - sweep over K and discuss
       - sweep over np and discuss
   - discuss time breakdown (need some terminated runs):
     + dblp
     + Friendster
     points to discuss:
     + update phi: compute time vs memory access time
     + update pi: idem, compare to update phi
     + update beta: idem, compare to update phi
     + perplexity: idem, compare to update phi

\end{comment}

\subsection{Distributed-memory evaluation}

The implementation started out using the std::unordered\_set in the C++
standard library as the datastructure to store $G$, $H$, the held-out set,
and test set. For the large networks like Friendster, this led to immediate
storage problems. The std::unordered\_set in gcc's implementation uses
between 24 and 32 extra bytes of storage per item~\cite{g++ unordered_set
memory overhead}, while a vertex in $G$ is represented by a 4-byte integer,
a 6 to 8-fold penalty. We switched our set implementation to use Google's
sparse-hash-set implementation~\cite{google-sparse-hash}, which has in the
order of~1\% space overhead. This set implementation is reported to be slower,
but in our application, the extra time was not measurable.

The evaluation of the distributed implementation of the MCMC a-MMSB solver
focuses on very large networks.

Use the best sampler strategy, RNL/Sx/BF

\subsubsection{Scalability over the number of machines}

Using dblp with -K 1K, -m 1K, -n 256:
\begin{itemize}
\item Compare sequential performance penalty against sequential C++ impl. and
	against OpenCL impl
\item analyze the time to load $\pi$ vs. compute time, for update\_phi and
	cal\_perplexity; check whether np is relevant; also check how $N*K/np$
	is relevant, using multiple networks (dblp, youtube, soc-lj)
\item sweep over np
\end{itemize}

\subsubsection{Scalability over the number of communities}

Present the Friendster data, incl some analysis: upsweep for K=256

### 2017-05-03 gpu-05-05-perplexity.tex
Import all texts from IPDPS2016 and ParLearning2016

\begin{comment}
\begin{figure}[t]	% [htb]
  \centering
  \epsfig{file=plots/ppx-gpu.eps, width=\columnwidth}
  \caption{Perplexity convergence of com-DBLP on the GTX~Titan-X~GPU. Number
  of communities chosen to maximally fill the GPU memory.}
  \label{fig-ppx-gpu}
\end{figure}

\begin{figure}[t]	% [htb]
  \centering
  \epsfig{file=plots/ppx-cpu.eps, width=\columnwidth}
  \caption{Perplexity convergence of com-LiveJournal on the multi-core~CPU.
  Number of communities chosen to maximally fill the CPU memory.}
  \label{fig-ppx-cpu}
\end{figure}
\end{comment}

\begin{figure*}[t]	% [htb]
  \centering
  \epsfig{file=plots/ppx-gpu-cpu.eps, width=\textwidth}
  \caption{Perplexity convergence. (a)~com-DBLP on the GTX~Titan-X~CPU.
  (b)~com-LiveJournal on the multi-core~CPU. Number of communities chosen to
  maximally fill the memory of GPU and CPU respectively.}
  \label{fig-ppx-gpu-cpu}
\end{figure*}

\subsection{Algorithm Convergence}

This section presents the evolution of the perplexity, i.e.\ the metric for
the solution quality, as a function of time for
two datasets, whose properties are given in Table~\ref{table-snap}. This
analysis shows that our implementations of the algorithm indeed achieve
convergence, and it gives an impression of the time required to achieve
convergence for the largest datasets and community sizes that fit into the memory
of a GPU and a CPU respectively. The dominant customer of memory is the
$\pi$ matrix, sized $N{\times}K$ for $K$ communities and $N$ vertices in the
graph. The runtime per iteration is independent of~$N$, but it depends on the
product of~$K$, the mini-batch size~$m$, and the neighbor sample size~$n$. On
the other hand, the number of iterations required to achieve sufficient
convergence does grow with the size of the graph.

Figure~\ref{fig-ppx-gpu-cpu}(a) shows the perplexity development for the dataset
com-DBLP, with $N=317,080$. It is computed on the GTX Titan-X GPU with input
parameter set $K=4096$, $m=4096$, $n=32$. Convergence is reached after 6~hours.

Figure~\ref{fig-ppx-gpu-cpu}(b) shows the perplexity development convergence for
com-LiveJournal, with $N=3,997,962$, computed on a multi-core DAS5 node.
To fill the CPU core memory~(64GB), the input parameter set is $K$=3072,
$m$=4096, $n$=32. Sufficient convergence is reached after 18~hours. Note
that this network graph is so large that the GTX Titan-X would be unable to
process more than $K$=256 communities.

### 2017-05-03 gpu-05-evaluation.tex
Import all texts from IPDPS2016 and ParLearning2016

\section{Evaluation}
\label{sec-evaluation}

The following subsections discuss the performance evaluation of the algorithm
across its different stages. As the system underwent multiple design mutations,
a careful analysis is needed to identify the contributions of each change
independently.
%
First, an examination is provided detailing the construction of an efficient
sequential baseline version.
This spans
the code conversion from Python to C++ as well as the effects of applying
the localized optimizations discussed in
Section~\ref{sec-design-sequential}.
%
Second, we explore the performance benefits of parallelizing the computations
on a multi-core CPU using OpenCL.
%
Next, we assess the trade-offs associated with each of the GPU optimization
strategies and their performance effects on four GPUs spanning two chip
architecture generations.
%
Finally, we present the overall outcome of accelerating the algorithm by
displaying its fast convergence over large networks.

All experiments were conducted on the VU Amsterdam DAS5 cluster. The cluster
consists of 68 compute nodes each equipped with a dual 8-core Intel Xeon
E5-2630v3 CPU clocked at 2.40GHz, 64GB of memory and 8TB of storage.
Additionally, the cluster is fitted with a number of Nvidia GPUs including GTX
TitanX, GTX980, K40c and K20m; see Table~\ref{table-gpus} for an overview of
their properties. Table~\ref{table-snap} outlines the graph
properties of the networks used to evaluate the algorithm's performance. These
networks were obtained from the SNAP collection~\cite{snapnets}.

\begin{table}[tb]
\center\begin{tabular}{l r r c}
Network & \#Vertices & \#Edges \\
\hline
CA-HepPh        &    12,008 &    118,521 \\
com-DBLP        &   317,080 &  1,049,866 \\
% com-Youtube     & 1,134,890 &  2,987,624 \\
com-LiveJournal & 3,997,962 & 34,681,189 \\
\hline
\end{tabular}
\caption{Network graphs from the Stanford Snap collection}
\label{table-snap}
\end{table}

\input{05-02-sequential}
\input{05-03-accelerator}
\input{05-05-perplexity}

### 2017-05-03 gpu-06-conclusion.tex
Import all texts from IPDPS2016 and ParLearning2016

\section{Conclusion and Future Work}
Modern machine learning algorithms are empowering their
users to solve or approximate solutions to previously intractable problems.
However, these advancements come at the cost of significant computational
complexity and long running learning processes which hinder our ability to reap
such algorithms' benefits when applied to large problems. There is a clear need
to assess such algorithms and identify optimization opportunities in order to
scale their performance.

The SG-MCMC algorithm discussed in this paper posed additional computational
challenges due to its unique stochastic nature and high data intensity. Unlike
common machine learning algorithms, its data dependencies and memory access
patterns are nondeterministic. In this paper, we presented our methodology of
improving the algorithm's performance by fundamentally restructuring it to
cater for concurrency.

We showed that the algorithm's state can be reduced by 75\% after a thorough
analysis of the computational patterns and data structures which significantly
reduces its data intensity. Further, we navigated the complex optimization
landscape by dynamically generating kernel code and testing different
combinations of optimizations. The outcome of these efforts culminated in
significant speedup factors of 21 and 86 using a multi-core CPU and a GPU
respectively. These speedup numbers where achieved in comparison to an already
optimized sequential implementation. Finally, we evaluated the performance of
the parallel algorithm across several GPUs, highlighting the difference between
their optimal configurations.

The outcome of this study reinforces the significance of avoiding premature
optimization as it can lead to unexpected results. In particular, the success
of common GPU optimizations depends on the particular device in use and the
problem it is applied to.

This paper focused on the acceleration of the SG-MCMC algorithm on a
single node. For future work, we are designing and implementing a distributed
version of the algorithm that uses the collective memory of a cluster, which will enable
processing of much larger datasets. These changes in context fundamentally
change the optimization landscape's boundaries and priorities.

### 2017-05-03 plots/collect-v.sh
Place the chunks of text unmodified into one document

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+function Find {
+  RET=$(grep -a PHI ${1}*W${2}* \
+    | grep -v LOG \
+    | awk '{print $1 " " $7}' \
+    | sed 's/\(:[^ ]\+\)//' \
+    | sed 's/.*-\([^-]\+\)-PHI\([0-9]\+\)/\1 \2/' \
+    | sed 's/-W\([0-9]\+\)-/ \1 /' \
+    | sed 's/-W\([0-9]\+\)/ \1 -/' \
+    | awk '{print $1" "$2" "$3" "$5" "$4}' \
+    | sort -nk 4 \
+    | head -n 1 )
+  echo $RET
+}
+# gets a line of the form:
+# GEN 64 1 3.13794 false-true-true
+# NAIVE 64 1 3.13794 -
+
+# find best config, get the W{1,2,4} for all
+function Collect {
+  Find $1 \
+    | awk "{if (\$5 == \"-\") {print \"${1}\"\$1\"-PHI\"\$2\"-W*\"} else {print \"${1}\"\$1\"-PHI\"\$2\"-W*-\"\$5} }" \
+    | xargs echo echo \
+    | bash \
+    | xargs ./collect.sh 
+}
+
+echo "GPU NO-VECTORIZATION VECTOR-WIDTH=2 VECTOR-WIDTH=4"
+Collect dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4}" | xargs echo TitanX
+Collect dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4 }" | xargs echo GTX980
+Collect dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4 }" | xargs echo K40c
+Collect dblp-K1024/TeslaK20m-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4 }" | xargs echo K20m

### 2017-05-03 plots/collect.sh

@@ -1,0 +1,35 @@
+#!/bin/bash
+F1=$$.tmp.1
+F2=$$.tmp.2
+
+grep -aH PHI "$@" | grep -v LOG \
+  | awk '{print $1 " " $7}' \
+  | sed 's/.*-\([^-]\+\)-PHI\([0-9]\+\)-W\([0-9]\+\)\(.*\)/\1 \2 \3 \4/' \
+  | sed 's/-/ /g' \
+  | sed 's/\(.*\):[^ ]\+\(.*\)/\1 \2/' \
+  | sed 's/false/0/g' \
+  | sed 's/true/1/g' \
+  | sed 's/\([^ ]\+\) \+\([^ ]\+\) \+\([^ ]\+\) \+\([^ ]\+\) \+\([^ ]\+\) \+\([^ ]\+\) \+\([^ ]\+\)/\1 \2 \3 \7 \4\5\6/' \
+  | tr -s "  " " " \
+  | sort -k 1 > $F1
+
+
+
+(echo 'obase=10;ibase=2;' && (cat $F1 | grep GEN | awk '{print $5}')) \
+  | bc \
+  | paste $F1 - \
+  | sed 's/\(NAIVE.*\)/\1 NAIVE -2/' \
+  | sed 's/\(SHARED.*\)/\1 SHARED -1/' \
+  | awk '{print $1 " " $2 " " $3 " " $4 " " $5 " " $6}' \
+  | sed 's/\([^ ]\+ [^ ]\+ [^ ]\+ [^ ]\+\) 0\([01][01]\) \([^ ]\+\)/\1 R\2 \3/' \
+  | sed 's/\([^ ]\+ [^ ]\+ [^ ]\+ [^ ]\+\) 1\([01][01]\) \([^ ]\+\)/\1 S\2 \3/' \
+  | sed 's/\([^ ]\+ [^ ]\+ [^ ]\+ [^ ]\+\) \([SR]\)0\([01]\) \([^ ]\+\)/\1 \2R\3 \4/' \
+  | sed 's/\([^ ]\+ [^ ]\+ [^ ]\+ [^ ]\+\) \([SR]\)1\([01]\) \([^ ]\+\)/\1 \2S\3 \4/' \
+  | sed 's/\([^ ]\+ [^ ]\+ [^ ]\+ [^ ]\+\) \([SR][SR]\)0 \([^ ]\+\)/\1 \2R \3/' \
[…]

### 2017-05-03 plots/plot-all-gpu-detail.sh

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+./plot-gpu-sweep.sh titanx-1024-strategies-w1 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
+./plot-gpu-sweep.sh titanx-1024-strategies-w2 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
+./plot-gpu-sweep.sh titanx-1024-strategies-w4 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
+
+./plot-gpu-sweep.sh gtx980-1024-strategies-w1 `ls dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
+./plot-gpu-sweep.sh gtx980-1024-strategies-w2 `ls dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
+./plot-gpu-sweep.sh gtx980-1024-strategies-w4 `ls dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
+
+./plot-gpu-sweep.sh k40-1024-strategies-w1 `ls dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
+./plot-gpu-sweep.sh k40-1024-strategies-w2 `ls dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
+./plot-gpu-sweep.sh k40-1024-strategies-w4 `ls dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
+
+./plot-gpu-sweep.sh k20-1024-strategies-w1 `ls dblp-K1024/TeslaK20m-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
+./plot-gpu-sweep.sh k20-1024-strategies-w2 `ls dblp-K1024/TeslaK20m-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
[…]

### 2017-05-03 plots/plot-cpu.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+cat <<EOF | gnuplot --persist
+
+set terminal postscript eps enhanced color font 'Helvetica,10' size 3.3,1.6
+set output 'cpu-vector.eps'
+
+set boxwidth 0.5
+set style fill solid
+
+set xlabel 'Vector Width'
+set ylabel 'Time (seconds)'
+
+set yrange[0:]
+
+set title 'CPU performance varying vector width'
+plot 'cpu-data' u 2:xtic(1) w boxes notitle
+
+EOF

### 2017-05-03 plots/plot-gpu-sweep.sh

@@ -1,0 +1,143 @@
+#!/bin/bash
+
+# splot '< ./collect.sh build-cu/log-phi-sweep-phi-*' u 1:3:4 w p
+
+NAME=$1
+shift
+
+ARGS="$@"
+
+TFILE="$$.tmp"
+
+./collect.sh $ARGS > $TFILE
+
+MIN=`cat $TFILE \
+  | awk 'NR == 1 {min = $4}
+         NR > 1 && $4 < min {min=$4}
+         END {print min}'`
+MIN=`perl -e "print $MIN - 1"`
+MAX=`perl -e "print $MIN * 2 + 1"`
+MID=128
+
+#cat $TFILE
+
+cat <<EOF | gnuplot --persist
+w = 0.6
+h = 0.4
+k = 0.05
+nr = 1
+nc = 2
+row(x) = ((nr-x-1)*h)+k
+col(x) = x*w
+
+set terminal postscript eps enhanced color font ',8'
+set size (nc*w),(nr*h+k)
+set output '$NAME.eps'
+set multiplot layout nr,nc
+
+#set logscale x 2
+#set logscale y
+set ytics
+set mytics
+set xtics
+set grid ytics xtics
+
+set xlabel "Block Size"
+set ylabel "Time (seconds)"
+
+set xrange [0:]
+set yrange [0:]
+#set key top left
+#set key below
+unset key
+
+set origin col(0),row(0)
+set size w,h
+set title '(a) Full sweep range'
+plot \
+  '< cat $TFILE | grep NAIVE' u 2:4 w lp title 'NAIVE', \
+  '< cat $TFILE | grep SHARED' u 2:4 w lp title 'SHARED', \
+  '< cat $TFILE | grep SSS' u 2:4 w lp title 'GEN-SSS', \
+  '< cat $TFILE | grep SSR' u 2:4 w lp title 'GEN-SSR', \
+  '< cat $TFILE | grep SRS' u 2:4 w lp title 'GEN-SRS', \
+  '< cat $TFILE | grep SRR' u 2:4 w lp title 'GEN-SRR', \
+  '< cat $TFILE | grep RSS' u 2:4 w lp title 'GEN-RSS', \
+  '< cat $TFILE | grep RSR' u 2:4 w lp title 'GEN-RSR', \
+  '< cat $TFILE | grep RRS' u 2:4 w lp title 'GEN-RRS', \
[…]

### 2017-05-03 plots/plot-gpus.sh

@@ -1,0 +1,37 @@
+#!/bin/bash
+
+TFILE=$$.tmp
+
+BASELINE=374.667
+
+
+echo > $TFILE
+
+X=1
+for i in GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m Intel-R-Xeon; do
+  v=`grep -a TOTAL dblp-K1024/${i}* | sort -nk 7 | head -n 1 | awk '{print $7}'`
+  v2=`perl -e "print $BASELINE / $v"`
+  echo $X $i $v2 >> $TFILE
+  X=`expr $X + 1`
+done
+
+cat $TFILE
+
+cat <<EOF | gnuplot --persist
+
+set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+set output 'gpus.eps'
+
+set boxwidth 0.5
+set style fill solid
+
+set ylabel 'Speedup over baseline'
+
+set yrange[0:]
+
+set title 'Performance of different compute devices'
+plot '$TFILE' u 1:3:xtic(2) w boxes notitle
+
+EOF
+
+rm $TFILE

### 2017-05-03 plots/plot-ppx.sh

@@ -1,0 +1,35 @@
+#!/bin/bash
+
+cat << EOF | gnuplot --persist
+w = 0.6
+h = 0.4
+k = 0.0
+nr = 1
+nc = 2
+row(x) = ((nr-x-1)*h)+k
+col(x) = x*w
+
+set terminal postscript eps enhanced color font ',8'
+set size (nc*w),(nr*h+k)
+set output 'ppx-gpu-cpu.eps'
+set multiplot layout nr,nc
+
+set origin col(0),row(0)
+set size w,h
+
+set xlabel 'Time (hours)'
+set ylabel 'Perplexity'
+
+set xrange [-0.15:6.15]
+set yrange [62:95]
+set title '(a) Com-DBLP K=4K m=4K n=32'
+plot "< perl ./get_time.perl ppx-DBLP-K4K-m4K-n32 " using (\$1/3600):3 notitle with lines
+
+set origin col(1),row(0)
+set size w,h
+
+set xrange [-0.3:18.3]
+set yrange [53:83]
+set title '(b) Com-LiveJournal K=3K m=4K n=32'
+plot "< perl ./get_time.perl ppx-LJ-K3K-m4K-n32 " using (\$1/3600):3 notitle with lines
+EOF

### 2017-05-03 plots/plot-vector-sweep.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+NAME=gpu-vector-sweep.eps
+shift
+
+TFILE="$$.tmp"
+
+./collect-v.sh > $TFILE
+
+cat <<EOF | gnuplot --persist
+set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+set output '$NAME'
+
+set style data histogram
+set style histogram cluster gap 1
+
+set xlabel 'GPU'
+set ylabel 'Time (normalized over non-vectorized version)'
+set key below
+
+set style fill solid noborder
+set auto x
+set ytics
+set mytics
+set grid ytics mytics
+
+set yrange [0:*]
+plot \
+  '$TFILE' u (\$2/\$2):xtic(1) title col, \
+  '$TFILE' u (\$3/\$2):xtic(1) title col, \
+  '$TFILE' u (\$4/\$2):xtic(1) title col
+
+EOF
+
+rm $TFILE
+

### 2017-05-04 02-background.tex
Start working on integration: block out double algorithm, unify symbol usage

--- 2017-05-03 distr-02-related.tex
@@
 \section{Background}
+\label{sec-background}
 % In this section, we briefly review the Bayesian model we are considering in this paper.
@@
     \STATE Sample mini-batch of vertices $\Neighbors$ from $\Vertices$
-    \STATE Update $\phi_{a}$ using Eq.~\ref{eqn:local_update}
+    \STATE \texttt{update\_phi:} Update $\phi_{a}$ using Eq.~\ref{eqn:local_update} \ENDFOR
   %  $\phi_{a}^* \law \left| \phi_{a} + \f{\ep_{\phi}}{2} \left( \al - \phi_{a} + \f{N}{|\Neighbors|} \sum_{b \in \Neighbors} G_{\phi_{a}}\right) + (\phi_{a})^{\ha} \xi_{a}\right|$
-    \STATE Obtain $\pi_a$ from $\phi_a^*$ \ENDFOR
+\FOR {each vertex in \Minibatch}
+    \STATE \texttt{update\_pi:} Obtain $\pi_a$ from $\phi_a^*$ \ENDFOR
 \FOR {$k = 1,\dots,K$}
-    \STATE Update $\ta_{k}$ using Eq.~\ref{eqn:global_update}
+    \STATE \texttt{update\_theta:} Update $\ta_{k}$ using Eq.~\ref{eqn:global_update} \ENDFOR
+\FOR {$k = 1,\dots,K$}
    % $\ta_{k}^* \law \left| \ta_{k} + \f{\ep_\ta}{2} \left( \eta - \ta_{k} + h(\cE_t) \sum_{(a,b) \in \cE_t} G_{\ta_{k}}(a,b) \right) + (\ta_{k})^{\ha}\xi_{k} \right|$
-    \STATE Obtain $\bt_k$ from $\ta_k^*$
-\ENDFOR
+    \STATE \texttt{update\_beta:} Obtain $\bt_k$ from $\ta_k^*$ \ENDFOR
 \ENDWHILE
@@
 \hline
+\\[-3ex]
 \end{tabular}
@@
 \end{comment}
-
-\subsection{Related work on parallel community detection}
-
-Among previous work on parallelizing MMSB algorithms, we mention the Online
-Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13} on GPUs, and our previous
-parallel implementation of SG-MCMC on GPUs and multi-core CPUs (submitted for
-publication). There are a few projects that did a distributed implementation
-for community detection on very large network graphs. In contrast to our work,
-they all detect non-overlapping communities: \cite{Bu2013246} investigate a
-large number of networks; \cite{2015arXiv150302115L} investigate hierarchical
-stochastic blockmodels; \cite{Prat-Perez:2014:HQS:2566486.2568010} use
-multi-core machines.
-
-\begin{comment}
-In this paper, we describe our custom RDMA D-KV (Distributed Key-Value)
-store. Current RMDA D-KV store implementations are RamCloud~\cite{RamCloud},
-Pilaf~\cite{Pilaf}, Herd~\cite{Herd} and FaRM~\cite{FaRM}. All these systems
-use RDMA to implement a D-KV store. However, all of them are far more powerful
-than our custom implementation -- and this power comes at a cost that we
-can avoid. They implement a generic D-KV store that controls concurrency,
-supports dynamic inserts and deletes, supports variable-sized values
-(whose size may change at an update), and keys of arbitrary type. Because
-of the nature of our distributed algorithm, we have to deal with none of
-these issues. For us, values are fixed-size, allocated only at the initial
-population, and remain alive forever. We have no concurrency between writes
-and reads or other writes. Our keys are a contiguous range of integers. All
-these properties together allow an extremely low-overhead implementation
-that does not involve the remote host in any transaction.
-\end{comment}

### 2017-05-04 02-related.tex
Start working on integration: block out double algorithm, unify symbol usage

--- 2016-01-20 02-related.tex
@@
-\section{Background}
-% In this section, we briefly review the Bayesian model we are considering in this paper.
+\section{Related Work}
 
-% In this section, we briefly review the Bayesian model we are considering in this paper.
-\subsection{Assortative Mixed-membership Stochastic Blockmodels (a-MMSB)}
-The assortative mixed membership stochastic blockmodel (a-MMSB)~\cite{gopalan2012scalable} is a special case of MMSB~\cite{airoldi2009mixed} that models the group structure in a network of $N$ vertices.
-Consider a set  $\cV$ containing all the vertices in the graph, and a set $\cE$ containing all the linked edges between pairs of vertices.
-Each vertex $a$ in the vertex set $\cV$ has a $K$-dimensional probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
-For every possible peer $b$ in the network, each vertex $a$ randomly draws a community $z_{ab}$.
-If a pair of vertices ($a,b$) in the edge set $\cE$ are in the same community, i.e., $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
-Otherwise this probability is small.
-Each community has its connection strength $\beta_{k} \in (0,1)$ which reflects how likely its members are linked to each other.
-The whole generative process of a-MMSB is then described by:
-\benum
-\item For each community $k$, draw community strength $\bt_k \sim \mbox{Beta}(\eta)$
-\item For each vertex $a$, draw community memberships $\pi_a \sim \mbox{Dirichlet}(\al)$
-\item For each pair of vertices $a$ and $b$,
-\benum
-	\item Draw interaction indicator $z_{ab} \sim \pi_a$
-    \item Draw interaction indicator $z_{ba} \sim \pi_b$
-    \item Draw link $y_{ab} \sim \mbox{Bernoulli}(r)$, where $r = \bt_k$ if $z_{ab}=z_{ba}=k$, and $r=\dt$ otherwise.
-\eenum
-\eenum
-$\eta$ and $\alpha$ are parameters to the Beta and Dirichlet distribution functions. $\dt$ (usually small) is a parameter for the algorithm.
+\subsection{Related work on Accelerators for Bayesian methods}
 
-\newcommand{\Vertices}{\ensuremath{\cV{}}\xspace}
-\newcommand{\Edges}{\ensuremath{\cE{}}\xspace}
-\newcommand{\Heldout}{\ensuremath{\cE_h{}}\xspace}
-\newcommand{\SGLDMinibatch}{\ensuremath{\cD_n{}}\xspace}
-\newcommand{\Minibatch}{\ensuremath{\cE_n{}}\xspace}
-\newcommand{\Minibatcht}{\ensuremath{\cE_n{}}\xspace}
-\newcommand{\Neighbors}{\ensuremath{\cV_n{}}\xspace}
+In this section, we discuss related machine learning projects that use
+accelerators.
+Immediately when GPUs evolved into generally programmable accelerators, machine
+learners recognized the potential of GPUs for their science.
+Much research
+has been devoted to deep-learning implementations on GPUs. However, Mixed
+Membership deduction (MMSB) is usually considered untractable for deep-learning
+algorithms. In contrast, our algorithm belongs to the family of Bayesian
+(approximative, stochastic) methods, where efficient algorithms for large-scale
+data have recently been developed. A broad survey of Bayesian methods is given
+in~\cite{DBLP:journals/corr/Zhu0H14}. Broadly speaking, Bayesian approximative
+algorithms are subdivided into variational and Monte Carlo methods; Markov
+Chain Monte Carlo (MCMC) is a subgroup of the latter. GPUs have been put to
+use in MCMC algorithms, e.g.\ \cite{journals/bioinformatics/MedlarGSBK13}
+uses GPUs to analyse parental linkage patterns in a biology context and
+\cite{DBLP:journals/csda/WhiteP14} does the same to model terrorist activity.
+Variational inference on MMSB is described
+in~\cite{DBLP:conf/nips/GopalanMGFB12}, but they don't use GPUs. Latent
+Dirichlet Allocation, another variety of Bayesian Approximation, is used on
+GPUs by~\cite{DBLP:conf/nips/YanXQ09}.
 
-\subsection{Stochastic Gradient Markov Chain Monte Carlo}
-The algorithm we are working on in this paper is based on stochastic gradient Langevin dynamics (SGLD)~\cite{welling2011bayesian}.
-SGLD applies the following update rule to obtain samples from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ of $N$ i.i.d.
-data points $\cX=\{x_i\}_{i=1}^{N}$:
-\bea
-\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\SGLDMinibatch)\right) + \xi_t, \label{eqn:sgld_update}
-\eea
-where $\xi_t \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\SGLDMinibatch$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\SGLDMinibatch)$ the mean stochastic gradient, i.e., $\frac{1}{|\SGLDMinibatch|}\sum_{x\in \SGLDMinibatch}^{}\grad_{\ta} \log p(x|\ta)$.
-As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution.
-One benefit of using SGLD is that we do not need the Metropolis-Hastings (MH) accept-reject tests since the rejection probability goes to zero as the step size tends to zero.
-Although in practice we often do not reduce the stepsize to zero to obtain better mixing (thus resulting in some bias), we obtain good performance by drawing many more samples per unit time.
+Still, for high-dimension problems, the standard MCMC algorithms fall short in
+performance. Recently, varieties of MCMC algorithms have been developed where
+gradient information is used to speed up convergence. Langevin and Hamiltonian
+dynamics are representatives of these
+varieties~\cite{Girolami_riemannmanifold}.  Our algorithm uses Riemann Manifold
+Langevin dynamics. In \cite{beam2014fast}, GPUs are used to perform Hamiltonian
+descent, using Python interfaces to access the standard cuBLAS
+library~\cite{cuBLAS}. They limit their optimizations for the GPU to
+saving on data transfers between host and device memory. They do not apply
+their framework to MMSB.
 
-SGLD originated from the Langevin Monte Carlo (LMC)~\cite{girolami2011riemann}, where unlike SGLD the gradient is computed exactly by using all data points.
-Then, Metropolis-Hastings accept-reject tests are applied.
-Comparing to LMC, SGLD only requires to process a mini-batch $\SGLDMinibatch$ at each iteration and ignores the MH test, and thus the computation complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
-
-Stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex.
-By applying Riemannian geometry~\cite{girolami2011riemann} and using the mini-batch-based estimator in Eqn.
-\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA)~\cite{blei2003latent}.
-In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$.
-Then, the update rule of SGRLD is:
-\bea
-\ta_{k}^* \law \left| \ta_{k} + \f{\ep_t}{2} \left( \al - \ta_{k} + \f{N}{|\SGLDMinibatch|} \sum_{d \in \SGLDMinibatch} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{t} \right|, \label{eqn:sgrld_update}
-\eea
-here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \SGLDMinibatch$.
-
-\subsection{Scalable MCMC for a-MMSB}
-This section describes the SG-MCMC algorithm for a-MMSB.
-Please refer \cite{LiAW15} for more details.
-The algorithm iterates updating local parameter $\pi$ and global parameter $\beta$.
-Since both parameters lie on the probability simplex, SGRLD is applied to make the sampling process more efficient.
-Here, the parameters $\phi$ and $\ta$ are used to re-parameterize $\pi$ and $\beta$ respectively.
-After updating $\phi$ and $\ta$, we can obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$, respectively.
-In the following, we briefly sketch the iterative update steps.
-
-\paragraph{\textbf{Sampling global parameters}} The update rule for global parameter $\theta$ is
-\bea
-\ta_{ki}^* \law \left| \ta_{ki} + \f{\ep}{2} \left( \eta - \ta_{ki} + h(\Minibatcht)\sum_{(a,b) \in \Minibatcht}g_{ab}(\ta_{ki})\right) \right.\nn \\ \left.+ (\ta_{ki})^{\ha}\xi_{ki} \right|, \label{eqn:global_update}
-\eea
-where the gradient in $\theta$
-\bea
-g_{ab}(\ta_{ki})
-&=& \f{f_{ab}^{(y)}(k,k)}{Z_{ab}^{(y)}} \left(\f{|1-i-y|}{\ta_{ki}} - \f{1}{\sum_j\ta_{kj}} \right),
-\eea
-here $\Minibatcht$ is a mini-batch of $n_t$ vertex pairs sampled from $\Edges$; $h(\Minibatcht)$ is a weight factor to scale the effect of a mini-batch towards the full network; and 
-\bea
-f_{ab}^{(y)}(k,l)=
-\begin{cases}
-\bt_k^y(1-\bt_k)^{(1-y)}\pi_{ak}\pi_{bk}, & \mbox{if } k=l\\
-\dt^{y}(1-\dt)^{(1-y)} \pi_{ak} \pi_{bl}, & \mbox{if } k \neq l \nn.
-\end{cases}
-\label{eqn:case}
-\eea
-$Z_{ab}^{(y)}$ is the normalization constant which we can compute in $\cO(K)$ time \cite{LiAW15}.
-
-\paragraph{\textbf{Sampling local parameters}}
-The update rule for the local parameters $\phi$ is
-\bea
-\phi_{ak}^* \law \left| \phi_{ak} + \f{\ep}{2} \left( \al - \phi_{ak} + \f{N}{|\Neighbors|} \sum_{b \in \Neighbors} g_{ab}(\phi_{ak})\right) \right. \nn \\ \left. + (\phi_{ak})^{\ha} \xi_{ak}\right|,
-\label{eqn:local_update}
-\eea
-where the gradient in $\phi$
-\bea
-g_{ab}(\phi_{ak}) = \f{f_{ab}^{(y)}(k)}{Z_{ab}^{(y)}\phi_{ak}} - \f{1}{\sum_j\phi_{aj}}.
-\eea
-Here, $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV$. Note that $|\Neighbors| \ll |\cV|=N$.
-
-As is common in machine learning problems, the edges of the graph are divided
-into two sets, the training set and a held-out set \Heldout.
-As the performance metric we use perplexity, which is defined as the exponential
-of the negative average log-likelihood of the held-out set \Heldout. Given a collection of $T$
-samples of the model parameters $\{\bt_t\}$ and $\{\pi_t\}$, the averaged
-perplexity on the held-out test set $\Heldout$ is 
-\begin{align}
-&\text{perp}_{\text{avg}}(\Heldout|\{\beta_t\},\{\pi_t\})\nonumber \\=&
-\exp\left(-\frac{\sum_{(a,b)\in \Heldout}^{}\log
-\{(1/T)\sm{t}{T}p(y_{ab}|\beta_t,\pi_t)\}}{|\cE_{h}|}\right)
-\end{align}
-The perplexity is not evaluated at every iteration, but at regular intervals.
-
-The pseudo-code of the sequential algorithm is presented in Algorithm~\ref{algorithm}.
-
-\begin{algorithm}[t]
-\caption{Sequential version of SG-MCMC for a-MMSB}\label{alg}
-\begin{algorithmic}[1]
-\STATE Initialize $\pi , \bt, \phi, \ta$
-\WHILE {sampling}
-\STATE Sample a mini-batch of vertex pairs, \Minibatch, from $\cE$
-\FOR {each vertex in \Minibatch}
-	\STATE Sample a mini-batch of vertices, $\Neighbors$, from $\Vertices$
-    \STATE Update $\phi_{a}$ using Eqn~\ref{eqn:local_update}
-  %  $\phi_{a}^* \law \left| \phi_{a} + \f{\ep_{\phi}}{2} \left( \al - \phi_{a} + \f{N}{|\Neighbors|} \sum_{b \in \Neighbors} G_{\phi_{a}}\right) + (\phi_{a})^{\ha} \xi_{a}\right|$
-    \STATE Obtain $\pi_a$ from $\phi_a^*$ \ENDFOR
-\FOR {$k = 1,\dots,K$}
-    \STATE Update $\ta_{k}$ using Eqn~\ref{eqn:global_update}
-   % $\ta_{k}^* \law \left| \ta_{k} + \f{\ep_\ta}{2} \left( \eta - \ta_{k} + h(\cE_t) \sum_{(a,b) \in \cE_t} G_{\ta_{k}}(a,b) \right) + (\ta_{k})^{\ha}\xi_{k} \right|$
-    \STATE Obtain $\bt_k$ from $\ta_k^*$
-\ENDFOR
-\ENDWHILE
-\end{algorithmic}
-\label{algorithm}
-\end{algorithm}
-
-\addtolength{\tabcolsep}{-3pt}
-\renewcommand{\arraystretch}{1.5}
-\begin{table}[b] % [htbp]
-\center
-\begin{tabular}{c C{0.075\textwidth} c L{0.26\textwidth}}
-\textbf{symbol}
-	 & \textbf{type} & \textbf{size} & \textbf{description} \\
-         % & \multicolumn{1}{c}{\textbf{type}}
-                           % & \multicolumn{1}{c}{\textbf{description}} \\
-\hline
-$\cK$      &              & $K$        & set of communities \\
-\Vertices  & \{vertex\}   & $N$          & vertices in the graph \\
-\Edges     & \{edge\}     &              & linked edges in the graph \\
-$E$        & \{edge\}     &              & $\Vertices\times\Vertices$: linked and nonlinked edges \\
-\Heldout   & \{edge\}     &              & held-out subset of the graph \\
-\Minibatch & \{edge\}     &              & sampled mini-batch of edges in~$E$ \\
-$M$        &              &              & number of vertices in \Minibatch \\
-\Neighbors & \{vertex\}   &              & sampled neighbor set for a vertex in~\Minibatch \\
-$\theta$   & float~vector 2-D & $K\times{}2$ & global latent variables for community strength \\
-$\beta$    & float vector & $K$          & $\beta[k] = \theta[k][0] / \sum_j\theta[k][j]$ is the community strength\\
-$\phi$     & float~vector 2-D & $N\times{}K$ & local latent variables; $\phi[i][k]$ reflects
-                            probability that vertex $i$ is in community~$k$ \\
-$\pi$      & float~vector 2-D & $N\times{}K$ &
-					$\pi[i][k]$~is~probability that $i$ is in $k$
-					$\pi[i][k] = \phi[i][k] / \sum_j\phi[i][j]$ \\
-\hline
-\end{tabular}
-\caption{Definition of most important symbols}
-\label{tbl-symbols}
-\end{table}
-\addtolength{\tabcolsep}{3pt}
-\renewcommand{\arraystretch}{1.0}
+A different MMSB algorithm with stochastic gradient descent on the GPU is
+the Online Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13}. They deduce
+overlapping communities for many of the same datasets as this paper, and they
+report fast convergence. Their implementation uses the cuBLAS
+library, and, unlike our work, there is no attempt to hand-optimize the GPU
+kernels. Additionally, the algorithmic approach to solve the problem is
+fundamentally different.
+Since they target GPUs only, the datasets they
+can handle are limited by the device memory of the GPU. Our implementation converges
+quickly on the GPU but can also function with reduced speedup on a multicore
+CPU. In the latter case, the dataset
+size is only limited by the memory of the host machine.
 
 \begin{comment}
-Assortative mixed-membership stochastic blockmodel (a-MMSB) [1] is a special case of MMSB [8] that models the group-structure in a network of $N$ nodes.
-In particular, each node $a$ in the node set $\cV$ has a $K$-dimension probability distribution $\pi_a$ of participating in the $K$ members of the community set $\cK$.
-For every possible peer $b$ in the network, each node $a$ randomly draws a community $z_{ab}$.
-If a pair of nodes ($a,b$) in the edge set $\cE$ are in the same community: $z_{ab}=z_{ba} = k$, then they have a significant probability $\bt_k$ to connect, i.e., $y_{ab}=1$.
-Otherwise this probability is small.
-Each community has its connection strength $\beta_{k} \in (0,1)$ which explains how likely its members are linked to each other.
+Since our SG-MCMC algorithm is completely different from deep-learning
+algorithms, it makes no real sense to compare to the learning frameworks
+Theano, Caffe, cuDNN, or NVidia Digits, even though they target GPUs.
 
-The generative process of a-MMSB is then given by,
+SG-MCMC work NOT on GPUs:
 
-\benum
-\item For each community $k$, draw community strength $\bt_k \sim \mbox{Beta}(\eta)$
-\item For each node $a$, draw community memberships $\pi_a \sim \mbox{Dirichlet}(\al)$
-\item For each pair of nodes $a$ and $b$,
-\benum
-	\item Draw interaction indicator $z_{ab} \sim \pi_a$
-    \item Draw interaction indicator $z_{ba} \sim \pi_b$
-    \item Draw link $y_{ab} \sim \mbox{Bernoulli}(r)$, where $r = \bt_k$ if $z_{ab}=z_{ba}=k$, and $r=\dt$ otherwise.
-\eenum
-\eenum
+GPU work:
+MCMC Hamiltonian:
+	1. Andrew L. Beam, Sujit K. Ghosh, Jon Doyle
+		Fast Hamiltonian Monte Carlo Using GPU Computing
+		http://arxiv.org/pdf/1402.4089.pdf
+MMSB tensor:
 
-Unlike the a-MMSB, the original MMSB maintains pair-wise community strength $\bt_{k,k'}$ for all pairs of the communities.
-Note that it is trivial to extend the results that we obtain in this paper to the general MMSB model.
-The joint probability of the above process can be written as:
-\begin{align}
-p(y,z,\pi ,\bt | \al, \eta) &= \prod_{a=1}^{N} \prod_{b > a}^{N} p(y_{ab} | z_{ab}, z_{ba}, \bt) p(z_{ab}|\pi_a)\nonumber \\ &p(z_{ba}|\pi_b) \pd{a}{N}p(\pi_a|\al) \pd{k}{K} p(\bt_k|\eta).\label{eqn:joint}
-\end{align}
-
-Both variational inference [1,4,16] and collapsed Gibbs sampling algorithms [11] have been used successfully for small to medium scale problems.
-However, the $\cO(N^2)$ computational complexity per update prevents it from being applied to large scale networks.
-A stochastic variational algorithm was developed in [1] to address this issue, where each update only depends on a small mini-batch of the nodes in the network.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Stochastic Gradient MCMC Algorithms}
-Our algorithm will be based on the stochastic gradient Langevin dynamics (SGLD) [3].
-To sample from a posterior distribution $p(\ta|\cX) \propto p(\cX|\ta) p(\ta)$ given $N$ i.i.d.
-data points $\cX=\{x_i\}_{i=1}^{N}$, SGLD applies the following update rule:
-\bea
-\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\SGLDMinibatch)\right) + \xi, \label{eqn:sgld_update}
-\eea
-where $\xi \sim \cN(0,\epsilon_t )$ with $\ep_t$ the step size, $\SGLDMinibatch$ a mini-batch of size $n$ sampled from $\cX$, and $\bar{g}(\ta;\SGLDMinibatch) = \frac{1}{|\SGLDMinibatch|}\sum_{x\in \SGLDMinibatch}^{}\grad_{\ta} \log p(x|\ta)$.
-As the step size goes to zero by a schedule satisfying $\sm{t}{\infty}\ep_t = \infty$ and $\sm{t}{\infty}\ep_t^2 < \infty$, SGLD samples from the true posterior distribution.
-In SGLD, the Metropolis-Hastings (MH) accept-reject tests are ignored since the rejection probability goes to zero as the step size collapses to zero.
-While for a finite step size this results in some bias, the overall error is reduced by the reduction of variance due to the ability to draw many more samples per unit time.
-
-SGLD originated from the Langevin Monte Carlo (LMC) [15] where, unlike SGLD, the gradient is computed exactly using all data points and then a Metropolis-Hastings accept-reject test is applied.
-Because at each iteration SGLD requires to process only a mini-batch $\SGLDMinibatch$ and ignores the MH test, the computation complexity per iteration is only $\cO(n)$ as opposed to $\cO(N)$ of LMC.
-Any mini-batch sampling algorithm in the form of Eqn.
-\eqref{eqn:sgld_update} is called valid SGLD as long as it guarantees the gradient estimator to be unbiased, i.e.,
-$\eE_{\SGLDMinibatch} \left[N\bar{g}(\ta;\SGLDMinibatch)\right] = \grad_{\ta} \log p(\cX|\ta)\label{eqn:sgld_grad_estm}$ and the variance to be finite [5].
-
-%%%%%%%%%%%%%%
-%\subsection{Stochastic Gradient Riemannian Langevin Dynamics (SGRLD)}
-
-The stochastic gradient Riemannian Langevin dynamics (SGRLD) [2] is a subclass of SGLD which is developed to sample from the probability simplex.
-By applying Riemannian geometry [15] and using the mini-batch estimator in Eqn.
-\ref{eqn:sgld_update}, it achieved state-of-the-art performance for latent Dirichlet allocation (LDA).
-In particular, for a $K$-dimensional probability simplex $\pi$, it uses the \textit{expanded-mean} re-parameterization trick, where the probability of a category $k$ is given by $\pi_k = \ta_k / \sum_{j=1}^K \ta_j$ with $\ta_k \sim \text{Gamma}(\al,1)$ and $\al$ a hyperparameter of the Dirichlet distribution $p(\pi | \al )$.
-Then, the update rule becomes
-
-%==============================================================
-
-\bea
-\ta_{k}^* \law \left| \ta_{k} + \f{\ep}{2} \left( \al - \ta_{k} + \f{N}{|\SGLDMinibatch|} \sum_{d \in \SGLDMinibatch} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{} \right|. \label{eqn:sgrld_update}
-\eea
-here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \SGLDMinibatch$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Scalable MCMC for a-MMSB}
-Our algorithm iterates updating local parameters $\pi$ and a global parameter $\beta$.
-Because both parameters lie on the probability simplex, we start from the SGRLD and modify it to be more efficient.
-Also, we introduce parameters $\phi$ and $\ta$ to re-parameterize $\pi$ and $\beta$ respectively.
-Thus, we alternatingly sample in the $\phi$ and $\ta$ spaces, and then obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$.
-From Eqn.
-\ref{eqn:joint}, summing over the latent variable $z$, we obtain the following joint probability,
-\bea
-&p(y, \pi , \bt | \al , \eta) = \prod_{a} p(\pi_a | \al ) + \prod_{a} p(\bt_k | \eta)\nn \\
-&+ \prod_{a} \prod_{b > a}\sum_{z_{ab}, z_{ba}} p(y_{ab}, z_{ab}, z_{ba} | \bt , \pi_a, \pi_b)
-\label{eqn:log_joint}.
-\eea
-
-%==============================================================
-\subsection{Sampling the global parameter}
-
-By the re-parameterization, we have $\beta_k=\ta_{k1}/(\ta_{k0}+\ta_{k1})$, where $\ta_{ki}\sim$ Gamma$(\eta)\propto \ta_{ki}^{\eta-1}e^{-\ta_{ki}}$.
-Because $p(y, \pi , \bt | \al , \eta)$ decomposes into $p(y,\bt|\pi,\eta) p(\pi|\al)$, replacing $\bt$ by $\ta$, we compute the derivative of log of Eqn.
-\ref{eqn:log_joint} w.r.t.
-$\ta_{ki}$ for $i=\{0,1\}$ as follows:
-\bea
-\fp{\ln p(y ,\ta | \pi, \eta )}{\ta_{ki}} =
-\fp{}{\ta_{ki}} \ln p(\ta_{ki} | \eta) + \sum_{a}\sum_{b>a} g_{ab}(\ta_{ki}),
-\label{eqn:grad_bt}
-\eea
-where $g_{ab}(\ta_{ki}) = \fp{}{\ta_{ki}} \ln \sum_{z_{ab},z_{ba}} p(y_{ab},z_{ab},z_{ba} | \ta , \pi_a, \pi_b)$ which, similar to SGRLD for LDA [2], we can rewrite as
-\bea
-g_{ab}(\ta_{ki}) = \eE \left[ \eI[z_{ab}=z_{ba}=k] \left( \f{|1-i-y_{ab}|}{\ta_{ki}} - \f{1}{\ta_{k}} \right)\right]. \label{eqn:G}
-\eea
-where $\ta_{k}=\sum_i \ta_{ki}$ and $\eI[S]$ is equal to $1$ if a condition $S$ is TRUE and $0$ otherwise.
-The expectation is w.r.t.
-the posterior distribution of latent variables $z_{ab}$ and $z_{ba}$,
-\bea
-&&p(z_{ab}=k,z_{ba}=l | y_{ab}, \pi_a, \pi_b, \bt )\\ &&\propto f_{ab}^{(y)}(k,l)=
-\begin{cases}
-\bt_k^y(1-\bt_k)^{(1-y)}\pi_{ak}\pi_{bk}, & \mbox{if } k=l\\
-\dt^{y}(1-\dt)^{(1-y)} \pi_{ak} \pi_{bl}, & \mbox{if } k \neq l \nn
-\end{cases}
-\label{eqn:case}
-\eea
-here we used simple notation $y$ instead of $y_{ab}$.
-Unlike the SGRLD for LDA [2], we compute the expectation in Eqn.
-\eqref{eqn:G} analytically by computing the normalization constant $Z_{ab}^{(y)} = \sm{k}{K}\sm{l}{K} f_{ab}^{(y)}(k,l)$ which can be reduced to $\cO(K)$ computation as follows
-\begin{align}
-&Z_{ab}^{(y)} = \dt^{y}(1-\dt)^{(1-y)} \nonumber\\&+\sm{k}{K} \left( \bt_k^{y}(1-\bt_k)^{(1-y)} - \dt^{y}(1-\dt)^{(1-y)}\right)\pi_{ak}\pi_{bk}
-\end{align}
-
-Then Eqn. \eqref{eqn:G} becomes
-\bea
-g_{ab}(\ta_{ki}) 
-&=& \f{f_{ab}^{(y)}(k,k)}{Z_{ab}^{(y)}} \left(\f{|1-i-y|}{\ta_{ki}} - \f{1}{\ta_k} \right).
-\eea
-Plugging this into Eqn. \ref{eqn:sgrld_update}, we obtain the update rule for the global parameter,
-
-\bea
-\ta_{ki}^* \law \left| \ta_{ki} + \f{\ep}{2} \left\{ \eta - \ta_{ki} + h(\Minibatcht)\sum_{(a,b) \in \Minibatcht}g_{ab}(\ta_{ki})\right\} \right.\nn \\ \left.+ (\ta_{ki})^{\ha}\xi_{ki} \right|, \label{eqn:global_update}
-\eea
-here $\Minibatcht$ is a mini-batch of $n_t$ node pairs sampled from $\cE$ for which we use the following strategy.
-
-\textbf{Stratified sampling:} considering that the number of links is much smaller than that of non-links, we can reduce the variance of the gradient using stratified sampling, similar to the method used in [1].
-For this, at every iteration we first randomly select a node $a$ and then toss a coin with probability 0.5 to decide whether to sample link edges or non-link edges for node $a$.
-If it is a link, we assign all of the link edges of node $a$ to $\cE_n$.
-Otherwise, i.e.
-if it is non-link, we uniformly sample a mini-batch of $N/m$ non-link edges from the entire set of non-link edges and assign it to $\cE_n$.
-Here, the $m$ is a hyper-parameter.
-Note that the size of $|\cE_n|$ will thus be much smaller than the total number of $N(N-1)/2$ edges when $m$ is reasonably large.
-Then, to ensure that the gradient is unbiased, a \textit{scaling parameter} $h(\cE_n)$ is multiplied.
-Specifically, $h(\cE_n)$ is set to $N$ when $\cE_n$ is a set of link edges and to $mN$ otherwise.
-
-Because the global parameters $\{\beta_k\}$ does not change very fast compared to the local parameters $\{\pi_{ak}\}$, in practice we update only a random subset of the $\{\beta_k\}$ at each iteration.
-
-\subsection{Sampling the local parameters}
-
-Similar to the global parameter, we re-parameterize the local parameter $\pi_a$ such that $\pi_{ak} = \phi_{ak} / \sum_{j=1}^{K}\phi_{aj}$, with $\phi_{ak}\sim$ Gamma$(\alpha)\propto \phi_{ak}^{\alpha-1}e^{-\phi_{ak}}$.
-Then, taking the derivative of the log of Eqn.
-\ref{eqn:log_joint} w.r.t.
-$\phi_{ak}$, we obtain
-\bea
-\fp{\ln p(y , \phi | \bt, \al)}{\phi_{ak}} = \fp{}{\phi_{ak}} \ln p(\phi_{ak} | \al) +
-\sum_{b} g_{ab}(\phi_{ak})
-\eea
-where $g_{ab}(\phi_{ak}) = \fp{}{\phi_{ak}} \ln \sum_{z_{ab}, z_{ba}} p(y_{ab}, z_{ab}, z_{ba} | \bt , \phi_a, \phi_b)$ which can be written as 
-\bea 
-g_{ab}(\phi_{ak}) = \eE\left[ \f{\eI [z_{ab} = k]}{ \phi_{ak}} - \f{1}{\phi_{a\cdot}} \right].
-\label{eqn:grad_local}
-\eea
-Here the expectation is w.r.t.
-the distribution in Eqn.
-\eqref{eqn:case}.
-To compute the expectation analytically, we first integrate out $z_{ba}$ from Eqn.
-\eqref{eqn:case} because the expectation depends only on $z_{ab}$, and obtain the following probability up to a normalization constant
-
-\begin{align}
-&f_{ab}^{(y)}(k) = \sum_{l=1}^K f_{ab}^{(y)}(k,l)\nonumber \\ =& \pi_{ak}
-\left\{ \bt_k^{y}(1-\bt_k)^{(1-y)} \pi_{bk} + \dt^{y}(1-\dt)^{(1-y)} (1- \pi_{bk}) \right\}.\label{eqn:local_f}
-\end{align}
-
-Then we obtain the normalization term by $Z_{ab}^{(y)} = \sm{k}{K}f_{ab}^{(y)}(k)$. 
-Integrating out the expectation in Eqn. \eqref{eqn:grad_local}, we obtain 
-\bea
-g_{ab}(\phi_{ak}) = \f{f_{ab}^{(y)}(k)}{Z_{ab}^{(y)}\phi_{ak}} - \f{1}{\phi_a}.
-\eea 
-Plugging this to Eqn. \ref{eqn:sgrld_update}, we obtain the SGRLD update rule for the local parameter $\phi_{ak}$
-\bea
-\phi_{ak}^* \law \left| \phi_{ak} + \f{\ep}{2} \left( \al - \phi_{ak} + \f{N}{|\Neighbors|} \sum_{b \in \Neighbors} g_{ab}(\phi_{ak})\right) \right.
-\nn \\ \left.
-+ (\phi_{ak})^{\ha} \xi_{ak}\right|.\label{eqn:local_update}
-\eea
-Here, the $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from $\cV$. Note that $|\Neighbors| \ll |\cV|=N$.
-
+MCMC (not SG):
+    1. Alan Medlar, Dorota Głowacka, Horia Stanescu, Kevin Bryson, Robert Kleta
+       SwiftLink: Parallel MCMC linkage analysis utilising multicore CPU and GPU
+       http://bioinformatics.oxfordjournals.org/content/early/2012/12/13/bioinformatics.bts704.full.pdf
+    2. Marc Suchard, Chris Holmes, Mike West
+       Some of the What?, Why?, How?, Who?  and Where?  of Graphics Processing Unit Computing for Bayesian Analysis
+       https://stat.duke.edu/gpustatsci/GPU-ISBABull2010.pdf
 \end{comment}
-
-\subsection{Related work on parallel community detection}
-
-Among previous work on parallelizing MMSB algorithms, we mention the Online
-Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13} on GPUs, and our previous
-parallel implementation of SG-MCMC on GPUs and multi-core CPUs (submitted for
-publication). There are a few projects that did a distributed implementation
-for community detection on very large network graphs. In contrast to our work,
-they all detect non-overlapping communities: \cite{Bu2013246} investigate a
-large number of networks; \cite{2015arXiv150302115L} investigate hierarchical
-stochastic blockmodels; \cite{Prat-Perez:2014:HQS:2566486.2568010} use
-multi-core machines.
 
@@
 \end{comment}
+
+\subsection{Related work on parallel community detection}
+
+Among previous work on parallelizing MMSB algorithms, we mention the Online
+Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13} on GPUs.
+There are a few projects that did a distributed implementation
+for community detection on very large network graphs. In contrast to our work,
+they all detect non-overlapping communities: \cite{Bu2013246} investigate a
+large number of networks; \cite{2015arXiv150302115L} investigate hierarchical
+stochastic blockmodels; \cite{Prat-Perez:2014:HQS:2566486.2568010} use
+multi-core machines.
+
+\begin{comment}
+\subsection{Related work on RDMA Key-Value stores}
+In this paper, we describe our custom RDMA D-KV (Distributed Key-Value)
+store. Current RMDA D-KV store implementations are RamCloud~\cite{RamCloud},
+Pilaf~\cite{Pilaf}, Herd~\cite{Herd} and FaRM~\cite{FaRM}. All these systems
+use RDMA to implement a D-KV store. However, all of them are far more powerful
+than our custom implementation -- and this power comes at a cost that we
+can avoid. They implement a generic D-KV store that controls concurrency,
+supports dynamic inserts and deletes, supports variable-sized values
+(whose size may change at an update), and keys of arbitrary type. Because
+of the nature of our distributed algorithm, we have to deal with none of
+these issues. For us, values are fixed-size, allocated only at the initial
+population, and remain alive forever. We have no concurrency between writes
+and reads or other writes. Our keys are a contiguous range of integers. All
+these properties together allow an extremely low-overhead implementation
+that does not involve the remote host in any transaction.
+\end{comment}

### 2017-05-04 distr-03-design.tex
Start working on integration: block out double algorithm, unify symbol usage

--- 2016-03-07 03-design.tex
@@
-\section{System Design}
+\subsection{System Design}
 \label{design-section}
@@
 
-This section describes how we parallelized the algorithm, which had been
-implemented sequentially in~C++.
-The usage of multi-threaded parallelism
-is straightforward, therefore, we will discuss its details only where appropriate. The
+This section describes how we created a distributed version of the C++
+algorithm. We didn't use the advanced intra-node parallelisation techniques
+described in the previous section, but found that straightforward
+multi-threaded parallelization techniques sufficed: the distributed performance
+is dominated by (remote) memory latency, not by node parallelism.
+The details of the multi-threaded parallelism
+will be discussed only where appropriate. The
 distributed design follows a master-worker paradigm, where
@@
 
-% horrible hack to move the table up
-\begin{table*}
-  \centering
-  \begin{tabular}{l r r r l}
-    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
-                    &            &               & communities    &             \\
-    \hline
-    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
-    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
-    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
-    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
-    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
-    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
-    \hline
-  \end{tabular}
-  \caption{Summary of SNAP graph data sets used for evaluation.}
-  \label{tab-datasets}
-\end{table*}
-
 % each of the stages

### 2017-05-04 distr-04-evaluation.tex
Start working on integration: block out double algorithm, unify symbol usage

--- 2016-03-07 04-evaluation.tex
@@
-\section{Evaluation}
+\subsection{Evaluation}
 
@@
 publicly available graphs from the Stanford Large Network Dataset
-Collection~(SNAP) \cite{snapnets}. Table~\ref{tab-datasets} lists the
+Collection~(SNAP) \cite{snapnets}. Table~\ref{table-snap} lists the
 collection of data sets used.
@@
 %################################################
-\subsection{Weak Scaling}
+\subsubsection{Weak Scaling}
 
@@
 %################################################
-\subsection{Strong Scaling}
+\subsubsection{Strong Scaling}
 
@@
 %################################################
-\subsection{Pipeline efficiency}
+\subsubsection{Pipeline efficiency}
 \label{eval-pipeline}
@@
     \hline
+    \\[0.2ex]
   \end{tabular}
@@
 %################################################
-\subsection{Horizontal vs. Vertical Scalability}
+\subsubsection{Horizontal vs. Vertical Scalability}
 \begin{figure}[t] % [htb]
@@
 %################################################
-\subsection{Cluster Communication Efficiency}
+\subsubsection{Cluster Communication Efficiency}
 
@@
 %################################################
-\subsection{Convergence of Large Datasets}
+\subsubsection{Convergence of Large Datasets}
 
@@
 convergence time to assess the system's utility. Figure~\ref{fig-ppx} presents
-the convergence time of the data sets listed in Table~\ref{tab-datasets}.
+the convergence time of the data sets listed in Table~\ref{table-snap}.
 %

### 2017-05-04 gpu-04-02-sequential.tex
Start working on integration: block out double algorithm, unify symbol usage

@@
 
-The initial implementation of the algorithm was developed by its authors in
-Python as is common with machine learning researchers and practitioners. It
+Initially, the algorithm was implemented in Python~\cite{LiAW15}, as is common
+with machine learning researchers and practitioners. It
 relied on numpy~\cite{numpy} to perform the numerical computations without the
@@
 
-Initially, the Python code base was ported into equivalent C++ code. Namely,
-the program structure remained unchanged and the Python data structures were
+Initially, the Python code base was ported into equivalent C++ code.
+The program structure remained unchanged and the Python data structures were
 simply substituted with their corresponding alternatives in C++.
@@
 was an expression of the form $a^y b^{1-y}$, implemented with floating-point
-exponentiation. A study of the algorithm showed that $y$ is a variable that can
+exponentiation, where $y$ is a variable that can
 only hold values of 0 or~1. Therefore, the expression is simply a choice

### 2017-05-04 gpu-04-03-accelerator.tex
Start working on integration: block out double algorithm, unify symbol usage

@@
 Second, the predominant kernel, \textit{update\_phi} is invoked to calculate the
-gradients for each vertex $i$ in the mini-batch $m$ and updates the value of $ \phi_i
-\mid i \in m$. Next, the \textit{update\_pi} kernel is invoked to normalize the
+gradients for each vertex $i$ in the mini-batch \Minibatch and updates the value of $ \phi_i
+\mid i \in \Minibatch$. Next, the \textit{update\_pi} kernel is invoked to normalize the
 individual $\phi_i$ and store the result in the corresponding~$\pi_i$.
@@
 matrices is redundant as $\pi$ is a row-normalized copy of~$\phi$, see
-Equation~\ref{eqn-pi} in Section~\ref{sec-algorithm}. The full storage for
+Table~\ref{tbl-symbols} in Section~\ref{sec-background}. The full storage for
 $\phi$ is discarded; $\phi_{ik}$ values are calculated as
@@
 \textit{update\_pi}. Both kernels access $\phi_i$ only for vertices~$i$ in the
-mini-batch~$m$, so for each iteration, the calculated values for $\phi$ are cached
+mini-batch~\Minibatch, so for each iteration, the calculated values for $\phi$ are cached
 in a smaller temporary matrix of size
-${|m|}\times{K}$.
+$M\times{K}$.
 %
@@
 reduced to a vector $\phi_{sum}$ of length~$N$ and a comparatively small
-${|m|}\times{K}$ matrix. For sufficiently large~$K$, this transformation roughly
+$M\times{K}$ matrix. For sufficiently large~$K$, this transformation roughly
 halves the memory footprint of the algorithm.
@@
 reflect the changes that were accumulated in $Grads_i$ for each vertex $i$ in the
-mini-batch~$m$.
+mini-batch~\Minibatch.
 
@@
 %
-The read-only access of $\pi_j$ of the randomly generated neighbors are unique
+The read-only accesses of $\pi_j$ of the randomly generated neighbors are unique
 with a high probability. More precisely, each $vertex_i$ in the mini-batch
@@
 % Naive strategy
-\subsubsection*{The naive strategy} simply allocates temporary vectors in thread
+\paragraph*{\textbf{The naive strategy}} simply allocate temporary vectors in thread
 local memory to store $Probs$ and $Grads$ which physically resides in
@@
 % Shared memory caching strategy
-\subsubsection*{The shared memory strategy} allocates the temporary vectors
+\paragraph*{\textbf{The shared memory strategy}} allocate the temporary vectors
 $Probs_i$ and $Grads_i$
@@
 % Register and shared memory using code generation and loop unrolling 
-\subsubsection*{The code generation strategy} dynamically generates the code of the kernel to
+\paragraph*{\textbf{The code generation strategy}} dynamically generate the code of the kernel to
 custom tailor its properties. For example, this strategy can control whether a

### 2017-05-04 gpu-05-02-sequential.tex
Start working on integration: block out double algorithm, unify symbol usage

@@
-\subsection{Achieving an Efficient Baseline Sequential Version}
+\subsubsection{Achieving an Efficient Baseline Sequential Version}
 
-This section presents the results of transforming the algorithm implementation
+This subsection presents the results of transforming the algorithm implementation
 from Python to C++, and then removing obvious inefficiencies, to achieve a
@@
 
-The mini-batch sampling as described in Section~\ref{sec-algorithm} randomly
+The mini-batch sampling as described in Section~\ref{sec-background} randomly
 chooses either a mini-batch of link edges, whose size is the degree of one
@@
 \hline
+\\[-1ex]
 \end{tabular}
@@
 We present timings for the compute stages of the algorithm described in
-Section~\ref{sec-algorithm}, with the exception of
+Section~\ref{sec-background}, with the exception of
 \textit{update\_theta\_beta} which combines \textit{update\_theta}

### 2017-05-04 gpu-05-03-accelerator.tex
Start working on integration: block out double algorithm, unify symbol usage

@@
-\subsection{Analysis of CPU Parallelism}
+\subsubsection{Analysis of CPU Parallelism}
 
@@
 \cline{1-1}\cline{3-6}
+\\[-1.5ex]
 \end{tabular}
@@
 
-\subsection{Analysis of GPU Parallelism}
+\subsubsection{Analysis of GPU Parallelism}
 As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the
@@
 
-\subsection{Comparison of Compute Devices}
+\subsubsection{Comparison of Compute Devices}
 Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the

### 2017-05-04 gpu-05-05-perplexity.tex
Start working on integration: block out double algorithm, unify symbol usage

@@
 
-\subsection{Algorithm Convergence}
+\subsubsection{Algorithm Convergence}
 
@@
 graph. The runtime per iteration is independent of~$N$, but it depends on the
-product of~$K$, the mini-batch size~$m$, and the neighbor sample size~$n$. On
+product of~$K$, the mini-batch size~$M$, and the neighbor sample size~|\Neighbors|. On
 the other hand, the number of iterations required to achieve sufficient
@@
 To fill the CPU core memory~(64GB), the input parameter set is $K$=3072,
-$m$=4096, $n$=32. Sufficient convergence is reached after 18~hours. Note
+$M$=4096, |\Neighbors|=32. Sufficient convergence is reached after 18~hours. Note
 that this network graph is so large that the GTX Titan-X would be unable to

### 2017-05-04 gpu-05-evaluation.tex
Place the chunks of text unmodified into one document
Start working on integration: block out double algorithm, unify symbol usage

@@
-\section{Evaluation}
+\subsection{Evaluation}
 \label{sec-evaluation}
@@
 
+\begin{comment}
 \begin{table}[tb]
@@
 \end{table}
+\end{comment}
 
-\input{05-02-sequential}
-\input{05-03-accelerator}
-\input{05-05-perplexity}
+\begin{table*}
+  \centering
+  \begin{tabular}{l r r r l}
+    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
+                    &            &               & communities    &             \\
+    \hline
+    CA-HepPh        &    12,008  &    118,521    & n/a            & Paper citation network \\
+    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
+    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
+    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
+    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
+    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
+    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
+    \hline
+    \\[-1ex]
+  \end{tabular}
+  \caption{Summary of SNAP graph data sets used for evaluation.}
+  \label{table-snap}
+\end{table*}
+
+\input{gpu-05-02-sequential}
+\input{gpu-05-03-accelerator}
+\input{gpu-05-05-perplexity}

### 2017-05-04 gpu-06-conclusion.tex
Start working on integration: block out double algorithm, unify symbol usage

@@
-\section{Conclusion and Future Work}
+\section{Conclusion}
 Modern machine learning algorithms are empowering their
@@
 problem it is applied to.
-
-This paper focused on the acceleration of the SG-MCMC algorithm on a
-single node. For future work, we are designing and implementing a distributed
-version of the algorithm that uses the collective memory of a cluster, which will enable
-processing of much larger datasets. These changes in context fundamentally
-change the optimization landscape's boundaries and priorities.

### 2017-05-09 02-background.tex
Some nitpicks

@@
 
-\paragraph{\textbf{Sampling global parameters}} The update rule for global parameter $\theta$ is
+\paragraph*{\textbf{Sampling global parameters}} the update rule for global parameter $\theta$ is
 \bea
@@
 
-\paragraph{\textbf{Sampling local parameters}}
-The update rule for the local parameters $\phi$ is
+\paragraph*{\textbf{Sampling local parameters}}
+the update rule for the local parameters $\phi$ is
 \bea

### 2017-05-09 gpu-03-problem-statement.tex
Place the chunks of text unmodified into one document
Some nitpicks

@@
 %
-Kindly refer to~\cite{DBLP:journals/corr/LiAW15} for a detailed explanation of
+Kindly refer to~\cite{LiAW15} for a detailed explanation of
 the statistical concepts of the algorithm.
@@
 \begin{figure}[tb]
-\input{03-02-algorithm}
+\input{gpu-03-02-algorithm}
 \end{figure}
@@
 The first stage (line~2 in the listing) randomly draws a mini-batch $M$,
-using the "stratified random node" strategy~\cite{DBLP:journals/corr/LiAW15}.
+using the "stratified random node" strategy~\cite{LiAW15}.
 In this strategy, a coin toss is used to decide between two sample types.
@@
 The
-perplexity, as defined in~\cite{DBLP:journals/corr/LiAW15}, is the exponential
+perplexity, as defined in~\cite{LiAW15}, is the exponential
 of the average over time of the negative

### 2017-05-10 01-intro.tex
Initial merge of intro, conclusions

@@
 
-In this paper, we focus on the SG-MCMC method applied to the a-MMSB model.
-Recent work\cite{LiAW15} showed that this methodology is faster and more accurate than the SVB method.
-\cite{LiAW15} further proposes a heuristic to scale up the number of communities at the cost of less precision; the work in this paper scales the algorithm without that heuristic with a custom high-performance implementation.
-To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem.
+In this paper, we focus on the SG-MCMC method applied to the a-MMSB model. Recent
+work~\cite{LiAW15} showed that this methodology is faster and more accurate than the
+SVB method. \cite{LiAW15} further proposes a heuristic to scale up the number of
+communities at the cost of less precision; the work in this paper considers the
+algorithm without that heuristic.
+
+From a computational point of view, our SG-MCMC a-MMSB algorithm differs
+from widespread machine learning algorithms, like deep learning, in several
+ways. First, it is highly data-intensive which makes parallel acceleration
+particularly challenging. Second, owing to the algorithm's stochastic
+nature, the majority of its memory access patterns and data dependencies
+are non-deterministic. As a result, straightforward optimization attempts
+of the memory access patterns either fail or lead to non-intuitive results. Third,
+the size of its intermediate data structures makes it difficult to scale to very large
+communities.
+
+Starting from an existing implementation in Python/Numpy, we will present
+how this algorithm can be implemented much more efficiently in C++. Then,
+we will describe our two key contributions: first, how to scale up our
+implementation to run on GPUs and multicore CPUs, second, to scale out
+our implementation to overcome the memory bounds posed by GPUs or general
+single-machine systems.
+% To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem.
 In particular, we use a mixture of OpenMP, MPI and RDMA in order to efficiently scale and accelerate the algorithm's computation.
 
-Achieving this goal necessitated overcoming several challenges.
-First, the algorithm's state grows rapidly with larger graphs and number of latent communities.
-Since the full state information is too large to fit in a single machine's memory, it is partitioned and distributed across a cluster of machines.
-Second, to access the full state, each cluster node must read remote memory hosted by its peers.
-We leveraged RDMA to limit the high latency of such operations and increase the communication bandwidth.
-To reduce the latency further, we pipelined the algorithm's computations such that data can be fetched in advance over the network.
-Finally, the algorithm's computation is effectively distributed across the cluster nodes and parallelized further within each node by exploiting their multi-core CPUs.
+The focus of the first part of our work is on the computational efficiency and parallel
+performance of the SG-MCMC algorithm by
+developing aggressive optimizations targeting
+multi-core CPUs and GPUs. The GPU implementation achieves speedup factors up to~86,
+compared to the
+well-tuned sequential C++ program which itself is a factor~1000-1500 faster than
+the original Python/Numpy program.
+%
+Through careful analysis of the computation and data structures we show that
+the algorithm's full state can be reduced by roughly 75\%. Compressing the
+state significantly reduces the data intensity and allows for tackling larger
+problems. % while maintaining all state in memory.
+
+\begin{comment}
+Further, by cataloguing and accounting for the various load and store
+operations, we identified the highest priority locations of data reuse. In
+order to circumvent the unclear optimization landscape, we developed an
+effective kernel code generation mechanism that explores the benefits of
+exploiting all permutations of the available optimization opportunities. These
+optimizations include caching in shared memory, caching in the register file,
+loop unrolling and explicit vectorization.
+\end{comment}
+
+The second part of our work, scaling out to the largest publicly available
+community graphs, also necessitated overcoming several challenges. First,
+the algorithm's state grows rapidly with larger graphs and number of latent
+communities.  Since the full state for the larger graphs does not fit in a
+single machine's memory, it is partitioned and distributed across a cluster
+of machines. Second, to access the full state, each cluster node must read
+remote memory hosted by its peers. We leveraged RDMA to limit the high latency
+of such operations and increase the communication bandwidth. To reduce the
+latency further, we pipelined the algorithm's computations such that data can
+be fetched in advance over the network. Finally, the algorithm's computation
+is effectively distributed across the cluster nodes and parallelized further
+within each node by exploiting their multi-core CPUs.
 
 The remainder of this paper is organized as follows.
-Section~2 provides an overview of the algorithm and its theoretical foundation.
-The design and implementation of the parallel and distributed solution is presented in Section~3.
-Section~4 evaluates the efficacy of the system and analyzes its performance.
-Finally, Section~5 provides concluding remarks.
+Section~\ref{sec-background} provides an overview of the
+algorithm and its theoretical foundation, with related work in
+Section~\ref{sec-related}. Section~\ref{sec-gpu} delves
+into the parallelization optimizations for GPUs and multicore CPUs.
+Section~\ref{sec-distr} describes the design, implementation and evaluation of
+the scalable distributed solution. Finally, Section~\ref{sec-conclusion}
+provides concluding remarks.
 

### 2017-05-10 06-conclusion.tex
Initial merge of intro, conclusions

--- 2017-05-04 gpu-06-conclusion.tex
@@
-\section{Conclusion}
+\section{Conclusions}
+\label{sec-conclusion}
+
 Modern machine learning algorithms are empowering their
@@
 analysis of the computational patterns and data structures which significantly
-reduces its data intensity. Further, we navigated the complex optimization
+reduces its data intensity.
+
+In our optimization for manycore processors, we navigated the complex optimization
 landscape by dynamically generating kernel code and testing different
@@
 significant speedup factors of 21 and 86 using a multi-core CPU and a GPU
-respectively. These speedup numbers where achieved in comparison to an already
+respectively. These speedup numbers were achieved in comparison to an already
 optimized sequential implementation. Finally, we evaluated the performance of
@@
 their optimal configurations.
-
+%
 The outcome of this study reinforces the significance of avoiding premature
@@
 problem it is applied to.
+
+We further created a highly scalable implementation to handle the largest existing
+community graphs.
+This design faced
+several obstacles in order to achieve high performance.
+The algorithm was restructured to facilitate its
+parallelization. We overlapped computation
+with communication to hide latency.  We use a
+mixture of MPI and RDMA primitives to speedup the communication between cluster
+nodes.
+%
+Further, we conducted a thorough empirical evaluation of the system to study its strong
+and weak scalability on 65 cluster nodes using large data sets.
+% Additionally, we assessed the efficiency of the algorithm's resource utilization.
+Finally, a
+demonstration of the implementation's utility was provided by processing 6
+different organic data sets.
+To the best of
+our knowledge, this is the first time that the problem of deducing overlapping
+communities has been learned for problems of such a large scale.

### 2017-05-10 01-intro.tex
Begin working on picture refactor

@@
 implementation to run on GPUs and multicore CPUs, second, to scale out
-our implementation to overcome the memory bounds posed by GPUs or general
-single-machine systems.
+our implementation to overcome the memory bounds posed by GPUs or
+single-machine systems in general.
 % To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem.
@@
 The second part of our work, scaling out to the largest publicly available
-community graphs, also necessitated overcoming several challenges. First,
-the algorithm's state grows rapidly with larger graphs and number of latent
+community graphs, also necessitated overcoming several challenges.
+The algorithm's state grows rapidly with larger graphs and number of latent
 communities.  Since the full state for the larger graphs does not fit in a
 single machine's memory, it is partitioned and distributed across a cluster
-of machines. Second, to access the full state, each cluster node must read
-remote memory hosted by its peers. We leveraged RDMA to limit the high latency
-of such operations and increase the communication bandwidth. To reduce the
-latency further, we pipelined the algorithm's computations such that data can
-be fetched in advance over the network. Finally, the algorithm's computation
+of machines. Hence, to access the full state, each cluster node must read
+remote memory hosted by its peers. We leveraged RDMA to improve the latency and
+bandwidth
+of such operations. To reduce the
+latency further, we pipelined the algorithm's computations by fetching
+data in advance over the network.
+\begin{comment} Finally, the algorithm's computation
 is effectively distributed across the cluster nodes and parallelized further
 within each node by exploiting their multi-core CPUs.
+\end{comment}
 

### 2017-05-10 02-background.tex
Begin working on picture refactor

@@
 Then, Metropolis-Hastings accept-reject tests are applied.
-Comparing to LMC, SGLD only requires to process a mini-batch $\SGLDMinibatch$ at each iteration and ignores the MH test, and thus the computation complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
+Comparing to LMC, SGLD only requires to process a mini-batch $\SGLDMinibatch$ at each iteration and ignores the MH test, and thus computation complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
 
+\looseness=-1
 Stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex.
@@
 here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \SGLDMinibatch$.
+
+\addtolength{\tabcolsep}{-3pt}
+\renewcommand{\arraystretch}{1.5}
+\begin{table}[b] % [htbp]
+\center
+\begin{tabular}{c C{0.075\textwidth} c L{0.26\textwidth}}
+\textbf{symbol}
+	 & \textbf{type} & \textbf{size} & \textbf{description} \\
+         % & \multicolumn{1}{c}{\textbf{type}}
+                           % & \multicolumn{1}{c}{\textbf{description}} \\
+\hline
+$\cK$      &              & $K$        & set of communities \\
+\Vertices  & \{vertex\}   & $N$          & vertices in the graph \\
+\Edges     & \{edge\}     &              & linked edges in the graph \\
+$E$        & \{edge\}     &              & $\Vertices\times\Vertices$: linked and nonlinked edges \\
+\Heldout   & \{edge\}     &              & held-out subset of the graph \\
+\Minibatch & \{edge\}     &              & sampled mini-batch of edges in~$E$ \\
+$M$        &              &              & number of vertices in \Minibatch \\
+\Neighbors & \{vertex\}   &              & sampled neighbor set for a vertex in~\Minibatch \\
+$\theta$   & float~vector 2-D & $K\times{}2$ & reparameterization of $\beta$.
+					$\beta[k] = \theta[k][1] / \sum_j\theta[k][j]$ \\
+$\beta$    & float vector & $K$          & community strength \\
+$\phi$     & float~vector 2-D & $N\times{}K$ & reparameterization of $\pi$.
+					$\pi[i][k] = \phi[i][k] / \sum_j\phi[i][j]$ \\
+$\pi$      & float~vector 2-D & $N\times{}K$ &
+					$\pi[i][k]$~is~probability that vertex $i$ is in community $k$ \\
+\hline
+\\[-3ex]
+\end{tabular}
+\caption{Definition of most important symbols}
+\label{tbl-symbols}
+\end{table}
+\addtolength{\tabcolsep}{3pt}
+\renewcommand{\arraystretch}{1.0}
 
@@
 \end{algorithm}
-
-\addtolength{\tabcolsep}{-3pt}
-\renewcommand{\arraystretch}{1.5}
-\begin{table}[b] % [htbp]
-\center
-\begin{tabular}{c C{0.075\textwidth} c L{0.26\textwidth}}
-\textbf{symbol}
-	 & \textbf{type} & \textbf{size} & \textbf{description} \\
-         % & \multicolumn{1}{c}{\textbf{type}}
-                           % & \multicolumn{1}{c}{\textbf{description}} \\
-\hline
-$\cK$      &              & $K$        & set of communities \\
-\Vertices  & \{vertex\}   & $N$          & vertices in the graph \\
-\Edges     & \{edge\}     &              & linked edges in the graph \\
-$E$        & \{edge\}     &              & $\Vertices\times\Vertices$: linked and nonlinked edges \\
-\Heldout   & \{edge\}     &              & held-out subset of the graph \\
-\Minibatch & \{edge\}     &              & sampled mini-batch of edges in~$E$ \\
-$M$        &              &              & number of vertices in \Minibatch \\
-\Neighbors & \{vertex\}   &              & sampled neighbor set for a vertex in~\Minibatch \\
-$\theta$   & float~vector 2-D & $K\times{}2$ & reparameterization of $\beta$.
-					$\beta[k] = \theta[k][1] / \sum_j\theta[k][j]$ \\
-$\beta$    & float vector & $K$          & community strength \\
-$\phi$     & float~vector 2-D & $N\times{}K$ & reparameterization of $\pi$.
-					$\pi[i][k] = \phi[i][k] / \sum_j\phi[i][j]$ \\
-$\pi$      & float~vector 2-D & $N\times{}K$ &
-					$\pi[i][k]$~is~probability that vertex $i$ is in community $k$ \\
-\hline
-\\[-3ex]
-\end{tabular}
-\caption{Definition of most important symbols}
-\label{tbl-symbols}
-\end{table}
-\addtolength{\tabcolsep}{3pt}
-\renewcommand{\arraystretch}{1.0}
 

### 2017-05-10 02-related.tex
Initial merge of intro, conclusions
Begin working on picture refactor

@@
 \section{Related Work}
+\label{sec-related}
 
-\subsection{Related work on Accelerators for Bayesian methods}
+\textbf{\textit{\large Need to rewrite to better fit the Background Section}}
 
-In this section, we discuss related machine learning projects that use
-accelerators.
 Immediately when GPUs evolved into generally programmable accelerators, machine
@@
 
-\subsection{Related work on parallel community detection}
-
-Among previous work on parallelizing MMSB algorithms, we mention the Online
-Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13} on GPUs.
 There are a few projects that did a distributed implementation

### 2017-05-10 distr-03-design.tex
Some nitpicks
Begin working on picture refactor

@@
-\subsection{System Design}
+% \subsection{System Design}
 \label{design-section}
@@
 
-The SG-MCMC algorithm described in the previous section has an abundance of
+The SG-MCMC a-MMSB algorithm has an abundance of
 opportunities for parallelism, for the multi-threaded shared-memory type as
@@
 machines. The downside of a distributed implementation is that it requires
-considerably more programming effort.
+considerable implementation effort.
 
@@
 multi-threaded parallelization techniques sufficed: the distributed performance
-is dominated by (remote) memory latency, not by node parallelism.
+is dominated by (remote) memory latency, not by node parallelism. As our
+analysis will show, our pipelining optimizations will fully overlap the
+parallel calculations with (pre)loading the data from remote hosts.
 The details of the multi-threaded parallelism

### 2017-05-10 distr-04-evaluation.tex
Begin working on picture refactor

@@
   \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\columnwidth}
-  \caption{(a) Execution time of 2048 algorithm iterations for the
+  \caption{(a)~Execution time of 2048 algorithm iterations for the
   same problem size (com-Friendster, K=1024, M=16384 n=32) across different
-  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
+  cluster sizes. (b)~Speedup achieved for same experiments in (a) with respect to
   8 nodes}

### 2017-05-10 gpu-05-03-accelerator.tex
Begin working on picture refactor

@@
 % CPU
-\begin{figure}[htb]
+\begin{figure}[tb]
 \centering
@@
 shared strategy and 8 variations of the code generation strategy. This
-section investigates the effectiveness of each flavor on Nvidia's GTX TitanX,
+section investigates the effectiveness of each flavor on Nvidia's GTX~\mbox{Titan-X},
 GTX980, K40c and K20m.
@@
   \centering
-  \epsfig{file=plots/titanx-1024-strategies-w1.eps, width=\textwidth}
-  \caption{Execution time of 1000 \textit{update\_phi} invocations using the TitanX GPU,
+  \setlength\tabcolsep{0em}
+  \begin{tabular}{c c c}
+  \epsfig{file=plots/titanx-1024-strategies-w1.eps, width=0.33\textwidth} &
+  \epsfig{file=plots/titanx-2048-strategies-w1.eps, width=0.33\textwidth} &
+  \epsfig{file=plots/k40-1024-strategies-w2.eps, width=0.33\textwidth}
+  \end{tabular}
+  \caption{Execution time of 1000 \textit{update\_phi} invocations using the
+  \mbox{Titan-X} GPU,
   without explicit kernel vectorization, across a sweep of
@@
 
+\begin{comment}
 % TITANX K=2K
@@
   \epsfig{file=plots/titanx-2048-strategies-w1.eps, width=\textwidth}
-  \caption{Execution time of 1000 \textit{update\_phi} invocations using the TitanX GPU,
+  \caption{Execution time of 1000 \textit{update\_phi} invocations using the
+  \mbox{Titan-X} GPU,
   without explicit kernel vectorization, across a sweep of
@@
 \end{figure*}
+\end{comment}
 
-Figure \ref{titanx-w1-sweep}(a) presents the performance of the GTX TitanX GPU
+Figure \ref{titanx-w1-sweep}(a) presents the performance of the GTX~\mbox{Titan-X} GPU
 for all 10 strategy flavors, without explicitly vectorizing the kernels. The
@@
 performance as they unroll internal loops as well as cache data in
-registers. The TitanX obtains the best performance with the \textit{GEN-RSS}
+registers. The \mbox{Titan-X} obtains the best performance with the \textit{GEN-RSS}
 strategy and a thread block size of 64. The results of explicitly vectorized kernels
-are omitted as they obtain worse performance on the TitanX.
+are omitted as they obtain worse performance on the \mbox{Titan-X}.
 
@@
 
+\begin{comment}
 % K40
@@
 \end{figure*}
+\end{comment}
 
@@
 Figures~\ref{k40-w2-sweep} (a) and (b) present the performance of the K40c GPU
-for the same experimental configuration as shown for the TitanX in
+for the same experimental configuration as shown for the \mbox{Titan-X} in
 Figure~\ref{titanx-w1-sweep} with one exception. Namely, the code version used
@@
 
-The performance results obtained from the GTX TitanX and Tesla K40c GPUs
+The performance results obtained from the GTX~\mbox{Titan-X} and Tesla~K40c GPUs
 reinforce the importance of customizing compute kernels to each GPU's specific
@@
 Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
-GTX TitanX, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
+GTX~\mbox{Titan-X}, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
 sequential C++ version. These results are consistent with the relative
 capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
-the TitanX achieves the highest speedup of 86 relative to the baseline. At the
+the \mbox{Titan-X} achieves the highest speedup of 86 relative to the baseline. At the
 other end, the parallel CPU version attains a speedup factor of~20.9.
@@
 normalized over the non-vectorized kernel version for each GPU. For instance,
-the TitanX exhibits an overhead factor of approximately 1.8 when using a vector
+the \mbox{Titan-X} exhibits an overhead factor of approximately 1.8 when using a vector
 width of~4. On the other hand, the K40c improves its performance by roughly

### 2017-05-10 gpu-05-05-perplexity.tex
Begin working on picture refactor

@@
   \epsfig{file=plots/ppx-gpu.eps, width=\columnwidth}
-  \caption{Perplexity convergence of com-DBLP on the GTX~Titan-X~GPU. Number
+  \caption{Perplexity convergence of com-DBLP on the GTX~\mbox{Titan-X}~GPU. Number
   of communities chosen to maximally fill the GPU memory.}
@@
 
-\begin{figure*}[t]	% [htb]
+\begin{figure}[t]	% [htb]
   \centering
-  \epsfig{file=plots/ppx-gpu-cpu.eps, width=\textwidth}
-  \caption{Perplexity convergence. (a)~com-DBLP on the GTX~Titan-X~CPU.
+  \epsfig{file=plots/ppx-gpu-cpu.eps, width=0.45\textwidth}
+  \caption{Perplexity convergence. (a)~com-DBLP on the GTX~\mbox{Titan-X}~CPU.
   (b)~com-LiveJournal on the multi-core~CPU. Number of communities chosen to
@@
   \label{fig-ppx-gpu-cpu}
-\end{figure*}
+\end{figure}
 
@@
 Figure~\ref{fig-ppx-gpu-cpu}(a) shows the perplexity development for the dataset
-com-DBLP, with $N=317,080$. It is computed on the GTX Titan-X GPU with input
+com-DBLP, with $N=317,080$. It is computed on the GTX~\mbox{Titan-X} GPU with input
 parameter set $K=4096$, $m=4096$, $n=32$. Convergence is reached after 6~hours.
 
-Figure~\ref{fig-ppx-gpu-cpu}(b) shows the perplexity development convergence for
+Figure~\ref{fig-ppx-gpu-cpu}(b) shows the perplexity convergence for
 com-LiveJournal, with $N=3,997,962$, computed on a multi-core DAS5 node.
 To fill the CPU core memory~(64GB), the input parameter set is $K$=3072,
-$M$=4096, |\Neighbors|=32. Sufficient convergence is reached after 18~hours. Note
-that this network graph is so large that the GTX Titan-X would be unable to
+$M$=4096, $|\Neighbors|$=32. Sufficient convergence is reached after 18~hours. Note
+that this network graph is so large that the GTX~\mbox{Titan-X} would be unable to
 process more than $K$=256 communities.

### 2017-05-10 gpu-05-evaluation.tex
Begin working on picture refactor

--- 2017-05-03 gpu-05-evaluation.tex
@@
 E5-2630v3 CPU clocked at 2.40GHz, 64GB of memory and 8TB of storage.
-Additionally, the cluster is fitted with a number of Nvidia GPUs including GTX
-TitanX, GTX980, K40c and K20m; see Table~\ref{table-gpus} for an overview of
+Additionally, the cluster is fitted with a number of Nvidia GPUs including
+GTX~\mbox{Titan-X}, GTX980, K40c and K20m; see Table~\ref{table-gpus} for an overview of
 their properties. Table~\ref{table-snap} outlines the graph
@@
 \end{table}
+\end{comment}
 
-\input{05-02-sequential}
-\input{05-03-accelerator}
-\input{05-05-perplexity}
+\begin{table}
+  \centering
+  \def\tabcolsep{0.2em}
+  \begin{tabular}{l r r r p{9em}}
+    Name            & \#Vertices &       \#Edges & \multicolumn{1}{c}{\#Ground-} & Description \\
+                    &            &               & \multicolumn{1}{c}{truths}    &             \\
+    \hline
+    CA-HepPh        &    12,008  &    118,521    & \multicolumn{1}{c}{n/a}       & Paper citation network \\
+    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
+    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
+    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
+    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
+    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
+    com-Amazon      &    334,863 &       925,872 & 75,149         & Product copurchasing network \\
+    \hline
+    \\[-1ex]
+  \end{tabular}
+  \caption{Summary of SNAP graph data sets used for evaluation.}
+  \label{table-snap}
+\end{table}
+
+\input{gpu-05-02-sequential}
+\input{gpu-05-03-accelerator}
+\input{gpu-05-05-perplexity}

### 2017-05-10 plots/plot-gpu-sweep.sh

@@ -22,11 +22,11 @@ MID=128
 #cat $TFILE
 
 cat <<EOF | gnuplot --persist
-w = 0.6
-h = 0.4
+w = 0.3
+h = 0.32
 k = 0.05
-nr = 1
-nc = 2
+nr = 2
+nc = 1
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
@@ -66,7 +66,7 @@ plot \
   '< cat $TFILE | grep RRS' u 2:4 w lp title 'GEN-RRS', \
   '< cat $TFILE | grep RRR' u 2:4 w lp title 'GEN-RRR'
 
-set origin col(1),row(0)
+set origin col(0),row(1)
 set size w,h
 set title '(b) Focused on optimal settings'
 set xrange [32:128]

### 2017-05-10 plots/plot-hpc-cloud.sh

@@ -83,16 +83,17 @@ set xlabel "Number of Communities (K)"
 set ylabel "Time per iteration (milliseconds)"
 #set xrange [256:]
 set grid ytics
-set key below
+# set key below
+set key left
 
 
 set origin col(0),row(0)
 set size w,h
 set title '(a) HPC Cloud (16 and 40 cores) vs. 1 DAS5 node (16 cores) using com-DBLP'
 plot \
-     "$$.tmp.hpc40" u 1:(1000*\$2) w lp ls 1 t 'HPC Cloud 40-cores', \
-     "$$.tmp.hpc16" u 1:(1000*\$2) w lp ls 2 t 'HPC Cloud 16-cores', \
-     "$$.tmp.4" u 1:(1000*\$2)     w lp ls 3 t 'Single DAS5 node 16-cores'
+     "$$.tmp.hpc40" u 1:(1000*\$2) w lp ls 1 t 'HPC Cloud 40 cores', \
+     "$$.tmp.hpc16" u 1:(1000*\$2) w lp ls 2 t 'HPC Cloud 16 cores', \
+     "$$.tmp.4" u 1:(1000*\$2)     w lp ls 3 t 'Single DAS5 node 16 cores'
 
 set origin col(0),row(1)
 set size w,h

### 2017-05-10 plots/plot-ml-curves.sh

@@ -8,15 +8,15 @@ set ylabel "Perplexity"
 set nokey
 set grid ytics
 
-w = 0.6
-h = 0.4
+w = 0.32
+h = 0.28
 k = 0.0
-nr = 3
-nc = 2
+nr = 2
+nc = 3
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
-set terminal postscript eps enhanced color font ',10'
+set terminal postscript eps enhanced color font ',8.5'
 set size (nc*w),(nr*h+k)
 set output 'ppx.eps'
 set multiplot layout nr,nc
@@ -39,7 +39,7 @@ set title "(b) com-LiveJournal | 192K communities on 64+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-epst5e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 set yrange [*:]
 
-set origin col(0),row(1)
+set origin col(2),row(0)
 set size w,h
 
 set xrange [:64]
@@ -47,20 +47,20 @@ set title "(c) com-Orkut | 256K communities on 64+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst3e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 set xrange [:*]
 
-set origin col(1),row(1)
+set origin col(0),row(1)
 set size w,h
 
 set title "(d) com-Youtube | 8,385 communities on 12+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-youtube-K8385-m1K-n32-msx1K-epst1e-05-np13/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
-set origin col(0),row(2)
+set origin col(1),row(1)
 set size w,h
 
 set yrange [110:]
 set title "(e) com-DBLP | 13,477 communities on 20+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-dblp-K13477-flt32-m1K-n32-msx1K-x10M-epst5e-05-np21/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
-set origin col(1),row(2)
+set origin col(2),row(1)
 set size w,h
 
 set yrange [*:]

### 2017-05-10 plots/plot-ppx.sh

@@ -1,11 +1,11 @@
 #!/bin/bash
 
 cat << EOF | gnuplot --persist
-w = 0.6
-h = 0.4
+w = 0.4
+h = 0.3
 k = 0.0
-nr = 1
-nc = 2
+nr = 2
+nc = 1
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
@@ -25,7 +25,7 @@ set yrange [62:95]
 set title '(a) Com-DBLP K=4K m=4K n=32'
 plot "< perl ./get_time.perl ppx-DBLP-K4K-m4K-n32 " using (\$1/3600):3 notitle with lines
 
-set origin col(1),row(0)
+set origin col(0),row(1)
 set size w,h
 
 set xrange [-0.3:18.3]

### 2017-05-10 plots/plot-sweep-K-proportional-np.sh

@@ -15,7 +15,7 @@ cat $$.tmp
 cat <<EOF | gnuplot --persist
 
 w = 0.6
-h = 0.4
+h = 0.3
 
 set terminal postscript eps enhanced color font ',10'
 set output 'sweep-over-K-proportional-np.eps'

### 2017-05-10 plots/plot-sweep-np-fixed-K.sh

@@ -59,8 +59,8 @@ set size w,h
 set xlabel "Number of Nodes"
 set xrange [0:]
 set yrange[0:]
-set key below
-set grid ytics
+# set key below
+# set grid ytics
 
 set style line 1 lt 1 lw 2 linecolor rgb "red"
 set style line 2 lt 2 lw 2 linecolor rgb "blue"
@@ -73,18 +73,19 @@ set ylabel "Time (Seconds)"
 set title "(a) Execution time"
 plot \
      "$$.tmp"   u (\$2-1):3 w lp axis x1y1 ls 1 t 'total', \
-     "$$.tmp.2" u (\$2-1):3 w lp axis x1y1 ls 2 t 'deploy minibatch', \
      "$$.tmp.3" u (\$2-1):3 w lp axis x1y1 ls 3 t 'update\_phi\_pi', \
+     "$$.tmp.2" u (\$2-1):3 w lp axis x1y1 ls 2 t 'deploy minibatch', \
      "$$.tmp.4" u (\$2-1):3 w lp axis x1y1 ls 4 t 'update\_beta\_theta'
 
 set origin col(0),row(1)
 set size w,h
 set ylabel "Speedup w.r.t. 8-node time"
 set title "(b) Speedup w.r.t. 8-node time"
+set key left
 plot \
      "$$.tmp" u (\$2-1):(${BASE}/\$3)    w lp axis x1y1 ls 1 t "total", \
-     "$$.tmp.2" u (\$2-1):(${BASE2}/\$3) w lp axis x1y1 ls 2 t 'deploy minibatch', \
      "$$.tmp.3" u (\$2-1):(${BASE3}/\$3) w lp axis x1y1 ls 3 t 'update\_phi\_pi', \
+     "$$.tmp.2" u (\$2-1):(${BASE2}/\$3) w lp axis x1y1 ls 2 t 'deploy minibatch', \
      "$$.tmp.4" u (\$2-1):(${BASE4}/\$3) w lp axis x1y1 ls 4 t 'update\_beta\_theta'
 
 EOF

### 2017-05-11 gpu-05-03-accelerator.tex
Refactor the figures, esp. the GPU block size sweep

@@
   \centering
+  \epsfig{file=plots/gpu-sweep-legends.eps, width=1.0\textwidth}
   \setlength\tabcolsep{0em}
@@
   \end{tabular}
-  \caption{Execution time of 1000 \textit{update\_phi} invocations using the
-  \mbox{Titan-X} GPU,
-  without explicit kernel vectorization, across a sweep of
-  \textit{update\_phi} thread block
-  sizes. Relevant model parameters: K=1024, M=4096, n=32.}
-  \label{titanx-w1-sweep}
+  \caption{Execution time of 1000 \textit{update\_phi} invocations across a sweep of
+  	\textit{update\_phi} thread block sizes. The lower figures are a zoom-in into
+	the optimal block range of the upper figures.
+	Other relevant model parameters: M=4096, n=32.
+	% (a) \mbox{Titan-X} GPU, without explicit kernel vectorization;
+	% Relevant model parameters: K=1024, M=4096, n=32.
+  	% (b) \mbox{Titan-X} GPU, without explicit kernel vectorization;
+	% K=2048, M=4096, n=32.
+	% (c) K40c GPU, with explicit kernel vectorization of width 2;
+	% K=1024, M=4096, n=32.
+  }
+  % \label{titanx-w1-sweep}
+  \label{gpu-sweep}
 \end{figure*}
@@
 
-Figure \ref{titanx-w1-sweep}(a) presents the performance of the GTX~\mbox{Titan-X} GPU
+Figure \ref{gpu-sweep}(a) presents the performance of the GTX~\mbox{Titan-X} GPU
 for all 10 strategy flavors, without explicitly vectorizing the kernels. The
@@
 A key model parameter that affects the behavior of the optimization strategies
-is the number of communities $K$. Figure~\ref{titanx-w1-sweep-2k} presents the
-same model configuration as in Figure~\ref{titanx-w1-sweep} but $K=2048$
+is the number of communities $K$. Figure~\ref{gpu-sweep}(b) presents the
+same model configuration as in Figure~\ref{gpu-sweep}(a) but $K=2048$
 instead of $K=1024$. One key difference between the two figures is the optimal
@@
 % \looseness=-1
-Figures~\ref{k40-w2-sweep} (a) and (b) present the performance of the K40c GPU
+Figures~\ref{gpu-sweep}(c) present the performance of the K40c GPU
 for the same experimental configuration as shown for the \mbox{Titan-X} in
-Figure~\ref{titanx-w1-sweep} with one exception. Namely, the code version used
+Figure~\ref{gpu-sweep}(a) with one exception. Namely, the code version used
 for this plot is explicitly vectorized with a vector width of~2. The results
@@
 %
-Surprisingly, Figure~\ref{k40-w2-sweep} shows that the NAIVE strategy
+Surprisingly, Figure~\ref{gpu-sweep}(c) shows that the NAIVE strategy
 outperforms SHARED and some of the GEN strategies. This can be explained by the

### 2017-05-11 gpu-05-05-perplexity.tex
Refactor the figures, esp. the GPU block size sweep

@@
 graph. The runtime per iteration is independent of~$N$, but it depends on the
-product of~$K$, the mini-batch size~$M$, and the neighbor sample size~|\Neighbors|. On
+product of~$K$, the mini-batch size~$M$, and the neighbor sample size~$|\Neighbors|$. On
 the other hand, the number of iterations required to achieve sufficient

### 2017-05-11 plots/plot-all-gpu-detail.sh

@@ -1,33 +1,33 @@
 #!/bin/bash
 
-./plot-gpu-sweep.sh titanx-1024-strategies-w1 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
-./plot-gpu-sweep.sh titanx-1024-strategies-w2 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
-./plot-gpu-sweep.sh titanx-1024-strategies-w4 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
+./plot-gpu-sweep.sh --title "(a) Titan-X K=1024 no vectorization" titanx-1024-strategies-w1 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
+./plot-gpu-sweep.sh --title "T(a) itan-X K=1024 vector width=2" titanx-1024-strategies-w2 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
+./plot-gpu-sweep.sh --title "T(a) itan-X K=1024 vector width=4" titanx-1024-strategies-w4 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
 
-./plot-gpu-sweep.sh gtx980-1024-strategies-w1 `ls dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
-./plot-gpu-sweep.sh gtx980-1024-strategies-w2 `ls dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
-./plot-gpu-sweep.sh gtx980-1024-strategies-w4 `ls dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
[…]

### 2017-05-11 plots/plot-gpu-sweep.sh

@@ -2,6 +2,26 @@
 
 # splot '< ./collect.sh build-cu/log-phi-sweep-phi-*' u 1:3:4 w p
 
+PLOT_TITLE="Put your title here"
+while true ; do
+	case "$1" in
+		--title)
+			shift
+			PLOT_TITLE="$1"
+			shift
+			;;
+		--*)
+			echo "Unknown flag: $1"
+			exit 37
+			;;
+		*)
+			break
+			;;
+	esac
+done
+
+# echo "Remaining args $@"
+
 NAME=$1
 shift
 
@@ -24,14 +44,13 @@ MID=128
 cat <<EOF | gnuplot --persist
 w = 0.3
 h = 0.32
-k = 0.05
 nr = 2
 nc = 1
-row(x) = ((nr-x-1)*h)+k
+row(x) = ((nr-x-1)*h)
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',8'
-set size (nc*w),(nr*h+k)
+set size (nc*w),(nr*h)
 set output '$NAME.eps'
 set multiplot layout nr,nc
 
@@ -53,7 +72,7 @@ unset key
 
 set origin col(0),row(0)
 set size w,h
-set title '(a) Full sweep range'
+set title '$PLOT_TITLE'
 plot \
   '< cat $TFILE | grep NAIVE' u 2:4 w lp title 'NAIVE', \
   '< cat $TFILE | grep SHARED' u 2:4 w lp title 'SHARED', \
@@ -68,7 +87,7 @@ plot \
 
 set origin col(0),row(1)
 set size w,h
-set title '(b) Focused on optimal settings'
+set title 'Focus on optimal settings'
 set xrange [32:128]
 set autoscale y
 plot \
@@ -100,6 +119,9 @@ plot \
 #  '< cat $TFILE | grep RRS' u 2:4 w lp title 'GEN-RRS', \
 #  '< cat $TFILE | grep RRR' u 2:4 w lp title 'GEN-RRR'
 
+unset multiplot
+set output 'gpu-sweep-legends.eps'
+
 unset xtics
 unset ytics
 unset y2tics
@@ -111,11 +133,13 @@ unset y2label
 unset title
 unset border
 
+w = 0.8
 set origin 0,0
+k = 0.05
 set size (nc*w),k
 set key center horizontal
 set xrange[0:1]
-set yrange[0:1]
+set yrange[0:3]
 set y2range[0:1]
[…]

### 2017-05-11 plots/plot-ml-curves.sh

@@ -27,14 +27,14 @@ set size w,h
 
 set title "(a) com-Friendster | 12K communities on 64+1 nodes"
 set yrange [105:]
-set xrange [:20]
+set xrange [-0.5:20.5]
 plot '< grep average_count ../data/ml-curves/log-Friendster-K12K-flt32-m16K-n32-mxs-16K-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
-set xrange [:*]
 
 set origin col(1),row(0)
 set size w,h
 
 set yrange [420:]
+set xrange [-2.0:66.0]
 set title "(b) com-LiveJournal | 192K communities on 64+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-epst5e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 set yrange [*:]
@@ -42,14 +42,14 @@ set yrange [*:]
 set origin col(2),row(0)
 set size w,h
 
-set xrange [:64]
+set xrange [-2.0:66.0]
 set title "(c) com-Orkut | 256K communities on 64+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst3e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
-set xrange [:*]
 
 set origin col(0),row(1)
 set size w,h
 
+set xrange [-0.2:8.2]
 set title "(d) com-Youtube | 8,385 communities on 12+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-youtube-K8385-m1K-n32-msx1K-epst1e-05-np13/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
@@ -57,6 +57,7 @@ set origin col(1),row(1)
 set size w,h
 
 set yrange [110:]
+set xrange [-0.2:8.2]
 set title "(e) com-DBLP | 13,477 communities on 20+1 nodes"
 plot '< grep average_count ../data/ml-curves/log-dblp-K13477-flt32-m1K-n32-msx1K-x10M-epst5e-05-np21/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
 
@@ -64,6 +65,7 @@ set origin col(2),row(1)
 set size w,h
 
 set yrange [*:]
[…]

### 2017-05-11 plots/plot-ppx.sh

@@ -23,7 +23,7 @@ set ylabel 'Perplexity'
 set xrange [-0.15:6.15]
 set yrange [62:95]
 set title '(a) Com-DBLP K=4K m=4K n=32'
-plot "< perl ./get_time.perl ppx-DBLP-K4K-m4K-n32 " using (\$1/3600):3 notitle with lines
+plot "< perl ./get_time.perl ppx-DBLP-K4K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 5
 
 set origin col(0),row(1)
 set size w,h
@@ -31,5 +31,5 @@ set size w,h
 set xrange [-0.3:18.3]
 set yrange [53:83]
 set title '(b) Com-LiveJournal K=3K m=4K n=32'
-plot "< perl ./get_time.perl ppx-LJ-K3K-m4K-n32 " using (\$1/3600):3 notitle with lines
+plot "< perl ./get_time.perl ppx-LJ-K3K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 5
 EOF

### 2017-05-12 02-background.tex
Gone over figures etc. Still need to redo related work.

@@
 
-SGLD originated from the Langevin Monte Carlo (LMC)~\cite{girolami2011riemann}, where unlike SGLD the gradient is computed exactly by using all data points.
+SGLD originated from Langevin Monte Carlo (LMC)~\cite{girolami2011riemann}, where unlike SGLD the gradient is computed exactly by using all data points.
 Then, Metropolis-Hastings accept-reject tests are applied.
@@
 \renewcommand{\arraystretch}{1.5}
-\begin{table}[b] % [htbp]
+\begin{table}[tb] % [htbp]
 \center

### 2017-05-12 gpu-05-evaluation.tex
Gone over figures etc. Still need to redo related work.

@@
     com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
-    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
+    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography network \\
     com-Amazon      &    334,863 &       925,872 & 75,149         & Product copurchasing network \\

### 2017-05-12 plots/plot-cpu.sh

@@ -4,6 +4,7 @@ cat <<EOF | gnuplot --persist
 
 set terminal postscript eps enhanced color font 'Helvetica,10' size 3.3,1.6
 set output 'cpu-vector.eps'
+set colorsequence classic
 
 set boxwidth 0.5
 set style fill solid

### 2017-05-12 plots/plot-gpu-sweep.sh

@@ -50,6 +50,7 @@ row(x) = ((nr-x-1)*h)
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',8'
+set colorsequence classic
 set size (nc*w),(nr*h)
 set output '$NAME.eps'
 set multiplot layout nr,nc
@@ -87,7 +88,7 @@ plot \
 
 set origin col(0),row(1)
 set size w,h
-set title 'Focus on optimal settings'
+set title 'As above, zoom-in around optimal settings'
 set xrange [32:128]
 set autoscale y
 plot \

### 2017-05-12 plots/plot-gpus.sh

@@ -21,6 +21,7 @@ cat <<EOF | gnuplot --persist
 
 set terminal postscript eps enhanced color font ',8' size 3.3,1.6
 set output 'gpus.eps'
+set colorsequence classic
 
 set boxwidth 0.5
 set style fill solid

### 2017-05-12 plots/plot-hpc-cloud.sh

@@ -69,6 +69,7 @@ row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',10'
+set colorsequence classic
 set size (nc*w),(nr*h+k)
 set output 'hpc-cloud.eps'
 set multiplot layout nr,nc

### 2017-05-12 plots/plot-ml-curves.sh

@@ -17,6 +17,7 @@ row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',8.5'
+set colorsequence classic
 set size (nc*w),(nr*h+k)
 set output 'ppx.eps'
 set multiplot layout nr,nc
@@ -28,7 +29,7 @@ set size w,h
 set title "(a) com-Friendster | 12K communities on 64+1 nodes"
 set yrange [105:]
 set xrange [-0.5:20.5]
-plot '< grep average_count ../data/ml-curves/log-Friendster-K12K-flt32-m16K-n32-mxs-16K-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
+plot '< grep average_count ../data/ml-curves/log-Friendster-K12K-flt32-m16K-n32-mxs-16K-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 4
 
 set origin col(1),row(0)
 set size w,h
@@ -36,7 +37,7 @@ set size w,h
 set yrange [420:]
 set xrange [-2.0:66.0]
 set title "(b) com-LiveJournal | 192K communities on 64+1 nodes"
-plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-epst5e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
+plot '< grep average_count ../data/ml-curves/log-com-LJ-K192K-flt32-m1K-n32-mxs1K-x1M-epst5e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 4
 set yrange [*:]
 
 set origin col(2),row(0)
@@ -44,14 +45,14 @@ set size w,h
 
 set xrange [-2.0:66.0]
 set title "(c) com-Orkut | 256K communities on 64+1 nodes"
-plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst3e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 5
+plot '< grep average_count ../data/ml-curves/log-orkut-flt32-K256K-m1K-n32-x10M-epst3e-05-np65/log.0' u (\$8 / 3600):(exp(\$14)) w l lw 4
 
 set origin col(0),row(1)
 set size w,h
 
 set xrange [-0.2:8.2]
[…]

### 2017-05-12 plots/plot-ppx.sh

@@ -10,6 +10,7 @@ row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',8'
+set colorsequence classic
 set size (nc*w),(nr*h+k)
 set output 'ppx-gpu-cpu.eps'
 set multiplot layout nr,nc
@@ -23,7 +24,7 @@ set ylabel 'Perplexity'
 set xrange [-0.15:6.15]
 set yrange [62:95]
 set title '(a) Com-DBLP K=4K m=4K n=32'
-plot "< perl ./get_time.perl ppx-DBLP-K4K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 5
+plot "< perl ./get_time.perl ppx-DBLP-K4K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 4
 
 set origin col(0),row(1)
 set size w,h
@@ -31,5 +32,5 @@ set size w,h
 set xrange [-0.3:18.3]
 set yrange [53:83]
 set title '(b) Com-LiveJournal K=3K m=4K n=32'
-plot "< perl ./get_time.perl ppx-LJ-K3K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 5
+plot "< perl ./get_time.perl ppx-LJ-K3K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 4
 EOF

### 2017-05-12 plots/plot-qperf.sh

@@ -1,6 +1,7 @@
 cat <<EOI | gnuplot --persist
 
 set terminal postscript eps enhanced color font ',10' size 3.3,1.6
+set colorsequence classic
 set output 'qperf.eps'
 
 set style line 1 lt 1 lw 2 linecolor rgb "red"

### 2017-05-12 plots/plot-sweep-K-fixed-np.sh

@@ -23,6 +23,7 @@ grep iteration ../data/sweep-over-K-fixed-np/*/* \
 cat <<EOF | gnuplot --persist
 
 set terminal postscript eps enhanced color font ',10' size 3.3,1.6
+set colorsequence classic
 
 set xlabel "Number of Communities (K)"
 set ylabel "Time (seconds)"

### 2017-05-12 plots/plot-sweep-K-proportional-np.sh

@@ -18,6 +18,7 @@ w = 0.6
 h = 0.3
 
 set terminal postscript eps enhanced color font ',10'
+set colorsequence classic
 set output 'sweep-over-K-proportional-np.eps'
 set size w,h
 

### 2017-05-12 plots/plot-sweep-np-fixed-K.sh

@@ -51,6 +51,7 @@ row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',10'
+set colorsequence classic
 set size (nc*w),(nr*h+k)
 set output 'sweep-over-np-fixed-K.eps'
 set multiplot layout nr,nc

### 2017-05-12 plots/plot-vector-sweep.sh

@@ -9,6 +9,7 @@ TFILE="$$.tmp"
 
 cat <<EOF | gnuplot --persist
 set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+set colorsequence classic
 set output '$NAME'
 
 set style data histogram

### 2017-05-18 01-intro.tex
Gone over figures etc. Still need to redo related work.
Polish a bit on GPU plots

@@
 
-The problem of detecting overlapping communities is modeled by the mixed membership stochastic blockmodels (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of MMSB, called assortative MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
+The problem of detecting overlapping communities is modeled by the mixed membership stochastic block model (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of MMSB, called assortative MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
 The MMSB model is a probabilistic graphical model~\cite{koller2009probabilistic} that represents a convenient paradigm for modeling complex relationships between a potentially large number of random variables.
@@
 the size of its intermediate data structures makes it difficult to scale to very large
-communities.
+community graphs.
 

### 2017-05-18 03-related.tex
Polish a bit on GPU plots

--- 2017-05-10 02-related.tex
@@
 
-\textbf{\textit{\large Need to rewrite to better fit the Background Section}}
-
-Immediately when GPUs evolved into generally programmable accelerators, machine
-learners recognized the potential of GPUs for their science.
-Much research
-has been devoted to deep-learning implementations on GPUs. However, Mixed
-Membership deduction (MMSB) is usually considered untractable for deep-learning
-algorithms. In contrast, our algorithm belongs to the family of Bayesian
-(approximative, stochastic) methods, where efficient algorithms for large-scale
-data have recently been developed. A broad survey of Bayesian methods is given
-in~\cite{DBLP:journals/corr/Zhu0H14}. Broadly speaking, Bayesian approximative
-algorithms are subdivided into variational and Monte Carlo methods; Markov
-Chain Monte Carlo (MCMC) is a subgroup of the latter. GPUs have been put to
-use in MCMC algorithms, e.g.\ \cite{journals/bioinformatics/MedlarGSBK13}
-uses GPUs to analyse parental linkage patterns in a biology context and
+We consider related work for both parts of this paper: acceleration and scaling
+of our MCMC-MMSB algorithm. Acceleration of deep learning algorithms on GPUs
+is an ongoing success story. Although approximative Baysian algorithms are
+not such a natural fit for GPUs, a number of projects explores this terrain.
+E.g.~\cite{journals/bioinformatics/MedlarGSBK13} uses GPUs with their
+MCMC approach to analyse parental linkage patterns in a biology context and
 \cite{DBLP:journals/csda/WhiteP14} does the same to model terrorist activity.
-Variational inference on MMSB is described
-in~\cite{DBLP:conf/nips/GopalanMGFB12}, but they don't use GPUs. Latent
-Dirichlet Allocation, another variety of Bayesian Approximation, is used on
-GPUs by~\cite{DBLP:conf/nips/YanXQ09}.
-
-Still, for high-dimension problems, the standard MCMC algorithms fall short in
-performance. Recently, varieties of MCMC algorithms have been developed where
-gradient information is used to speed up convergence. Langevin and Hamiltonian
-dynamics are representatives of these
-varieties~\cite{Girolami_riemannmanifold}.  Our algorithm uses Riemann Manifold
-Langevin dynamics. In \cite{beam2014fast}, GPUs are used to perform Hamiltonian
-descent, using Python interfaces to access the standard cuBLAS
-library~\cite{cuBLAS}. They limit their optimizations for the GPU to
-saving on data transfers between host and device memory. They do not apply
-their framework to MMSB.
+Latent Dirichlet Allocation, another variety of Bayesian Approximation,
+is used on GPUs by~\cite{DBLP:conf/nips/YanXQ09}. There is also related
+work on the variants of MCMC algorithms that use the gradient to speed
+up convergence. Langevin and Hamiltonian dynamics are representatives of
+these varieties~\cite{Girolami_riemannmanifold}. Our algorithm uses Riemann
+Manifold Langevin dynamics. In \cite{beam2014fast}, GPUs are used to perform
+Hamiltonian descent, using Python interfaces to access the standard cuBLAS
+library~\cite{cuBLAS}. They limit their optimizations for the GPU to saving
+on data transfers between host and device memory. They do not apply their
+framework to MMSB.
 
@@
 can handle are limited by the device memory of the GPU. Our implementation converges
-quickly on the GPU but can also function with reduced speedup on a multicore
-CPU. In the latter case, the dataset
-size is only limited by the memory of the host machine.
+quickly on the GPU but can also use multicore CPUs. In the latter case, speedup is
+limited but the larger memory allows tackling larger datasets.
+
+\cite{liakos-bigdata16} also addresses overlapping community detection, but their
+algorithm uses a local dispersion-aware approach. They find communities for 5000
+nodes in each of the large SNAP datasets, see Table~\ref{table-snap}.
 
@@
 
-There are a few projects that did a distributed implementation
+There are a few projects that implemented a distributed algorithm
 for community detection on very large network graphs. In contrast to our work,
@@
 stochastic blockmodels; \cite{Prat-Perez:2014:HQS:2566486.2568010} use
-multi-core machines.
+multi-core machines. The results of our distributed MCMC-aMMSB implementation
+have been previously published at the 2016 ParLearning
+workshop~\cite{DBLP:conf/ipps/El-HelwHLAWB16}, where it earned a best-paper
+award.
 
+\textbf{\emph{QUESTION: Re-insert related work on RDMA KV stores?}}
 \begin{comment}

### 2017-05-18 distr-03-design.tex
Polish a bit on GPU plots

@@
 largest distributed experiment, com-Friendster with $N$=64M and $K$=12K, each
-requires~3TB. We decided to store only $\pi$ and $\sum\phi$, and recalculate
+requires~3TB. Analogously to the single-machine optimization described in
+Section~\ref{sec-gpu}, we store only $\pi$ and
+$\sum\phi$, and recalculate
 $\phi$ from these whenever appropriate. Here, we trade a substantial gain
@@
 %\looseness=-1
-These updates are used by \textit{update\_pi} (line 7 in
+These updates are used by \textit{update\_pi} (line 9 in
 Algorithm~\ref{algorithm}) locally, besides the value of
@@
 memory consistency, this stage awaits completion of \textit{update\_phi} with
-an MPI barrier. After the calculation, the updated values of $\pi+\sum\phi$
+an MPI barrier. After the calculation, the updated values of $\pi,~\sum\phi$
 are written to the DKV store.
 
-\textit{update\_beta} (lines 10 and 11 in Algorithm~\ref{algorithm}) requires $|\Minibatch| \times K$ operations. It is preceded by an MPI
+\textit{update\_beta} (lines 12 and 15 in Algorithm~\ref{algorithm}) requires $|\Minibatch| \times K$ operations. It is preceded by an MPI
 barrier to ensure that up-to-date $\pi$ values are read. \textit{update\_beta}

### 2017-05-18 distr-04-evaluation.tex
Gone over figures etc. Still need to redo related work.
Polish a bit on GPU plots

@@
 
-This section presents an in-depth empirical analysis of the design discussed in
-Section~\ref{design-section}. To this end, we assess the scalability and
+This section presents an in-depth empirical analysis of our distributed design.
+To this end, we assess the scalability and
 efficiency of the solution using multiple criteria.
@@
 
-Empirical results for the distributed implementation were obtained from experiments on the VU and
-Leiden University
-DAS5 clusters which consist of 68 and 24 compute nodes respectively. Each
-compute node is equipped with a dual 8-core Intel Xeon E5-2630v3 CPU clocked
-at 2.40GHz, 64GB of memory and 8TB of storage. Moreover, the compute nodes of
-each site are interconnected by
-FDR~InfiniBand. The MPI implementation used is OpenMPI, which supports
+As noted in Section~\ref{sec-experimental}, experimental results
+were obtained using the DAS5 clusters.
+The MPI implementation used is OpenMPI, which supports
 user-space Infiniband~\cite{gabriel04:_open_mpi}.
 
-We compared the distributed performance against a shared-memory virtual machine
+Moreover, we compared the distributed performance against a shared-memory virtual machine
 at SURFsara's HPC Cloud system with 40~cores and 1TB of memory; its underlying
@@
 
-\begin{figure}[t] % [htbp]
+\begin{figure}[tb] % [htbp]
   \centering
@@
 
-\begin{figure}[t] % [htb]
+\begin{figure}[tb] % [htb]
   \centering
@@
 
-\begin{table}[t] % [htbp]
+\begin{table}[tb] % [htbp]
   \centering
@@
 \subsubsection{Horizontal vs. Vertical Scalability}
-\begin{figure}[t] % [htb]
+\begin{figure}[tb] % [htb]
   \centering

### 2017-05-18 gpu-04-02-sequential.tex
Polish a bit on GPU plots

@@
-\subsection{An Efficient Sequential Baseline Version}
-\label{sec-design-sequential}
-
-Initially, the algorithm was implemented in Python~\cite{LiAW15}, as is common
+The MCMC-aMMSB algorithm that is the subject of this paper was initially
+implemented in Python~\cite{LiAW15}, as is common
 with machine learning researchers and practitioners. It
@@
 overhead associated with Python's interpreter. However, the algorithm also
-depended on native Python data structures such as sets, dictionaries and
+depended on Python data structures such as sets, dictionaries and
 lists, which are far less efficient.
-This section presents how the Python implementation was
+This section describes how the Python implementation was
 converted into an efficient sequential C++ application which serves as a
-baseline for the parallel implementations.
+baseline for the parallel and distributed implementations.
 
-Initially, the Python code base was ported into equivalent C++ code.
+First, the Python code base was ported into equivalent C++ code.
 The program structure remained unchanged and the Python data structures were

### 2017-05-18 gpu-04-03-accelerator.tex
Polish a bit on GPU plots

@@
-\subsection{Restructuring for Parallelism}
+% \section{Accelerating on one Many-core Machine}
 
@@
 % Naive strategy
-\paragraph*{\textbf{The naive strategy}} simply allocate temporary vectors in thread
-local memory to store $Probs$ and $Grads$ which physically resides in
+\paragraph*{\textbf{Naive strategy}} simply allocate temporary vectors in thread
+local memory to store $Probs$ and $Grads$ which physically reside in
 device memory. All memory accesses are coalesced to achieve the highest
@@
 % Shared memory caching strategy
-\paragraph*{\textbf{The shared memory strategy}} allocate the temporary vectors
+\paragraph*{\textbf{Shared memory strategy}} allocate the temporary vectors
 $Probs_i$ and $Grads_i$
@@
 % Register and shared memory using code generation and loop unrolling 
-\paragraph*{\textbf{The code generation strategy}} dynamically generate the code of the kernel to
+\paragraph*{\textbf{Code generation strategy}} dynamically generate the code of the kernel to
 custom tailor its properties. For example, this strategy can control whether a

### 2017-05-18 gpu-04-design.tex
Place the chunks of text unmodified into one document
Start working on integration: block out double algorithm, unify symbol usage
Some nitpicks
Gone over figures etc. Still need to redo related work.
Polish a bit on GPU plots

@@
-\section{System Design and Implementation}
+% \subsection{System Design and Implementation}
 
-The process of designing an accelerated version of the SG-MCMC MMSB algorithm
-involved multiple transformations. First, an efficient C++ baseline version
-was created. Further, changes to the algorithm and the data structures
+Based on the efficient C++ baseline version,
+changes to the algorithm and the data structures
 were carried out, for both efficient resource utilization and parallelization.
@@
 phases, identifying the key contributions and differences between consecutive
-states.
+states. This section concludes with an in-depth evaluation of the performance
+of our various optimizations, on multi-core CPUs as well as a number of GPU
+accelerators.
 
-\input{04-02-sequential}
+% \input{gpu-04-02-sequential}
 
-\input{04-03-accelerator}
+\input{gpu-04-03-accelerator}

### 2017-05-18 gpu-05-02-sequential.tex
Gone over figures etc. Still need to redo related work.
Polish a bit on GPU plots

@@
-\subsubsection{Achieving an Efficient Baseline Sequential Version}
+\subsection{Performance of Sequential Versions}
 
@@
 %
-Namely, we use the CA-HepPh dataset; the number of
+Namely, we use the small CA-HepPh dataset from the SNAP graph collection, see
+Table~\ref{table-snap}; the number of
 iterations is 1000.
@@
 
-\begin{table}[b]
+\begin{table}[tb]
 \center\begin{tabular}{l d{5.1} d{2.1} d{2.1} d{2.1}}
@@
 \end{tabular}
-\caption{Performance comparison between Python, C++ that follows the Python
-idioms, and our baseline C++ implementations. 1000 iterations, times in
+\caption{Performance comparison between Python; C++ that follows the Python
+idioms; and our baseline C++ implementations. 1000 iterations, times in
 \textrm{ms} per iteration; *perplexity time divided by 100 iterations.}
@@
 
+A first multi-threaded version was created by annotating the
+baseline implementation with OpenMP primitives in a straightforward
+fashion. This accelerated the stages \textit{update\_phi} and
+\textit{update\_pi} with a factor~4 for the very small problem instance
+in Table~\ref{table-python-comparison}. For larger problem instances, the
+speedup for these dominant phases grew to a factor~10. The vectorizing OpenCL
+backend described in the next Section again significantly improves the
+performance, but at the cost of substantial coding effort.
+\textbf{\emph{Question: speak about the other stages? they are pipelined.}}
+
 \begin{comment}

### 2017-05-18 gpu-05-evaluation.tex
Polish a bit on GPU plots

--- 2017-05-04 gpu-05-evaluation.tex
@@
 
-The following subsections discuss the performance evaluation of the algorithm
-across its different stages. As the system underwent multiple design mutations,
-a careful analysis is needed to identify the contributions of each change
-independently.
-%
-First, an examination is provided detailing the construction of an efficient
-sequential baseline version.
-This spans
-the code conversion from Python to C++ as well as the effects of applying
-the localized optimizations discussed in
-Section~\ref{sec-design-sequential}.
-%
-Second, we explore the performance benefits of parallelizing the computations
+The following sections provide a detailed experimental performance analysis
+of the optimizations described above. We start by discussing the performance
+effects of parallelizing the computations
 on a multi-core CPU using OpenCL.
@@
 Finally, we present the overall outcome of accelerating the algorithm by
-displaying its fast convergence over large networks.
+displaying its fast convergence over the largest networks that fit in the
+memory of one GPU or CPU.
 
-All experiments were conducted on the VU Amsterdam DAS5 cluster. The cluster
-consists of 68 compute nodes each equipped with a dual 8-core Intel Xeon
-E5-2630v3 CPU clocked at 2.40GHz, 64GB of memory and 8TB of storage.
-Additionally, the cluster is fitted with a number of Nvidia GPUs including GTX
-TitanX, GTX980, K40c and K20m; see Table~\ref{table-gpus} for an overview of
-their properties. Table~\ref{table-snap} outlines the graph
-properties of the networks used to evaluate the algorithm's performance. These
-networks were obtained from the SNAP collection~\cite{snapnets}.
+As noted in Section~\ref{sec-experimental}, all experiments were conducted on the VU Amsterdam DAS5 cluster.
+Table~\ref{table-gpus} presents the properties of the various GPUs in
+the DAS5 cluster.
 
-\begin{comment}
-\begin{table}[tb]
-\center\begin{tabular}{l r r c}
-Network & \#Vertices & \#Edges \\
-\hline
-CA-HepPh        &    12,008 &    118,521 \\
-com-DBLP        &   317,080 &  1,049,866 \\
-% com-Youtube     & 1,134,890 &  2,987,624 \\
-com-LiveJournal & 3,997,962 & 34,681,189 \\
-\hline
-\\[-1ex]
-\end{tabular}
-\caption{Network graphs from the Stanford Snap collection}
-\label{table-snap}
-\end{table}
-\end{comment}
-
-\begin{table*}
-  \centering
-  \begin{tabular}{l r r r l}
-    Name            & \#Vertices &       \#Edges & \#Ground-truth & Description \\
-                    &            &               & communities    &             \\
-    \hline
-    CA-HepPh        &    12,008  &    118,521    & n/a            & Paper citation network \\
-    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
-    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
-    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
-    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
-    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography collaboration network \\
-    com-Amazon      &    334,863 &       925,872 & 75,149         & Product co-purchasing network \\
-    \hline
-    \\[-1ex]
-  \end{tabular}
-  \caption{Summary of SNAP graph data sets used for evaluation.}
-  \label{table-snap}
-\end{table*}
-
-\input{gpu-05-02-sequential}
+% \input{gpu-05-02-sequential}
 \input{gpu-05-03-accelerator}

### 2017-05-18 plots/plot-gpus.sh

@@ -11,7 +11,8 @@ X=1
 for i in GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m Intel-R-Xeon; do
   v=`grep -a TOTAL dblp-K1024/${i}* | sort -nk 7 | head -n 1 | awk '{print $7}'`
   v2=`perl -e "print $BASELINE / $v"`
-  echo $X $i $v2 >> $TFILE
+  device=`echo $i | sed 's/GeForceGTXTITANX/Titan-X/;s/GeForceGTX980/GTX980/;s/TeslaK40c/K40c/;s/TeslaK20m/K20m/;s/Intel-R-Xeon/Xeon/'`
+  echo $X $device $v2 >> $TFILE
   X=`expr $X + 1`
 done
 
@@ -19,7 +20,7 @@ cat $TFILE
 
 cat <<EOF | gnuplot --persist
 
-set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+set terminal postscript eps enhanced color font ',10' size 3.3,1.6
 set output 'gpus.eps'
 set colorsequence classic
 

### 2017-05-18 plots/plot-vector-sweep.sh

@@ -8,7 +8,7 @@ TFILE="$$.tmp"
 ./collect-v.sh > $TFILE
 
 cat <<EOF | gnuplot --persist
-set terminal postscript eps enhanced color font ',8' size 3.3,1.6
+set terminal postscript eps enhanced color font ',9' size 3.3,1.6
 set colorsequence classic
 set output '$NAME'
 
@@ -17,13 +17,14 @@ set style histogram cluster gap 1
 
 set xlabel 'GPU'
 set ylabel 'Time (normalized over non-vectorized version)'
-set key below
+# set key below
 
 set style fill solid noborder
 set auto x
-set ytics
-set mytics
-set grid ytics mytics
+set ytics 0,0.2
+# set mytics
+# set grid ytics mytics
+set grid ytics
 
 set yrange [0:*]
 plot \

### 2017-05-18 plots/plot-vector-sweep.sh
Polish a bit on GPU plots

@@ -15,7 +15,7 @@ set output '$NAME'
 set style data histogram
 set style histogram cluster gap 1
 
-set xlabel 'GPU'
+# set xlabel 'GPU'
 set ylabel 'Time (normalized over non-vectorized version)'
 # set key below
 

### 2017-05-18 04-experimental-setup.tex
Refactor experimental setup into separate section

\section{Experimental setup}
\label{sec-experimental}

Unless indicated otherwise, the experiments described in the next sections
were conducted on the VU~Amsterdam and Leiden DAS5 clusters.  The clusters
consists of 68 and 24 compute nodes respectively. Each node is equipped with
a dual 8-core Intel Xeon E5-2630v3 CPU clocked at 2.40GHz, 64GB of memory and
8TB of storage. The nodes of each site are interconnected by FDR~InfiniBand.
Additionally, the VU cluster is fitted with a number of Nvidia GPUs including
GTX~\mbox{Titan-X}, GTX980, K40c and K20m.

To assess the performance of our implementations, we use publicly available
community networks from the SNAP collection~\cite{snapnets}, including the largest
networks. Table~\ref{table-snap}
outlines the graph properties of the networks used.

\begin{table}
  \centering
  \def\tabcolsep{0.2em}
  \begin{tabular}{l r r r p{9em}}
    Name            & \#Vertices &       \#Edges & \multicolumn{1}{c}{\#Ground-} & Description \\
                    &            &               & \multicolumn{1}{c}{truths}    &             \\
    \hline
    CA-HepPh        &    12,008  &    118,521    & \multicolumn{1}{c}{n/a}       & Paper citation network \\
    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography network \\
    com-Amazon      &    334,863 &       925,872 & 75,149         & Product copurchasing network \\
    \hline
    \\[-1ex]
  \end{tabular}
  \caption{Summary of SNAP graph data sets used for evaluation.}
  \label{table-snap}
\end{table}

### 2017-05-23 gpu-04-02-sequential.tex
Move exp. setup into GPU section

@@
 an efficient sequential baseline version, which is 1000..1500 times faster
-than the original Python code, see Section~\ref{sec-evaluation}.
+than the original Python code, see the next Section.
 All transformations were rigorously tested to ensure they do

### 2017-05-23 01-intro.tex
Work on Henri's comments for 1. Introduction

@@
 In this paper, we focus on the SG-MCMC method applied to the a-MMSB model. Recent
-work~\cite{LiAW15} showed that this methodology is faster and more accurate than the
-SVB method. \cite{LiAW15} further proposes a heuristic to scale up the number of
+work of the authors of this paper~\cite{LiAW15} showed that this methodology is faster and more accurate than the
+SVB method. That work further proposes a heuristic to scale up the number of
 communities at the cost of less precision; the work in this paper considers the
@@
 From a computational point of view, our SG-MCMC a-MMSB algorithm differs
-from widespread machine learning algorithms, like deep learning, in several
-ways. First, it is highly data-intensive which makes parallel acceleration
+from widespread machine learning algorithms like deep learning in several
+ways. First, it is \emph{highly data-intensive} which makes parallel acceleration
 particularly challenging. Second, owing to the algorithm's stochastic
@@
 
-Starting from an existing implementation in Python/Numpy, we will present
-how this algorithm can be implemented much more efficiently in C++. Then,
-we will describe our two key contributions: first, how to scale up our
-implementation to run on GPUs and multicore CPUs, second, to scale out
-our implementation to overcome the memory bounds posed by GPUs or
-single-machine systems in general.
-% To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem.
-In particular, we use a mixture of OpenMP, MPI and RDMA in order to efficiently scale and accelerate the algorithm's computation.
-
-The focus of the first part of our work is on the computational efficiency and parallel
-performance of the SG-MCMC algorithm by
-developing aggressive optimizations targeting
-multi-core CPUs and GPUs. The GPU implementation achieves speedup factors up to~86,
-compared to the
-well-tuned sequential C++ program which itself is a factor~1000-1500 faster than
-the original Python/Numpy program.
+Starting from an existing implementation in Python/Numpy, we create an efficient
+baseline version in C++ to serve as the baseline for the parallel versions.
+This already improves performance by a factor of 1000 to 1500 compared to
+Python. Then, our key contributions are:
+\begin{itemize}
+\item
+our implementation is scaled up to run on GPUs.
+We developed aggressive optimizations to solve the memory bandwidth issues for
+GPUs. This implementation achieves speedup factors up to~86,
+compared to the baseline.
 %
@@
 state significantly reduces the data intensity and allows for tackling larger
-problems. % while maintaining all state in memory.
+problems; % while maintaining all state in memory.
+
+\item
+our implementation is scaled out to overcome the memory bounds posed by GPUs or
+single-machine systems in general, to solve the largest publicly available
+community graphs. % This necessitated overcoming several challenges.
+% The algorithm's state grows rapidly with larger graphs and number of latent
+% communities.
+Since the full state for the larger graphs does not fit in a
+single machine's memory, it is partitioned across a cluster
+of machines. Hence, the cluster nodes must read
+remote memory hosted by their peers. Since the algorithm is memory-bound, we
+substantially optimized remote memory accesses by implementing
+a custom RDMA store.
+On top of that, we pipelined the algorithm stages by fetching
+data in advance over the network;
+\begin{comment} Finally, the algorithm's computation
+is effectively distributed across the cluster nodes and parallelized further
+within each node by exploiting their multi-core CPUs.
+\end{comment}
+
+\item we argue that it is valid to use a CPU version with straightforward
+intra-node thread parallelism for the distribibuted implementation, and
+that it is pointless to combine the GPU work into it.
+
+\end{itemize}
+% To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem.
+In particular, we use a mixture of OpenMP, MPI and RDMA in order to efficiently scale and accelerate the algorithm's computation.
 
@@
 
-The second part of our work, scaling out to the largest publicly available
-community graphs, also necessitated overcoming several challenges.
-The algorithm's state grows rapidly with larger graphs and number of latent
-communities.  Since the full state for the larger graphs does not fit in a
-single machine's memory, it is partitioned and distributed across a cluster
-of machines. Hence, to access the full state, each cluster node must read
-remote memory hosted by its peers. We leveraged RDMA to improve the latency and
-bandwidth
-of such operations. To reduce the
-latency further, we pipelined the algorithm's computations by fetching
-data in advance over the network.
-\begin{comment} Finally, the algorithm's computation
-is effectively distributed across the cluster nodes and parallelized further
-within each node by exploiting their multi-core CPUs.
-\end{comment}
+A bird's eye-view of the work in this paper, without any technical detail, has been
+published at CCGrid-2016~\cite{10.1109/CCGrid.2016.98}.
+The results of our distributed MCMC-aMMSB implementation
+have been previously published at the 2016 ParLearning
+workshop~\cite{DBLP:conf/ipps/El-HelwHLAWB16}, where it earned a best-paper
+award.
 
@@
 algorithm and its theoretical foundation, with related work in
-Section~\ref{sec-related}. Section~\ref{sec-gpu} delves
+Section~\ref{sec-related}. Section~\ref{sec-design-sequential} describes how we
+arrived at an efficient sequential version that serves as the baseline for the
+parallel implementations. Section~\ref{sec-gpu} delves
 into the parallelization optimizations for GPUs and multicore CPUs.

### 2017-05-24 01-intro.tex
Remove deprecated files. Do an aspell check

@@
 \item we argue that it is valid to use a CPU version with straightforward
-intra-node thread parallelism for the distribibuted implementation, and
+intra-node thread parallelism for the distributed implementation, and
 that it is pointless to combine the GPU work into it.
@@
 parallel implementations. Section~\ref{sec-gpu} delves
-into the parallelization optimizations for GPUs and multicore CPUs.
+into the parallelization optimizations for GPUs and multi-core CPUs.
 Section~\ref{sec-distr} describes the design, implementation and evaluation of

### 2017-05-24 02-background.tex
Remove deprecated files. Do an aspell check

@@
 One benefit of using SGLD is that we do not need the Metropolis-Hastings (MH) accept-reject tests since the rejection probability goes to zero as the step size tends to zero.
-Although in practice we often do not reduce the stepsize to zero to obtain better mixing (thus resulting in some bias), we obtain good performance by drawing many more samples per unit time.
+Although in practice we often do not reduce the step size to zero to obtain better mixing (thus resulting in some bias), we obtain good performance by drawing many more samples per unit time.
 

### 2017-05-24 06-conclusion.tex
Remove deprecated files. Do an aspell check

@@
 
-In our optimization for manycore processors, we navigated the complex optimization
+In our optimization for many-core processors, we navigated the complex optimization
 landscape by dynamically generating kernel code and testing different

### 2017-05-24 distr-03-design.tex
Remove deprecated files. Do an aspell check

@@
 \begin{comment}
-- distributed parallelisation:
+- distributed parallelization:
   cut up into stages like the CC-Grid paper:
@@
 This section describes how we created a distributed version of the C++
-algorithm. We didn't use the advanced intra-node parallelisation techniques
+algorithm. We didn't use the advanced intra-node parallelization techniques
 described in the previous section, but found that straightforward
@@
 and~$\phi$, see Table~\ref{tbl-symbols}.
-For the largest dataset in this paper, com-Friendster, \Edges has 1.8~billion
+For the largest data set in this paper, com-Friendster, \Edges has 1.8~billion
 undirected edges. In our representation with directed edges, this takes

### 2017-05-24 distr-04-evaluation.tex
Remove deprecated files. Do an aspell check

@@
 All of the experiments reported in this section utilized
-publicly available graphs from the Stanford Large Network Dataset
+publicly available graphs from the Stanford Large Network Data set
 Collection~(SNAP) \cite{snapnets}. Table~\ref{table-snap} lists the
@@
 column has pipelining enabled. Within \textit{update\_phi},
-loading $\pi$ from the DKV store is dominant.
+loading $\pi$ from the D-KV store is dominant.
 The pipelining optimization significantly overlaps loading $\pi$ both with generation and
@@
   \epsfig{file=plots/qperf.eps, width=\columnwidth}
-  \caption{Performance of DKV store vs.\ \textit{qperf} and MPI roundtrip.}
+  \caption{Performance of D-KV store vs.\ \textit{qperf} and MPI round-trip.}
   \label{fig-qperf}
@@
 read operations between one server and one client for a range of payload sizes,
-and compare these to the bandwidth achieved by an MPI roundtrip test,
+and compare these to the bandwidth achieved by an MPI round-trip test,
 and by \textit{qperf}~\cite{qperf-mellanox} which reports
@@
 lopsided: \textit{qperf} and MPI are low-level point-to-point benchmarks without
-any DKV store implementation on top.
+any D-KV store implementation on top.
 The performance presented in this Figure
 is ample justification for using a low-level RDMA
-implementation for the DKV store.
-The bandwidth achieved by our DKV store falls short
+implementation for the D-KV store.
+The bandwidth achieved by our D-KV store falls short
 of the \textit{qperf} performance for packets less than 4KB; this is attributed
-to additional per-request overhead for the DKV store. For the largest packet
-size, the DKV store performance is hampered by the fact that its values
+to additional per-request overhead for the D-KV store. For the largest packet
+size, the D-KV store performance is hampered by the fact that its values
 are spread over a larger memory area, whereas \textit{qperf} and MPI always read
-from the same memory locations. For packets between 8KB and 512KB, our DKV
+from the same memory locations. For packets between 8KB and 512KB, our D-KV
 store achieves performance very close to \textit{qperf}.
@@
 %################################################
-\subsubsection{Convergence of Large Datasets}
+\subsubsection{Convergence of Large Data sets}
 

### 2017-05-24 gpu-05-03-accelerator.tex
Gone over figures etc. Still need to redo related work.
Remove deprecated files. Do an aspell check

@@
 \begin{comment}
-  \begin{table}[t]
+  \begin{table}[tb]
     \centering
@@
 
-\begin{table}[b]	% [htb]
+\begin{table}[tb]	% [htb]
   \centering
@@
   \caption{Performance break-down of multi-core CPU version without
-  vectorization. Model parameters: dataset com-DBLP; K=1024, m=4096, n=32.}
+  vectorization. Model parameters: data set com-DBLP; K=1024, m=4096, n=32.}
   \label{TABLE-CPU}
@@
 
-\begin{table}[b]	% [htb]
+\begin{table}[tb]	% [htb]
 % \center\begin{tabular}{l c c c c c c c c d{3.1}}
@@
 % TITANX
-\begin{figure*}[t]	% [htb]
+\begin{figure*}[tb]	% [htb]
   \centering
@@
 % TITANX K=2K
-\begin{figure*}[t]	% [htb]
+\begin{figure*}[tb]	% [htb]
   \centering
@@
 The wide performance range inhibits the readability of the figure,
-therefore, a more focused Figure~\ref{titanx-w1-sweep}(b) is provided. As would
+therefore, a more focused Figure~\ref{gpu-sweep}(a), bottom figure, is provided. As would
 be expected, the naive strategy exhibits the worst performance over all
@@
 % K40
-\begin{figure*}[t]	% [hbt]
+\begin{figure*}[tb]	% [hbt]
   \centering
@@
 
-\begin{figure}[b]	% [htb]
+\begin{figure}[tb]	% [htb]
   \centering

### 2017-05-24 01-intro.tex
Merged cleanup from master

@@
 Starting from an existing implementation in Python/Numpy, we create an efficient
-baseline version in C++ to serve as the baseline for the parallel versions.
+C++ version to serve as the baseline for the parallel versions.
 This already improves performance by a factor of 1000 to 1500 compared to
-Python. Then, our key contributions are:
+Python. Then, our key contributions to the algorithm are:
+
 \begin{itemize}
-\item
-our implementation is scaled up to run on GPUs.
-We developed aggressive optimizations to solve the memory bandwidth issues for
-GPUs. This implementation achieves speedup factors up to~86,
-compared to the baseline.
+\item \textbf{scaling up to run on GPUs.}
+The application is not a natural fit for GPUs, because it is memory-bound and
+irregular in its data accesses. Nevertheless, we developed aggressive
+memory allocation optimizations to solve the data bandwidth issues for GPUs.
+This implementation achieves up to 86$\times$ speedup compared to the baseline.
 %
-Through careful analysis of the computation and data structures we show that
-the algorithm's full state can be reduced by roughly 75\%. Compressing the
-state significantly reduces the data intensity and allows for tackling larger
+We managed to reduce the algorithm's memory requirements by roughly 75\%, which
+significantly reduces the data intensity and allows for tackling larger
 problems; % while maintaining all state in memory.
 
-\item
-our implementation is scaled out to overcome the memory bounds posed by GPUs or
-single-machine systems in general, to solve the largest publicly available
+\item \textbf{scaling out to overcome the memory bounds posed by GPUs} or,
+generally, single-machine systems, to solve the largest publicly available
 community graphs. % This necessitated overcoming several challenges.
@@
 
-\item we argue that it is valid to use a CPU version with straightforward
-intra-node thread parallelism for the distributed implementation, and
-that it is pointless to combine the GPU work into it.
+\item \textbf{the use of pipelining} allows to use straightforward intra-node
+thread parallelism in the distributed implementation, and it is pointless
+to combine the GPU work into it. The cause is that the computation can be
+fully overlapped with fulfilling the data requests for the next iteration.
 
@@
 parallel implementations. Section~\ref{sec-gpu} delves
-into the parallelization optimizations for GPUs and multi-core CPUs.
+into the parallelization optimizations for GPUs.
 Section~\ref{sec-distr} describes the design, implementation and evaluation of

### 2017-05-24 gpu-05-02-sequential.tex
Remove deprecated files. Do an aspell check
Begin compacting python->C++

@@
 
-This subsection presents the results of transforming the algorithm implementation
-from Python to C++, and then removing obvious inefficiencies, to achieve a
-good baseline implementation for comparison with the parallel implementation.
+This subsection presents the results of transforming the algorithm
+implementation from Python to efficient C++.
+Our experiments use a DAS5 compute node, see Section~\ref{sec-experimental}.
 %
@@
 32-bit floating point numbers in stead of 64-bit doubles.
-This hardly reduces the computation intensity for the Xeon, but it very much
-affects data intensity because the data sizes are halved.
+This hardly reduces the computation intensity for a DAS5 Xeon machine, but
+it certainly affects data intensity because the data sizes are halved.
 %
@@
 Due to the performance limitations of the Python implementation, the
-experimental configuration of this section is minimal.
+experimental configuration of this section is very small.
 %
-Namely, we use the small CA-HepPh dataset from the SNAP graph collection, see
-Table~\ref{table-snap}; the number of
-iterations is 1000.
+We use the CA-HepPh dataset from the SNAP graph collection, see
+Table~\ref{table-snap} in Section~\ref{sec-experimental}. The number of
+iterations is~1000;
 %Note that these trial runs stop very far before convergence; using a small
@@
 %during the lifetime of the algorithm.
-The selected number of communities is~1024, the mini-batch size~32, the
-neighbor sample size~32.
+The number of communities is $K=1024$, the mini-batch size is fixed at $M=32$, the
+neighbor sample size $\Neighbors=32$.
 
+\begin{comment}
+******** Unnecessary detail?
 The mini-batch sampling as described in Section~\ref{sec-background} randomly
@@
 is fully consistent.
+\end{comment}
 
@@
 \caption{Performance comparison between Python; C++ that follows the Python
-idioms; and our baseline C++ implementations. 1000 iterations, times in
+idioms; and our baseline C++ implementations. Times in
 \textrm{ms} per iteration; *perplexity time divided by 100 iterations.}
@@
 We present timings for the compute stages of the algorithm described in
-Section~\ref{sec-background}, with the exception of
+Algorithm~\ref{algorithm}, with the exception of
 \textit{update\_theta\_beta} which combines \textit{update\_theta}
@@
 %
-The numbers are times in ms. For all stages except perplexity, the times
-are per iteration.
+% The numbers are times in ms. For all stages except perplexity, the times
+% are per iteration.
 %
-Once every so many iterations (hundreds or thousands in production runs),
-the perplexity calculation is invoked. The perplexity time in the table
-has been divided by 100 iterations; per perplexity invocation, it is large in
-comparison with the iteration stage times, but it is amortized over those
-iterations.
+% Once every so many iterations (hundreds or thousands in production runs),
+% the perplexity calculation is invoked. The perplexity time in the table
+% has been divided by 100 iterations; per perplexity invocation, it is large in
+% comparison with the iteration stage times, but it is amortized over those
+% iterations.
+In production runs, the perplexity calculation is invoked once every hunderds
+or thousands iterations. Hence, the perplexity timings are amortized over
+100~iterations.
 
@@
 The translation from Python into 'Python Idiom' C++ speeds up this stage
-by a factor~171, even though the Python implementation uses numpy for its data
-structures. Comparison of the C++ table entries for 'Python idiom' and
-64-bit float baseline shows that our C++ optimizations speed up this stage
+by a factor~171. % , even though the Python implementation uses numpy for its data
+% structures. 
+Our C++ optimizations speed up this stage
 by another factor of~6. Reducing the floating-point precision to 32~bits
-doubles this factor, which shows that the algorithm is very much data-dominant.
+doubles this factor, which again shows that the algorithm is very much data-dominant.
 %
 \textit{update\_pi} has a disproportionately large speed
-gain after removing inefficiencies. This was attained by limiting the
-update to $\pi$ to only those values of $\pi$ that have changed in this
-iteration, in stead of the full $\pi$ array (which, incidentally, is handled
-rather efficiently by numpy).
+gain after removing inefficiencies. The numpy implementation updated the full
+$\pi$ array (incidentally, rather efficiently), whereas our C++ optimization
+limits the updates to only those values of $\pi$ that have changed.
+%
 \textit{update\_theta\_beta} gave fewer opportunities for C++ optimization,
@@
 
-A first multi-threaded version was created by annotating the
+\subsection{Multi-threading}
+
+A multi-threaded version was created by annotating loops in the
 baseline implementation with OpenMP primitives in a straightforward

### 2017-05-24 gpu-04-02-sequential.tex

@@
 implemented in Python~\cite{LiAW15}, as is common
-with machine learning researchers and practitioners. It
-relied on numpy~\cite{numpy} to perform the numerical computations without the
-overhead associated with Python's interpreter. However, the algorithm also
-depended on Python data structures such as sets, dictionaries and
-lists, which are far less efficient.
-This section describes how the Python implementation was
-converted into an efficient sequential C++ application which serves as a
-baseline for the parallel and distributed implementations.
+with machine learning researchers and practitioners.
+We converted this implementation into an efficient sequential C++ application
+to serve as a baseline for the parallel and distributed implementations.
+This is certainly not one of our contributions to the parallelization
+of the algorithm, but for a fair analysis of the parallel implementations
+a good baseline is necessary.
+
+The Python implementation
+relied on numpy~\cite{numpy} to perform numerical computations efficiently.
+Besides, the algorithm
+made heavy use of Python data structures such as sets, dictionaries and
+lists, which have no efficient numpy equivalent.
 
@@
 The program structure remained unchanged and the Python data structures were
-simply substituted with their corresponding alternatives in C++.
-
-The next step was to remove a number of inefficiencies in the C++
-implementation. For example, one recurring idiom in the Python implementation
-was an expression of the form $a^y b^{1-y}$, implemented with floating-point
-exponentiation, where $y$ is a variable that can
-only hold values of 0 or~1. Therefore, the expression is simply a choice
-between $a$ for $y=1$ and $b$ for $y=0$. However, as the compiler cannot infer
-the value of $y$, both expressions $a$ and $b$ were computed and their
-resulting values were raised to the power $y$ and $1-y$ respectively.
+substituted with their direct C++ alternatives.
+%
+The next step was to remove a number of inefficiencies.
+For example, one recurring idiom in the Python implementation
+was an expression of the form $a^y b^{1-y}$, where $y$ is either 0 or~1.
+% Therefore, the expression is simply a choice
+% between $a$ for $y=1$ and $b$ for $y=0$. However, as the compiler cannot infer
+% the value of $y$, both expressions $a$ and $b$ were computed and their
+% resulting values were raised to the power $y$ and $1-y$ respectively.
 We transformed such expressions into conditional expressions which
-compute either the expression $a$ or $b$ and avoid exponentiation.
+compute either $a$ or $b$, and avoid floating-point exponentiation.
 %
@@
 reduction and common subexpression lifting.
-
-In summary, a multitude of localized optimizations were applied to achieve
-an efficient sequential baseline version, which is 1000..1500 times faster
-than the original Python code, see the next Section.
+%
+% In summary, a multitude of localized optimizations were applied to achieve
+% an efficient sequential baseline version, which is 1000..1500 times faster
+% than the original Python code, see the next Section.
 All transformations were rigorously tested to ensure they do
 not modify the behavior of the algorithm.
-
+%
 \begin{comment}

### 2017-05-24 gpu-04-03-accelerator.tex
Move remark on 32-bit to sequential

@@
 Another modification to the original algorithm is migrating all of its
-computation from double to single precision floating-point.
+computation from double to single precision floating-point, analogous to the change
+for the baseline implementation in Section~\ref{sec-seq-performance}.
 %
-This change affects the algorithm in multiple ways. For instance, it reduces
+This change affects the algorithm in multiple ways. For instance, on a GPU it reduces
 both the computation and data intensity. It again reduces the memory
-footprint of the algorithm and frees registers, enabling more
-effective data reuse.
-%
-It has been previously shown that stochastic learning algorithms do not require
-high precision in the presence of statistical approximations and the addition
-of random noise~\cite{DBLP:journals/corr/GuptaAGN15}.
+footprint of the algorithm and frees registers.
 

### 2017-05-24 03-related.tex
Work on Henri's comments for 1. Introduction
Begin polishing the GPU part. On the way, start removing CPU stuff

@@
-\section{Related Work}
+\subsection{Related Work}
 \label{sec-related}
@@
 E.g.~\cite{journals/bioinformatics/MedlarGSBK13} uses GPUs with their
-MCMC approach to analyse parental linkage patterns in a biology context and
+MCMC approach to analyze parental linkage patterns in a biology context and
 \cite{DBLP:journals/csda/WhiteP14} does the same to model terrorist activity.
@@
 the Online Tensor approach~\cite{DBLP:journals/corr/HuangNHVA13}. They deduce
-overlapping communities for many of the same datasets as this paper, and they
+overlapping communities for many of the same data sets as this paper, and they
 report fast convergence. Their implementation uses the cuBLAS
@@
 fundamentally different.
-Since they target GPUs only, the datasets they
+Since they target GPUs only, the data sets they
 can handle are limited by the device memory of the GPU. Our implementation converges
-quickly on the GPU but can also use multicore CPUs. In the latter case, speedup is
-limited but the larger memory allows tackling larger datasets.
+quickly on the GPU but can also use multi-core CPUs. In the latter case, speedup is
+limited but the larger memory allows tackling larger data sets.
 
@@
 algorithm uses a local dispersion-aware approach. They find communities for 5000
-nodes in each of the large SNAP datasets, see Table~\ref{table-snap}.
+nodes in each of the large SNAP data sets, see Table~\ref{table-snap}.
 
@@
 stochastic blockmodels; \cite{Prat-Perez:2014:HQS:2566486.2568010} use
-multi-core machines. The results of our distributed MCMC-aMMSB implementation
-have been previously published at the 2016 ParLearning
-workshop~\cite{DBLP:conf/ipps/El-HelwHLAWB16}, where it earned a best-paper
-award.
+multi-core machines.
 

### 2017-05-24 gpu-04-03-accelerator.tex
Begin polishing the GPU part. On the way, start removing CPU stuff

@@
 % \section{Accelerating on one Many-core Machine}
+
+\subsection{Design of GPU Optimizations}
 
@@
 crucial modifications to allow for efficient parallelization. This section
-discusses the key design contributions to attain two optimized
-parallel versions, catering for multi-core CPUs and GPUs. First, we present
-common structural changes that provide a foundation for compute device specific
-optimizations. Next, we provide an in-depth description of the optimized CPU
-and GPU versions respectively.
+discusses the key design contributions to attain optimized
+parallel versions. The state represented by $\pi$ and $\phi$ is permanently
+stored in the GPU memory, because staging of the relevant portions of these
+data structures in each iteration is too costly compared to the amount
+of computation. The graph is stored at the host, because mini-batch
+selection is cheap, and the data required for a minibatch is small.
+
+Since the algorithm is memory-bound, many of the optimizations concern
+memory usage. We offer a fast, memory-efficient hash table taylored towards
+the algorithm's usage patterns. Some optimizations reduce the total memory
+footprint of the algorithm, following analysis of the algorithm. Other
+optimizations explore the opportunities offered by the faster regions of the
+GPU memory hierarchy: data can be put into thread-local memory, in shared
+GPU memory, or in registers which requires a more complicated analysis. One
+other optimization exploits the vectorization possibilities.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Using OpenCL and CUDA}
+% \subsection*{Using OpenCL and CUDA}
 %OpenCL/CUDA
+\textit{\bf !!! Remove the history approach !!!}
 Initially, the system was developed using OpenCL~\cite{opencl} in order to
 accelerate the computations of the algorithm. OpenCL was chosen as it provides
-a common abstraction for a variety of compute devices which fulfilled our
-requirements of employing CPUs or GPUs. However, Nvidia's OpenCL SDK limits the
+a common abstraction for a variety of compute devices.
+However, NVidia's OpenCL SDK limits the
 total memory allocations within a context to 4GB which severely limits the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Fast Lookup of Graph Edges}
+\subsubsection{Fast Lookup of Graph Edges}
 % Set/Cuckoo Hash
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Parallelization \& Data Dependencies}
+\subsubsection{Parallelization \& Data Dependencies}
 % Division into kernels
-The original algorithm was structurally reorganized into 4 main sections with
-one or more kernels in each depending on data dependencies and the applicable
-synchronization requirements.
+The original algorithm was structurally reorganized into 4 main sections, each with
+one or more kernels, data dependencies and synchronization requirements.
 %
@@
 Second, the predominant kernel, \textit{update\_phi} is invoked to calculate the
-gradients for each vertex $i$ in the mini-batch \Minibatch and updates the value of $ \phi_i
+gradients for each vertex $i$ in the mini-batch \Minibatch and update the value of $\phi_i
 \mid i \in \Minibatch$. Next, the \textit{update\_pi} kernel is invoked to normalize the
@@
 %
-Finally, the perplexity calculation is performed in a dedicated kernel.
+Finally, perplexity calculation is performed in a dedicated kernel.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Memory Footprint Reduction}
+\subsubsection{Memory Footprint Reduction}
 % Memory footprint halved
@@
 %
-This change affects the algorithm in multiple ways. For instance, on a GPU it reduces
-both the computation and data intensity. It again reduces the memory
+This change affects the algorithm in multiple ways. For instance, on some GPUs it
+reduces
+the computation intensity besides the data intensity. It again halves the memory
 footprint of the algorithm and frees registers.
@@
 
-\subsection*{CPU-Specific Optimization}
-The multi-core
-CPU version of the algorithm uses OpenCL to perform its computations. The
-work decomposition of the CPU kernels ensures that each thread performs
+\subsubsection{Edge-centric and Vertex-centric kernels}
+The work decomposition of the kernels ensures that each thread performs
 independent computation in order to avoid expensive synchronization operations.
-Edge-centric kernels, which operate over mini-batch edges perform computations
+Edge-centric kernels, which operate over mini-batch edges, perform computations
 over every edge in parallel.  Similarly, vertex-centric kernels that operate
@@
 % vectorization
-Additionally, the kernels were vectorized to decrease instruction overhead and
-utilize the SSE capabilities of the CPU cores.
+Additionally, the kernels were vectorized to decrease instruction overhead.
 
-\subsection*{GPU-Specific Optimization}
+\subsubsection{Memory Hierarchy Optimizations}
 \label{gpu-design}
-The GPU implementation builds on top of the CPU work decomposition scheme.
-However, instead of having every thread perform independent computations, each
-block of threads shares the work associated with the single edge or vertex for
-edge-centric and vertex-centric kernels.
+% The GPU implementation builds on top of the CPU work decomposition scheme.
+% However, instead of having every thread perform independent computations, each
+% block of threads shares the work associated with the single edge or vertex for
+% edge-centric and vertex-centric kernels.
 

### 2017-05-24 gpu-04-design.tex
Begin polishing the GPU part. On the way, start removing CPU stuff

@@
 
-Based on the efficient C++ baseline version,
-changes to the algorithm and the data structures
-were carried out, for both efficient resource utilization and parallelization.
+Based on the efficient C++ baseline version, a version for GPU accelaration
+was constructed.
+This involved optimizations for both efficient resource utilization and
+parallelization.
 This section provides an overview of the system's evolution in incremental
@@
 states. This section concludes with an in-depth evaluation of the performance
-of our various optimizations, on multi-core CPUs as well as a number of GPU
+of our various optimizations on a number of GPU
 accelerators.

### 2017-05-24 gpu-05-02-sequential.tex
Begin polishing the GPU part. On the way, start removing CPU stuff

@@
 \subsection{Performance of Sequential Versions}
+\label{sec-seq-performance}
 
@@
 it certainly affects data intensity because the data sizes are halved.
+It has been previously shown that stochastic learning algorithms do not require
+high precision in the presence of statistical approximations and the addition
+of random noise~\cite{DBLP:journals/corr/GuptaAGN15}.
 %
@@
 in Table~\ref{table-python-comparison}. For larger problem instances, the
-speedup for these dominant phases grew to a factor~10. The vectorizing OpenCL
-backend described in the next Section again significantly improves the
-performance, but at the cost of substantial coding effort.
+speedup for these dominant phases grew to a factor~10.
+% The vectorizing OpenCL
+% backend described in the next Section again significantly improves the
+% performance, but at the cost of substantial coding effort.
 \textbf{\emph{Question: speak about the other stages? they are pipelined.}}

### 2017-05-25 03-related.tex
Make related work a subsection of Background

@@
-\subsection{Related Work}
+\subsection{Related Parallelization Work}
 \label{sec-related}
 
-We consider related work for both parts of this paper: acceleration and scaling
-of our MCMC-MMSB algorithm. Acceleration of deep learning algorithms on GPUs
-is an ongoing success story. Although approximative Baysian algorithms are
-not such a natural fit for GPUs, a number of projects explores this terrain.
+We consider related work for both parallelization contributions of this paper: acceleration and scaling
+of our MCMC-MMSB algorithm. While acceleration of deep learning algorithms on GPUs
+is an ongoing success story, approximative Baysian algorithms are
+not such a natural fit for GPUs. Nevertheless, a number of projects explores this terrain.
 E.g.~\cite{journals/bioinformatics/MedlarGSBK13} uses GPUs with their

### 2017-05-25 03-sequential-evaluation.tex
Rename the background/related/sequential files

--- 2017-05-24 gpu-05-02-sequential.tex
@@
 %during the lifetime of the algorithm.
-The number of communities is $K=1024$, the mini-batch size is fixed at $M=32$, the
+the number of communities is $K=1024$, the mini-batch size is fixed at $M=32$, the
 neighbor sample size $\Neighbors=32$.

### 2017-05-25 06-conclusion.tex
Start removing the OpenCL CPU work

@@
 combinations of optimizations. The outcome of these efforts culminated in
-significant speedup factors of 21 and 86 using a multi-core CPU and a GPU
-respectively. These speedup numbers were achieved in comparison to an already
+significant speedup factor of 86 for the GTX-\mbox{Titan-X} GPU.
+This speedup was achieved in comparison to an already
 optimized sequential implementation. Finally, we evaluated the performance of

### 2017-05-25 gpu-05-03-accelerator.tex
Start removing the OpenCL CPU work

@@
-\subsubsection{Analysis of CPU Parallelism}
-
-\begin{comment}
-  \begin{table}[tb]
-    \centering
-    \begin{tabular}{l l c c c c}
-      Comparison & Data Set        & K    & M    & n  & Iterations \\
-      \hline
-      Python/Baseline    & CA-HepPh        & 1024 & 32   & 32 & 1000 \\
-      Parallel CPU/GPU   & com-DBLP        & 1024 & 4096 & 32 & 1000 \\
-      Parallel CPU       & com-LiveJournal & 1024 & 4096 & 32 & 1000 \\
-    \end{tabular}
-    \caption{Experiment parameter reference.}
-    \label{exp-params}
-  \end{table}
-\end{comment}
-
-\begin{table}[tb]	% [htb]
-  \centering
-  \begin{tabular}{l d{2.8} d{2.8}}
-    Kernel & Time (seconds) \\
-    \hline
-    PPX CALC    &  0.0364737 \\
-    PPX ACCUM   &  0.083 \\
-    SAMPLING    &  0.535599 \\
-    UPDATE\_PHI & 25.6598 \\
-    UPDATE\_PI  &  0.645875 \\
-    THETA SUM   &  0.0483902 \\
-    GRADS PAR   &  1.92919 \\
-    GRADS SUM   &  9.31122 \\
-    UPDATE THETA&  0.0548013 \\
-    NORM THETA  &  0.001 \\
-    \hline
-    TOTAL    & 38.5858 \\
-  \end{tabular}
-  \caption{Performance break-down of multi-core CPU version without
-  vectorization. Model parameters: data set com-DBLP; K=1024, m=4096, n=32.}
-  \label{TABLE-CPU}
-\end{table}
-
-This section discusses the use of the multi-core CPU available on the DAS5
-cluster. The parallel OpenCL version divides the work across the CPU cores and
-performs independent calculations concurrently. As shown in
-Table~\ref{TABLE-CPU}, the dominant kernel in the computation is
-\textit{update\_phi},
-which accounts for 66.5\% of the computation time. Without exploiting
-the dual 8-core processor's vectorization capabilities, the speedup relative to
-the baseline sequential C++ version is~9.8.
-
-% CPU
-\begin{figure}[tb]
-\centering
-\epsfig{file=plots/cpu-vector.eps, width=\columnwidth}
-\caption{Performance of CPU for varying vector width.}
-\label{FIG-CPU-VECTOR}
-\end{figure}
-
-In addition to applying computations in parallel, we investigated the use of
-the CPU's SIMD instructions to maximize its resource utilization.
-Figure~\ref{FIG-CPU-VECTOR} presents the performance obtained by vectorizing
-the kernels with varying vector widths. A key aspect in this figure is the
-diminishing performance benefit for higher vector widths. As the computational
-performance increases, the memory throughput becomes the leading performance
-bottleneck. Moreover, using 16-wide SIMD instructions gave a slight performance
-penalty compared to 8-wide SIMD. The 8-wide vector version improves the speedup
-relative to the baseline version from 9.8 to~20.9.
-
 \begin{table}[tb]	% [htb]
@@
 % GTX 980     & Maxwell &  2816 & 1126 (1216) & 4612 &  144 &      &  4 & 336.5 \\
+
 % GTX Titan-X & Maxwell &  3072 & 1000 (1075) & 6144 &  192 & 2048 & 12 & 336.5 \\
@@
 Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
-GTX~\mbox{Titan-X}, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
+GTX~\mbox{Titan-X}, GTX980, K40c and K20m relative to the baseline
 sequential C++ version. These results are consistent with the relative
 capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
-the \mbox{Titan-X} achieves the highest speedup of 86 relative to the baseline. At the
-other end, the parallel CPU version attains a speedup factor of~20.9.
+the \mbox{Titan-X} achieves the highest speedup of 86 relative to the baseline.
 

### 2017-05-25 gpu-05-05-perplexity.tex
Gone over figures etc. Still need to redo related work.
Start removing the OpenCL CPU work

@@
 \begin{comment}
-\begin{figure}[t]	% [htb]
+\begin{figure}[tb]	% [htb]
   \centering
@@
 
-\begin{figure}[t]	% [htb]
+\begin{figure}[tb]	% [htb]
   \centering
@@
 
-\begin{figure}[t]	% [htb]
+\begin{figure}[tb]	% [htb]
   \centering
   \epsfig{file=plots/ppx-gpu-cpu.eps, width=0.45\textwidth}
-  \caption{Perplexity convergence. (a)~com-DBLP on the GTX~\mbox{Titan-X}~CPU.
-  (b)~com-LiveJournal on the multi-core~CPU. Number of communities chosen to
-  maximally fill the memory of GPU and CPU respectively.}
+  \caption{Perplexity convergence. com-DBLP on the GTX~\mbox{Titan-X}~GPU.
+  Number of communities chosen to
+  maximally fill the GPU memory.}
   \label{fig-ppx-gpu-cpu}
@@
 convergence for the largest datasets and community sizes that fit into the memory
-of a GPU and a CPU respectively. The dominant customer of memory is the
+of a GPU. The dominant customer of memory is the
 $\pi$ matrix, sized $N{\times}K$ for $K$ communities and $N$ vertices in the
@@
 com-DBLP, with $N=317,080$. It is computed on the GTX~\mbox{Titan-X} GPU with input
-parameter set $K=4096$, $m=4096$, $n=32$. Convergence is reached after 6~hours.
-
-Figure~\ref{fig-ppx-gpu-cpu}(b) shows the perplexity convergence for
-com-LiveJournal, with $N=3,997,962$, computed on a multi-core DAS5 node.
-To fill the CPU core memory~(64GB), the input parameter set is $K$=3072,
-$M$=4096, $|\Neighbors|$=32. Sufficient convergence is reached after 18~hours. Note
-that this network graph is so large that the GTX~\mbox{Titan-X} would be unable to
-process more than $K$=256 communities.
+parameter set $K=4096$, $M=4096$, $|\Neighbors|=32$. Convergence is reached after 6~hours.
+%
+Note that the GTX~\mbox{Titan-X} would be unable to
+process more than $K$=256 communities for the much larger com-LiveJournal dataset.

### 2017-05-25 gpu-05-evaluation.tex
Move exp. setup into GPU section
Start removing the OpenCL CPU work

@@
 The following sections provide a detailed experimental performance analysis
-of the optimizations described above. We start by discussing the performance
-effects of parallelizing the computations
-on a multi-core CPU using OpenCL.
+of the optimizations described above.
 %
-Next, we assess the trade-offs associated with each of the GPU optimization
+We assess the trade-offs associated with each of the GPU optimization
 strategies and their performance effects on four GPUs spanning two chip
@@
 Finally, we present the overall outcome of accelerating the algorithm by
-displaying its fast convergence over the largest networks that fit in the
-memory of one GPU or CPU.
+displaying its fast convergence over the largest sensible combination of
+community network and number of communities, so that the algorithm data fit in the
+memory of a GPU.
 
-As noted in Section~\ref{sec-experimental}, all experiments were conducted on the VU Amsterdam DAS5 cluster.
-Table~\ref{table-gpus} presents the properties of the various GPUs in
-the DAS5 cluster.
+\input{04-experimental-setup}
+
+% As noted in Section~\ref{sec-experimental}, all experiments were conducted on the VU Amsterdam DAS5 cluster.
+% Table~\ref{table-gpus} presents the properties of the various GPUs in
+% the DAS5 cluster.
 

### 2017-05-25 plots/plot-gpus.sh

@@ -8,7 +8,8 @@ BASELINE=374.667
 echo > $TFILE
 
 X=1
-for i in GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m Intel-R-Xeon; do
+# for i in GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m Intel-R-Xeon; do
+for i in GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m ; do
   v=`grep -a TOTAL dblp-K1024/${i}* | sort -nk 7 | head -n 1 | awk '{print $7}'`
   v2=`perl -e "print $BASELINE / $v"`
   device=`echo $i | sed 's/GeForceGTXTITANX/Titan-X/;s/GeForceGTX980/GTX980/;s/TeslaK40c/K40c/;s/TeslaK20m/K20m/;s/Intel-R-Xeon/Xeon/'`

### 2017-05-25 plots/plot-ppx.sh

@@ -4,7 +4,7 @@ cat << EOF | gnuplot --persist
 w = 0.4
 h = 0.3
 k = 0.0
-nr = 2
+nr = 1
 nc = 1
 row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
@@ -26,11 +26,11 @@ set yrange [62:95]
 set title '(a) Com-DBLP K=4K m=4K n=32'
 plot "< perl ./get_time.perl ppx-DBLP-K4K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 4
 
-set origin col(0),row(1)
-set size w,h
+# set origin col(0),row(1)
+# set size w,h
 
-set xrange [-0.3:18.3]
-set yrange [53:83]
-set title '(b) Com-LiveJournal K=3K m=4K n=32'
-plot "< perl ./get_time.perl ppx-LJ-K3K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 4
+# set xrange [-0.3:18.3]
+# set yrange [53:83]
+# set title '(b) Com-LiveJournal K=3K m=4K n=32'
+# plot "< perl ./get_time.perl ppx-LJ-K3K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 4
 EOF

### 2017-05-29 01-intro.tex
Some wording in the intro

@@
 This already improves performance by a factor of 1000 to 1500 compared to
-Python. Then, our key contributions to the algorithm are:
+Python.
+% This transformation is not considered one of our key contributions.
+Then, our key contributions to the algorithm are:
 
@@
 \item \textbf{scaling up to run on GPUs.}
-The application is not a natural fit for GPUs, because it is memory-bound and
+The application is not a natural fit for GPUs, because it is data-bound and
 irregular in its data accesses. Nevertheless, we developed aggressive
@@
 of machines. Hence, the cluster nodes must read
-remote memory hosted by their peers. Since the algorithm is memory-bound, we
+remote memory hosted by their peers. Since the algorithm is data-bound, we
 substantially optimized remote memory accesses by implementing
@@
 
-\item \textbf{the use of pipelining} allows to use straightforward intra-node
-thread parallelism in the distributed implementation, and it is pointless
-to combine the GPU work into it. The cause is that the computation can be
+\item \textbf{the use of pipelining} \emph{in the distributed implementation}
+justifies straightforward intra-node thread parallelism, and it is pointless
+to combine the GPU implementations
+into it. The cause is that computing for an iteration takes less
+time than fetching the data from remote memory, and the computation is
 fully overlapped with fulfilling the data requests for the next iteration.
+Hence, it is useless to improve the computation time.
 

### 2017-05-29 04-experimental-setup.tex
Move exp. setup into GPU section
Tiny refactor experimental wording
Some wording in DAS section

@@
-\section{Experimental setup}
+\subsection{Experimental setup}
 \label{sec-experimental}
 
-Unless indicated otherwise, the experiments described in the next sections
+Unless indicated otherwise, all experiments described in this paper
 were conducted on the VU~Amsterdam and Leiden DAS5 clusters.  The clusters
@@
 8TB of storage. The nodes of each site are interconnected by FDR~InfiniBand.
-Additionally, the VU cluster is fitted with a number of Nvidia GPUs including
+Additionally, the VU cluster is fitted with a number of NVidia GPUs including
 GTX~\mbox{Titan-X}, GTX980, K40c and K20m.
 
-To assess the performance of our implementations, we use publicly available
-community networks from the SNAP collection~\cite{snapnets}, including the largest
-networks. Table~\ref{table-snap}
+To evaluate our implementations, we use publicly available
+community networks from the SNAP collection~\cite{snapnets}.
+Table~\ref{table-snap}
 outlines the graph properties of the networks used.

### 2017-05-29 distr-03-design.tex
Wording, fit with previous intro into distr

@@
 
-The SG-MCMC a-MMSB algorithm has an abundance of
-opportunities for parallelism, for the multi-threaded shared-memory type as
-well as for distributed-memory parallelism. The benefit of multi-threaded
-parallelism is the speedup of the computation. Additionally, a distributed implementation
-allows the storage of data in the collective memory of a cluster
-and increases aggregate memory bandwidth. Both of these parameters scale with the number of
-machines. The downside of a distributed implementation is that it requires
-considerable implementation effort.
+The implementations in the previous section put a limit to the size of
+community networks and number of communities that can be processed: the
+state must fit into GPU memory. To tackle networks of much larger size, we
+created a distributed implementation that stores the data in the
+collective memory of a cluster. The downside of a distributed implementation
+is that the burden of memory access is increased for our already data-dominant
+algorithm.
 
@@
 This section describes how we created a distributed version of the C++
-algorithm. We didn't use the advanced intra-node parallelization techniques
-described in the previous section, but found that straightforward
-multi-threaded parallelization techniques sufficed: the distributed performance
-is dominated by (remote) memory latency, not by node parallelism. As our
-analysis will show, our pipelining optimizations will fully overlap the
-parallel calculations with (pre)loading the data from remote hosts.
-The details of the multi-threaded parallelism
-will be discussed only where appropriate. The
-distributed design follows a master-worker paradigm, where
-the master controls the sequence of parallel operations and the workers perform the
-calculations. For thread parallelism, we annotate the program with
-OpenMP~\cite{OpenMPSpec}. For distributed communication, we use
-MPI~\cite{Forum:1994:MMI:898758}.
+algorithm. We used the multicore CPUs in the DAS5 with straightforward
+intra-node thread parallelisation, not the GPU implementations described
+in the previous section. The justification is twofold: in our DAS5 system,
+it is pointless to speed up the intra-node performance beyond straightforward
+multithreading, because distributed performance is dominated by remote memory
+latency; the analysis in Section~\ref{eval-pipeline} shows that the time for
+(pre)loading the data from remote hosts greatly exceeds the calculation time.
+The calculation time is fully overlapped in our pipelining implementation.
+Moreover, usage of GPUs in the distributed setting would necessitate staging of
+the relevant parts of $\pi/\phi$ from the host to GPU memory in each iteration,
+which would eat most of the performance gain offered by the GPU because
+the algorithm is so data-intensive.  % The details of the multi-threaded
+parallelism % will be discussed only where appropriate.  The distributed design
+follows a master-worker paradigm, where the master controls the sequence of
+parallel operations and the workers perform the calculations. For thread
+parallelism, we annotate the program with OpenMP~\cite{OpenMPSpec}. For
+distributed communication, we use MPI~\cite{Forum:1994:MMI:898758}.
 

### 2017-05-29 gpu-04-03-accelerator.tex
Remove the history approach from OpenCL. Remove OpenCL

@@
 % Accelerator friendliness
-The design of an accelerated version of the MCMC algorithm necessitated several
-crucial modifications to allow for efficient parallelization. This section
-discusses the key design contributions to attain optimized
-parallel versions. The state represented by $\pi$ and $\phi$ is permanently
-stored in the GPU memory, because staging of the relevant portions of these
-data structures in each iteration is too costly compared to the amount
+% The design of an accelerated version of the MCMC algorithm necessitated several
+% crucial modifications to allow for efficient parallelization. This section
+% discusses the key design contributions to attain optimized
+% parallel versions.
+
+Since the algorithm is memory-bound, many of the optimizations we implemented
+are concerned with
+memory usage. For starters, we offer a fast, memory-efficient hash table on the GPU,
+taylored towards
+the algorithm's usage patterns. Some optimizations reduce the total memory
+footprint of the algorithm. Other
+optimizations explore the opportunities offered by the faster regions of the
+GPU memory hierarchy: data can be put into GPU shared memory, into thread-local
+memory, or in registers which requires a more complicated analysis. One
+other optimization is exploitation of the vectorization possibilities.
+
+The state for the combination of $\pi$ and $\phi$ is permanently
+stored in GPU memory, because copying of the relevant portions of these
+data structures to the GPU in each iteration is too costly compared to the amount
 of computation. The graph is stored at the host, because mini-batch
 selection is cheap, and the data required for a minibatch is small.
-
-Since the algorithm is memory-bound, many of the optimizations concern
-memory usage. We offer a fast, memory-efficient hash table taylored towards
-the algorithm's usage patterns. Some optimizations reduce the total memory
-footprint of the algorithm, following analysis of the algorithm. Other
-optimizations explore the opportunities offered by the faster regions of the
-GPU memory hierarchy: data can be put into thread-local memory, in shared
-GPU memory, or in registers which requires a more complicated analysis. One
-other optimization exploits the vectorization possibilities.
 
@@
 %OpenCL/CUDA
+\begin{comment}
 \textit{\bf !!! Remove the history approach !!!}
@@
 CUDA's primitive types, address spaces and functions.
+\end{comment}
+The GPU code and interfacing was developed with CLCudaAPI~\cite{claduc}, which
+is an API that targets both CUDA for NVidia GPUs and OpenCL for other GPUs and
+multicore CPUs. It was not feasible to directly use OpenCL~\cite{opencl} since
+NVidia's OpenCL SDK limits the total memory allocations to 4GB, which is
+significantly smaller than the memory available in our GPUs.
 
@@
 The data structures that occupy most memory are $\pi$ and~$\phi$, both of
-which are matrices of dimensions ${N}\times{K}$. However, the storage of both
-matrices is redundant as $\pi$ is a row-normalized copy of~$\phi$, see
+which are matrices of dimension ${N}\times{K}$. However, storing both
+is redundant as $\pi$ is a row-normalized copy of~$\phi$, see
 Table~\ref{tbl-symbols} in Section~\ref{sec-background}. The full storage for

### 2017-05-29 02-02-related.tex
Rename the background/related/sequential files
Chance citation: not the subject

--- 2017-05-24 03-related.tex
@@
-\subsection{Related Work}
+\section{Related Work}
 \label{sec-related}
@@
 not such a natural fit for GPUs, a number of projects explores this terrain.
-E.g.~\cite{journals/bioinformatics/MedlarGSBK13} uses GPUs with their
+Medlar~c.s.~\cite{journals/bioinformatics/MedlarGSBK13} use GPUs with their
 MCMC approach to analyze parental linkage patterns in a biology context and
-\cite{DBLP:journals/csda/WhiteP14} does the same to model terrorist activity.
+White and Porter~\cite{DBLP:journals/csda/WhiteP14} do the same to model terrorist activity.
 Latent Dirichlet Allocation, another variety of Bayesian Approximation,
-is used on GPUs by~\cite{DBLP:conf/nips/YanXQ09}. There is also related
+is used on GPUs by Yan~c.s.~\cite{DBLP:conf/nips/YanXQ09}. There is also related
 work on the variants of MCMC algorithms that use the gradient to speed
@@
 these varieties~\cite{Girolami_riemannmanifold}. Our algorithm uses Riemann
-Manifold Langevin dynamics. In \cite{beam2014fast}, GPUs are used to perform
+Manifold Langevin dynamics. Beam~c.s.~\cite{beam2014fast} use GPUs to perform
 Hamiltonian descent, using Python interfaces to access the standard cuBLAS
@@
 
-\cite{liakos-bigdata16} also addresses overlapping community detection, but their
+Liakos~c.s.~\cite{liakos-bigdata16} also address overlapping community detection, but their
 algorithm uses a local dispersion-aware approach. They find communities for 5000
@@
 for community detection on very large network graphs. In contrast to our work,
-they all detect non-overlapping communities: \cite{Bu2013246} investigate a
-large number of networks; \cite{2015arXiv150302115L} investigate hierarchical
-stochastic blockmodels; \cite{Prat-Perez:2014:HQS:2566486.2568010} use
+they all detect non-overlapping communities: Bu~c.s.~\cite{Bu2013246} investigate a
+large number of networks; Lyzinski~c.s.~\cite{2015arXiv150302115L} investigate hierarchical
+stochastic blockmodels; Prat-P{\'e}rez~c.s.~\cite{Prat-Perez:2014:HQS:2566486.2568010} use
 multi-core machines.

### 2017-05-29 02-02-related.tex

@@
-\section{Related Work}
+\subsection{Related Parallelization Work}
 \label{sec-related}
 
-We consider related work for both parts of this paper: acceleration and scaling
-of our MCMC-MMSB algorithm. Acceleration of deep learning algorithms on GPUs
-is an ongoing success story. Although approximative Baysian algorithms are
-not such a natural fit for GPUs, a number of projects explores this terrain.
+We consider related work for both parallelization contributions of this paper: acceleration and scaling
+of our MCMC-MMSB algorithm. While acceleration of deep learning algorithms on GPUs
+is an ongoing success story, approximative Baysian algorithms are
+not such a natural fit for GPUs. Nevertheless, a number of projects explores this terrain.
 Medlar~c.s.~\cite{journals/bioinformatics/MedlarGSBK13} use GPUs with their

### 2017-05-30 02-02-related.tex
Replace D-KV with DKV

--- 2017-05-29 02-02-related.tex
@@
 \begin{comment}
-In this paper, we describe our custom RDMA D-KV (Distributed Key-Value)
-store. Current RMDA D-KV store implementations are RamCloud~\cite{RamCloud},
+In this paper, we describe our custom RDMA DKV (Distributed Key-Value)
+store. Current RMDA DKV store implementations are RamCloud~\cite{RamCloud},
 Pilaf~\cite{Pilaf}, Herd~\cite{Herd} and FaRM~\cite{FaRM}. All these systems
-use RDMA to implement a D-KV store. However, all of them are far more powerful
+use RDMA to implement a DKV store. However, all of them are far more powerful
 than our custom implementation -- and this power comes at a cost that we
-can avoid. They implement a generic D-KV store that controls concurrency,
+can avoid. They implement a generic DKV store that controls concurrency,
 supports dynamic inserts and deletes, supports variable-sized values
@@
 \subsection{Related work on RDMA Key-Value stores}
-In this paper, we describe our custom RDMA D-KV (Distributed Key-Value)
-store. Current RMDA D-KV store implementations are RamCloud~\cite{RamCloud},
+In this paper, we describe our custom RDMA DKV (Distributed Key-Value)
+store. Current RMDA DKV store implementations are RamCloud~\cite{RamCloud},
 Pilaf~\cite{Pilaf}, Herd~\cite{Herd} and FaRM~\cite{FaRM}. All these systems
-use RDMA to implement a D-KV store. However, all of them are far more powerful
+use RDMA to implement a DKV store. However, all of them are far more powerful
 than our custom implementation -- and this power comes at a cost that we
-can avoid. They implement a generic D-KV store that controls concurrency,
+can avoid. They implement a generic DKV store that controls concurrency,
 supports dynamic inserts and deletes, supports variable-sized values

### 2017-05-30 distr-04-evaluation.tex
Replace D-KV with DKV

--- 2017-05-18 distr-04-evaluation.tex
@@
   \epsfig{file=plots/qperf.eps, width=\columnwidth}
-  \caption{Performance of DKV store vs.\ \textit{qperf} and MPI roundtrip.}
+  \caption{Performance of DKV store vs.\ \textit{qperf} and MPI round-trip.}
   \label{fig-qperf}

### 2017-05-30 distr-03-design.tex
Tiny textual fixes

@@
 \subsection{Pipelining of computation and loading $\pi$}
+\label{sec-pipelining}
 
@@
 loading $\pi$ for the next chunk. The rationale for implementing pipeline
-parallelism in these places of the program, is that loading $\pi$ is the
+parallelism in these places of the program is that loading $\pi$ is the
 dominant contribution within the dominant stage \textit{update\_phi}, as

### 2017-05-30 gpu-04-03-accelerator.tex
Tiny textual fixes

@@
 data-intensive, a reduction of the state is accompanied by a decrease in its
-data-intensity.
+data intensity.
 %
@@
 %
-This change affects the algorithm in multiple ways. For instance, on some GPUs it
+This change affects the algorithm in multiple ways.
+It again halves the memory
+footprint of the algorithm and frees registers.
+Besides, on our GPUs it
 reduces
-the computation intensity besides the data intensity. It again halves the memory
-footprint of the algorithm and frees registers.
+the computation intensity besides the data intensity.
 
@@
 
-\subsubsection{Edge-centric and Vertex-centric kernels}
+% \subsubsection{Edge-centric and Vertex-centric kernels}
+\subsubsection{Memory Hierarchy Optimizations}
+\label{gpu-design}
+
 The work decomposition of the kernels ensures that each thread performs
@@
 
-\subsubsection{Memory Hierarchy Optimizations}
-\label{gpu-design}
 % The GPU implementation builds on top of the CPU work decomposition scheme.
@@
 Since all of the kernels are data-bound, several memory organization strategies
-of varying complexities were investigated to speedup the algorithm by
+of varying complexity were investigated to speed up the algorithm by
 exploiting the GPU's memory hierarchy.
@@
 % Naive strategy
-\paragraph*{\textbf{Naive strategy}} simply allocate temporary vectors in thread
+\paragraph*{\textbf{Naive strategy}} allocate temporary vectors in thread
 local memory to store $Probs$ and $Grads$ which physically reside in
@@
 Hence, this strategy allows for 8 possible configurations denoted by
-three letters each of which is a choice between $R$ or $S$ to represent
+three letters, each of which is a choice between $R$ or $S$ to represent
 $Register$ or $Shared$ respectively. For example, $SSR$ denotes that $Probs_i$,
-$Grads_i$ and $\pi_i$ were placed in $Shared$, $Shared$ and $Register$
+$Grads_i$ and $\pi_i$ are placed in $Shared$, $Shared$ and $Register$
 respectively.

### 2017-05-30 gpu-05-evaluation.tex
Tiny textual fixes

@@
 Finally, we present the overall outcome of accelerating the algorithm by
-displaying its fast convergence over the largest sensible combination of
-community network and number of communities, so that the algorithm data fit in the
-memory of a GPU.
+displaying its convergence speed over the largest sensible combination of
+community network and number of communities while keeping the algorithm data
+in the memory of a GPU.
 

### 2017-05-30 plots/plot-all-gpu-detail.sh

@@ -1,16 +1,20 @@
 #!/bin/bash
 
+./plot-gpu-sweep.sh --title "(a) Titan-X/Pascal K=1024 no vectorization" titanx-1024-strategies-w1 `ls dblp-K1024/TITANX\(Pascal\)-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
+./plot-gpu-sweep.sh --title "(a) Titan-X/Pascal K=1024 vector width=2" titanx-1024-strategies-w2 `ls dblp-K1024/TITANX\(Pascal\)-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
+./plot-gpu-sweep.sh --title "(a) Titan-X/Pascal K=1024 vector width=4" titanx-1024-strategies-w4 `ls dblp-K1024/TITANX\(Pascal\)-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
+
 ./plot-gpu-sweep.sh --title "(a) Titan-X K=1024 no vectorization" titanx-1024-strategies-w1 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
-./plot-gpu-sweep.sh --title "T(a) itan-X K=1024 vector width=2" titanx-1024-strategies-w2 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
-./plot-gpu-sweep.sh --title "T(a) itan-X K=1024 vector width=4" titanx-1024-strategies-w4 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
+./plot-gpu-sweep.sh --title "(a) Titan-X K=1024 vector width=2" titanx-1024-strategies-w2 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
+./plot-gpu-sweep.sh --title "(a) Titan-X K=1024 vector width=4" titanx-1024-strategies-w4 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
 
[…]

### 2017-05-30 plots/plot-gpus.sh

@@ -9,10 +9,11 @@ echo > $TFILE
 
 X=1
 # for i in GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m Intel-R-Xeon; do
-for i in GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m ; do
+for i in 'TITANX(Pascal)' GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m ; do
+# for i in 'TITANX(Pascal)' ; do
   v=`grep -a TOTAL dblp-K1024/${i}* | sort -nk 7 | head -n 1 | awk '{print $7}'`
   v2=`perl -e "print $BASELINE / $v"`
-  device=`echo $i | sed 's/GeForceGTXTITANX/Titan-X/;s/GeForceGTX980/GTX980/;s/TeslaK40c/K40c/;s/TeslaK20m/K20m/;s/Intel-R-Xeon/Xeon/'`
+  device=`echo $i | sed 's+TITANX\(Pascal\)+Titan-X/Pascal+'`
   echo $X $device $v2 >> $TFILE
   X=`expr $X + 1`
 done

### 2017-05-31 gpu-05-03-accelerator.tex
Spell NVidia iso Nvidia
Add a few Titan-X/Pascal measurements

--- 2017-05-24 gpu-05-03-accelerator.tex
@@
 % GTX980 & TITANX: https://en.wikipedia.org/wiki/List_of_Nvidia_graphics_processing_units
-\center\begin{tabular}{lc c c c c}
-                         && \multicolumn{4} {c}{Device} \\
-\cline{3-6}
-Specification            && K20m  & K40c  & GTX 980 & GTX Titan-X \\
-\cline{1-1}\cline{3-6}
-Number of Cores          && 2496  & 2880  & 2048    & 3072  \\
-Clock (MHz)              && 706   & 745   & 1126    & 1000  \\
-GFlops (single)          && 3520  & 4290  & 4612    & 6144  \\
-GFlops (double)          && 1170  & 1430  & 144     & 192   \\
-Memory (GB)              && 5     & 12    & 4       & 12    \\
-Bandwidth (GB/s)         && 208   & 288   & 224     & 336.5 \\
-\cline{1-1}\cline{3-6}
+\setlength\tabcolsep{0.4em}
+\center\begin{tabular}{lc c c c c c}
+                         && \multicolumn{5}{c}{Device} \\
+\cline{3-7}
+                         && K20m  & K40c  & GTX 980 & \multicolumn{2}{c}{GTX Titan-X} \\
+Specification            &&       &       &         & Maxwell     & Pascal \\
+\cline{1-1}\cline{3-7}
+Number of Cores          && 2496  & 2880  & 2048    & 3072        & 3584 \\
+Clock (MHz)              && 706   & 745   & 1126    & 1000        & 1417 \\
+GFlops (single)          && 3520  & 4290  & 4612    & 6144        & 10157 \\
+GFlops (double)          && 1170  & 1430  & 144     & 192         & 317 \\
+Memory (GB)              && 5     & 12    & 4       & 12          & 12 \\
+Bandwidth (GB/s)         && 208   & 288   & 224     & 336.5       & 480 \\
+\cline{1-1}\cline{3-7}
 \\[-1.5ex]
@@
 shared strategy and 8 variations of the code generation strategy. This
-section investigates the effectiveness of each flavor on Nvidia's GTX~\mbox{Titan-X},
+section investigates the effectiveness of each flavor on Nvidia's GTX~\mbox{Titan-X}
+Maxwell and Pascal,
 GTX980, K40c and K20m.
@@
 % \looseness=-1
-Figures~\ref{gpu-sweep}(c) present the performance of the K40c GPU
+Figure~\ref{gpu-sweep}(c) presents the performance of the K40c GPU
 for the same experimental configuration as shown for the \mbox{Titan-X} in
-Figure~\ref{gpu-sweep}(a) with one exception. Namely, the code version used
-for this plot is explicitly vectorized with a vector width of~2. The results
+Figure~\ref{gpu-sweep}(a) except that the code
+is explicitly vectorized with a vector width of~2. The results
 for the other code versions with vector width 4 and no vectorization are

### 2017-05-31 plots/collect-v.sh

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 function Find {
-  RET=$(grep -a PHI ${1}*W${2}* \
+  RET=$(grep -a PHI "${1}"*W${2}* \
     | grep -v LOG \
     | awk '{print $1 " " $7}' \
     | sed 's/\(:[^ ]\+\)//' \
@@ -19,15 +19,16 @@ function Find {
 
 # find best config, get the W{1,2,4} for all
 function Collect {
-  Find $1 \
-    | awk "{if (\$5 == \"-\") {print \"${1}\"\$1\"-PHI\"\$2\"-W*\"} else {print \"${1}\"\$1\"-PHI\"\$2\"-W*-\"\$5} }" \
+  Find "$1" \
+    | awk "{if (\$5 == \"-\") {print \""${1}"\"\$1\"-PHI\"\$2\"-W*\"} else {print \""${1}"\"\$1\"-PHI\"\$2\"-W*-\"\$5} }" \
     | xargs echo echo \
     | bash \
     | xargs ./collect.sh 
 }
 
 echo "GPU NO-VECTORIZATION VECTOR-WIDTH=2 VECTOR-WIDTH=4"
-Collect dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4}" | xargs echo TitanX
+Collect dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4}" | xargs echo '"GTX Titan-X\nPascal"'
+Collect dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4}" | xargs echo '"GTX Titan-X\nMaxwell"'
 Collect dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4 }" | xargs echo GTX980
 Collect dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4 }" | xargs echo K40c
 Collect dblp-K1024/TeslaK20m-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4 }" | xargs echo K20m

### 2017-05-31 plots/collect.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
-F1=$$.tmp.1
-F2=$$.tmp.2
+F1="$$.tmp.1"
+F2="$$.tmp.2"
 
 grep -aH PHI "$@" | grep -v LOG \
   | awk '{print $1 " " $7}' \

### 2017-05-31 plots/plot-all-gpu-detail.sh

@@ -8,9 +8,9 @@
 ./plot-gpu-sweep.sh --title "GTX980 K=1024 vector width=2" gtx980-1024-strategies-w2 `ls dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
 ./plot-gpu-sweep.sh --title "GTX980 K=1024 vector width=4" gtx980-1024-strategies-w4 `ls dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
 
-./plot-gpu-sweep.sh --title "K40c K=1024 no vectorization" k40-1024-strategies-w1 `ls dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
-./plot-gpu-sweep.sh --title "K40c K=1024 vector width=2" k40-1024-strategies-w2 `ls dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
-./plot-gpu-sweep.sh --title "K40c K=1024 vector width=4" k40-1024-strategies-w4 `ls dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
+./plot-gpu-sweep.sh --title "(c) K40c K=1024 no vectorization" k40-1024-strategies-w1 `ls dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
+./plot-gpu-sweep.sh --title "(c) K40c K=1024 vector width=2" k40-1024-strategies-w2 `ls dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
+./plot-gpu-sweep.sh --title "(c) K40c K=1024 vector width=4" k40-1024-strategies-w4 `ls dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
 
 ./plot-gpu-sweep.sh --title "K20m K=1024 no vectorization" k20-1024-strategies-w1 `ls dblp-K1024/TeslaK20m-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
[…]

### 2017-05-31 plots/plot-gpus.sh

@@ -8,11 +8,11 @@ BASELINE=374.667
 echo > $TFILE
 
 X=1
-for i in GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m Intel-R-Xeon; do
+for i in 'TITANX-Pascal' GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m Intel-R-Xeon; do
   v=`grep -a TOTAL dblp-K1024/${i}* | sort -nk 7 | head -n 1 | awk '{print $7}'`
   v2=`perl -e "print $BASELINE / $v"`
-  device=`echo $i | sed 's/GeForceGTXTITANX/Titan-X/;s/GeForceGTX980/GTX980/;s/TeslaK40c/K40c/;s/TeslaK20m/K20m/;s/Intel-R-Xeon/Xeon/'`
-  echo $X $device $v2 >> $TFILE
+  device=`echo $i | sed 's/TITANX-Pascal/GTX Titan-X\\\nPascal/;s/GeForceGTXTITANX/GTX Titan-X\\\nMaxwell/;s/GeForceGTX980/GTX980/;s/TeslaK40c/K40c/;s/TeslaK20m/K20m/;s/Intel-R-Xeon/Xeon/'`
+  echo $X \"$device\" $v2 >> $TFILE
   X=`expr $X + 1`
 done
 
@@ -30,6 +30,7 @@ set style fill solid
 set ylabel 'Speedup over baseline'
 
 set yrange[0:]
+set bmargin 3.0
 
 set title 'Performance of different compute devices'
 plot '$TFILE' u 1:3:xtic(2) w boxes notitle

### 2017-05-31 plots/plot-vector-sweep.sh

@@ -26,6 +26,7 @@ set ytics 0,0.2
 # set grid ytics mytics
 set grid ytics
 
+set bmargin 3.0
 set yrange [0:*]
 plot \
   '$TFILE' u (\$2/\$2):xtic(1) title col, \
@@ -34,5 +35,5 @@ plot \
 
 EOF
 
-rm $TFILE
+# rm $TFILE
 

### 2017-05-31 plots/collect-v.sh
Import the K=2048 Titan-X/Pascal data files too

@@ -27,8 +27,8 @@ function Collect {
 }
 
 echo "GPU NO-VECTORIZATION VECTOR-WIDTH=2 VECTOR-WIDTH=4"
-Collect dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4}" | xargs echo '"GTX Titan-X\nPascal"'
-Collect dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4}" | xargs echo '"GTX Titan-X\nMaxwell"'
+Collect dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4}" | xargs echo '"Titan-X\nPascal"'
+Collect dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4}" | xargs echo '"Titan-X\nMaxwell"'
 Collect dblp-K1024/GeForceGTX980-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4 }" | xargs echo GTX980
 Collect dblp-K1024/TeslaK40c-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4 }" | xargs echo K40c
 Collect dblp-K1024/TeslaK20m-K1024-M4096-N32-Snode-R0.01-A0.001-C0- | awk "{print \$4 }" | xargs echo K20m

### 2017-05-31 plots/plot-gpus.sh

@@ -11,7 +11,7 @@ X=1
 for i in 'TITANX-Pascal' GeForceGTXTITANX GeForceGTX980 TeslaK40c TeslaK20m Intel-R-Xeon; do
   v=`grep -a TOTAL dblp-K1024/${i}* | sort -nk 7 | head -n 1 | awk '{print $7}'`
   v2=`perl -e "print $BASELINE / $v"`
-  device=`echo $i | sed 's/TITANX-Pascal/GTX Titan-X\\\nPascal/;s/GeForceGTXTITANX/GTX Titan-X\\\nMaxwell/;s/GeForceGTX980/GTX980/;s/TeslaK40c/K40c/;s/TeslaK20m/K20m/;s/Intel-R-Xeon/Xeon/'`
+  device=`echo $i | sed 's/TITANX-Pascal/Titan-X\\\nPascal/;s/GeForceGTXTITANX/Titan-X\\\nMaxwell/;s/GeForceGTX980/GTX980/;s/TeslaK40c/K40c/;s/TeslaK20m/K20m/;s/Intel-R-Xeon/Xeon/'`
   echo $X \"$device\" $v2 >> $TFILE
   X=`expr $X + 1`
 done
@@ -20,7 +20,7 @@ cat $TFILE
 
 cat <<EOF | gnuplot --persist
 
-set terminal postscript eps enhanced color font ',10' size 3.3,1.6
+set terminal postscript eps enhanced color font ',12' size 3.3,1.6
 set output 'gpus.eps'
 set colorsequence classic
 

### 2017-05-31 plots/plot-vector-sweep.sh

@@ -8,7 +8,7 @@ TFILE="$$.tmp"
 ./collect-v.sh > $TFILE
 
 cat <<EOF | gnuplot --persist
-set terminal postscript eps enhanced color font ',9' size 3.3,1.6
+set terminal postscript eps enhanced color font ',12' size 3.3,1.6
 set colorsequence classic
 set output '$NAME'
 

### 2017-05-31 04-experimental-setup.tex
Add a few Titan-X/Pascal measurements
Reorder tables both for GPUs and SNAP sets

--- 2017-05-18 04-experimental-setup.tex
@@
-\section{Experimental setup}
+\subsection{Experimental setup}
 \label{sec-experimental}
@@
 Additionally, the VU cluster is fitted with a number of Nvidia GPUs including
-GTX~\mbox{Titan-X}, GTX980, K40c and K20m.
+GTX~\mbox{Titan-X/Pascal},
+GTX~\mbox{Titan-X/Maxwell}, GTX980, K40c and K20m.
 
@@
     CA-HepPh        &    12,008  &    118,521    & \multicolumn{1}{c}{n/a}       & Paper citation network \\
+    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography network \\
+    com-Amazon      &    334,863 &       925,872 & 75,149         & Product copurchasing network \\
+    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
+    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
     com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
     com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
-    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
-    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
-    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography network \\
-    com-Amazon      &    334,863 &       925,872 & 75,149         & Product copurchasing network \\
     \hline

### 2017-05-31 gpu-05-03-accelerator.tex
Reorder tables both for GPUs and SNAP sets

--- 2017-05-31 gpu-05-03-accelerator.tex
@@
 \cline{3-7}
-                         && K20m  & K40c  & GTX 980 & \multicolumn{2}{c}{GTX Titan-X} \\
-Specification            &&       &       &         & Maxwell     & Pascal \\
+              && \multicolumn{2}{c}{GTX Titan-X} & GTX 980 & K40c  & K20m  \\
+Specification            && Pascal & Maxwell     &         &       &       \\
 \cline{1-1}\cline{3-7}
-Number of Cores          && 2496  & 2880  & 2048    & 3072        & 3584 \\
-Clock (MHz)              && 706   & 745   & 1126    & 1000        & 1417 \\
-GFlops (single)          && 3520  & 4290  & 4612    & 6144        & 10157 \\
-GFlops (double)          && 1170  & 1430  & 144     & 192         & 317 \\
-Memory (GB)              && 5     & 12    & 4       & 12          & 12 \\
-Bandwidth (GB/s)         && 208   & 288   & 224     & 336.5       & 480 \\
+Number of Cores          && 3584   & 3072        & 2048    & 2880  & 2496  \\
+Clock (MHz)              && 1417   & 1000        & 1126    & 745   & 706   \\
+GFlops (single)          && 10157  & 6144        & 4612    & 4290  & 3520  \\
+GFlops (double)          && 317    & 192         & 144     & 1430  & 1170  \\
+Memory (GB)              && 12     & 12          & 4       & 12    & 5     \\
+Bandwidth (GB/s)         && 480    & 336.5       & 224     & 288   & 208   \\
 \cline{1-1}\cline{3-7}

### 2017-06-01 04-experimental-setup.tex
Refactor to accomodate the Pascal. Still need strategy graph, ppx line for the Pascal.

@@
 Additionally, the VU cluster is fitted with a number of Nvidia GPUs including
-GTX~\mbox{Titan-X/Pascal},
-GTX~\mbox{Titan-X/Maxwell}, GTX980, K40c and K20m.
+GTX~\mbox{Titan-X(Pascal)},
+GTX~\mbox{Titan-X(Maxwell)}, GTX980, K40c and K20m.
 

### 2017-06-01 gpu-05-03-accelerator.tex
Refactor to accomodate the Pascal. Still need strategy graph, ppx line for the Pascal.

@@
   \caption{Performance break-down of multi-core CPU version without
-  vectorization. Model parameters: data set com-DBLP; K=1024, m=4096, n=32.}
+  vectorization. Model parameters: data set com-DBLP; K=1024, m=4096, $|\Neighbors|$=32.}
   \label{TABLE-CPU}
@@
                 && \multicolumn{2}{c}{GTX Titan-X} & GTX 980 & K40c  & K20m  \\
-Specification            && Pascal & Maxwell     &         &       &       \\
+Specification            && (Pascal) & (Maxwell)   &         &       &       \\
 \cline{1-1}\cline{3-7}
@@
 	the optimal block range of the upper figures.
-	Other relevant model parameters: M=4096, n=32.
+	Other relevant model parameters: M=4096, $|\Neighbors|$=32.
 	% (a) \mbox{Titan-X} GPU, without explicit kernel vectorization;
-	% Relevant model parameters: K=1024, M=4096, n=32.
+	% Relevant model parameters: K=1024, M=4096, $|\Neighbors|$=32.
   	% (b) \mbox{Titan-X} GPU, without explicit kernel vectorization;
-	% K=2048, M=4096, n=32.
+	% K=2048, M=4096, $|\Neighbors|$=32.
 	% (c) K40c GPU, with explicit kernel vectorization of width 2;
-	% K=1024, M=4096, n=32.
+	% K=1024, M=4096, $|\Neighbors|$=32.
   }
@@
   \caption{Execution time of 1000 \textit{update\_phi} invocations using the
-  \mbox{Titan-X} GPU,
+  \mbox{Titan-X(Maxwell)} GPU,
   without explicit kernel vectorization, across a sweep of
   \textit{update\_phi} thread block
-  sizes. Relevant model parameters: K=2048, M=4096, n=32.}
+  sizes. Relevant model parameters: K=2048, M=4096, $|\Neighbors|$=32.}
   \label{titanx-w1-sweep-2k}
@@
 
-Figure \ref{gpu-sweep}(a) presents the performance of the GTX~\mbox{Titan-X} GPU
+Figure \ref{gpu-sweep}(a) presents the performance of the Maxwell \mbox{Titan-X} GPU
 for all 10 strategy flavors, without explicitly vectorizing the kernels. The
@@
 performance as they unroll internal loops as well as cache data in
-registers. The \mbox{Titan-X} obtains the best performance with the \textit{GEN-RSS}
-strategy and a thread block size of 64. The results of explicitly vectorized kernels
-are omitted as they obtain worse performance on the \mbox{Titan-X}.
+registers. The Pascal Titan-X is fastest with the \textit{GEN-RSR} strategy and a
+block size of 96 and a vector width of~2, whereas
+the Maxwell Titan-X obtains the best performance with the \textit{GEN-RSS}
+strategy and a thread block size of 64 and no vectorization. The results of other
+vector widths
+are omitted as they obtain lower performance.
 
@@
   \textit{update\_phi} thread block
-  sizes. Relevant model parameters: K=1024, M=4096, n=32.}
+  sizes. Relevant model parameters: K=1024, M=4096, $|\Neighbors|$=32.}
   \label{k40-w2-sweep}
@@
 Figure~\ref{gpu-sweep}(c) presents the performance of the K40c GPU
-for the same experimental configuration as shown for the \mbox{Titan-X} in
+for the same experimental configuration as shown for the Maxwel \mbox{Titan-X} in
 Figure~\ref{gpu-sweep}(a) except that the code
@@
 
-The performance results obtained from the GTX~\mbox{Titan-X} and Tesla~K40c GPUs
+The performance results obtained from the Maxwell \mbox{Titan-X} and Tesla~K40c GPUs
 reinforce the importance of customizing compute kernels to each GPU's specific
@@
   each compute device normalized over baseline C++ version. Relevant model
-  parameters: K=1024, M=4096, n=32.}
+  parameters: K=1024, M=4096, $|\Neighbors|$=32.}
   \label{gpus-fig}
@@
   vector width, normalized over the non-vectorized kernel. Relevant model
-  parameters: K=1024, M=4096, n=32.}
+  parameters: K=1024, M=4096, $|\Neighbors|$=32.}
   \label{gpus-v-sweep}
@@
 Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
-GTX~\mbox{Titan-X}, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
+GTX~\mbox{Titan-X} Maxwell and Pascal, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
 sequential C++ version. These results are consistent with the relative
 capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
-the \mbox{Titan-X} achieves the highest speedup of 86 relative to the baseline. At the
+the \mbox{Titan-X(Pascal)} achieves the highest speedup of 120 relative to the baseline. At the
 other end, the parallel CPU version attains a speedup factor of~20.9.
@@
 As vectorization is a popular GPU optimization, it is important to note that
-both GTX GPUs achieved their highest performance with non-vectorized kernels.
-On the other hand, the Tesla GPUs obtained the best performance with a vector
+the Maxwell \mbox{Titan-X} and GTX980 achieve their highest performance with non-vectorized kernels.
+On the other hand, the Pascal \mbox{Titan-X} and Tesla GPUs obtain the best performance with a vector
 width of~2. Figure~\ref{gpus-v-sweep} presents the execution time of the best
 performing strategy for each of the 4 GPUs. In this figure, the performance is
-normalized over the non-vectorized kernel version for each GPU. For instance,
-the \mbox{Titan-X} exhibits an overhead factor of approximately 1.8 when using a vector
+normalized over the non-vectorized kernel version for each GPU. At one extreme,
+the Maxwell \mbox{Titan-X} exhibits an overhead factor of nearly 1.8 when using a vector
 width of~4. On the other hand, the K40c improves its performance by roughly

### 2017-06-01 gpu-05-05-perplexity.tex
Refactor to accomodate the Pascal. Still need strategy graph, ppx line for the Pascal.

--- 2017-05-11 gpu-05-05-perplexity.tex
@@
 \begin{comment}
-\begin{figure}[t]	% [htb]
+\begin{figure}[tb]	% [htb]
   \centering
   \epsfig{file=plots/ppx-gpu.eps, width=\columnwidth}
-  \caption{Perplexity convergence of com-DBLP on the GTX~\mbox{Titan-X}~GPU. Number
+  \caption{Perplexity convergence of com-DBLP on the Maxwell \mbox{Titan-X}~GPU. Number
   of communities chosen to maximally fill the GPU memory.}
@@
 
-\begin{figure}[t]	% [htb]
+\begin{figure}[tb]	% [htb]
   \centering
   \epsfig{file=plots/ppx-gpu-cpu.eps, width=0.45\textwidth}
-  \caption{Perplexity convergence. (a)~com-DBLP on the GTX~\mbox{Titan-X}~CPU.
+  \caption{Perplexity convergence. (a)~com-DBLP on the Maxwell \mbox{Titan-X}~GPU.
   (b)~com-LiveJournal on the multi-core~CPU. Number of communities chosen to
-  maximally fill the memory of GPU and CPU respectively.}
+  maximally fill the memory of GPU and CPU respectively. M=4K, $|\Neighbors|$=32.}
   \label{fig-ppx-gpu-cpu}
@@
 Figure~\ref{fig-ppx-gpu-cpu}(a) shows the perplexity development for the dataset
-com-DBLP, with $N=317,080$. It is computed on the GTX~\mbox{Titan-X} GPU with input
+com-DBLP, with $N=317,080$. It is computed on the Maxwell \mbox{Titan-X} GPU with input
 parameter set $K=4096$, $m=4096$, $n=32$. Convergence is reached after 6~hours.

### 2017-06-01 plots/plot-ppx.sh

@@ -23,7 +23,8 @@ set ylabel 'Perplexity'
 
 set xrange [-0.15:6.15]
 set yrange [62:95]
-set title '(a) Com-DBLP K=4K m=4K n=32'
+# set title '(a) Com-DBLP K=4K M=4K |Neighbor sample|=32'
+set title '(a) Com-DBLP K=4K'
 plot "< perl ./get_time.perl ppx-DBLP-K4K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 4
 
 set origin col(0),row(1)
@@ -31,6 +32,7 @@ set size w,h
 
 set xrange [-0.3:18.3]
 set yrange [53:83]
-set title '(b) Com-LiveJournal K=3K m=4K n=32'
+# set title '(b) Com-LiveJournal K=3K M=4K |Neighbor sample|=32'
+set title '(b) Com-LiveJournal K=3K'
 plot "< perl ./get_time.perl ppx-LJ-K3K-m4K-n32 " using (\$1/3600):3 notitle with lines lw 4
 EOF

### 2017-06-01 plots/plot-vector-sweep.sh
Forgot debug uncomment

@@ -35,5 +35,5 @@ plot \
 
 EOF
 
-# rm $TFILE
+rm $TFILE
 

### 2017-06-01 01-intro.tex
Remove contribution 'no GPU in distr' from intro
Refactor to accomodate the Pascal. Still need strategy graph, ppx line for the Pascal.

@@
 memory allocation optimizations to solve the data bandwidth issues for GPUs.
-This implementation achieves up to 86$\times$ speedup compared to the baseline.
+This implementation achieves up to 120$\times$ speedup compared to the baseline.
 %
@@
 On top of that, we pipelined the algorithm stages by fetching
-data in advance over the network;
+data in advance over the network.
 \begin{comment} Finally, the algorithm's computation
@@
 
+\begin{comment}
 \item \textbf{the use of pipelining} \emph{in the distributed implementation}
@@
 Hence, it is useless to improve the computation time.
+\end{comment}
 

### 2017-06-01 04-experimental-setup.tex

--- 2017-05-31 04-experimental-setup.tex
@@
 
-Unless indicated otherwise, the experiments described in the next sections
+Unless indicated otherwise, all experiments described in this paper
 were conducted on the VU~Amsterdam and Leiden DAS5 clusters.  The clusters
@@
 8TB of storage. The nodes of each site are interconnected by FDR~InfiniBand.
-Additionally, the VU cluster is fitted with a number of Nvidia GPUs including
-GTX~\mbox{Titan-X/Pascal},
-GTX~\mbox{Titan-X/Maxwell}, GTX980, K40c and K20m.
+Additionally, the VU cluster is fitted with a number of NVidia GPUs including
+GTX~\mbox{Titan-X(Pascal)},
+GTX~\mbox{Titan-X(Maxwell)}, GTX980, K40c and K20m.
 
-To assess the performance of our implementations, we use publicly available
-community networks from the SNAP collection~\cite{snapnets}, including the largest
-networks. Table~\ref{table-snap}
+To evaluate our implementations, we use publicly available
+community networks from the SNAP collection~\cite{snapnets}.
+Table~\ref{table-snap}
 outlines the graph properties of the networks used.

### 2017-06-01 gpu-05-03-accelerator.tex

@@
-\subsubsection{Analysis of CPU Parallelism}
-
-\begin{comment}
-  \begin{table}[tb]
-    \centering
-    \begin{tabular}{l l c c c c}
-      Comparison & Data Set        & K    & M    & n  & Iterations \\
-      \hline
-      Python/Baseline    & CA-HepPh        & 1024 & 32   & 32 & 1000 \\
-      Parallel CPU/GPU   & com-DBLP        & 1024 & 4096 & 32 & 1000 \\
-      Parallel CPU       & com-LiveJournal & 1024 & 4096 & 32 & 1000 \\
-    \end{tabular}
-    \caption{Experiment parameter reference.}
-    \label{exp-params}
-  \end{table}
-\end{comment}
-
-\begin{table}[tb]	% [htb]
-  \centering
-  \begin{tabular}{l d{2.8} d{2.8}}
-    Kernel & Time (seconds) \\
-    \hline
-    PPX CALC    &  0.0364737 \\
-    PPX ACCUM   &  0.083 \\
-    SAMPLING    &  0.535599 \\
-    UPDATE\_PHI & 25.6598 \\
-    UPDATE\_PI  &  0.645875 \\
-    THETA SUM   &  0.0483902 \\
-    GRADS PAR   &  1.92919 \\
-    GRADS SUM   &  9.31122 \\
-    UPDATE THETA&  0.0548013 \\
-    NORM THETA  &  0.001 \\
-    \hline
-    TOTAL    & 38.5858 \\
-  \end{tabular}
-  \caption{Performance break-down of multi-core CPU version without
-  vectorization. Model parameters: data set com-DBLP; K=1024, m=4096, $|\Neighbors|$=32.}
-  \label{TABLE-CPU}
-\end{table}
-
-This section discusses the use of the multi-core CPU available on the DAS5
-cluster. The parallel OpenCL version divides the work across the CPU cores and
-performs independent calculations concurrently. As shown in
-Table~\ref{TABLE-CPU}, the dominant kernel in the computation is
-\textit{update\_phi},
-which accounts for 66.5\% of the computation time. Without exploiting
-the dual 8-core processor's vectorization capabilities, the speedup relative to
-the baseline sequential C++ version is~9.8.
-
-% CPU
-\begin{figure}[tb]
-\centering
-\epsfig{file=plots/cpu-vector.eps, width=\columnwidth}
-\caption{Performance of CPU for varying vector width.}
-\label{FIG-CPU-VECTOR}
-\end{figure}
-
-In addition to applying computations in parallel, we investigated the use of
-the CPU's SIMD instructions to maximize its resource utilization.
-Figure~\ref{FIG-CPU-VECTOR} presents the performance obtained by vectorizing
-the kernels with varying vector widths. A key aspect in this figure is the
-diminishing performance benefit for higher vector widths. As the computational
-performance increases, the memory throughput becomes the leading performance
-bottleneck. Moreover, using 16-wide SIMD instructions gave a slight performance
-penalty compared to 8-wide SIMD. The 8-wide vector version improves the speedup
-relative to the baseline version from 9.8 to~20.9.
-
 \begin{table}[tb]	% [htb]
@@
 Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
-GTX~\mbox{Titan-X} Maxwell and Pascal, GTX980, K40c, K20m and the Intel Xeon CPU relative to the baseline
+GTX~\mbox{Titan-X} Maxwell and Pascal, GTX980, K40c and K20m relative to the baseline
 sequential C++ version. These results are consistent with the relative
 capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
-the \mbox{Titan-X(Pascal)} achieves the highest speedup of 120 relative to the baseline. At the
-other end, the parallel CPU version attains a speedup factor of~20.9.
+the \mbox{Titan-X(Pascal)} achieves the highest speedup of 120 relative to the baseline.
 

### 2017-06-01 gpu-05-05-perplexity.tex

--- 2017-05-25 gpu-05-05-perplexity.tex
@@
   \epsfig{file=plots/ppx-gpu.eps, width=\columnwidth}
-  \caption{Perplexity convergence of com-DBLP on the GTX~\mbox{Titan-X}~GPU. Number
+  \caption{Perplexity convergence of com-DBLP on the Maxwell \mbox{Titan-X}~GPU. Number
   of communities chosen to maximally fill the GPU memory.}
@@
   \epsfig{file=plots/ppx-gpu-cpu.eps, width=0.45\textwidth}
-  \caption{Perplexity convergence. com-DBLP on the GTX~\mbox{Titan-X}~GPU.
+  \caption{Perplexity convergence. com-DBLP on the Maxwell \mbox{Titan-X}~GPU.
   Number of communities chosen to
-  maximally fill the GPU memory.}
+  maximally fill the GPU memory.
+  M=4K, $|\Neighbors|$=32.}
   \label{fig-ppx-gpu-cpu}
@@
 Figure~\ref{fig-ppx-gpu-cpu}(a) shows the perplexity development for the dataset
-com-DBLP, with $N=317,080$. It is computed on the GTX~\mbox{Titan-X} GPU with input
+com-DBLP, with $N=317,080$. It is computed on the Maxwell \mbox{Titan-X} GPU with input
 parameter set $K=4096$, $M=4096$, $|\Neighbors|=32$. Convergence is reached after 6~hours.

### 2017-06-01 01-intro.tex
Tiny typos etc

--- 2017-05-24 01-intro.tex
@@
 nature, the majority of its memory access patterns and data dependencies
-are non-deterministic. As a result, straightforward optimization attempts
+are \emph{non-deterministic}. As a result, straightforward optimization attempts
 of the memory access patterns either fail or lead to non-intuitive results. Third,
@@
 Starting from an existing implementation in Python/Numpy, we create an efficient
-C++ version to serve as the baseline for the parallel versions.
+baseline version in C++ to serve as the baseline for the parallel versions.
 This already improves performance by a factor of 1000 to 1500 compared to
-Python. Then, our key contributions to the algorithm are:
-
+Python. Then, our key contributions are:
 \begin{itemize}
-\item \textbf{scaling up to run on GPUs.}
-The application is not a natural fit for GPUs, because it is memory-bound and
-irregular in its data accesses. Nevertheless, we developed aggressive
-memory allocation optimizations to solve the data bandwidth issues for GPUs.
-This implementation achieves up to 86$\times$ speedup compared to the baseline.
+\item
+our implementation is scaled up to run on GPUs.
+We developed aggressive optimizations to solve the memory bandwidth issues for
+GPUs. This implementation achieves speedup factors up to~120,
+compared to the baseline.
 %
-We managed to reduce the algorithm's memory requirements by roughly 75\%, which
-significantly reduces the data intensity and allows for tackling larger
+Through careful analysis of the computation and data structures we show that
+the algorithm's full state can be reduced by roughly 75\%. Compressing the
+state significantly reduces the data intensity and allows for tackling larger
 problems; % while maintaining all state in memory.
 
-\item \textbf{scaling out to overcome the memory bounds posed by GPUs} or,
-generally, single-machine systems, to solve the largest publicly available
+\item
+our implementation is scaled out to overcome the memory bounds posed by GPUs or
+single-machine systems in general, to solve the largest publicly available
 community graphs. % This necessitated overcoming several challenges.
@@
 
-\item \textbf{the use of pipelining} allows to use straightforward intra-node
-thread parallelism in the distributed implementation, and it is pointless
-to combine the GPU work into it. The cause is that the computation can be
-fully overlapped with fulfilling the data requests for the next iteration.
+\item we argue that it is valid to use a CPU version with straightforward
+intra-node thread parallelism for the distributed implementation, and
+that it is pointless to combine the GPU work into it.
 
@@
 parallel implementations. Section~\ref{sec-gpu} delves
-into the parallelization optimizations for GPUs.
+into the parallelization optimizations for GPUs and multi-core CPUs.
 Section~\ref{sec-distr} describes the design, implementation and evaluation of

### 2017-06-01 03-sequential-evaluation.tex
Tiny typos etc

@@
 In production runs, the perplexity calculation is invoked once every hunderds
-or thousands iterations. Hence, the perplexity timings are amortized over
+or thousands of iterations. Hence, the perplexity timings are amortized over
 100~iterations.
@@
 The translation from Python into 'Python Idiom' C++ speeds up this stage
-by a factor~171. % , even though the Python implementation uses numpy for its data
+by a factor~171. % , even though the Python implementation uses Numpy for its data
 % structures. 
@@
 \textit{update\_pi} has a disproportionately large speed
-gain after removing inefficiencies. The numpy implementation updated the full
+gain after removing inefficiencies. The Numpy implementation updated the full
 $\pi$ array (incidentally, rather efficiently), whereas our C++ optimization

### 2017-06-01 distr-03-design.tex
Refactor distr so there is a design section again

--- 2017-05-24 distr-03-design.tex
@@
 
+\subsection{Design of distributed implementation}
 % intro parallelisation, distribution
 
-This section describes how we created a distributed version of the C++
+This section describes how we create and evaluate a distributed version of the C++
 algorithm. We didn't use the advanced intra-node parallelization techniques
@@
 
-\subsection{Data distribution}
+\subsubsection{Data distribution}
 
@@
 
-\subsection{RDMA Distributed Key-Value Store}
+\subsubsection{RDMA Distributed Key-Value Store}
 
@@
 of important aspects. First, the number of keys and values to be stored are
-constant throughout the execution of the algorithm. For instance, no records
-are dynamically inserted or deleted from the DKV after their initial insertion,
-although the value associated with each key can be updated. Additionally, all
-keys are of the same type and represent a unique vertex in the graph.
-Therefore, keys can be represented as integer identifiers. Moreover, each key
-points to a vector of $K+1$ floats as its value.
+constant throughout the execution of the algorithm: no records
+are dynamically inserted or deleted from the DKV store after their initial insertion,
+although the value for any key can be updated. Additionally, all
+keys are of the same type and represent a unique vertex in the graph
+so, they can be represented as integer identifiers. Each key
+has a vector of $K+1$ floats as its value.
 %
@@
 computation is partitioned into stages that are separated by barriers. A
-single stage either reads values or updates values. The algorithm ensures that updates
-are always targeting unique elements,
-therefore, there are no read/write or write/write hazards. The existing DKV
+single stage either reads values or updates values, and the algorithm ensures that updates
+are always targeting unique elements.
+Therefore, there are no read/write or write/write hazards. The existing DKV
 stores (RamCloud~\cite{Ousterhout:2015:RSS:2818727.2806887}, FaRM~\cite{179767},
@@
 
-\subsection{Implementation of distributed parallelism}
+\subsubsection{Implementation of distributed parallelism}
 
-This section describes the parallelization of each of the stages of the
-algorithm's main loop. All but the first stage justify parallelization for
+This section describes the parallelization of the stages of the
+algorithm's main loop. All but the first stage,
+mini-batch selection (line 3 in Algorithm~\ref{algorithm}),
+justify parallelization for
 high values of combinations of the parameters mini-batch size $|\Minibatch|$,
 number of vertices in the mini-batch $M$,
-number of communities $K$, neighbor sample size $|\Neighbors|$, and held-out set
-size $|\cE_h|$.
+number of communities $K$, and sizes of neighbor sample set~$|\Neighbors|$ and held-out
+set~$|\cE_h|$.
 
-The first stage, mini-batch selection (line 3 in Algorithm~\ref{algorithm}),
-is done by the master and it is
+Mini-batch selection is done by the master and it is
 not itself parallelized. However, the distributed implementation overlaps
@@
 $\pi/\phi$ for the mini-batch vertices. The update to $\pi$ requires
-$M \times K$ operations, and it is done in parallel over mini-batch vertices. Because of
+$M \times K$ operations, and it is done in parallel over mini-batch vertices. For
 memory consistency, this stage awaits completion of \textit{update\_phi} with
@@
 barrier to ensure that up-to-date $\pi$ values are read. \textit{update\_beta}
-is split into four steps. The first stage partitions the mini-batch across
+is split into four steps. The first step partitions the mini-batch across
 machines to calculate contributions to $g_{ab}(\theta)$. The values of $\pi$ for
@@
 
-\subsection{Pipelining of computation and loading $\pi$}
+\subsubsection{Pipelining of computation and loading $\pi$}
 

### 2017-06-01 distr-04-evaluation.tex
Tiny textual fixes
Refactor distr so there is a design section again

@@
 All of the experiments reported in this section utilized
-publicly available graphs from the Stanford Large Network Data set
-Collection~(SNAP) \cite{snapnets}. Table~\ref{table-snap} lists the
-collection of data sets used.
+publicly available graphs from the SNAP collection, see Table~\ref{table-snap}.
 
@@
 %
-In addition to the total execution time, Figure~\ref{fig-strong-scaling}-a
+In addition to the total execution time, \mbox{Figure~\ref{fig-strong-scaling}(a)}
 presents the cumulative time spent in individual computational phases across
-iterations. Moreover, Figure~\ref{fig-strong-scaling}-b presents the speedup
+iterations. Moreover, \mbox{Figure~\ref{fig-strong-scaling}(b)} presents the speedup
 achieved for the same experiments reported in
-Figure~\ref{fig-strong-scaling}-a. As is clearly shown, the dominant phase of
+Figure~\ref{fig-strong-scaling}(a). As is clearly shown, the dominant phase of
 the execution is \textit{update\_phi\_pi}.
@@
 Naturally, increasing the number of communities causes a proportional increase
-in execution time. However, when pipelining is enabled, some of the
-incurred network latency is hidden by overlapping it with computation.
+in execution time. However, when pipelining is enabled, the
+incurred network latency is partly hidden by overlapping it with computation.
 Moreover, since both computation time and network latency increase with larger

### 2017-06-01 03-sequential-design.tex

--- 2017-05-24 gpu-04-02-sequential.tex
@@
 implemented in Python~\cite{LiAW15}, as is common
-with machine learning researchers and practitioners.
+in the machine learning community.
 We converted this implementation into an efficient sequential C++ application
 to serve as a baseline for the parallel and distributed implementations.
-This is certainly not one of our contributions to the parallelization
+We certainly do not consider this to be a contribution to the parallelization
 of the algorithm, but for a fair analysis of the parallel implementations
@@
 The Python implementation
-relied on numpy~\cite{numpy} to perform numerical computations efficiently.
+relied on Numpy~\cite{numpy} to perform numerical computations efficiently.
 Besides, the algorithm
 made heavy use of Python data structures such as sets, dictionaries and
-lists, which have no efficient numpy equivalent.
+lists, which have no efficient Numpy equivalent.
 
@@
 For example, one recurring idiom in the Python implementation
-was an expression of the form $a^y b^{1-y}$, where $y$ is either 0 or~1.
+was a choice expression of the form $a^y b^{1-y}$, where $y$ is either 0 or~1.
 % Therefore, the expression is simply a choice

### 2017-06-01 distr-03-design.tex

@@
 This section describes how we create and evaluate a distributed version of the C++
-algorithm. We didn't use the advanced intra-node parallelization techniques
-described in the previous section, but found that straightforward
-multi-threaded parallelization techniques sufficed: the distributed performance
-is dominated by (remote) memory latency, not by node parallelism. As our
-analysis will show, our pipelining optimizations will fully overlap the
-parallel calculations with (pre)loading the data from remote hosts.
-The details of the multi-threaded parallelism
-will be discussed only where appropriate. The
-distributed design follows a master-worker paradigm, where
-the master controls the sequence of parallel operations and the workers perform the
-calculations. For thread parallelism, we annotate the program with
-OpenMP~\cite{OpenMPSpec}. For distributed communication, we use
-MPI~\cite{Forum:1994:MMI:898758}.
+algorithm. We used the multicore CPUs in the DAS5 with straightforward
+intra-node thread parallelisation, not the GPU implementations described
+in the previous section. The justification is twofold: in our DAS5 system,
+it is pointless to speed up the intra-node performance beyond straightforward
+multithreading, because distributed performance is dominated by remote memory
+latency; the analysis in Section~\ref{eval-pipeline} shows that the time for
+(pre)loading the data from remote hosts greatly exceeds the calculation time.
+The calculation time is fully overlapped in our pipelining implementation.
+Moreover, usage of GPUs in the distributed setting would necessitate staging of
+the relevant parts of $\pi/\phi$ from the host to GPU memory in each iteration,
+which would eat most of the performance gain offered by the GPU because
+the algorithm is so data-intensive.
+% The details of the multi-threaded parallelism
+% will be discussed only where appropriate.
+The distributed design
+follows a master-worker paradigm, where the master controls the sequence of
+parallel operations and the workers perform the calculations. For thread
+parallelism, we annotate the program with OpenMP~\cite{OpenMPSpec}. For
+distributed communication, we use MPI~\cite{Forum:1994:MMI:898758}.
 
@@
 \subsubsection{Pipelining of computation and loading $\pi$}
+\label{sec-pipelining}
 

### 2017-06-01 06-conclusion.tex
Refactor to accomodate the Pascal. Still need strategy graph, ppx line for the Pascal.
Some text redaction

--- 2017-05-10 06-conclusion.tex
@@
 The SG-MCMC algorithm discussed in this paper posed additional computational
-challenges due to its unique stochastic nature and high data intensity. Unlike
+challenges due to its unique stochastic nature and \emph{high data intensity}. Unlike
 common machine learning algorithms, its data dependencies and memory access
-patterns are nondeterministic. In this paper, we presented our methodology of
-improving the algorithm's performance by fundamentally restructuring it to
-cater for concurrency.
+patterns are \emph{nondeterministic}. In this paper, we presented our methodology of
+improving the algorithm's performance by restructuring it to
+cater for concurrent and distributed parallelism.
 
 We showed that the algorithm's state can be reduced by 75\% after a thorough
-analysis of the computational patterns and data structures which significantly
-reduces its data intensity.
+analysis of the computational patterns and data structures, thus significantly
+reducing its data intensity.
 
-In our optimization for manycore processors, we navigated the complex optimization
-landscape by dynamically generating kernel code and testing different
+In our implementation for many-core processors, we navigated the complex optimization
+landscape by dynamically generating kernel code and investigating different
 combinations of optimizations. The outcome of these efforts culminated in
-significant speedup factors of 21 and 86 using a multi-core CPU and a GPU
+significant speedup factors of 21 and 120 using a multi-core CPU and a GPU
 respectively. These speedup numbers were achieved in comparison to an already
-optimized sequential implementation. Finally, we evaluated the performance of
-the parallel algorithm across several GPUs, highlighting the difference between
-their optimal configurations.
+optimized sequential implementation. The evaluation of the performance
+across several GPUs yielded one striking find. No single optimization strategy
+works best for all GPUs: different choices for memory allocation strategy,
+kernel block size and vector width were required for optimal results for
+different GPUs and application instances.
 %
-The outcome of this study reinforces the significance of avoiding premature
-optimization as it can lead to unexpected results. In particular, the success
-of common GPU optimizations depends on the particular device in use and the
-problem it is applied to.
+Hence, the outcome of this study reinforces the significance of avoiding premature
+optimization. In particular, the success
+of common GPU optimizations depends on the device and the
+problem size it is applied to.
 
 We further created a highly scalable implementation to handle the largest existing
-community graphs.
-This design faced
-several obstacles in order to achieve high performance.
-The algorithm was restructured to facilitate its
-parallelization. We overlapped computation
+community graphs on a distributed cluster machine.
+This design solved
+several problems to achieve scalability and high performance.
+We overlapped computation
 with communication to hide latency.  We use a
-mixture of MPI and RDMA primitives to speedup the communication between cluster
+mixture of MPI and RDMA primitives to speed up the communication between cluster
 nodes.
@@
 demonstration of the implementation's utility was provided by processing 6
-different organic data sets.
+large real-world data sets.
 To the best of

### 2017-06-01 06-conclusion.tex

@@
 combinations of optimizations. The outcome of these efforts culminated in
-significant speedup factors of 21 and 120 using a multi-core CPU and a GPU
-respectively. These speedup numbers were achieved in comparison to an already
+significant speedup factor of 120 for the \mbox{Titan-X(Pascal)} GPU.
+This speedup was achieved in comparison to an already
 optimized sequential implementation. The evaluation of the performance

### 2017-06-01 distr-04-evaluation.tex
Some text redaction

@@
 
-As discussed in Section~\ref{design-section}, we implemented pipeline
+As discussed in Section~\ref{sec-pipelining}, we implemented pipeline
 parallelism at
@@
 the targeted number of communities. To the best of our knowledge, the data sets
-used in this study are the largest publicly available organic graphs.
+used in this study are the largest publicly available real-world graphs.

### 2017-06-01 01-intro.tex
Text thingy in Related

@@
 
-Starting from an existing implementation in Python/Numpy, we create an efficient
+Starting from an existing implementation in Python/Numpy, we created an efficient
 baseline version in C++ to serve as the baseline for the parallel versions.

### 2017-06-01 02-02-related.tex
Text thingy in Related

--- 2017-05-30 02-02-related.tex
@@
 We consider related work for both parts of this paper: acceleration and scaling
-of our MCMC-MMSB algorithm. Acceleration of deep learning algorithms on GPUs
-is an ongoing success story. Although approximative Baysian algorithms are
-not such a natural fit for GPUs, a number of projects explores this terrain.
+of our MCMC-MMSB algorithm. Whereas acceleration of \emph{deep learning algorithms} on GPUs
+is an ongoing success story, approximative Baysian algorithms are
+not such a natural fit for GPUs. Nevertheless, a number of projects explores this terrain.
 Medlar~c.s.~\cite{journals/bioinformatics/MedlarGSBK13} use GPUs with their

### 2017-06-01 gpu-04-03-accelerator.tex
Restructure GPU design to use subsubsections

--- 2017-05-24 gpu-04-03-accelerator.tex
@@
 
-\subsection{Design of GPU Optimizations}
+\subsection{Design of Accelerator Optimizations}
 
@@
 crucial modifications to allow for efficient parallelization. This section
-discusses the key design contributions to attain optimized
-parallel versions. The state represented by $\pi$ and $\phi$ is permanently
-stored in the GPU memory, because staging of the relevant portions of these
-data structures in each iteration is too costly compared to the amount
-of computation. The graph is stored at the host, because mini-batch
-selection is cheap, and the data required for a minibatch is small.
-
-Since the algorithm is memory-bound, many of the optimizations concern
-memory usage. We offer a fast, memory-efficient hash table taylored towards
-the algorithm's usage patterns. Some optimizations reduce the total memory
-footprint of the algorithm, following analysis of the algorithm. Other
-optimizations explore the opportunities offered by the faster regions of the
-GPU memory hierarchy: data can be put into thread-local memory, in shared
-GPU memory, or in registers which requires a more complicated analysis. One
-other optimization exploits the vectorization possibilities.
+discusses the key design contributions to attain two optimized
+parallel versions, catering for multi-core CPUs and GPUs. First, we present
+common structural changes that provide a foundation for compute device specific
+optimizations. Next, we provide an in-depth description of the optimized CPU
+and GPU versions respectively.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \subsection*{Using OpenCL and CUDA}
+\subsubsection{Using OpenCL and CUDA}
 %OpenCL/CUDA
-\textit{\bf !!! Remove the history approach !!!}
 Initially, the system was developed using OpenCL~\cite{opencl} in order to
 accelerate the computations of the algorithm. OpenCL was chosen as it provides
-a common abstraction for a variety of compute devices.
-However, NVidia's OpenCL SDK limits the
+a common abstraction for a variety of compute devices which fulfilled our
+requirements of employing CPUs or GPUs. However, Nvidia's OpenCL SDK limits the
 total memory allocations within a context to 4GB which severely limits the
@@
 implementation was developed that restricts the set's features based on an
-observation of its usage patterns. For example, the set is used as a container
-for the graph edges which are known in advance. Therefore, the set's maximum
+observation of its usage patterns.
+The graph is immutable, hence its maximum
 size can be specified at creation time and need not be adjustable or provide
-thread-safety guarantees as the graph is immutable. Further, the set data
+thread-safety guarantees. The set data
 structure is built as a variant of a cuckoo
@@
 uses two hash functions to index their corresponding spaces. Additionally, the
-cuckoo hash allows for 4 keys per bucket. This hash design allows for a loading
-factor upwards of 90\% which reduces the hash space overhead
-significantly.
+cuckoo hash allows for 4 keys per bucket. This hash design allows for a load
+factor upwards of 90\% which allows a very economical memory usage.
 %% FIXME: Talk lookup hit/miss with 4 keys per bucket?
@@
 % Division into kernels
-The original algorithm was structurally reorganized into 4 main sections, each with
-one or more kernels, data dependencies and synchronization requirements.
+The original algorithm was structurally reorganized into 4 main sections with
+one or more kernels in each depending on data dependencies and the applicable
+synchronization requirements.
 %
@@
 Second, the predominant kernel, \textit{update\_phi} is invoked to calculate the
-gradients for each vertex $i$ in the mini-batch \Minibatch and update the value of $\phi_i
+gradients for each vertex $i$ in the mini-batch \Minibatch and updates the value of $ \phi_i
 \mid i \in \Minibatch$. Next, the \textit{update\_pi} kernel is invoked to normalize the
@@
 %
-Finally, perplexity calculation is performed in a dedicated kernel.
+Finally, the perplexity calculation is performed in a dedicated kernel.
 
@@
 %
-This change affects the algorithm in multiple ways. For instance, on some GPUs it
-reduces
-the computation intensity besides the data intensity. It again halves the memory
+This change affects the algorithm in multiple ways. For instance, on a GPU it reduces
+both the computation and data intensity. It again reduces the memory
 footprint of the algorithm and frees registers.
@@
 
-\subsubsection{Edge-centric and Vertex-centric kernels}
-The work decomposition of the kernels ensures that each thread performs
+\subsubsection{CPU-Specific Optimization}
+The multi-core
+CPU version of the algorithm uses OpenCL to perform its computations. The
+work decomposition of the CPU kernels ensures that each thread performs
 independent computation in order to avoid expensive synchronization operations.
-Edge-centric kernels, which operate over mini-batch edges, perform computations
+Edge-centric kernels, which operate over mini-batch edges perform computations
 over every edge in parallel.  Similarly, vertex-centric kernels that operate
@@
 % vectorization
-Additionally, the kernels were vectorized to decrease instruction overhead.
+Additionally, the kernels were vectorized to decrease instruction overhead and
+utilize the SSE capabilities of the CPU cores.
 
-\subsubsection{Memory Hierarchy Optimizations}
+\subsubsection{GPU-Specific Optimization}
 \label{gpu-design}
-% The GPU implementation builds on top of the CPU work decomposition scheme.
-% However, instead of having every thread perform independent computations, each
-% block of threads shares the work associated with the single edge or vertex for
-% edge-centric and vertex-centric kernels.
+The GPU implementation builds on top of the CPU work decomposition scheme.
+However, instead of having every thread perform independent computations, each
+block of threads shares the work associated with the single edge or vertex for
+edge-centric and vertex-centric kernels.
 

### 2017-06-01 gpu-04-03-accelerator.tex

--- 2017-05-30 gpu-04-03-accelerator.tex
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % \subsection*{Using OpenCL and CUDA}

### 2017-06-01 gpu-04-design.tex
Remove the history approach from OpenCL. Remove OpenCL
Rephrase GPU design intro for better overview

@@
 
-Based on the efficient C++ baseline version, a version for GPU accelaration
-was constructed.
-This involved optimizations for both efficient resource utilization and
-parallelization.
-This section provides an overview of the system's evolution in incremental
-phases, identifying the key contributions and differences between consecutive
-states. This section concludes with an in-depth evaluation of the performance
-of our various optimizations on a number of GPU
-accelerators.
+As the first contribution of this paper, we created optimized versions for
+accelerators of the MCMC-aMMSB
+algorithm, based on the efficient C++ baseline version.
+The algorithm is data-intensive and nondeterministic in its memory accesses,
+which makes an efficient GPU implementation highly nontrivial. Hence, profound
+changes to the algorithm and the data structures were necessary,
+both for efficient resource utilization and parallelization.
+% This section provides an overview of the system's evolution in incremental
+% phases, identifying the key contributions and differences between consecutive
+% states.
+This section describes the various optimization steps in technical detail.
+It concludes with an in-depth evaluation of the performance
+of our optimizations,
+on multi-core CPUs as well as a number of GPU accelerators.
 

### 2017-06-01 03-sequential-design.tex
Some text rephrasings

@@
 We converted this implementation into an efficient sequential C++ application
-to serve as a baseline for the parallel and distributed implementations.
-We certainly do not consider this to be a contribution to the parallelization
+to serve as a baseline for the parallel and distributed implementations that
+are the subject of the rest of this paper.
+We certainly do not consider this sequential implementation to be a contribution to the parallelization
 of the algorithm, but for a fair analysis of the parallel implementations

### 2017-06-01 04-experimental-setup.tex
Expand SNAP on first reference

--- 2017-06-01 04-experimental-setup.tex
@@
 
-To assess the performance of our implementations, we use publicly available
-community networks from the SNAP collection~\cite{snapnets}, including the largest
-networks. Table~\ref{table-snap}
+To evaluate our implementations, we use publicly available
+community networks from the Stanford Network Analysis Project (SNAP)
+collection~\cite{snapnets}, including the largest networks.
+Table~\ref{table-snap}
 outlines the graph properties of the networks used.

### 2017-06-01 04-experimental-setup.tex

--- 2017-06-01 04-experimental-setup.tex
@@
 To evaluate our implementations, we use publicly available
-community networks from the SNAP collection~\cite{snapnets}.
+community networks from the Stanford Network Analysis Project (SNAP)
+collection~\cite{snapnets}, including the largest networks.
 Table~\ref{table-snap}

### 2017-06-01 gpu-04-03-accelerator.tex
NVidia iso Nvidia, small rephrase

--- 2017-06-01 gpu-04-03-accelerator.tex
@@
 a common abstraction for a variety of compute devices which fulfilled our
-requirements of employing CPUs or GPUs. However, Nvidia's OpenCL SDK limits the
+requirements of employing CPUs or GPUs. However, NVidia's OpenCL SDK limits the
 total memory allocations within a context to 4GB which severely limits the

### 2017-06-01 gpu-05-evaluation.tex
NVidia iso Nvidia, small rephrase

--- 2017-05-18 gpu-05-evaluation.tex
@@
 Next, we assess the trade-offs associated with each of the GPU optimization
-strategies and their performance effects on four GPUs spanning two chip
+strategies and their performance effects on five GPUs spanning three chip
 architecture generations.
@@
 
-As noted in Section~\ref{sec-experimental}, all experiments were conducted on the VU Amsterdam DAS5 cluster.
-Table~\ref{table-gpus} presents the properties of the various GPUs in
-the DAS5 cluster.
+\input{04-experimental-setup}
+
+% As noted in Section~\ref{sec-experimental}, all experiments were conducted on the VU Amsterdam DAS5 cluster.
+% Table~\ref{table-gpus} presents the properties of the various GPUs in
+% the DAS5 cluster.
 

### 2017-06-01 gpu-05-evaluation.tex

--- 2017-05-30 gpu-05-evaluation.tex
@@
 We assess the trade-offs associated with each of the GPU optimization
-strategies and their performance effects on four GPUs spanning two chip
+strategies and their performance effects on five GPUs spanning three chip
 architecture generations.

### 2017-06-01 gpu-04-03-accelerator.tex
Some text thingies

--- 2017-06-01 gpu-04-03-accelerator.tex
@@
 are concerned with
-memory usage. For starters, we offer a fast, memory-efficient hash table on the GPU,
+memory usage. For starters, we describe our fast, memory-efficient hash table on the GPU,
 taylored towards
-the algorithm's usage patterns. Some optimizations reduce the total memory
+the algorithm's usage patterns. Then some optimizations reduce the total memory
 footprint of the algorithm. Other
@@
 data structures to the GPU in each iteration is too costly compared to the amount
-of computation. The graph is stored at the host, because mini-batch
-selection is cheap, and the data required for a minibatch is small.
+of computation. The graph is stored both at the host and in GPU memory. The
+host draws the mini-batches, because that operation is cheap,
+and the data required for a minibatch is small.
 

### 2017-06-01 06-conclusion.tex
Work on conclusion: try to better show the contributions

--- 2017-06-01 06-conclusion.tex
@@
 
-We showed that the algorithm's state can be reduced by 75\% after a thorough
-analysis of the computational patterns and data structures, thus significantly
+After a thorough
+analysis of the computational patterns and data structures,
+we showed that the algorithm's state can be reduced by 75\%,
+thus significantly
 reducing its data intensity.
 
-In our implementation for many-core processors, we navigated the complex optimization
-landscape by dynamically generating kernel code and investigating different
-combinations of optimizations. The outcome of these efforts culminated in
-significant speedup factors of 21 and 120 using a multi-core CPU and a GPU
+In our implementation for accelerators, we navigated the complex optimization
+landscape by dynamically generating kernel code for finest control of allocation in
+the GPU memory hierarchy, and by tuning parameters like kernel block size and vector
+width. These efforts resulted in
+a significant speedup of 21$\times$ and 120$\times$ using a multi-core CPU and a GPU
 respectively. These speedup numbers were achieved in comparison to an already
@@
 kernel block size and vector width were required for optimal results for
-different GPUs and application instances.
+different GPUs and problem instances. Counter-intuitively, some GPUs performed
+best with vectorization disabled.
 %
-Hence, the outcome of this study reinforces the significance of avoiding premature
+The outcome of this study therefore reinforces the significance of avoiding premature
 optimization. In particular, the success
@@
 community graphs on a distributed cluster machine.
-This design solved
-several problems to achieve scalability and high performance.
-We overlapped computation
-with communication to hide latency.  We use a
+We made several contributions to
+achieve scalability and high performance:
+the most important data structures are partitioned across the collective memory of the
+cluster nodes;
+we use a
 mixture of MPI and RDMA primitives to speed up the communication between cluster
-nodes.
+nodes; and
+we overlapped computation
+with communication to hide latency.
 %
-Further, we conducted a thorough empirical evaluation of the system to study its strong
-and weak scalability on 65 cluster nodes using large data sets.
+By evaluating the system empirically using 65 cluster nodes and large data sets,
+we show that its strong and weak scalability are both good. We also compared its
+performance against a 40-core shared-memory system with 1TB of memory.
+The distributed system has far stronger performance because it has more cores and
+more memory bandwidth, although it must fetch its data through the network.
 % Additionally, we assessed the efficiency of the algorithm's resource utilization.

### 2017-06-01 06-conclusion.tex

@@
 width. These efforts resulted in
-a significant speedup of 21$\times$ and 120$\times$ using a multi-core CPU and a GPU
-respectively. These speedup numbers were achieved in comparison to an already
+a significant speedup of 21$\times$ on a GPU.
+This speedup is achieved in comparison to an already
 optimized sequential implementation. The evaluation of the performance

### 2017-06-01 gpu-04-03-accelerator.tex
Rephrase 'allows' in Cuckoo hash

--- 2017-06-01 gpu-04-03-accelerator.tex
@@
 uses two hash functions to index their corresponding spaces. Additionally, the
-cuckoo hash allows for 4 keys per bucket. This hash design allows for a load
-factor upwards of 90\% which allows a very economical memory usage.
+cuckoo hash supports upt to 4~keys per bucket. This hash design allows for a load
+factor upwards of 90\%, a very economical memory usage.
 %% FIXME: Talk lookup hit/miss with 4 keys per bucket?

### 2017-06-01 02-02-related.tex
Nitpick in distr Related Work

@@
-\section{Related Work}
+\subsection{Related Parallelization Work}
 \label{sec-related}
 
-We consider related work for both parts of this paper: acceleration and scaling
+We consider related work for both parallelization contributions of this paper: acceleration and scaling
 of our MCMC-MMSB algorithm.  Whereas acceleration of \emph{deep learning algorithms} on GPUs
@@
 can handle are limited by the device memory of the GPU. Our implementation converges
-quickly on the GPU but can also use multi-core CPUs. In the latter case, speedup is
-limited but the larger memory allows tackling larger data sets.
+quickly on the GPU but can also use clusters of multi-core CPUs. In the latter case, speedup is
+lower but the distributed memory allows tackling much larger data sets.
 

### 2017-06-13 01-intro.tex
Process Henri's comments 1june (except Maxwell/Pascal analysis for GPU strategies)

--- 2017-06-01 01-intro.tex
@@
 \section{Introduction}
-The tremendous amount of data that we generate through our daily applications such as social networking services, online shopping, and news recommendations, provides us with an opportunity to extract hidden but useful, even invaluable information.
-Realizing this opportunity, however, requires a significant amount of effort because traditional machine learning algorithms often become extremely inefficient with large amounts of data.
+The tremendous amount of data that we generate through our daily applications such as
+social networking services, online shopping, and news recommendations, provides us
+with an opportunity to extract hidden but useful, even invaluable information.
+Realizing this opportunity, however, requires a significant amount of effort because
+traditional machine learning algorithms may become inefficient with large amounts of data.
 
-There have been two main approaches to resolving this issue; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic infrastructure systems which can be leveraged to construct machine learning solutions more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
+There have been two main approaches to resolving this issue; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic software infrastructure systems which can be leveraged to construct machine learning solutions more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
 However, the best possible performance is often achieved by carefully integrating both of these approaches in a single solution.
@@
 work of the authors of this paper~\cite{LiAW15} showed that this methodology is faster and more accurate than the
-SVB method. That work further proposes a heuristic to scale up the number of
-communities at the cost of less precision; the work in this paper considers the
-algorithm without that heuristic.
+SVB method.
 
@@
 memory allocation optimizations to solve the data bandwidth issues for GPUs.
-This implementation achieves up to 120$\times$ speedup compared to the baseline.
+This implementation achieves up to 120$\times$ speedup compared to the C++ baseline.
 %
-We managed to reduce the algorithm's memory requirements by roughly 75\%, which
+We reduced the algorithm's memory requirements by roughly 75\%, which
 significantly reduces the data intensity and allows for tackling larger
@@
 
-\item \textbf{scaling out to overcome the memory bounds posed by GPUs} or,
-generally, single-machine systems, to solve the largest publicly available
+\item \textbf{scaling out to overcome the memory bounds} posed by 
+single-machine systems, to be the first to solve the largest publicly available
 community graphs. % This necessitated overcoming several challenges.
@@
 
-\begin{comment}
-\item \textbf{the use of pipelining} \emph{in the distributed implementation}
-justifies straightforward intra-node thread parallelism, and it is pointless
-to combine the GPU implementations
-into it. The cause is that computing for an iteration takes less
+The use of pipelining makes it pointless to combine GPU acceleration
+into the distributed implementation: computing for an iteration takes less
 time than fetching the data from remote memory, and the computation is
 fully overlapped with fulfilling the data requests for the next iteration.
-Hence, it is useless to improve the computation time.
-\end{comment}
+% Hence, it is useless to improve the computation time.
 
@@
 
-A bird's eye-view of the work in this paper, without any technical detail, has been
+A bird's-eye view of the work in this paper, without any technical detail, has been
 published at CCGrid-2016~\cite{10.1109/CCGrid.2016.98}.

### 2017-06-13 06-conclusion.tex
Process Henri's comments 1june (except Maxwell/Pascal analysis for GPU strategies)

@@
 width. These efforts resulted in
-a significant speedup of 21$\times$ on a GPU.
+a significant speedup of 120$\times$ on a GPU.
 This speedup is achieved in comparison to an already
@@
 We further created a highly scalable implementation to handle the largest existing
-community graphs on a distributed cluster machine.
-We made several contributions to
-achieve scalability and high performance:
-the most important data structures are partitioned across the collective memory of the
-cluster nodes;
-we use a
-mixture of MPI and RDMA primitives to speed up the communication between cluster
-nodes; and
+community graphs on a distributed cluster machine. The most important data structures
+are partitioned across the collective memory of the cluster nodes, which
+significantly increases data access latency for our already data-dominant algorithm.
+We made several contributions to improve performance and still maintain scalabitiy:
+we created a custom distributed Key-Value store implemented with high-speed RDMA
+primitives that exploits the algorithm's structure to avoid concurrency issues; we
+use MPI for synchronization and work distribution between the master and worker nodes;
+and
 we overlapped computation
@@
 The distributed system has far stronger performance because it has more cores and
-more memory bandwidth, although it must fetch its data through the network.
+more collective memory bandwidth, even though it must fetch its data through the network.
 % Additionally, we assessed the efficiency of the algorithm's resource utilization.

### 2017-06-13 distr-03-design.tex
Process Henri's comments 1june (except Maxwell/Pascal analysis for GPU strategies)

@@
 This section describes how we create and evaluate a distributed version of the C++
-algorithm. We used the multicore CPUs in the DAS5 with straightforward
+algorithm. We used multicore CPUs with straightforward
 intra-node thread parallelisation, not the GPU implementations described
-in the previous section. The justification is twofold: in our DAS5 system,
+in the previous section. The justification is twofold:
 it is pointless to speed up the intra-node performance beyond straightforward
@@
 The calculation time is fully overlapped in our pipelining implementation.
-Moreover, usage of GPUs in the distributed setting would necessitate staging of
+Secondly, usage of GPUs in the distributed setting would necessitate staging of
 the relevant parts of $\pi/\phi$ from the host to GPU memory in each iteration,
-which would eat most of the performance gain offered by the GPU because
+which would consume most of the performance gain offered by the GPU because
 the algorithm is so data-intensive.
@@
 % will be discussed only where appropriate.
+
 The distributed design

### 2017-06-13 gpu-05-03-accelerator.tex
NVidia iso Nvidia, small rephrase
Process Henri's comments 1june (except Maxwell/Pascal analysis for GPU strategies)

@@
 shared strategy and 8 variations of the code generation strategy. This
-section investigates the effectiveness of each flavor on Nvidia's GTX~\mbox{Titan-X}
+section investigates the effectiveness of each flavor on NVidia's GTX~\mbox{Titan-X}
 Maxwell and Pascal,
@@
   \setlength\tabcolsep{0em}
-  \begin{tabular}{c c c}
-  \epsfig{file=plots/titanx-1024-strategies-w1.eps, width=0.33\textwidth} &
-  \epsfig{file=plots/titanx-2048-strategies-w1.eps, width=0.33\textwidth} &
-  \epsfig{file=plots/k40-1024-strategies-w2.eps, width=0.33\textwidth}
+  \begin{tabular}{c c c c}
+  \epsfig{file=plots/titanx-maxwell-1024-strategies-w2.eps, width=0.25\textwidth} &
+  \epsfig{file=plots/titanx-maxwell-2048-strategies-w2.eps, width=0.25\textwidth} &
+  \epsfig{file=plots/titanx-pascal-1024-strategies-w2.eps, width=0.25\textwidth} &
+  \epsfig{file=plots/k40-1024-strategies-w2.eps, width=0.25\textwidth}
   \end{tabular}
@@
 	Other relevant model parameters: M=4096, $|\Neighbors|$=32.
-	% (a) \mbox{Titan-X} GPU, without explicit kernel vectorization;
+	% (a) \mbox{Titan-X(Pascal)} GPU, with kernel vectorization of width 2;
 	% Relevant model parameters: K=1024, M=4096, $|\Neighbors|$=32.
 	% (b) \mbox{Titan-X} GPU, without explicit kernel vectorization;
+	% Relevant model parameters: K=1024, M=4096, $|\Neighbors|$=32.
+  	% (c) \mbox{Titan-X} GPU, without explicit kernel vectorization;
 	% K=2048, M=4096, $|\Neighbors|$=32.
-	% (c) K40c GPU, with explicit kernel vectorization of width 2;
+	% (d) K40c GPU, with explicit kernel vectorization of width 2;
 	% K=1024, M=4096, $|\Neighbors|$=32.
@@
   \centering
-  \epsfig{file=plots/titanx-2048-strategies-w1.eps, width=\textwidth}
+  \epsfig{file=plots/titanx-maxwell-2048-strategies-w1.eps, width=\textwidth}
   \caption{Execution time of 1000 \textit{update\_phi} invocations using the
@@
 
-Figure \ref{gpu-sweep}(a) presents the performance of the Maxwell \mbox{Titan-X} GPU
-for all 10 strategy flavors, without explicitly vectorizing the kernels. The
+Figure \ref{gpu-sweep}(a) presents the performance of the Pascal \mbox{Titan-X} GPU
+for all 10 strategy flavors, with an explicit vector width~2 in the kernels. The
 x-axis represents different \textit{update\_phi} thread block sizes while the y-axis presents
@@
 The wide performance range inhibits the readability of the figure,
-therefore, a more focused Figure~\ref{gpu-sweep}(a), bottom figure, is provided. As would
+therefore, a zoomed-in version of Figure~\ref{gpu-sweep}(a), is provided as the bottom figure. As would
 be expected, the naive strategy exhibits the worst performance over all
@@
 Both strategies cache $Grads_i$, $Probs_i$ and $\pi_i$ in shared memory but
-differ in one aspect. Namely, \textit{GEN-SSS} explicitly unrolls the internal
+differ in one aspect: \textit{GEN-SSS} explicitly unrolls the internal
 loops of the kernel. However, there is no significant performance difference
@@
 block size of 96 and a vector width of~2, whereas
-the Maxwell Titan-X obtains the best performance with the \textit{GEN-RSS}
+the Maxwell Titan-X in Figure~\ref{gpu-sweep}(b) obtains the best performance with the \textit{GEN-RSS}
 strategy and a thread block size of 64 and no vectorization. The results of other
@@
 A key model parameter that affects the behavior of the optimization strategies
-is the number of communities $K$. Figure~\ref{gpu-sweep}(b) presents the
-same model configuration as in Figure~\ref{gpu-sweep}(a) but $K=2048$
-instead of $K=1024$. One key difference between the two figures is the optimal
+is the number of communities~$K$. Figure~\ref{gpu-sweep}(c) presents the
+same model configuration as in Figure~\ref{gpu-sweep}(b) but $K=2048$
+instead of $K=1024$. The most important difference in performance between the two figures is the optimal
 thread block size. An increase in $K$ comes with a proportional increase in
@@
 capabilities of each device as listed in Table~\ref{table-gpus}. For instance,
-the \mbox{Titan-X(Pascal)} achieves the highest speedup of 120 relative to the baseline.
+the Pascal \mbox{Titan-X} achieves the highest speedup of 120 relative to the baseline.
 

### 2017-06-13 plots/plot-all-gpu-detail.sh

@@ -1,12 +1,16 @@
 #!/bin/bash
 
-./plot-gpu-sweep.sh --title "(a) Titan-X/Pascal K=1024 no vectorization" titanx-1024-strategies-w1 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
-./plot-gpu-sweep.sh --title "(a) Titan-X/Pascal K=1024 vector width=2" titanx-1024-strategies-w2 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
-./plot-gpu-sweep.sh --title "(a) Titan-X/Pascal K=1024 vector width=4" titanx-1024-strategies-w4 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
+./plot-gpu-sweep.sh --title "(a) Titan-X(Pascal) K=1024 no vectorization" titanx-pascal-1024-strategies-w1 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
+./plot-gpu-sweep.sh --title "(a) Titan-X(Pascal) K=1024 vector width=2" titanx-pascal-1024-strategies-w2 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
+./plot-gpu-sweep.sh --title "(a) Titan-X(Pascal) K=1024 vector width=4" titanx-pascal-1024-strategies-w4 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
 
-./plot-gpu-sweep.sh --title "(a) Titan-X K=1024 no vectorization" titanx-1024-strategies-w1 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
-./plot-gpu-sweep.sh --title "(a) Titan-X K=1024 vector width=2" titanx-1024-strategies-w2 `ls dblp-K1024/GeForceGTXTITANX-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
[…]

### 2017-06-13 plots/plot-gpu-sweep.sh

@@ -89,7 +89,7 @@ plot \
 set origin col(0),row(1)
 set size w,h
 set title 'As above, zoom-in around optimal settings'
-set xrange [32:128]
+set xrange [32:$MID]
 set autoscale y
 plot \
   "< cat $TFILE | grep NAIVE  | awk '{if (\$2 <= $MID) print \$0 }'" u 2:4 w lp title 'NAIVE', \

### 2017-06-14 06-conclusion.tex
Typos in conclusion

@@
 are partitioned across the collective memory of the cluster nodes, which
-significantly increases data access latency for our already data-dominant algorithm.
-We made several contributions to improve performance and still maintain scalabitiy:
+significantly hurts data access latency and bandwidth for our already data-dominant algorithm.
+We made several contributions to improve performance and still maintain scalability:
 we created a custom distributed Key-Value store implemented with high-speed RDMA
-primitives that exploits the algorithm's structure to avoid concurrency issues; we
+primitives, that exploits the algorithm's structure to avoid concurrency issues; we
 use MPI for synchronization and work distribution between the master and worker nodes;

### 2017-06-14 gpu-05-03-accelerator.tex
Maxwell in (a) and (b), Pascal in (c)

@@
 As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the
-GPU version of the algorithm. Specifically, there is the naive strategy, the
-shared strategy and 8 variations of the code generation strategy. This
+GPU version of the algorithm: the naive strategy, the
+shared strategy, and 8 variations of the code generation strategy. This
 section investigates the effectiveness of each flavor on NVidia's GTX~\mbox{Titan-X}
-Maxwell and Pascal,
-GTX980, K40c and K20m.
+(Maxwell and Pascal),
+GTX980, K40c and~K20m.
 
@@
   \begin{tabular}{c c c c}
-  \epsfig{file=plots/titanx-maxwell-1024-strategies-w2.eps, width=0.25\textwidth} &
-  \epsfig{file=plots/titanx-maxwell-2048-strategies-w2.eps, width=0.25\textwidth} &
+  \epsfig{file=plots/titanx-maxwell-1024-strategies-w1.eps, width=0.25\textwidth} &
+  \epsfig{file=plots/titanx-maxwell-2048-strategies-w1.eps, width=0.25\textwidth} &
   \epsfig{file=plots/titanx-pascal-1024-strategies-w2.eps, width=0.25\textwidth} &
@@
 
-Figure \ref{gpu-sweep}(a) presents the performance of the Pascal \mbox{Titan-X} GPU
+Figure \ref{gpu-sweep}(a) presents the performance of the Maxwell \mbox{Titan-X} GPU
 for all 10 strategy flavors, with an explicit vector width~2 in the kernels. The
@@
 The wide performance range inhibits the readability of the figure,
-therefore, a zoomed-in version of Figure~\ref{gpu-sweep}(a), is provided as the bottom figure. As would
+therefore, a zoomed-in version of Figure~\ref{gpu-sweep}(a) is provided as the bottom figure. As would
 be expected, the naive strategy exhibits the worst performance over all
@@
 performance as they unroll internal loops as well as cache data in
-registers. The Pascal Titan-X is fastest with the \textit{GEN-RSR} strategy and a
-block size of 96 and a vector width of~2, whereas
-the Maxwell Titan-X in Figure~\ref{gpu-sweep}(b) obtains the best performance with the \textit{GEN-RSS}
+registers.
+The Maxwell Titan-X in Figure~\ref{gpu-sweep}(b) obtains the best performance with the \textit{GEN-RSS}
 strategy and a thread block size of 64 and no vectorization. The results of other
@@
 
+The Pascal \mbox{Titan-X} in Figure~\ref{gpu-sweep}(c) is the fastest of the GPUs
+under investigation, note the y-axis range. The NAIVE strategy yields relatively
+worse performance in comparison to the other GPUs. In contrast to the Maxwell
+\mbox{Titan-X}, the Pascal shows very good performance with the SHARED strategy.
+This GPU performs best with the \textit{GEN-RSR} strategy, a block size of~96
+and a vector width of~2.
+
 \begin{comment}
@@
 % \looseness=-1
-Figure~\ref{gpu-sweep}(c) presents the performance of the K40c GPU
-for the same experimental configuration as shown for the Maxwel \mbox{Titan-X} in
-Figure~\ref{gpu-sweep}(a) except that the code
-is explicitly vectorized with a vector width of~2. The results
+Figure~\ref{gpu-sweep}(d) presents the performance of the K40c GPU
+for the same experimental configuration as shown for the Pascal \mbox{Titan-X} in
+Figure~\ref{gpu-sweep}(c), again with a vector width of~2. The results
 for the other code versions with vector width 4 and no vectorization are
@@
 
-The performance results obtained from the Maxwell \mbox{Titan-X} and Tesla~K40c GPUs
+These performance results for a range of GPUs
 reinforce the importance of customizing compute kernels to each GPU's specific

### 2017-06-14 plots/plot-all-gpu-detail.sh

@@ -1,24 +1,24 @@
 #!/bin/bash
 
-./plot-gpu-sweep.sh --title "(a) Titan-X(Pascal) K=1024 no vectorization" titanx-pascal-1024-strategies-w1 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
-./plot-gpu-sweep.sh --title "(a) Titan-X(Pascal) K=1024 vector width=2" titanx-pascal-1024-strategies-w2 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
-./plot-gpu-sweep.sh --title "(a) Titan-X(Pascal) K=1024 vector width=4" titanx-pascal-1024-strategies-w4 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
+./plot-gpu-sweep.sh --title "(c) Titan-X(Pascal) K=1024 no vectorization" titanx-pascal-1024-strategies-w1 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
+./plot-gpu-sweep.sh --title "(c) Titan-X(Pascal) K=1024 vector width=2" titanx-pascal-1024-strategies-w2 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
+./plot-gpu-sweep.sh --title "(c) Titan-X(Pascal) K=1024 vector width=4" titanx-pascal-1024-strategies-w4 `ls dblp-K1024/TITANX-Pascal-K1024-M4096-N32-Snode-R0.01-A0.001-C0-*-W4*`
 
-./plot-gpu-sweep.sh --title "(b) Titan-X(Pascal) K=2048 no vectorization" titanx-pascal-2048-strategies-w1 `ls dblp-K2048/TITANX-Pascal-K2048-M4096-N32-Snode-R0.01-A0.001-C0-*-W1*`
-./plot-gpu-sweep.sh --title "(b) Titan-X(Pascal) K=2048 vector width=2" titanx-pascal-2048-strategies-w2 `ls dblp-K2048/TITANX-Pascal-K2048-M4096-N32-Snode-R0.01-A0.001-C0-*-W2*`
[…]

### 2017-06-28 03-sequential-evaluation.tex
Process Henri's comments 1june (except Maxwell/Pascal analysis for GPU strategies)
Add decimal point in table 2 to more clearly show alignment

@@
 sample neighbor sets    &      8.5  &  0.5  &  0.4  & 0.35 \\
-update\_phi             &  9,014    & 52.7  &  8.9  & 4.9  \\
+update\_phi             &  9,014.   & 52.7  &  8.9  & 4.9  \\
 update\_pi              &     93.1  & 40.3  &  0.11 & 0.05 \\
-update\_theta\_beta     &    339    &  2.1  &  0.76 & 0.55 \\
-perplexity*             &    178    &  0.18 &  0.18 & 0.26 \\
+update\_theta\_beta     &    339.   &  2.1  &  0.76 & 0.55 \\
+perplexity*             &    178.   &  0.18 &  0.18 & 0.26 \\
 \hline
@@
 in Table~\ref{table-python-comparison}. For larger problem instances, the
-speedup for these dominant phases grew to a factor~10. The vectorizing OpenCL
-backend described in the next Section again significantly improves the
-performance, but at the cost of substantial coding effort.
-\textbf{\emph{Question: speak about the other stages? they are pipelined.}}
+speedup for these dominant phases grew to a factor~10.
+% The vectorizing OpenCL
+% backend described in the next Section again significantly improves the
+% performance, but at the cost of substantial coding effort.
+% \textbf{\emph{Question: speak about the other stages? they are pipelined.}}
 

### 2017-06-28 02-01-background.tex
Start removing the OpenCL CPU work
Process Henri's comments 1june (except Maxwell/Pascal analysis for GPU strategies)
Process Sungjin's comments (Section 2)

--- 2017-05-12 02-background.tex
@@
 
-\section{Background}
+\section{Background and Related Work}
 \label{sec-background}
@@
 \bea
-\ta^* \law \ta + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta;\SGLDMinibatch)\right) + \xi_t, \label{eqn:sgld_update}
+\ta^* \law \ta_t + \f{\epsilon_t}{2}\left(\nabla_{\ta}\log p(\ta_t)+N\bar{g}(\ta_t;\SGLDMinibatch)\right) + \xi_t, \label{eqn:sgld_update}
 \eea
@@
 One benefit of using SGLD is that we do not need the Metropolis-Hastings (MH) accept-reject tests since the rejection probability goes to zero as the step size tends to zero.
-Although in practice we often do not reduce the stepsize to zero to obtain better mixing (thus resulting in some bias), we obtain good performance by drawing many more samples per unit time.
+Although in practice we do not reduce the step size to zero to obtain better mixing (thus resulting in some bias), we obtain good performance by drawing many more samples per unit time.
 
-SGLD originated from Langevin Monte Carlo (LMC)~\cite{girolami2011riemann}, where unlike SGLD the gradient is computed exactly by using all data points.
-Then, Metropolis-Hastings accept-reject tests are applied.
+SGLD originated from Langevin Monte Carlo (LMC)~\cite{girolami2011riemann} where
+unlike SGLD the gradient is computed exactly by using all data points and Metropolis-Hastings accept-reject tests are applied.
 Comparing to LMC, SGLD only requires to process a mini-batch $\SGLDMinibatch$ at each iteration and ignores the MH test, and thus computation complexity substantially reduces from $\cO(N)$ to $\cO(n)$.
@@
 \bea
-\ta_{k}^* \law \left| \ta_{k} + \f{\ep_t}{2} \left( \al - \ta_{k} + \f{N}{|\SGLDMinibatch|} \sum_{d \in \SGLDMinibatch} g_d(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{t} \right|, \label{eqn:sgrld_update}
+\ta_{k}^* \law \left| \ta_{k} + \f{\ep_t}{2} \left( \al - \ta_{k} + \f{N}{|\SGLDMinibatch|} \sum_{x \in \SGLDMinibatch} g_x(\ta_{k}) \right) + (\ta_{k})^\ha \xi_{t} \right|, \label{eqn:sgrld_update}
 \eea
-here $g_d(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $d\in \SGLDMinibatch$.
+here $g_x(\ta_{k})$ is the gradient of the log posterior w.r.t. $\ta_{k}$ on a data point $x\in \SGLDMinibatch$.
 
@@
 Here, the parameters $\phi$ and $\ta$ are used to re-parameterize $\pi$ and $\beta$ respectively.
-After updating $\phi$ and $\ta$, we can obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$ respectively, see Table~\ref{tbl-symbols}.
+After updating $\phi$ and $\ta$, we can obtain $\pi$ and $\beta$ by normalizing $\phi$ and $\ta$ respectively, as shown in Table~\ref{tbl-symbols}.
 In the following, we briefly sketch the iterative update steps.
@@
 \eea
-where the gradient in $\theta$ is
+where the gradient w.r.t. $\theta$ is
 \bea
@@
 \eea
-where the gradient in $\phi$ is
+where the gradient w.r.t. $\phi$ is
 \bea
@@
 \eea
-Here, $\Neighbors$ is the neighbor set for a mini-batch node, another random mini-batch of $n$ nodes sampled from \Vertices. Note that $|\Neighbors| \ll |\Vertices|=N$.
+Here, $\Neighbors$ is the neighbor set for a node in the mini-batch $\Minibatch$. $\Neighbors$ is another mini-batch of nodes sampled from \Vertices. Note that $|\Neighbors| \ll |\Vertices|=N$.
 
-As is common in machine learning problems, the edges of the graph are divided
+For evaluation, the edges of the graph are divided
 into two sets, the training set and a held-out set \Heldout.
@@
 \end{algorithm}
+
+In another publication~\cite{LiAW15},
+authors of this paper
+propose a heuristic to scale up the number of
+communities at the cost of less precision; the work in this paper considers the
+algorithm without that heuristic.
 

### 2017-06-28 01-intro.tex
Process Alex Uta's comments

@@
 
-There have been two main approaches to resolving this issue; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic software infrastructure systems which can be leveraged to construct machine learning solutions more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
-However, the best possible performance is often achieved by carefully integrating both of these approaches in a single solution.
+There have been two main approaches to solving this issue; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic software infrastructure systems which can be leveraged to construct machine learning solutions more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
+However, the best performance is often achieved by carefully integrating both of these approaches in a single solution.
 
-One such big data problem is analyzing large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of million of vertices~\cite{yang2015defining}.
-In particular, we are interested in the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, given a set of vertices, the links between them (which are usually very sparse), and the number of latent communities.
+One such big data problem is analyzing large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of millions of vertices~\cite{yang2015defining}.
+In this paper, we address the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, given a set of vertices, the links between them (which are usually very sparse), and the number of latent communities.
 A community can be seen as a densely connected group of vertices that are only sparsely connected to the rest of the network.
@@
 
-The problem of detecting overlapping communities is modeled by the mixed membership stochastic block model (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of MMSB, called assortative MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
+The problem of detecting overlapping communities is modeled by the mixed membership stochastic block model (MMSB)~\cite{airoldi2009mixed} and in this paper we investigate a variant of MMSB, called assortative MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
 The MMSB model is a probabilistic graphical model~\cite{koller2009probabilistic} that represents a convenient paradigm for modeling complex relationships between a potentially large number of random variables.
 Bayesian graphical models, where we define priors and infer posteriors over parameters, also allow us to quantify model uncertainty and facilitate model selection and averaging.
-However, an increasingly urgent question is whether these models and their inference procedures will be up to the challenge of handling very large graphs.
+However, an increasingly urgent question is whether these models and their inference procedures will be up to the challenge of handling very large graphs, since their computation and memory demands increase quickly with the graph size.
 
@@
 Starting from an existing implementation in Python/Numpy, we create an efficient
-C++ version to serve as the baseline for the parallel versions.
+single-threaded C++ version to serve as the baseline for the parallel versions.
 This already improves performance by a factor of 1000 to 1500 compared to
@@
 \item \textbf{scaling up to run on GPUs.}
-The application is not a natural fit for GPUs, because it is data-bound and
+The application is not a natural fit for GPUs, because it is data-intensive and
 irregular in its data accesses. Nevertheless, we developed aggressive
@@
 \item \textbf{scaling out to overcome the memory bounds} posed by 
-single-machine systems, to be the first to solve the largest publicly available
+single-machine systems. Our implementation is the first to solve the largest publicly available
 community graphs. % This necessitated overcoming several challenges.
@@
 a custom RDMA store.
-On top of that, we pipelined the algorithm stages by fetching
-data in advance over the network.
+On top of that, we pipelined the algorithm stages by prefetching
+data over the network.
 \begin{comment} Finally, the algorithm's computation
@@
 
+\begin{comment}
 The use of pipelining makes it pointless to combine GPU acceleration
@@
 % Hence, it is useless to improve the computation time.
+\end{comment}
 
@@
 A bird's-eye view of the work in this paper, without any technical detail, has been
-published at CCGrid-2016~\cite{10.1109/CCGrid.2016.98}.
-The results of our distributed MCMC-aMMSB implementation
+published at CCGrid-2016~\cite{10.1109/CCGrid.2016.98}. The technical description and
+evaluation of our accelerator contributions, which make up half of this paper's
+content, have not been published before.
+The results of our distributed contribution
 have been previously published at the 2016 ParLearning

### 2017-06-28 06-conclusion.tex
Process Alex Uta's comments

@@
 %
-By evaluating the system empirically using 65 cluster nodes and large data sets,
-we show that its strong and weak scalability are both good. We also compared its
+We argued that it is sensible to use multicore CPUs in stead of GPUs in this
+distributed setting, because GPUs require staging the data piecemeal into GPU
+memory, thus counter-balancing any speed gain. Moreover, the computation time is
+fully hidden by the latency of remote data accesses.
+%
+Our empirical evaluation using 65 cluster nodes and large data sets
+shows that the system exhibits good strong and weak scalability. We also compared its
 performance against a 40-core shared-memory system with 1TB of memory.
-The distributed system has far stronger performance because it has more cores and
+The distributed system has superior performance because it has more cores and
 more collective memory bandwidth, even though it must fetch its data through the network.
 % Additionally, we assessed the efficiency of the algorithm's resource utilization.
-Finally, a
-demonstration of the implementation's utility was provided by processing 6
+Finally, we provide a
+demonstration of the implementation's utility by processing 6
 large real-world data sets.

### 2017-06-28 gpu-05-03-accelerator.tex
Process Alex Uta's comments

@@
 Figure~\ref{gpus-fig} compares the highest speedup achieved by each of the
-GTX~\mbox{Titan-X} Maxwell and Pascal, GTX980, K40c and K20m relative to the baseline
-sequential C++ version. These results are consistent with the relative
+GTX~\mbox{Titan-X} Maxwell and Pascal, GTX980, K40c and K20m relative to the
+single-threaded baseline C++ version. These results are consistent with the relative
 capabilities of each device as listed in Table~\ref{table-gpus}. For instance,

### 2017-07-17 01-intro.tex
Comments Henri on what's new&unpublished

@@
 A bird's-eye view of the work in this paper, without any technical detail, has been
-published at CCGrid-2016~\cite{10.1109/CCGrid.2016.98}. The technical description and
-evaluation of our accelerator contributions, which make up half of this paper's
-content, have not been published before.
+published at CCGrid-2016~\cite{10.1109/CCGrid.2016.98}. The technical descriptions
+in Section~\ref{sec-gpu-design} and evaluation of our accelerator contributions in
+Section~\ref{sec-gpu-evaluation} and Figures \ref{gpu-sweep}, \ref{gpus-fig}
+and~\ref{gpus-v-sweep} have not been published before.
 The results of our distributed contribution
@@
 workshop~\cite{DBLP:conf/ipps/El-HelwHLAWB16}, where it earned a best-paper
-award.
+award. In summary, nearly half of the work in this paper consists of new, unpublished
+material.
 

### 2017-07-17 gpu-05-03-accelerator.tex
Comments Henri on what's new&unpublished

@@
 \subsubsection{Analysis of GPU Parallelism}
+\label{sec-gpu-analysis}
+
 As discussed in Section~\ref{gpu-design}, there are 10 distinct flavors of the

### 2017-07-18 01-intro.tex
Process comments by Kees V and Ceriel (but don't oblige some of them)

@@
 with an opportunity to extract hidden but useful, even invaluable information.
-Realizing this opportunity, however, requires a significant amount of effort because
+Realizing this opportunity, however, requires significant efforts because
 traditional machine learning algorithms may become inefficient with large amounts of data.
@@
 
-One such big data problem is analyzing large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of millions of vertices~\cite{yang2015defining}.
+One big data problem is analysis of large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of millions of vertices~\cite{yang2015defining}.
 In this paper, we address the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, given a set of vertices, the links between them (which are usually very sparse), and the number of latent communities.
@@
 
-Starting from an existing implementation in Python/Numpy, we create an efficient
+Starting from an existing implementation in Python/Numpy, we created an efficient
 single-threaded C++ version to serve as the baseline for the parallel versions.

### 2017-07-18 02-01-background.tex
Process comments by Kees V and Ceriel (but don't oblige some of them)

@@
 
-\looseness=-1
 Stochastic gradient Riemannian Langevin dynamics (SGRLD)~\cite{patterson2013stochastic} is a subclass of SGLD which is developed to efficiently sample from the probability simplex.
@@
 g_{ab}(\phi_{ak}) = \f{f_{ab}^{(y)}(k)}{Z_{ab}^{(y)}\phi_{ak}} - \f{1}{\sum_j\phi_{aj}}.
+\label{eqn:local_update_grad}
 \eea

### 2017-07-18 03-sequential-evaluation.tex
Process comments by Kees V and Ceriel (but don't oblige some of them)

@@
 
-In all implementations, \textit{update\_phi} dominates the computation.
+In all implementations, \textit{update\_phi} as given in Equations \ref{eqn:local_update}
+and \ref{eqn:local_update_grad} dominates the computation.
 The translation from Python into 'Python Idiom' C++ speeds up this stage
@@
 in Table~\ref{table-python-comparison}. For larger problem instances, the
-speedup for these dominant phases grew to a factor~10.
+speedup for these dominant phases grew to a factor~10 on a DAS5 machine.
 % The vectorizing OpenCL

### 2017-07-18 06-conclusion.tex
Process comments by Kees V and Ceriel (but don't oblige some of them)

@@
 analysis of the computational patterns and data structures,
-we showed that the algorithm's state can be reduced by 75\%,
+we showed that the algorithm's state size can be reduced by 75\%,
 thus significantly
@@
 We argued that it is sensible to use multicore CPUs in stead of GPUs in this
-distributed setting, because GPUs require staging the data piecemeal into GPU
+distributed setting, because GPUs require piecemeal staging of the data into GPU
 memory, thus counter-balancing any speed gain. Moreover, the computation time is

### 2017-07-18 gpu-04-03-accelerator.tex
Process Alex Uta's comments
Comments Henri on what's new&unpublished
Process comments by Kees V and Ceriel (but don't oblige some of them)

@@
 \subsection{Design of GPU Optimizations}
+\label{sec-gpu-design}
 
@@
 uses two hash functions to index their corresponding spaces. Additionally, the
-cuckoo hash supports upt to 4~keys per bucket. This hash design allows for a load
+cuckoo hash supports up to 4~keys per bucket. This hash design allows for a load
 factor upwards of 90\%, a very economical memory usage.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Parallelization \& Data Dependencies}
+\subsubsection{Parallelization and Data Dependencies}
 % Division into kernels
@@
 reduces
-the computation intensity besides the data intensity.
+the computation intensity as well as the data intensity.
 

### 2017-07-18 gpu-05-05-perplexity.tex
Process comments by Kees V and Ceriel (but don't oblige some of them)

@@
 the solution quality, as a function of time for
-two datasets, whose properties are given in Table~\ref{table-snap}. This
-analysis shows that our implementations of the algorithm indeed achieve
+the DBLP dataset whose properties is given in Table~\ref{table-snap}. This
+analysis shows that our implementation of the algorithm indeed achieves
 convergence, and it gives an impression of the time required to achieve
@@
 
-Figure~\ref{fig-ppx-gpu-cpu}(a) shows the perplexity development for the dataset
+Figure~\ref{fig-ppx-gpu-cpu} shows the perplexity development for the dataset
 com-DBLP, with $N=317,080$. It is computed on the Maxwell \mbox{Titan-X} GPU with input

### 2017-07-20 01-intro.tex
Polish newness for journal, author bios. Make system descr a subsubsection

@@
 A bird's-eye view of the work in this paper, without any technical detail, has been
-published at CCGrid-2016~\cite{10.1109/CCGrid.2016.98}. The technical descriptions
-in Section~\ref{sec-gpu-design} and evaluation of our accelerator contributions in
-Section~\ref{sec-gpu-evaluation} and Figures \ref{gpu-sweep}, \ref{gpus-fig}
+published at CCGrid-2016~\cite{10.1109/CCGrid.2016.98}. The technical
+description and evaluation of our contributions
+in Sections \ref{sec-design-sequential} and~\ref{sec-gpu},
+Table~\ref{table-python-comparison}
+and Figures \ref{gpu-sweep}, \ref{gpus-fig}
 and~\ref{gpus-v-sweep} have not been published before.
@@
 workshop~\cite{DBLP:conf/ipps/El-HelwHLAWB16}, where it earned a best-paper
-award. In summary, nearly half of the work in this paper consists of new, unpublished
+award. In summary, about 40\% of the work in this paper consists of new, unpublished
 material.

### 2017-07-22 01-intro.tex
Rewrite 2nd bulleted contrib in intro

@@
 \item \textbf{scaling out to overcome the memory bounds} posed by
-single-machine systems. Our implementation is the first to solve the largest publicly available
-community graphs. % This necessitated overcoming several challenges.
-% The algorithm's state grows rapidly with larger graphs and number of latent
-% communities.
-Since the full state for the larger graphs does not fit in a
-single machine's memory, it is partitioned across a cluster
-of machines. Hence, the cluster nodes must read
-remote memory hosted by their peers. Since the algorithm is data-bound, we
-substantially optimized remote memory accesses by implementing
-a custom RDMA store.
+GPUs which are highly effective when the problem state
+fits in GPU memory. In order to address larger graphs and numbers of communities,
+we created another implementation
+that partitions the state across a cluster of machines. This allows processing
+much larger problem instances at the cost of (memory) speed, because the
+cluster nodes must read remote memory hosted by their peers. Since the
+algorithm is data-bound, we substantially optimized remote memory accesses
+by implementing a custom RDMA store.
 On top of that, we pipelined the algorithm stages by prefetching
@@
 \end{comment}
+This implementation is the first to solve the largest publicly available
+community graphs.
 

### 2017-07-24 plots/plot-cpu.sh
Do a version test on gnuplot for compatibility

@@ -4,7 +4,7 @@ cat <<EOF | gnuplot --persist
 
 set terminal postscript eps enhanced color font 'Helvetica,10' size 3.3,1.6
 set output 'cpu-vector.eps'
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 
 set boxwidth 0.5
 set style fill solid

### 2017-07-24 plots/plot-gpu-sweep.sh

@@ -50,7 +50,7 @@ row(x) = ((nr-x-1)*h)
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',8'
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 set size (nc*w),(nr*h)
 set output '$NAME.eps'
 set multiplot layout nr,nc

### 2017-07-24 plots/plot-gpus.sh

@@ -23,7 +23,7 @@ cat <<EOF | gnuplot --persist
 
 set terminal postscript eps enhanced color font ',12' size 3.3,1.6
 set output 'gpus.eps'
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 
 set boxwidth 0.5
 set style fill solid

### 2017-07-24 plots/plot-hpc-cloud.sh

@@ -69,7 +69,7 @@ row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',10'
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 set size (nc*w),(nr*h+k)
 set output 'hpc-cloud.eps'
 set multiplot layout nr,nc

### 2017-07-24 plots/plot-ml-curves.sh

@@ -17,7 +17,7 @@ row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',8.5'
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 set size (nc*w),(nr*h+k)
 set output 'ppx.eps'
 set multiplot layout nr,nc

### 2017-07-24 plots/plot-ppx.sh

@@ -10,7 +10,7 @@ row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',8'
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 set size (nc*w),(nr*h+k)
 set output 'ppx-gpu-cpu.eps'
 set multiplot layout nr,nc

### 2017-07-24 plots/plot-qperf.sh

@@ -1,7 +1,7 @@
 cat <<EOI | gnuplot --persist
 
 set terminal postscript eps enhanced color font ',10' size 3.3,1.6
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 set output 'qperf.eps'
 
 set style line 1 lt 1 lw 2 linecolor rgb "red"

### 2017-07-24 plots/plot-sweep-K-fixed-np.sh

@@ -23,7 +23,7 @@ grep iteration ../data/sweep-over-K-fixed-np/*/* \
 cat <<EOF | gnuplot --persist
 
 set terminal postscript eps enhanced color font ',10' size 3.3,1.6
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 
 set xlabel "Number of Communities (K)"
 set ylabel "Time (seconds)"

### 2017-07-24 plots/plot-sweep-K-proportional-np.sh

@@ -18,7 +18,7 @@ w = 0.6
 h = 0.3
 
 set terminal postscript eps enhanced color font ',10'
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 set output 'sweep-over-K-proportional-np.eps'
 set size w,h
 

### 2017-07-24 plots/plot-sweep-np-fixed-K.sh

@@ -51,7 +51,7 @@ row(x) = ((nr-x-1)*h)+k
 col(x) = x*w
 
 set terminal postscript eps enhanced color font ',10'
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 set size (nc*w),(nr*h+k)
 set output 'sweep-over-np-fixed-K.eps'
 set multiplot layout nr,nc

### 2017-07-24 plots/plot-vector-sweep.sh

@@ -9,7 +9,7 @@ TFILE="$$.tmp"
 
 cat <<EOF | gnuplot --persist
 set terminal postscript eps enhanced color font ',12' size 3.3,1.6
-set colorsequence classic
+if (GPVAL_VERSION >= 5.0) { set colorsequence classic }
 set output '$NAME'
 
 set style data histogram
